### 2023-02-04 products/manuscript/Manuscript.qmd
Part 1
Part 1 of the MADA Semester Long Project. Edited the All of my edits were made in the Manuscript.qmd file located in the manuscript folder located in the products folder.  In the manuscript.qmd file I outlined Part 1 of the project. I also edited the processingcode.R file but that will be subject to change.

---
title: "Manuscript for MADA Project"
subtitle: ""
author: Jacob Matta
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---
# Part 1

## Description of data and data source

The data is a count of residents with at least one COVID-19 vaccination by demographic in the state of Georgia released by the Georgia Department of Public Health. The data has 1962 observations. The data is uploaded weekly.

You can download the data from this link: <https://www.arcgis.com/home/item.html?id=17a8a4af79e5453a9132aaa29fb4ad22>

## Why did I choose this data?

As a graduate research assistant, I am participating in an evaluation being conducted on a COVID-19 health literacy and vaccine hesitancy program in Albany, GA. This secondary data will be used to determine whether the program will produce a significant increase in COVID-19 vaccinations within the target zipcodes of 31701, 31705 and 31707.

##Loading the Data

```{r}
#looading packages
library(readxl) #for loading Excel files
library(here) #to set paths
library(dplyr) #for description summary of data 
```

```{r}
#loading data
rawdata_location <- here::here("data","raw_data","rawdata.xlsx")
rawdata <- readxl::read_excel(rawdata_location)

#loading data code book for reference 
codebook_location <-here::here("data","raw_data","Code_book_Vaccination_Demog_.xlsx")

codebook <- readxl::read_excel(codebook_location)
```

##Summary Descriptions of Data
```{r}
dplyr::glimpse(rawdata)
dplyr::summary(rawdata)
```

## Questions/Hypotheses to be addressed
Using this data, I would like to explore whether the COVID-19 health literacy and vaccine hesitancy program in Albany, GA is causing a noticeable increase in vaccinations within the zipcodes 31701, 31705, 31707.

##Predictors and Outcomes
My goal is to stratify the data by race and age and observe the difference due to those demographics.

##Overall Goal 
Since the data is updated weekly by the Georgia Department of Public Health, I am seeking to establish an efficient method of processing and cleaning the data to produce a data set with the target zip codes. Then I intend to analyze in hopes of observing the amount of vaccinations in the given week. I can then apply that data cleaning and analyzing process to each proceeding week to compare results and examine whether there is an increase in vaccination rates within the time of the health literacy program.

##Important Notes
I have been downloading this data weekly since Septmember 17th and I will continue doing so until the end of the intervention in Albany.

### 2023-02-04 code/processing_code/processingcode.R

@@ -12,27 +12,24 @@
 ## ---- packages --------
 #load needed packages. make sure they are installed.
 library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-library(skimr) #for nice visualization of data 
 library(here) #to set paths
+library(dplyr) #for description summary of data 
 
 ## ---- loaddata --------
 #path to data
 #note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw_data","exampledata.xlsx")
+#loading data
+rawdata_location <- here::here("data","raw_data","rawdata.xlsx")
+rawdata <- readxl::read_excel(rawdata_location)
 
-#load data. 
-#note that for functions that come from specific packages (instead of base R)
-# I often specify both package and function like so
-#package::function() that's not required one could just call the function
-#specifying the package makes it clearer where the function "lives",
-#but it adds typing. You can do it either way.
-rawdata <- readxl::read_excel(data_location)
+#loading data code book for reference 
+codebook_location <-here::here("data","raw_data","Code_book_Vaccination_Demog_.xlsx")
 
-# We can look in the Codebook (second tab) for a variable explanation
-codebook <- readxl::read_excel(data_location, sheet ="Codebook")
-print(codebook)
+codebook <- readxl::read_excel(codebook_location)
+
+#Summary of data
+dplyr::glimpse(rawdata)
+dplyr::summary(rawdata)
 
 
 ## ---- exploredata --------

### 2023-02-09 products/manuscript/Manuscript.qmd
Error found by Dr. Handel Fixed
Error caused by dplyr::summary(rawdata)

@@
 ```{r}
-dplyr::glimpse(rawdata)
-dplyr::summary(rawdata)
+summary(rawdata)
 ```
@@
 ##Important Notes
-I have been downloading this data weekly since Septmember 17th and I will continue doing so until the end of the intervention in Albany.
+I have been downloading this data weekly since September 17th and I will continue doing so until the end of the intervention in Albany.

### 2023-02-09 code/processing_code/processingcode.R

@@ -28,13 +28,13 @@ codebook_location <-here::here("data","raw_data","Code_book_Vaccination_Demog_.x
 codebook <- readxl::read_excel(codebook_location)
 
 #Summary of data
-dplyr::glimpse(rawdata)
-dplyr::summary(rawdata)
+glimpse(rawdata)
+summary(rawdata)
 
 
 ## ---- exploredata --------
 #take a look at the data
-dplyr::glimpse(rawdata)
+glimpse(rawdata)
 
 #another way to look at the data
 summary(rawdata)

### 2023-02-27 Overview.qmd
Part 2

--- 2023-02-09 products/manuscript/Manuscript.qmd
@@
 ---
-title: "Manuscript for MADA Project"
-subtitle: ""
-author: Jacob Matta
-date: "`r Sys.Date()`"
-format:
-  docx:
-    toc: false
-    number-sections: true
-    highlight-style: github
-bibliography: ../dataanalysis_template_references.bib
-csl: ../apa.csl
+title: "Overview"
+format: html
+editor: visual
 ---
+
 # Part 1
+
+## Background of project
 
@@
 
-You can download the data from this link: <https://www.arcgis.com/home/item.html?id=17a8a4af79e5453a9132aaa29fb4ad22>
+You can download the data from [here](https://www.arcgis.com/home/item.html?id=17a8a4af79e5453a9132aaa29fb4ad22)
+
+A further description of the data can be found [here](https://outlookuga-my.sharepoint.com/personal/jrthapa_uga_edu/_layouts/15/onedrive.aspx?FolderCTID=0x012000EA7894D56E25AD4AA0284A71A1248429&id=%2Fpersonal%2Fjrthapa%5Fuga%5Fedu%2FDocuments%2FProject%5FHealth%20Literacy%20Program%20Evaluation)
 
@@
 
-##Loading the Data
+Loading the Data
 
@@
 #loading data
-rawdata_location <- here::here("data","raw_data","rawdata.xlsx")
+rawdata_location <- here::here("data","raw_data","Feb_16.xlsx")
 rawdata <- readxl::read_excel(rawdata_location)
@@
 
-##Summary Descriptions of Data
-```{r}
-summary(rawdata)
-```
+## Questions/Hypotheses to be addressed
 
-## Questions/Hypotheses to be addressed
 Using this data, I would like to explore whether the COVID-19 health literacy and vaccine hesitancy program in Albany, GA is causing a noticeable increase in vaccinations within the zipcodes 31701, 31705, 31707.
 
-##Predictors and Outcomes
-My goal is to stratify the data by race and age and observe the difference due to those demographics.
+Predictors and Outcomes My goal is to stratify the data by race and age and observe the difference due to those demographics.
 
-##Overall Goal 
+## Overall Goal
+
 Since the data is updated weekly by the Georgia Department of Public Health, I am seeking to establish an efficient method of processing and cleaning the data to produce a data set with the target zip codes.
@@
 
-##Important Notes
-I have been downloading this data weekly since September 17th and I will continue doing so until the end of the intervention in Albany.
+Important Notes I have been downloading this data weekly since September 17th and I will continue doing so until the end of the intervention in Albany.

### 2023-02-27 code/processing_code/processing_code.qmd
Part 2

---
title: "An example cleaning script"
author: "Andreas Handel"
date: "2023-01-03"
output: html_document
---

# Processing script

This contains the same code and comments/information as `processingcode.R`.

This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.

# Setup

Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(here) #used for setting a path 
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

# Raw Data loading

```{r}
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept17 = read_excel(rawdata_location)
glimpse(Feb_16)
```

# Codebook loading

```{r}
#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

# Summary of raw data

```{r}
glimpse(Sept_17)
summary(Sept_17)
```

# Filtering the data by Doughert County
### I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
```{r}
Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Week1 [Dougherty_Vax_Week1 < 0] <- 0  
```

# Cleaning

By inspecting the data as done above, we find some problems that need addressing:

First, there is an entry for height which says "sixty" instead of a number. Does that mean it should be a numeric 60? It somehow doesn't make sense since the weight is 60kg, which can't happen for a 60cm person (a baby). Since we don't know how to fix this, we might decide to remove the person. This "sixty" entry also turned all Height entries into characters instead of numeric. That conversion to character also means that our summary function isn't very meaningful. So let's fix that first.

```{r}
d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
                  dplyr::mutate(Height = as.numeric(Height))
skimr::skim(d1)
hist(d1$Height)
```

Now we see that there is one person with a height of 6. That could be a typo, or someone mistakenly entered their height in feet. Since we unfortunately don't know, we might need to remove this person, which we'll do here.

```{r}
d2 <- d1 %>% dplyr::mutate( Height = replace(Height, Height=="6",round(6*30.48,0)) )
skimr::skim(d2)
```

Height values seem ok now.

Now let's look at the `Weight` variable. There is a person with weight of 7000, which is impossible, and one person with missing weight. To be able to analyze the data, we'll remove those individuals as well.

```{r}
d3 <- d2 %>%  dplyr::filter(Weight != 7000) %>% tidyr::drop_na()
skimr::skim(d3)
```

Now checking the `Gender` variable. Gender should be a categorical/factor variable but is loaded as character. We can fix that with simple base R code to mix things up.

```{r}
d3$Gender <- as.factor(d3$Gender)  
skimr::skim(d3)
```

Now we see that there is another NA, but it's not `NA` from R, instead it was loaded as character and is now considered as a category. Well proceed here by removing that individual with that NA entry. Since this keeps an empty category for Gender, I'm also using droplevels() to get rid of it.

```{r}
d4 <- d3 %>% dplyr::filter( !(Gender %in% c("NA","N")) ) %>% droplevels()
skimr::skim(d4)
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- d4
```

# Save data

Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. This preserves coding like factors, characters, numeric, etc. If you save as CSV, that information would get lost. However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

### 2023-02-27 products/manuscript/Manuscript.qmd
Part 2

@@
 ---
-# Part 1
 
-## Description of data and data source
+# Abstract
 
-The data is a count of residents with at least one COVID-19 vaccination by demographic in the state of Georgia released by the Georgia Department of Public Health.
-The data has 1962 observations.
-The data is uploaded weekly.
+# Introduction
 
-You can download the data from this link: <https://www.arcgis.com/home/item.html?id=17a8a4af79e5453a9132aaa29fb4ad22>
+# Methods
 
-## Why did I choose this data?
+# Results
 
-As a graduate research assistant, I am participating in an evaluation being conducted on a COVID-19 health literacy and vaccine hesitancy program in Albany, GA.
-This secondary data will be used to determine whether the program will produce a significant increase in COVID-19 vaccinations within the target zipcodes of 31701, 31705 and 31707.
+# References
 
-##Loading the Data
-
-```{r}
-#looading packages
-library(readxl) #for loading Excel files
-library(here) #to set paths
-library(dplyr) #for description summary of data 
-```
-
-```{r}
-#loading data
-rawdata_location <- here::here("data","raw_data","rawdata.xlsx")
-rawdata <- readxl::read_excel(rawdata_location)
-
-#loading data code book for reference 
-codebook_location <-here::here("data","raw_data","Code_book_Vaccination_Demog_.xlsx")
-
-codebook <- readxl::read_excel(codebook_location)
-```
-
-##Summary Descriptions of Data
-```{r}
-summary(rawdata)
-```
-
-## Questions/Hypotheses to be addressed
-Using this data, I would like to explore whether the COVID-19 health literacy and vaccine hesitancy program in Albany, GA is causing a noticeable increase in vaccinations within the zipcodes 31701, 31705, 31707.
-
-##Predictors and Outcomes
-My goal is to stratify the data by race and age and observe the difference due to those demographics.
-
-##Overall Goal 
-Since the data is updated weekly by the Georgia Department of Public Health, I am seeking to establish an efficient method of processing and cleaning the data to produce a data set with the target zip codes.
-Then I intend to analyze in hopes of observing the amount of vaccinations in the given week.
-I can then apply that data cleaning and analyzing process to each proceeding week to compare results and examine whether there is an increase in vaccination rates within the time of the health literacy program.
-
-##Important Notes
-I have been downloading this data weekly since September 17th and I will continue doing so until the end of the intervention in Albany.
+# Discussion

### 2023-02-27 code/processing_code/processingcode.R

@@ -1,148 +1,187 @@
 ###############################
 # processing script
-#
-#this script loads the raw data, processes and cleans it 
-#and saves it as Rds file in the processed_data folder
-#
-# Note the ## ---- name ---- notation
-# This is done so one can pull in the chunks of code into the Quarto document
-# see here: https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html
 
-
-## ---- packages --------
-#load needed packages. make sure they are installed.
+#library
+library(here)
 library(readxl) #for loading Excel files
-library(here) #to set paths
-library(dplyr) #for description summary of data 
+library(dplyr) #for data processing/cleaning
+library(tidyr) #for data processing/cleaning
 
-## ---- loaddata --------
-#path to data
-#note the use of the here() package and not absolute paths
-#loading data
-rawdata_location <- here::here("data","raw_data","rawdata.xlsx")
-rawdata <- readxl::read_excel(rawdata_location)
+#loading data and looking at the data using glimpse()
+
+rawdata_location = here("data","raw_data","Sept_17.xlsx")
+Sept_17 <- read_excel(rawdata_location)
+
 
 #loading data code book for reference 
-codebook_location <-here::here("data","raw_data","Code_book_Vaccination_Demog_.xlsx")
-
-codebook <- readxl::read_excel(codebook_location)
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
 
 #Summary of data
-glimpse(rawdata)
-summary(rawdata)
+glimpse(Sept_17)
+summary(Sept_17)
 
 
-## ---- exploredata --------
-#take a look at the data
-glimpse(rawdata)
+#filtering data by county becuase the data contains all counties of Georgia and for this 
+#project I am interested in Dougherty Count, which is where the city of Albany is 
+Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")  
 
-#another way to look at the data
-summary(rawdata)
+#replacing negative values with 0
[…]

### 2023-02-27 processingcode2.R

@@ -1,0 +1,37 @@
+#determining whether the data is per 100K or single counts of vaccinations 
+
+#the count of residents in GA with at least one dose of the covid-19 vaccines is 6,935,491
+
+#the count can be found here https://outlookuga-my.sharepoint.com/personal/jrthapa_uga_edu/_layouts/15/onedrive.aspx?FolderCTID=0x012000EA7894D56E25AD4AA0284A71A1248429&id=%2Fpersonal%2Fjrthapa%5Fuga%5Fedu%2FDocuments%2FProject%5FHealth%20Literacy%20Program%20Evaluation
+
+#library 
+library(here)
+library(readxl) #for loading Excel files
+library(dplyr) #for data processing/cleaning
+library(tidyr) #for data processing/cleaning
+
+#loading data and looking at the data using glimpse()
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
+
+#dropping columns that i don't need
+#keeping county name to group data together 
+Feb_26 <- Feb_26[-c(1:3)]
+
+#removing all of the population data 
+Feb_26 = Feb_26[-c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 37, 39, 41, 43, 45, 47, 49, 51, 53, 55, 57, 59, 61, 63, 65, 67, 69, 71, 73, 75, 77, 79, 81, 83, 85, 87, 89, 91, 93, 95, 97, 99, 101, 103, 105, 107, 109, 111, 113, 115, 117, 119, 121, 123, 125, 127, 129, 131, 133, 135, 137, 139, 141, 143, 145, 147, 149, 151, 153, 155, 157, 159, 161, 163, 165, 167, 169, 171, 173, 175, 177, 179, 181.)]
+Feb_26[-c(1)]
+#summing variables 
+Sum_feb26 = Feb_26 %>% rowwise() %>%  mutate(Total_vax = sum(c(UAsianNHPI_05_09_VAX)))
+
+colnames(Feb_26)
+
+## row wise sum using dplyr 
+library(dplyr)
+
+my_basket %>%
+  rowwise() %>%
+  mutate(
+    Total_price = sum(c(Price,Tax))
+  )
+

### 2023-03-06 Overview.qmd
Summing the vax totals from Feb 26th

@@
 
-A further description of the data can be found [here](https://outlookuga-my.sharepoint.com/personal/jrthapa_uga_edu/_layouts/15/onedrive.aspx?FolderCTID=0x012000EA7894D56E25AD4AA0284A71A1248429&id=%2Fpersonal%2Fjrthapa%5Fuga%5Fedu%2FDocuments%2FProject%5FHealth%20Literacy%20Program%20Evaluation)
+A further description of the data can be found [here](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c)
 

### 2023-03-06 code/processing_code/processingcode1.qmd
Summing the vax totals from Feb 26th

--- 2023-02-27 code/processing_code/processing_code.qmd
@@
 Sept17 = read_excel(rawdata_location)
-glimpse(Feb_16)
+glimpse(Sept17)
 ```
@@
 ```
-
-# Cleaning
-
-By inspecting the data as done above, we find some problems that need addressing:
-
-First, there is an entry for height which says "sixty" instead of a number.
-Does that mean it should be a numeric 60?
-It somehow doesn't make sense since the weight is 60kg, which can't happen for a 60cm person (a baby).
-Since we don't know how to fix this, we might decide to remove the person.
-This "sixty" entry also turned all Height entries into characters instead of numeric.
-That conversion to character also means that our summary function isn't very meaningful.
-So let's fix that first.
-
-```{r}
-d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
-                  dplyr::mutate(Height = as.numeric(Height))
-skimr::skim(d1)
-hist(d1$Height)
-```
-
-Now we see that there is one person with a height of 6.
-That could be a typo, or someone mistakenly entered their height in feet.
-Since we unfortunately don't know, we might need to remove this person, which we'll do here.
-
-```{r}
-d2 <- d1 %>% dplyr::mutate( Height = replace(Height, Height=="6",round(6*30.48,0)) )
-skimr::skim(d2)
-```
-
-Height values seem ok now.
-
-Now let's look at the `Weight` variable.
-There is a person with weight of 7000, which is impossible, and one person with missing weight.
-To be able to analyze the data, we'll remove those individuals as well.
-
-```{r}
-d3 <- d2 %>%  dplyr::filter(Weight != 7000) %>% tidyr::drop_na()
-skimr::skim(d3)
-```
-
-Now checking the `Gender` variable. Gender should be a categorical/factor variable but is loaded as character. We can fix that with simple base R code to mix things up.
-
-```{r}
-d3$Gender <- as.factor(d3$Gender)  
-skimr::skim(d3)
-```
-
-Now we see that there is another NA, but it's not `NA` from R, instead it was loaded as character and is now considered as a category.
-Well proceed here by removing that individual with that NA entry.
-Since this keeps an empty category for Gender, I'm also using droplevels() to get rid of it.
-
-```{r}
-d4 <- d3 %>% dplyr::filter( !(Gender %in% c("NA","N")) ) %>% droplevels()
-skimr::skim(d4)
-```
-
-All done, data is clean now.
-
-Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.
-
-```{r}
-processeddata <- d4
-```
-
-# Save data
-
-Finally, we save the clean data as RDS file.
-I suggest you save your processed and cleaned data as RDS or RDA/Rdata files.
-This preserves coding like factors, characters, numeric, etc.
-If you save as CSV, that information would get lost.
-However, CSV is better for sharing with others since it's plain text.
-If you do CSV, you might want to write down somewhere what each variable is.
-
-See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
-
-```{r}
-save_data_location <- here::here("data","processed_data","processeddata.rds")
-saveRDS(processeddata, file = save_data_location)
-```
-
-# Notes
-
-Removing anyone who had "faulty" or missing data is one approach.
-It's often not the best.
-based on your question and your analysis approach, you might want to do cleaning differently (e.g.
-keep individuals with some missing information).

### 2023-03-06 code/processing_code/processingcode2.qmd
Summing the vax totals from Feb 26th

--- 2023-03-06 code/processing_code/processingcode1.qmd
@@
 title: "An example cleaning script"
-author: "Andreas Handel"
+author: "Jacob Matta"
 date: "2023-01-03"
@@
 # Processing script
-
-This contains the same code and comments/information as `processingcode.R`.
-
-This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.
 
@@
 rawdata_location = here("data","raw_data","Sept_17.xlsx")
-Sept17 = read_excel(rawdata_location)
-glimpse(Sept17)
+Sept_17 = read_excel(rawdata_location)
+glimpse(Sept_17)
 ```

### 2023-03-06 code/processing_code/processingcode1.R

@@ -22,166 +22,87 @@ glimpse(Sept_17)
 summary(Sept_17)
 
 
-#filtering data by county becuase the data contains all counties of Georgia and for this 
+#filtering data by county becausese the data contains all counties of Georgia and for this 
 #project I am interested in Dougherty Count, which is where the city of Albany is 
 Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")  
 
 #replacing negative values with 0
 Dougherty_Vax_Week1 [Dougherty_Vax_Week1 < 0] <- 0  
 
-#organize the data. then separate into zips and add a zip code column   
-
-
+##organizing the data then separate into zips and add a zip code column   
 
 #total vax column that sums the rows across for each census tract  
-
 Dougherty_Vax_Week1 <- Dougherty_Vax_Week1 %>% rowwise() %>% mutate(total_vax_Week1 = sum(across(ends_with("VAX")), na.rm = T))  
 
-
-
 #total vax column for male and female  
-
 Dougherty_Vax_Week1 <- Dougherty_Vax_Week1  %>% rowwise() %>%   
-  
   mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
 
-
-
 Dougherty_Vax_Week1 <- Dougherty_Vax_Week1 %>% rowwise() %>%   
-  
   mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))  
 
-
-
 #total white, black, asian, other race male   
-
 Dougherty_Vax_Week1 <- Dougherty_Vax_Week1 %>% rowwise() %>%   
-  
   mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))  
 
-
-
 Dougherty_Vax_Week1 <- Dougherty_Vax_Week1 %>% rowwise() %>%   
-  
   mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))  
 
-
[…]

### 2023-03-06 code/processing_code/processingcode2.R

@@ -1,6 +1,6 @@
 #determining whether the data is per 100K or single counts of vaccinations 
 
-#the count of residents in GA with at least one dose of the covid-19 vaccines is 6,935,491
+#the count of residents in GA with at least one dose of the covid-19 vaccines is 6,935,491 on Feb_26th 2023
 
 #the count can be found here https://outlookuga-my.sharepoint.com/personal/jrthapa_uga_edu/_layouts/15/onedrive.aspx?FolderCTID=0x012000EA7894D56E25AD4AA0284A71A1248429&id=%2Fpersonal%2Fjrthapa%5Fuga%5Fedu%2FDocuments%2FProject%5FHealth%20Literacy%20Program%20Evaluation
 
@@ -21,17 +21,18 @@ Feb_26 <- Feb_26[-c(1:3)]
 #removing all of the population data 
 Feb_26 = Feb_26[-c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 37, 39, 41, 43, 45, 47, 49, 51, 53, 55, 57, 59, 61, 63, 65, 67, 69, 71, 73, 75, 77, 79, 81, 83, 85, 87, 89, 91, 93, 95, 97, 99, 101, 103, 105, 107, 109, 111, 113, 115, 117, 119, 121, 123, 125, 127, 129, 131, 133, 135, 137, 139, 141, 143, 145, 147, 149, 151, 153, 155, 157, 159, 161, 163, 165, 167, 169, 171, 173, 175, 177, 179, 181.)]
 Feb_26[-c(1)]
-#summing variables 
-Sum_feb26 = Feb_26 %>% rowwise() %>%  mutate(Total_vax = sum(c(UAsianNHPI_05_09_VAX)))
 
-colnames(Feb_26)
+#removing the negative values 
+Feb_26 [Feb_26 < 0] <- 0  
 
-## row wise sum using dplyr 
-library(dplyr)
+#Sum of multiple columns
+#summing by each column 
+Feb_26_sum_by_col = colSums(Feb_26[ , c(2:91)])     
 
-my_basket %>%
-  rowwise() %>%
-  mutate(
-    Total_price = sum(c(Price,Tax))
-  )
+Feb_26_sum_tib = tibble(Feb_26_sum_by_col)
 
+Overall_sum = sum(Feb_26_sum_tib) 
+Overall_sum
+
[…]

### 2023-03-07 code/processing_code/processingcode1.qmd
data processing file 2

@@
 ---
-title: "An example cleaning script"
-author: "Andreas Handel"
+title: "Processing Script 1"
+author: "Jacob Matta"
 date: "2023-01-03"
@@
 
-# Processing script
-
-This contains the same code and comments/information as `processingcode.R`.
-
-This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.
-
-# Setup
+# Processing script 1
 
@@
 ```{r}
-library(readxl) #for loading Excel files
-library(here) #used for setting a path 
+library(here)
 library(readxl) #for loading Excel files
@@
 
-# Raw Data loading
+### Raw Data & Codebook loading
 
 ```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
 rawdata_location = here("data","raw_data","Sept_17.xlsx")
-Sept17 = read_excel(rawdata_location)
-glimpse(Sept17)
-```
+Sept_17 <- read_excel(rawdata_location)
 
-# Codebook loading
-
-```{r}
 #loading data code book for reference 
@@
 
-# Summary of raw data
+### Summary of data
 
@@
 
-# Filtering the data by Doughert County
-### I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+### Cleaning
+
 ```{r}
+#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
 Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")

### 2023-03-07 code/processing_code/processingcode2.qmd
data processing file 2

@@
 ---
-title: "An example cleaning script"
+title: "Processing script 2"
 author: "Jacob Matta"
@@
 ---
+# Processing script 2: Understanding the rawdata and data source
 
-# Processing script
+## Determining the count of residents in GA with at least one dose of the covid-19 vaccines
 
-# Setup
+The count of residents in GA with at least one dose of the Covid-19 vaccines as of Feb 26th, 2023 is 6,935,491.
+This was found from the Georgia Department of [Public Health Vaccine Distribution Dashboard](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c) on Feb 26th/
 
-Load needed packages. make sure they are installed.
+## This processing script will look to see how the raw data used for this analysis compares to the overall number of 6,935,491 as of Feb 26th, 2023.
+
+The overall sum of residents in GA with at least one dose of the covid-19 vaccines as of Feb 26th, 2023 should equal 6,935,491
+
+### Load needed packages. make sure they are installed.
 
@@
 
-# Raw Data loading
+### Raw Data loading
 
 ```{r}
-rawdata_location = here("data","raw_data","Sept_17.xlsx")
-Sept_17 = read_excel(rawdata_location)
-glimpse(Sept_17)
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
 ```
 
-# Codebook loading
+### Codebook loading
 
@@
 
-# Summary of raw data
+### Cleaning
 
 ```{r}
-glimpse(Sept_17)
-summary(Sept_17)
+#dropping columns that i don't need
+#keeping county name to group data together 
+Feb_26 <- Feb_26[-c(1:3)]
+
+#removing all of the population data 
+Feb_26 = Feb_26[-c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 37, 39, 41, 43, 45, 47, 49, 51, 53, 55, 57, 59, 61, 63, 65, 67, 69, 71, 73, 75, 77, 79, 81, 83, 85, 87, 89, 91, 93, 95, 97, 99, 101, 103, 105, 107, 109, 111, 113, 115, 117, 119, 121, 123, 125, 127, 129, 131, 133, 135, 137, 139, 141, 143, 145, 147, 149, 151, 153, 155, 157, 159, 161, 163, 165, 167, 169, 171, 173, 175, 177, 179, 181.)]
+Feb_26[-c(1)]
+
+#removing the negative values 
+Feb_26 [Feb_26 < 0] <- 0
+
 ```
 
-# Filtering the data by Doughert County
-### I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+### Summing the vaccinations in each column
+
 ```{r}
-Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")
+Feb_26_sum_by_col = colSums(Feb_26[ , c(2:91)])     
+Feb_26_sum_by_col
+```
 
-#replacing negative values with 0
-Dougherty_Vax_Week1 [Dougherty_Vax_Week1 < 0] <- 0  
+### Turning the sums by column into a tibble that puts all of the sums of each column into one column
+
+```{r}
+Feb_26_sum_tib = tibble(Feb_26_sum_by_col)
+Feb_26_sum_tib
 ```
+
+### Summing all of the vaccinations together
+
+```{r}
+Overall_sum = sum(Feb_26_sum_tib) 
+Overall_sum
+```
+
+### Thoughts
+
+The overall total of residents in GA with at least one dose of the covid-19 vaccines is 5,654,635 which is more than 1 million should of the total about of GA residents with at least one dose of a Covid-19 vaccine that the Georgia Department of Public Health Vaccine Distribution Dashboard reported on Feb, 26th, 2023.
+
+The data that I am using for this project possibly excludes certain vaccines brands (eg.
+Pfizer, Moderna, Johnson & Johnson) whereas the overall vaccination count represented on the website includes all vaccination brands.
+This is a question I will attempt to figure out.
+For now, it is my understanding that there are 5,654,635 Ga residents by census tract who currently have one dose of the Covid-19 vaccine.
+The brands of Covid-19 vaccines included in the census tract data has yet to be specified.
+
+### Explanation for difference Covid-19 Vaccination totals
+
+1.  The data I downloaded is by census tract
+
+2.
+The data displayed on the Georgia Department of [Public Health Vaccine Distribution Dashboard](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c) on Feb 26th is by county level which does not align with census tracts
+
+3.
+Going forward for details regarding the raw data you should go to arcGIS insights dashboard for Covid-19 vaccinations among GA Residents with at least one dose found [here](https://insights.arcgis.com/#/view/08d2c1f995524053b1f9669dd0b0b995)
+
+4.
+The data dashboard for Covid-19 vaccinations among GA Residents with at least one dose (link above) states that the data downloaded by census tract will not add up to the county level and state level data
+
+### Moving forward
+If the vaccinations are not rates (eg.
+per 100,000) why are there negative values and why does it state that the where "-1 is indicated, either the numerator is less than five or the denominator is less than 25".
+-   What is the numerator and what is the denominator?

### 2023-03-07 code/processing_code/processingcode1.R

@@ -7,8 +7,13 @@ library(readxl) #for loading Excel files
 library(dplyr) #for data processing/cleaning
 library(tidyr) #for data processing/cleaning
 
-#loading data and looking at the data using glimpse()
+#library
+library(here)
+library(readxl) #for loading Excel files
+library(dplyr) #for data processing/cleaning
+library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
 
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
 rawdata_location = here("data","raw_data","Sept_17.xlsx")
 Sept_17 <- read_excel(rawdata_location)
 

### 2023-03-07 code/processing_code/processingcode2.R

@@ -1,4 +1,3 @@
-#determining whether the data is per 100K or single counts of vaccinations 
 
 #the count of residents in GA with at least one dose of the covid-19 vaccines is 6,935,491 on Feb_26th 2023
 

### 2023-03-13 code/processing_code/processingcode1.qmd
Part 2 cleaning and some analysis

@@
 ---
-title: "Processing Script 1"
+title: "Processing Code for September Data"
 author: "Jacob Matta"
@@
 
-### Raw Data & Codebook loading
+## Raw Data & Codebook loading
 
@@
 
-### Summary of data
+## Summary of data
 
@@
 glimpse(Sept_17)
-summary(Sept_17)
 ```
 
-### Cleaning
+## Cleaning
+
+### Separating the data by zip code and removing negative values
 
@@
 #by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
-Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")
+Dougherty_Vax_Sept <- filter(Sept_17, COUNTY_NAME == "Dougherty County")
 
 #replacing negative values with 0
-Dougherty_Vax_Week1 [Dougherty_Vax_Week1 < 0] <- 0  
+Dougherty_Vax_Sept [Dougherty_Vax_Sept < 0] <- 0  
 ```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_vax_Sept = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Dougherty_Vax_Sept<- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.
+
+```{r}
+#Vax total Asian Residents
+Dougherty_Vax_Sept$total_asian_vax <- Dougherty_Vax_Sept$total_masian_vax + Dougherty_Vax_Sept$total_fasian_vax
+
+#Vax total for African American Residents 
+Dougherty_Vax_Sept$total_black_vax <- Dougherty_Vax_Sept$total_mblack_vax + Dougherty_Vax_Sept$total_fblack_vax
+
+#Vax total for Other Race Residents
+Dougherty_Vax_Sept$total_orace_vax <- Dougherty_Vax_Sept$total_morace_vax + Dougherty_Vax_Sept$total_forace_vax
+
+#Vax total for White Residents
+Dougherty_Vax_Sept$total_white_vax <- Dougherty_Vax_Sept$total_mwhite_vax + Dougherty_Vax_Sept$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Dougherty_Vax_Sept_sums <- Dougherty_Vax_Sept[,c(1:4,185:204)]  
+```
+
+### Filtering the data by zip code and creating a
+
+```{r}
+#filtering data by zip: 31701  
+#All of the following FIPS or census tracts correspond to the 31701 zip code
+Zip_31701_Sept <-filter(Dougherty_Vax_Sept_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
+
+#filtering data by zip: 31705  
+#All of the following FIPS or census tracts correspond to the 31705 zip code
+Zip_31705_Sept <- filter(Dougherty_Vax_Sept_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
+
+#filtering data by zip: 31707  
+#All of the following FIPS or census tracts correspond to the 31707 zip code
+Zip_31707_Sept <- filter(Dougherty_Vax_Sept_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
+
+```
+
+### Change the column names for 31701 Sept
+
+```{r}
+colnames(Zip_31701_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")
+
+#change column names for 31705 Sept  
+colnames(Zip_31705_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")
+
+#change column names for 31707 Sept  
+colnames(Zip_31707_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")
+
+```
+
+# Cleaning is done, now lets save the processed data
+
+### Deleting excess environment objects and saving data as processed_data_sept
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Dougherty_Vax_Sept,Sept_17, codebook, Dougherty_Vax_Sept_sums)
+
+#saving clean data for 31701 to processed_data folder
+save_data_location = here("data", "processed_data", "Sept", "processed_data_31701_Sept.rds")
+
+saveRDS(Zip_31701_Sept, file = save_data_location)
+
+#saving clean data for 31705 to processed_data folder
+save_data_location2 = here("data", "processed_data", "Sept", "processed_data_31705_Sept.rds")
+
+saveRDS(Zip_31705_Sept, file = save_data_location2)
+
+#saving clean data for 31707 to processed_data folder
+save_data_location3 = here("data", "processed_data", "Sept", "processed_data_31707_Sept.rds")
+
+saveRDS(Zip_31707_Sept, file = save_data_location3)
+
+```

### 2023-03-13 code/processing_code/processingcode2.qmd
Part 2 cleaning and some analysis

--- 2023-03-06 code/processing_code/processingcode2.qmd
@@
 ---
-title: "An example cleaning script"
-author: "Jacob Matta"
-date: "2023-01-03"
-output: html_document
+title: "Processing Code for October Data"
+format: html
+editor: visual
 ---
-
-# Processing script
-
-# Setup
+# Processing script 2
 
@@
 
-# Raw Data loading
+## Raw Data & Codebook loading
 
 ```{r}
-rawdata_location = here("data","raw_data","Sept_17.xlsx")
-Sept_17 = read_excel(rawdata_location)
-glimpse(Sept_17)
-```
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Oct_28.xlsx")
+Oct_28 <- read_excel(rawdata_location)
 
-# Codebook loading
-
-```{r}
 #loading data code book for reference 
@@
 
-# Summary of raw data
+## Summary of data
 
 ```{r}
-glimpse(Sept_17)
-summary(Sept_17)
+glimpse(Oct_28)
 ```
 
-# Filtering the data by Doughert County
-### I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+## Cleaning
+
+### Separating the data by zip code and removing negative values
+
 ```{r}
-Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")
+#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
+Dougherty_Vax_Oct <- filter(Oct_28, COUNTY_NAME == "Dougherty County")
 
 #replacing negative values with 0
-Dougherty_Vax_Week1 [Dougherty_Vax_Week1 < 0] <- 0  
+Dougherty_Vax_Oct [Dougherty_Vax_Oct < 0] <- 0  
 ```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_vax_Oct = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Dougherty_Vax_Oct<- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.
+
+```{r}
+#Vax total Asian Residents
+Dougherty_Vax_Oct$total_asian_vax <- Dougherty_Vax_Oct$total_masian_vax + Dougherty_Vax_Oct$total_fasian_vax
+
+#Vax total for African American Residents 
+Dougherty_Vax_Oct$total_black_vax <- Dougherty_Vax_Oct$total_mblack_vax + Dougherty_Vax_Oct$total_fblack_vax
+
+#Vax total for Other Race Residents
+Dougherty_Vax_Oct$total_orace_vax <- Dougherty_Vax_Oct$total_morace_vax + Dougherty_Vax_Oct$total_forace_vax
+
+#Vax total for White Residents
+Dougherty_Vax_Oct$total_white_vax <- Dougherty_Vax_Oct$total_mwhite_vax + Dougherty_Vax_Oct$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Dougherty_Vax_Oct_sums <- Dougherty_Vax_Oct[,c(1:4,185:204)]  
+```
+
+### Filtering the data by zip code and creating a
+
+```{r}
+#filtering data by zip: 31701  
+#All of the following FIPS or census tracts correspond to the 31701 zip code
+Zip_31701_Oct <-filter(Dougherty_Vax_Oct_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
+
+#filtering data by zip: 31705  
+#All of the following FIPS or census tracts correspond to the 31705 zip code
+Zip_31705_Oct <- filter(Dougherty_Vax_Oct_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
+
+#filtering data by zip: 31707  
+#All of the following FIPS or census tracts correspond to the 31707 zip code
+Zip_31707_Oct <- filter(Dougherty_Vax_Oct_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
+
+```
+
+### Change the column names for 31701 Oct
+
+```{r}
+colnames(Zip_31701_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")
+
+#change column names for 31705 Oct  
+colnames(Zip_31705_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")
+
+#change column names for 31707 Oct  
+colnames(Zip_31707_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")
+
+```
+
+# Cleaning is done, now lets save the processed data
+
+### Deleting excess environment objects and saving data as processed_data_Oct
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Dougherty_Vax_Oct,Oct_28, codebook, Dougherty_Vax_Oct_sums)
+
+#saving clean data for 31701 to processed_data folder
+save_data_location = here("data", "processed_data","Oct", "processed_data_31701_Oct.rds")
+
+saveRDS(Zip_31701_Oct, file = save_data_location)
+
+#saving clean data for 31705 to processed_data folder
+save_data_location2 = here("data", "processed_data", "Oct", "processed_data_31705_Oct.rds")
+
+saveRDS(Zip_31705_Oct, file = save_data_location2)
+
+#saving clean data for 31707 to processed_data folder
+save_data_location3 = here("data", "processed_data", "Oct", "processed_data_31707_Oct.rds")
+
+saveRDS(Zip_31707_Oct, file = save_data_location3)
+
+```

### 2023-03-13 code/processing_code/processingcode3.qmd
Part 2 cleaning and some analysis

---
title: "Processing Code for November Data"
format: html
editor: visual
---

# Processing script 3

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Nov_24.xlsx")
Nov_24 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Nov_24)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Nov <- filter(Nov_24, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Nov [Dougherty_Vax_Nov < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_vax_Nov = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Nov <- Dougherty_Vax_Nov  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Nov<- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Nov$total_asian_vax <- Dougherty_Vax_Nov$total_masian_vax + Dougherty_Vax_Nov$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Nov$total_black_vax <- Dougherty_Vax_Nov$total_mblack_vax + Dougherty_Vax_Nov$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Nov$total_orace_vax <- Dougherty_Vax_Nov$total_morace_vax + Dougherty_Vax_Nov$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Nov$total_white_vax <- Dougherty_Vax_Nov$total_mwhite_vax + Dougherty_Vax_Nov$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Nov_sums <- Dougherty_Vax_Nov[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Nov <-filter(Dougherty_Vax_Nov_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Nov <- filter(Dougherty_Vax_Nov_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Nov <- filter(Dougherty_Vax_Nov_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Nov

```{r}
colnames(Zip_31701_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

#change column names for 31705 Nov  
colnames(Zip_31705_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

#change column names for 31707 Nov  
colnames(Zip_31707_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

```

# Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Nov

```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Nov,Nov_24, codebook, Dougherty_Vax_Nov_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Nov", "processed_data_31701_Nov.rds")

saveRDS(Zip_31701_Nov, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Nov", "processed_data_31705_Nov.rds")

saveRDS(Zip_31705_Nov, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Nov", "processed_data_31707_Nov.rds")

saveRDS(Zip_31707_Nov, file = save_data_location3)

```

### 2023-03-13 code/processing_code/processingcode4.qmd
Part 2 cleaning and some analysis

---
title: "Processing Code for December Data"
format: html
editor: visual
---

# Processing script 4

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Dec_29.xlsx")
Dec_29 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Dec_29)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Dec <- filter(Dec_29, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Dec [Dougherty_Vax_Dec < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_vax_Dec = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Dec <- Dougherty_Vax_Dec  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Dec<- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Dec$total_asian_vax <- Dougherty_Vax_Dec$total_masian_vax + Dougherty_Vax_Dec$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Dec$total_black_vax <- Dougherty_Vax_Dec$total_mblack_vax + Dougherty_Vax_Dec$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Dec$total_orace_vax <- Dougherty_Vax_Dec$total_morace_vax + Dougherty_Vax_Dec$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Dec$total_white_vax <- Dougherty_Vax_Dec$total_mwhite_vax + Dougherty_Vax_Dec$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Dec_sums <- Dougherty_Vax_Dec[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Dec <-filter(Dougherty_Vax_Dec_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Dec <- filter(Dougherty_Vax_Dec_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Dec <- filter(Dougherty_Vax_Dec_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Dec

```{r}
colnames(Zip_31701_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

#change column names for 31705 Dec  
colnames(Zip_31705_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

#change column names for 31707 Dec  
colnames(Zip_31707_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

```

# Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Dec

```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Dec,Dec_29, codebook, Dougherty_Vax_Dec_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Dec", "processed_data_31701_Dec.rds")

saveRDS(Zip_31701_Dec, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Dec", "processed_data_31705_Dec.rds")

saveRDS(Zip_31705_Dec, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Dec", "processed_data_31707_Dec.rds")

saveRDS(Zip_31707_Dec, file = save_data_location3)

```

### 2023-03-13 code/processing_code/processingcode5.qmd
Part 2 cleaning and some analysis

---
title: "Processing Code for January Data"
format: html
editor: visual
---

# Processing script 5

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Jan_26.xlsx")
Jan_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Jan_26)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Jan <- filter(Jan_26, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Jan [Dougherty_Vax_Jan < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_vax_Jan = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Jan <- Dougherty_Vax_Jan  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Jan<- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Jan$total_asian_vax <- Dougherty_Vax_Jan$total_masian_vax + Dougherty_Vax_Jan$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Jan$total_black_vax <- Dougherty_Vax_Jan$total_mblack_vax + Dougherty_Vax_Jan$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Jan$total_orace_vax <- Dougherty_Vax_Jan$total_morace_vax + Dougherty_Vax_Jan$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Jan$total_white_vax <- Dougherty_Vax_Jan$total_mwhite_vax + Dougherty_Vax_Jan$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Jan_sums <- Dougherty_Vax_Jan[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Jan <-filter(Dougherty_Vax_Jan_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Jan <- filter(Dougherty_Vax_Jan_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Jan <- filter(Dougherty_Vax_Jan_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Jan

```{r}
colnames(Zip_31701_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

#change column names for 31705 Jan  
colnames(Zip_31705_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

#change column names for 31707 Jan  
colnames(Zip_31707_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

```

# Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Jan
```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Jan,Jan_26, codebook, Dougherty_Vax_Jan_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Jan", "processed_data_31701_Jan.rds")

saveRDS(Zip_31701_Jan, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Jan", "processed_data_31705_Jan.rds")

saveRDS(Zip_31705_Jan, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Jan", "processed_data_31707_Jan.rds")

saveRDS(Zip_31707_Jan, file = save_data_location3)

```

### 2023-03-13 products/manuscript/Manuscript.qmd
Part 2 cleaning and some analysis

@@
     highlight-style: github
-bibliography: ../dataanalysis_template_references.bib
+bibliography: ../My_Library.bib
 csl: ../apa.csl
@@
 
-# Abstract
+# Summary
+Back in March 2020, COVID-19 began to disproportionately impact the citizens of the United States.
+Among African Americans sub-population in rural South Georgia, there was a call to increase access to health services, promote health literacy as it relates to COVID-19 and decrease vaccine hesitancy.
+One such program includes the partnership between the City of Albany, GA and Albany State University to reduce COVID-19 vaccine hesitancy in Albany through a $3.9 million grant from Health and Human Services and the Office of Minority Health.
+
+This study aims to conduct a secondary analysis of residents in Albany, GA that have received at least one dose of the COVID-19 vaccine and live in the one of the three zip codes 31701, 31705 and 31707.
+
+{{< pagebreak >}}
 
 # Introduction
+
+## General Background
+
+## Description of data and data source
+The data is a count of residents with at least one COVID-19 vaccination by demographic in the state of Georgia released by the Georgia Department of Public Health.
+The data has 1962 observations.
+The data is uploaded weekly.
+
+You can download the data from [here](https://www.arcgis.com/home/item.html?id=17a8a4af79e5453a9132aaa29fb4ad22)
+
+A further description of the data can be found [here](https://insights.arcgis.com/#/view/08d2c1f995524053b1f9669dd0b0b995)
+
+## Objective 
+I am seeking to establish an efficient method of processing and cleaning the data to produce a data set with the target zip codes.
+
+My goal is to compare monthly trends of vaccinations and examine whether there is an increase in vaccinations among African American residents during the time of the health literacy program, which is from September through April.
+
+In other words, I am looking to see if the COVID-19 vaccine hesitancy program will have any observable effects through secondary data.
+
+## Hypothesis 
+The city of Albany's population is predominately African American and by examining vaccination rates of African American residents in Albany, there will be an observable change in vaccinations over the course of the COVID-19 vaccine hesitany program.
 

### 2023-03-13 code/analysis_code/analysiscode1.R

@@ -1,0 +1,24 @@
+# Analysis script for 31701 zip code 
+
+#library
+library(here)
+library(readxl) #for loading Excel files
+library(dplyr) #for data processing/cleaning
+library(tidyr)
+
+# Loading the data
+
+## Loading September 31701
+data_location1 = here("data", "processed_data", "Sept", "processed_data_31701_sept.rds")
+prcessed_data_31701_sept = readRDS(data_location1)
+
+## Loading October 31701
+data_location1 = here("data", "processed_data", "Oct", "processed_data_31701_Oct.rds")
+prcessed_data_31701_Oct = readRDS(data_location1)
+
+## Loading December 31701
+data_location1 = here("data", "processed_data", "Oct", "processed_data_31701_Oct.rds")
+prcessed_data_31701_sept = readRDS(data_location1)
+
+filename <- file.choose()
+Canteen_clean <- readRDS(filename)

### 2023-03-13 code/processing_code/processingcode1.R

@@ -5,12 +5,6 @@
 library(here)
 library(readxl) #for loading Excel files
 library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-
-#library
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
 library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
 
 # Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
@@ -27,87 +21,179 @@ glimpse(Sept_17)
 summary(Sept_17)
 
 
-#filtering data by county becausese the data contains all counties of Georgia and for this 
+
+
+########################################################################################cleaning1
+#filtering data by county because the data contains all counties of Georgia and for this 
 #project I am interested in Dougherty Count, which is where the city of Albany is 
-Dougherty_Vax_Week1 <- filter(Sept_17, COUNTY_NAME == "Dougherty County")  
+Dougherty_Vax_Sept <- filter(Sept_17, COUNTY_NAME == "Dougherty County")  
 
 #replacing negative values with 0
-Dougherty_Vax_Week1 [Dougherty_Vax_Week1 < 0] <- 0  
+Dougherty_Vax_Sept [Dougherty_Vax_Sept < 0] <- 0  
 
+
+
+
+
+
+########################################################################################cleaning2
 ##organizing the data then separate into zips and add a zip code column   
 
-#total vax column that sums the rows across for each census tract  
-Dougherty_Vax_Week1 <- Dougherty_Vax_Week1 %>% rowwise() %>% mutate(total_vax_Week1 = sum(across(ends_with("VAX")), na.rm = T))  
+#Adding a column of the total vax for each census tract (for each observation)  
[…]

### 2023-03-13 code/processing_code/processingcode2.R

@@ -1,37 +1,198 @@
+# processing code 2 
 
-#the count of residents in GA with at least one dose of the covid-19 vaccines is 6,935,491 on Feb_26th 2023
-
-#the count can be found here https://outlookuga-my.sharepoint.com/personal/jrthapa_uga_edu/_layouts/15/onedrive.aspx?FolderCTID=0x012000EA7894D56E25AD4AA0284A71A1248429&id=%2Fpersonal%2Fjrthapa%5Fuga%5Fedu%2FDocuments%2FProject%5FHealth%20Literacy%20Program%20Evaluation
-
-#library 
+#library
 library(here)
 library(readxl) #for loading Excel files
 library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
+library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
 
-#loading data and looking at the data using glimpse()
-rawdata_location = here("data","raw_data","Feb_26.xlsx")
-Feb_26 <- read_excel(rawdata_location)
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Oct_28.xlsx")
+Oct_28 <- read_excel(rawdata_location)
 
-#dropping columns that i don't need
-#keeping county name to group data together 
-Feb_26 <- Feb_26[-c(1:3)]
 
-#removing all of the population data 
-Feb_26 = Feb_26[-c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 37, 39, 41, 43, 45, 47, 49, 51, 53, 55, 57, 59, 61, 63, 65, 67, 69, 71, 73, 75, 77, 79, 81, 83, 85, 87, 89, 91, 93, 95, 97, 99, 101, 103, 105, 107, 109, 111, 113, 115, 117, 119, 121, 123, 125, 127, 129, 131, 133, 135, 137, 139, 141, 143, 145, 147, 149, 151, 153, 155, 157, 159, 161, 163, 165, 167, 169, 171, 173, 175, 177, 179, 181.)]
-Feb_26[-c(1)]
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
 
-#removing the negative values 
[…]

### 2023-03-13 code/processing_code/processingcode3.R

@@ -1,0 +1,198 @@
+# processing code 3 
+
+#library
+library(here)
+library(readxl) #for loading Excel files
+library(dplyr) #for data processing/cleaning
+library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
+
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Nov_24.xlsx")
+Nov_24 <- read_excel(rawdata_location)
+
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+
+#Summary of data
+glimpse(Nov_24)
+summary(Nov_24)
+
+
+
+
+########################################################################################cleaning1
+#filtering data by county because the data contains all counties of Georgia and for this 
+#project I am interested in Dougherty Count, which is where the city of Albany is 
+Dougherty_Vax_Nov <- filter(Nov_24, COUNTY_NAME == "Dougherty County")  
+
+#replacing negative values with 0
+Dougherty_Vax_Nov [Dougherty_Vax_Nov < 0] <- 0  
+
+
+
+
+
+
+########################################################################################cleaning2
+##organizing the data then separate into zips and add a zip code column   
+
+#Adding a column of the total vax for each census tract (for each observation)  
+Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_vax_Nov = sum(across(ends_with("VAX")), na.rm = T))  
+
+#total vax column for male  
+Dougherty_Vax_Nov <- Dougherty_Vax_Nov  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
+
+#total vax column for female 
+Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
[…]

### 2023-03-13 code/processing_code/processingcode4.R

@@ -1,0 +1,198 @@
+#processing code 4
+
+#library
+library(here)
+library(readxl) #for loading Excel files
+library(dplyr) #for data processing/cleaning
+library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
+
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Dec_29.xlsx")
+Dec_29 <- read_excel(rawdata_location)
+
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+
+#Summary of data
+glimpse(Dec_29)
+summary(Dec_29)
+
+
+
+
+########################################################################################cleaning1
+#filtering data by county because the data contains all counties of Georgia and for this 
+#project I am interested in Dougherty Count, which is where the city of Albany is 
+Dougherty_Vax_Dec <- filter(Dec_29, COUNTY_NAME == "Dougherty County")  
+
+#replacing negative values with 0
+Dougherty_Vax_Dec [Dougherty_Vax_Dec < 0] <- 0  
+
+
+
+
+
+
+########################################################################################cleaning2
+##organizing the data then separate into zips and add a zip code column   
+
+#Adding a column of the total vax for each census tract (for each observation)  
+Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_vax_Dec = sum(across(ends_with("VAX")), na.rm = T))  
+
+#total vax column for male  
+Dougherty_Vax_Dec <- Dougherty_Vax_Dec  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
+
+#total vax column for female 
+Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))  
[…]

### 2023-03-13 code/processing_code/processingcode5.R

@@ -1,0 +1,198 @@
+#processing code 5
+
+#library
+library(here)
+library(readxl) #for loading Excel files
+library(dplyr) #for data processing/cleaning
+library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
+
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Jan_26.xlsx")
+Jan_26 <- read_excel(rawdata_location)
+
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+
+#Summary of data
+glimpse(Jan_26)
+summary(Jan_26)
+
+
+
+
+########################################################################################cleaning1
+#filtering data by county because the data contains all counties of Georgia and for this 
+#project I am interested in Dougherty Count, which is where the city of Albany is 
+Dougherty_Vax_Jan <- filter(Jan_26, COUNTY_NAME == "Dougherty County")  
+
+#replacing negative values with 0
+Dougherty_Vax_Jan [Dougherty_Vax_Jan < 0] <- 0  
+
+
+
+
+
+
+########################################################################################cleaning2
+##organizing the data then separate into zips and add a zip code column   
+
+#Adding a column of the total vax for each census tract (for each observation)  
+Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_vax_Jan = sum(across(ends_with("VAX")), na.rm = T))  
+
+#total vax column for male  
+Dougherty_Vax_Jan <- Dougherty_Vax_Jan  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
+
+#total vax column for female 
+Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))  
[…]

### 2023-03-13 code/processing_code/processingcode6.R

@@ -1,0 +1,198 @@
+#processing code 6
+
+#library
+library(here)
+library(readxl) #for loading Excel files
+library(dplyr) #for data processing/cleaning
+library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
+
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
+
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+
+#Summary of data
+glimpse(Feb_26)
+summary(Feb_26)
+
+
+
+
+########################################################################################cleaning1
+#filtering data by county because the data contains all counties of Georgia and for this 
+#project I am interested in Dougherty Count, which is where the city of Albany is 
+Dougherty_Vax_Feb <- filter(Feb_26, COUNTY_NAME == "Dougherty County")  
+
+#replacing negative values with 0
+Dougherty_Vax_Feb [Dougherty_Vax_Feb < 0] <- 0  
+
+
+
+
+
+
+########################################################################################cleaning2
+##organizing the data then separate into zips and add a zip code column   
+
+#Adding a column of the total vax for each census tract (for each observation)  
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_vax_Feb = sum(across(ends_with("VAX")), na.rm = T))  
+
+#total vax column for male  
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
+
+#total vax column for female 
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))  
[…]

### 2023-03-13 code/analysis_code/analysiscode1.R
analysis code 1
started writing code for analysis 1 where I intend to make visual for the trend of vax from sept to feb

@@ -9,16 +9,30 @@ library(tidyr)
 # Loading the data
 
 ## Loading September 31701
-data_location1 = here("data", "processed_data", "Sept", "processed_data_31701_sept.rds")
-prcessed_data_31701_sept = readRDS(data_location1)
+data_location1 = here("data", "processed_data", "Sept", "processed_data_31701_Sept.rds")
+processed_data_31701_Sept = readRDS(data_location1)
 
-## Loading October 31701
-data_location1 = here("data", "processed_data", "Oct", "processed_data_31701_Oct.rds")
-prcessed_data_31701_Oct = readRDS(data_location1)
+## Loading October 31701 and joining september and october 
+data_location2 = here("data", "processed_data", "Oct", "processed_data_31701_Oct.rds")
+processed_data_31701_Oct = readRDS(data_location2)
+processed_combined <- left_join(processed_data_31701_Sept, processed_data_31701_Oct, by = c("FIPS" = "FIPS"))
+
+## Loading November 31701
+data_location3 = here("data", "processed_data", "Nov", "processed_data_31701_Nov.rds")
+processed_data_31701_Nov = readRDS(data_location3)
 
 ## Loading December 31701
-data_location1 = here("data", "processed_data", "Oct", "processed_data_31701_Oct.rds")
-prcessed_data_31701_sept = readRDS(data_location1)
+data_location4 = here("data", "processed_data", "Dec", "processed_data_31701_Dec.rds")
+processed_data_31701_Dec = readRDS(data_location4)
+
+## Loading January 31701
+data_location5 = here("data", "processed_data", "Jan", "processed_data_31701_Jan.rds")
+processed_data_31701_Jan = readRDS(data_location5)
+
+## Loading February 31701
+data_location6 = here("data", "processed_data", "Feb", "processed_data_31701_Feb.rds")
+processed_data_31701_Feb = readRDS(data_location6)
+
 
 filename <- file.choose()
 Canteen_clean <- readRDS(filename)

### 2023-03-14 code/analysis_code/analysiscode1.R
analysis code 1

@@ -1,38 +1,70 @@
-# Analysis script for 31701 zip code 
+# Analysis script for 31701 zip code for the total African American Population
 
 #library
 library(here)
 library(readxl) #for loading Excel files
 library(dplyr) #for data processing/cleaning
 library(tidyr)
+library(ggplot2)
 
 # Loading the data
 
-## Loading September 31701
+# Loading September 31701
 data_location1 = here("data", "processed_data", "Sept", "processed_data_31701_Sept.rds")
 processed_data_31701_Sept = readRDS(data_location1)
 
-## Loading October 31701 and joining september and october 
+# Loading October 31701 and joining september and october 
 data_location2 = here("data", "processed_data", "Oct", "processed_data_31701_Oct.rds")
 processed_data_31701_Oct = readRDS(data_location2)
 processed_combined <- left_join(processed_data_31701_Sept, processed_data_31701_Oct, by = c("FIPS" = "FIPS"))
 
-## Loading November 31701
+# Loading November 31701
 data_location3 = here("data", "processed_data", "Nov", "processed_data_31701_Nov.rds")
 processed_data_31701_Nov = readRDS(data_location3)
+processed_combined <- left_join(processed_combined, processed_data_31701_Nov, by = c("FIPS" = "FIPS"))
 
-## Loading December 31701
+# Loading December 31701
 data_location4 = here("data", "processed_data", "Dec", "processed_data_31701_Dec.rds")
 processed_data_31701_Dec = readRDS(data_location4)
+processed_combined <- left_join(processed_combined, processed_data_31701_Dec, by = c("FIPS" = "FIPS"))
 
-## Loading January 31701
+# Loading January 31701
 data_location5 = here("data", "processed_data", "Jan", "processed_data_31701_Jan.rds")
 processed_data_31701_Jan = readRDS(data_location5)
+processed_combined <- left_join(processed_combined, processed_data_31701_Jan, by = c("FIPS" = "FIPS"))
 
-## Loading February 31701
+# Loading February 31701
 data_location6 = here("data", "processed_data", "Feb", "processed_data_31701_Feb.rds")
 processed_data_31701_Feb = readRDS(data_location6)
+processed_combined_31701 <- left_join(processed_combined, processed_data_31701_Feb, by = c("FIPS" = "FIPS"))
+
+# taking a look at the combined data that contains vaccinations by month for African Americans from September through February
[…]

### 2023-03-17 products/manuscript/supplement.qmd
Organization
establishing organized structure of the project

# Overview

This file describes all files in this project, the reproducibility procedure for this project, any addition information regarding the methods, results and discussion for this project.

# Description of Project Files

-   `gitignore` file: Tells Git which files and directories to ignore when you make a commit

-   `JM-MADA-Project.Rproj` file: Establishes file paths for project

-   `data` folder: Contains the `processed_data` and `raw_data` folders

    -   `processed_data` folder: Contains the data that has been processed by the processing code

        -   `Sept` folder: Contains processed data for September

            -   `processed_data_31701_Sept.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Sept.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Sept.rds`: Contains processed data for the 31707 zip code

        -   `Oct` folder: Contains processed data for October

            -   `processed_data_31701_Oct.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Oct.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Oct.rds`: Contains processed data for the 31707 zip code

        -   `Nov` folder: Contains processed data for November

            -   `processed_data_31701_Nov.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Nov.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Nov.rds`: Contains processed data for the 31707 zip code

        -   `Dec` folder: Contains processed data for December

            -   `processed_data_31701_Dec.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Dec.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Dec.rds`: Contains processed data for the 31707 zip code

        -   `Jan` folder: Contains processed data for January

            -   `processed_data_31701_Jan.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Jan.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Jan.rds`: Contains processed data for the 31707 zip code

        -   `Feb` folder: Contains processed data for February

            -   `processed_data_31701_Feb.rds`: Contains processed data for the 31701 zip code

            -   `processed_data_31705_Feb.rds`: Contains processed data for the 31705 zip code

            -   `processed_data_31707_Feb.rds`: Contains processed data for the 31707 zip code

    -   `raw_data` folder: Contains the raw data for this project

        -   `Codebook.xlsx` : Contains the codebook for the raw data

        -   `Sept_17.xlsx`: Contains raw data that was downloaded on September 17th, 2022

        -   `Oct_28.xlsx`: Contains raw data that was downloaded on October 28th, 2022

        -   `Nov_24.xlsx`: Contains raw data that was downloaded on November 24th, 2022

        -   `Dec_29.xlsx`: Contains raw data that was downloaded on December 29th, 2022

        -   `Jan_26.xlsx`: Contains raw data that was downloaded on January 26th, 2023

        -   `Feb_26.xlsx`: Contains raw data that was downloaded on February 26th, 2023

-   `code` folder: Contains `analysis_code` folder and `processing_code` folder which are used to process the raw data into processed data through a series of cleaning steps and then using the analysis code to analyze the data

    -   `analysis_code`

### 2023-03-19 products/Manuscript.qmd
Organization

---
title: "Manuscript for MADA Project"
subtitle: ""
author: Jacob Matta
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../My_Library.bib
csl: ../apa.csl
---

# Summary

Back in March 2020, COVID-19 began to disproportionately impact the citizens of the United States. Among African Americans sub-population in rural South Georgia, there was a call to increase access to health services, promote health literacy as it relates to COVID-19 and decrease vaccine hesitancy. One such program includes the partnership between the City of Albany, GA and Albany State University to reduce COVID-19 vaccine hesitancy in Albany through a $3.9 million grant from Health and Human Services and the Office of Minority Health.

This study aims to conduct a secondary analysis of residents in Albany, GA that have received at least one dose of the COVID-19 vaccine and live in the one of the three zip codes 31701, 31705 and 31707.

{{< pagebreak >}}

# Introduction

## General Background

## Description of data and data source

The data is a count of residents with at least one COVID-19 vaccination by demographic in the state of Georgia released by the Georgia Department of Public Health. The data has 1962 observations which are 1962 different census tracts in Georgia. There are 187 variables each are a different combination of gender, race and age. The data is uploaded weekly.

The data is downloaded from the Georgia Department of Public Health Distribution Dashboard ([citation for dashboard pasted here](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c)) which is a publicly available dashboard that displays the current vaccination administered and the up-to-date count of Georgia residents that have been vaccinated.

The data of interest from this dashboard is the data containing the count of residents with at least one dose of the COVID-19 vaccine by census tract by demographics. You can download the data from [the census tract demographic data](https://www.arcgis.com/home/item.html?id=17a8a4af79e5453a9132aaa29fb4ad22) and a description of the data by census tract demographic can be found on the separate dashboard found [here](https://insights.arcgis.com/#/view/08d2c1f995524053b1f9669dd0b0b995).

## Objective

I am seeking to establish an efficient method of processing and cleaning the data to produce a data set with the target zip codes.

My goal is to compare monthly trends of vaccinations and examine whether there is an increase in vaccinations among African American residents during the time of the health literacy program, which is from September through April.

In other words, I am looking to see if the COVID-19 vaccine hesitancy program will have any observable effects through secondary data.

## Hypothesis

The city of Albany's population is predominately African American and by examining vaccination rates of African American residents in Albany, there will be an observable change in vaccinations over the course of the COVID-19 vaccine hesitancy program.

# Methods

## Data Import

After downloading the data the excel file containing the vaccination counts by census tract was saved to the rawdata file within this product and uploaded to a processing file where the intent is to clean and prepare the data for analysis.

## Data Processing

In preparation for analysis, the raw data was first sub-setted by  Dougherty County, the county in which Albany, Georgia lies. This allowed me to work only with census tracts from Dougherty county. I then summed the data by demographic to get demographic totals

# Results

# References

# Discussion

### 2023-03-19 products/supplement.qmd
Organization

--- 2023-03-17 products/manuscript/supplement.qmd
@@
 
-    -   `analysis_code`
+    -   `processing_code`: Contains the code for data processing in both .R files and .qmd files
+
+        -   `processingcode1.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in September 2022
+
+        -   `processingcode1.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in September 2022
+
+        -   `processingcode2.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022
+
+        -   `processingcode2.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022
+
+        -   `processingcode3.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in November 2022
+
+        -   `processingcode3.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022
+
+        -   `processingcode4.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in December 2022
+
+        -   `processingcode4.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in December 2022
+
+        -   `processingcode5.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in January 2023
+
+        -   `processingcode5.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in January 2022
+
+        -   `processingcode6.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in February 2023
+
+        -   `processingcode6.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in February 2023
+
+        -   `readme.md`: provides a brief description of the files in `processing_code` folder
+
+    -   `analysis_code`: Contains the code for the analysis
+
+        -   `analysiscode1.R`: Contains the code an analysis of the total amount of African Americans with at least one COVID-19 vaccine in the 31701 zip code
+        -   `readme.md`: provides a brief description of the analysis files
+
+-   `products` folder: Contains files for the manuscript and supplementary material of this project
+
+    -   `apa.csl` file: APA format for citations
+
+    -   `Manuscript.qmd` file: Contains the manuscript for this project
+    -   `My_Library.bib`: Bibliography for this project
+
+    -   `readme.md`: Contains a description of the files in the `products` folder
+
+    -   `supplement.qmd`: Contains supplementart material relevent to reproducing this project and a brief description of every file in this project
+
+-   `results` folder:

### 2023-03-19 code/analysis_code/analysiscode1.qmd
Part 3

---
title: "Total Amount of Vaccinations Among African Americans in the Target Zip Codes"
format: html
editor: visual
echo: FALSE 
---

# Library

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr)
library(ggplot2)
library(reshape2) #reshaping data from wide to long format
library(plotly) #creating an interactive line plot 
```

# Analysis of the Amount of Vaccinations Among African Americans in Each Target Zip Code

------------------------------------------------------------------------

## Zip code: 31701

### Loading the processed data in order to begin analysis for the 31701 zip code

```{r}
# Loading September 31701
data_location1 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Sept.rds")
processed_data_31701_Sept = readRDS(data_location1)

# Loading October 31701 and joining september and october 
data_location2 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Oct.rds")
processed_data_31701_Oct = readRDS(data_location2)
processed_combined <- left_join(processed_data_31701_Sept, processed_data_31701_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31701
data_location3 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Nov.rds")
processed_data_31701_Nov = readRDS(data_location3)
processed_combined <- left_join(processed_combined, processed_data_31701_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31701
data_location4 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Dec.rds")
processed_data_31701_Dec = readRDS(data_location4)
processed_combined <- left_join(processed_combined, processed_data_31701_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31701
data_location5 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Jan.rds")
processed_data_31701_Jan = readRDS(data_location5)
processed_combined <- left_join(processed_combined, processed_data_31701_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31701
data_location6 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Feb.rds")
processed_data_31701_Feb = readRDS(data_location6)
processed_combined_31701 <- left_join(processed_combined, processed_data_31701_Feb, by = c("FIPS" = "FIPS"))
```

### Saving the combined data for African Americans for future use. Will be useful when stratifying the African American population by age and gender

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
saveRDS(processed_combined_31701, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all African Americans in the 31701 zip code

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of African Americans with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31701_total_black = processed_combined_31701 %>% select(starts_with("total_black_vax"))
```

### Summing each Month variable by census tract

```{r}
#Summing each column for each month 
sum(processed_combined_31701_total_black$total_black_vax_Sept_17) 
sum(processed_combined_31701_total_black$total_black_vax_Oct_28) 
sum(processed_combined_31701_total_black$total_black_vax_Nov_24) 
sum(processed_combined_31701_total_black$total_black_vax_Dec_29) 
sum(processed_combined_31701_total_black$total_black_vax_Jan_26) 
sum(processed_combined_31701_total_black$total_black_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Total = c(7797, 7780, 7755, 7773, 7792, 7817)) 
```

### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 31701 zip code 
```{r}
plot_31701_total_black_vax = Sum_by_month_31701 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot 
plot_31701_total_black_vax = plot_31701_total_black_vax + 
  ggtitle("Vaccination Totals for African Americans in the 31701 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
```

------------------------------------------------------------------------

## Zip code: 31705

### Loading the processed data in order to begin analysis for the 31705 zip code
```{r}
#Loading September 31705
data_location7 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Sept.rds")
processed_data_31705_Sept = readRDS(data_location7)

# Loading October 31705 and joining september and october 
data_location8 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Oct.rds")
processed_data_31705_Oct = readRDS(data_location8)
processed_combined <- left_join(processed_data_31705_Sept, processed_data_31705_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31705
data_location9 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Nov.rds")
processed_data_31705_Nov = readRDS(data_location9)
processed_combined <- left_join(processed_combined, processed_data_31705_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31705
data_location10= here("data", "processed_data", "Zip_31705", "processed_data_31705_Dec.rds")
processed_data_31705_Dec = readRDS(data_location10)
processed_combined <- left_join(processed_combined, processed_data_31705_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31705
data_location11 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Jan.rds")
processed_data_31705_Jan = readRDS(data_location11)
processed_combined <- left_join(processed_combined, processed_data_31705_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31705
data_location12 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Feb.rds")
processed_data_31705_Feb = readRDS(data_location12)
processed_combined_31705 <- left_join(processed_combined, processed_data_31705_Feb, by = c("FIPS" = "FIPS"))

```

### Saving the combined data for African Americans for future use 
```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
saveRDS(processed_combined_31705, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all African Americans in the 31705 zip code 
```{r}
processed_combined_31705_total_black = processed_combined_31705 %>% select(starts_with("total_black_vax"))
```

### Summing each column for each month 
```{r}

sum(processed_combined_31705_total_black$total_black_vax_Sept_17)

sum(processed_combined_31705_total_black$total_black_vax_Oct_28)

sum(processed_combined_31705_total_black$total_black_vax_Nov_24)

sum(processed_combined_31705_total_black$total_black_vax_Dec_29)

sum(processed_combined_31705_total_black$total_black_vax_Jan_26)

sum(processed_combined_31705_total_black$total_black_vax_Feb_26)
```

### Separating sums by month into a data fram consisting of two variables: Date and Total

### Date contains months September-February

### Total contains the vaccination totals for each month
```{r}
Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                                Total = c(10383, 10327, 10303, 10301, 10337, 10348)) 
```

### ggplot creating a line plot for the vaccination totals by month for the 31705 zip code African American population
```{r}
plot_31705_total_black_vax = Sum_by_month_31705 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot 
plot_31705_total_black_vax = plot_31705_total_black_vax + 
  ggtitle("Vaccination Totals for African Americans in the 31705 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
```

------------------------------------------------------------------------

## Zip code: 31707

### Loading the processed data in order to begin analysis for the 31707 zip code
```{r}
# Loading September 31707
data_location13 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Sept.rds")
processed_data_31707_Sept = readRDS(data_location13)

# Loading October 31707 and joining september and october 
data_location14 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Oct.rds")
processed_data_31707_Oct = readRDS(data_location14)
processed_combined <- left_join(processed_data_31707_Sept, processed_data_31707_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31707
data_location15 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Nov.rds")
processed_data_31707_Nov = readRDS(data_location15)
processed_combined <- left_join(processed_combined, processed_data_31707_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31707
data_location16= here("data", "processed_data", "Zip_31707", "processed_data_31707_Dec.rds")
processed_data_31707_Dec = readRDS(data_location16)
processed_combined <- left_join(processed_combined, processed_data_31707_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31707
data_location17 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Jan.rds")
processed_data_31707_Jan = readRDS(data_location17)
processed_combined <- left_join(processed_combined, processed_data_31707_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31707
data_location18 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Feb.rds")
processed_data_31707_Feb = readRDS(data_location18)
processed_combined_31707 <- left_join(processed_combined, processed_data_31707_Feb, by = c("FIPS" = "FIPS"))
```

### Saving the combined data for African Americans for future use 
```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
saveRDS(processed_combined_31707, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all African Americans in the 31707 zip code 
```{r}
processed_combined_31707_total_black = processed_combined_31707 %>% select(starts_with("total_black_vax"))

```

### Summing each column for each month 
```{r}
sum(processed_combined_31707_total_black$total_black_vax_Sept_17) 
sum(processed_combined_31707_total_black$total_black_vax_Oct_28) 
sum(processed_combined_31707_total_black$total_black_vax_Nov_24) 
sum(processed_combined_31707_total_black$total_black_vax_Dec_29) 
sum(processed_combined_31707_total_black$total_black_vax_Jan_26) 
sum(processed_combined_31707_total_black$total_black_vax_Feb_26)
```

### Separating sums by month into a data fram consisting of two variables: Date and Total 
### Date contains months September-February
### Total contains the vaccination totals for each month
```{r}
Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                                Total = c(9035, 8985, 8997, 9005, 9035, 9042))

```

###ggplot: creating a line plot for the vaccination totals by month for the 31707 zip code African American population
```{r}
plot_31707_total_black_vax = Sum_by_month_31707 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot
plot_31707_total_black_vax = plot_31707_total_black_vax + 
  ggtitle("Vaccination Totals for African Americans in the 31707 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
```

------------------------------------------------------------------------

## Comparing plots the 3 target zip codes

### Taking a look at each of the line plots for the different zip codes and comparing the total vaccinations by month in the African American population 
```{r}
plot_31701_total_black_vax
plot_31705_total_black_vax
plot_31707_total_black_vax
```

### Combining all plots into one
```{r}
## Creating a new dataframe called "Total_black_vax" by combining the vaccination totals from 31701 and 3105 by using leftjoin() by the Date variable 
Total_black_vax = left_join(Sum_by_month_31701, Sum_by_month_31705, by = c("Date" = "Date"))

```

### Combining the new dataframe "Total_black_vax" with the 31707 zip vaccination totals so all three target zip codes are in the same data frame 
```{r}
Total_black_vax = left_join(Total_black_vax, Sum_by_month_31707, by = c("Date" = "Date"))
```

### Renaming column totals to match the zip
### Haing to do this because the the total variable names are Total.x, Total.y and Total
### Need to specificy each total variable by zip using the following code
```{r}
Total_black_vax <- Total_black_vax %>% 
  rename("31701" = "Total.x") %>% 
  rename("31705" = "Total.y") %>%
  rename("31707" = "Total")
```

### Converting data to long format in order to create a line plot using ggplot
```{r}
Total_black_vax_long = reshape2::melt(Total_black_vax, id = "Date")
```

### Changing the variable name "variable" to "Zip_code" and "value" to "Vaccination_total" for clarity
### Doing this because when the data was converted to a long format the column names we changed to "variable" which refers to the zip code and "value" which refers to the amount of vaccinations
### Changing "variable" to "Zip_code" and "value" to "Vaccination_total"
```{r}
Total_black_vax_long <- Total_black_vax_long %>% 
  rename("Zip_code" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### ggplot: Making a line graph that compares the vaccination totals for each African American population in each zip code
```{r}
plot_Total_black_vax = Total_black_vax_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Zip_code, 
                                color = Zip_code)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for African Americans in Target Zip Codes") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Total_black_vax
```

### Turning ggplot into an interactive plot
```{r}
plotly_Total_black_vax = ggplotly(plot_Total_black_vax)
plotly_Total_black_vax
```

### 2023-03-19 code/processing_code/processingcode1.qmd
Part 3

@@
 #saving clean data for 31701 to processed_data folder
-save_data_location = here("data", "processed_data", "Sept", "processed_data_31701_Sept.rds")
+save_data_location = here("data", "processed_data", "Zip_31701", "processed_data_31701_Sept.rds")
 
@@
 #saving clean data for 31705 to processed_data folder
-save_data_location2 = here("data", "processed_data", "Sept", "processed_data_31705_Sept.rds")
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Sept.rds")
 
@@
 #saving clean data for 31707 to processed_data folder
-save_data_location3 = here("data", "processed_data", "Sept", "processed_data_31707_Sept.rds")
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Sept.rds")
 

### 2023-03-19 code/processing_code/processingcode2.qmd
Part 3

@@
 ---
+
 # Processing script 2
@@
 ### Deleting excess environment objects and saving data as processed_data_Oct
+
 ```{r}
@@
 #saving clean data for 31701 to processed_data folder
-save_data_location = here("data", "processed_data","Oct", "processed_data_31701_Oct.rds")
+save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Oct.rds")
 
@@
 #saving clean data for 31705 to processed_data folder
-save_data_location2 = here("data", "processed_data", "Oct", "processed_data_31705_Oct.rds")
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Oct.rds")
 
@@
 #saving clean data for 31707 to processed_data folder
-save_data_location3 = here("data", "processed_data", "Oct", "processed_data_31707_Oct.rds")
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Oct.rds")
 

### 2023-03-19 code/processing_code/processingcode3.qmd
Part 3

@@
 #saving clean data for 31701 to processed_data folder
-save_data_location = here("data", "processed_data","Nov", "processed_data_31701_Nov.rds")
+save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Nov.rds")
 
@@
 #saving clean data for 31705 to processed_data folder
-save_data_location2 = here("data", "processed_data", "Nov", "processed_data_31705_Nov.rds")
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Nov.rds")
 
@@
 #saving clean data for 31707 to processed_data folder
-save_data_location3 = here("data", "processed_data", "Nov", "processed_data_31707_Nov.rds")
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Nov.rds")
 

### 2023-03-19 code/processing_code/processingcode4.qmd
Part 3

@@
 #saving clean data for 31701 to processed_data folder
-save_data_location = here("data", "processed_data","Dec", "processed_data_31701_Dec.rds")
+save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Dec.rds")
 
@@
 #saving clean data for 31705 to processed_data folder
-save_data_location2 = here("data", "processed_data", "Dec", "processed_data_31705_Dec.rds")
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Dec.rds")
 
@@
 #saving clean data for 31707 to processed_data folder
-save_data_location3 = here("data", "processed_data", "Dec", "processed_data_31707_Dec.rds")
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Dec.rds")
 

### 2023-03-19 code/processing_code/processingcode5.qmd
Part 3

@@
 ### Deleting excess environment objects and saving data as processed_data_Jan
+
 ```{r}
@@
 #saving clean data for 31701 to processed_data folder
-save_data_location = here("data", "processed_data","Jan", "processed_data_31701_Jan.rds")
+save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Jan.rds")
 
@@
 #saving clean data for 31705 to processed_data folder
-save_data_location2 = here("data", "processed_data", "Jan", "processed_data_31705_Jan.rds")
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Jan.rds")
 
@@
 #saving clean data for 31707 to processed_data folder
-save_data_location3 = here("data", "processed_data", "Jan", "processed_data_31707_Jan.rds")
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Jan.rds")
 

### 2023-03-19 code/processing_code/processingcode6.qmd
Part 2 cleaning and some analysis
Part 3

--- 2023-03-13 code/processing_code/processingcode5.qmd
@@
 ---
-title: "Processing Code for January Data"
+title: "Processing Code for February Data"
 format: html
@@
 
-# Processing script 5
+# Processing script 6
 
@@
 # Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
-rawdata_location = here("data","raw_data","Jan_26.xlsx")
-Jan_26 <- read_excel(rawdata_location)
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
 
@@
 ```{r}
-glimpse(Jan_26)
+glimpse(Feb_26)
 ```
@@
 #by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
-Dougherty_Vax_Jan <- filter(Jan_26, COUNTY_NAME == "Dougherty County")
+Dougherty_Vax_Feb <- filter(Feb_26, COUNTY_NAME == "Dougherty County")
 
 #replacing negative values with 0
-Dougherty_Vax_Jan [Dougherty_Vax_Jan < 0] <- 0  
+Dougherty_Vax_Feb [Dougherty_Vax_Feb < 0] <- 0  
 ```
@@
 #Adding a column of the total vax for each census tract (for each observation)  
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_vax_Jan = sum(across(ends_with("VAX")), na.rm = T))
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_vax_Feb = sum(across(ends_with("VAX")), na.rm = T))
 
 #total vax column for all males
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan  %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb  %>% rowwise() %>%   
   mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for all females
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for Asian male residents 
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for African American male residents
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for other race male residents
-Dougherty_Vax_Jan<- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb<- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for White male residents 
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for Asian female residents    
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for African American female residents
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for other race female residents
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
@@
 #total vax column for White female residents 
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
   mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
@@
 #Vax total Asian Residents
-Dougherty_Vax_Jan$total_asian_vax <- Dougherty_Vax_Jan$total_masian_vax + Dougherty_Vax_Jan$total_fasian_vax
+Dougherty_Vax_Feb$total_asian_vax <- Dougherty_Vax_Feb$total_masian_vax + Dougherty_Vax_Feb$total_fasian_vax
 
 #Vax total for African American Residents 
-Dougherty_Vax_Jan$total_black_vax <- Dougherty_Vax_Jan$total_mblack_vax + Dougherty_Vax_Jan$total_fblack_vax
+Dougherty_Vax_Feb$total_black_vax <- Dougherty_Vax_Feb$total_mblack_vax + Dougherty_Vax_Feb$total_fblack_vax
 
 #Vax total for Other Race Residents
-Dougherty_Vax_Jan$total_orace_vax <- Dougherty_Vax_Jan$total_morace_vax + Dougherty_Vax_Jan$total_forace_vax
+Dougherty_Vax_Feb$total_orace_vax <- Dougherty_Vax_Feb$total_morace_vax + Dougherty_Vax_Feb$total_forace_vax
 
 #Vax total for White Residents
-Dougherty_Vax_Jan$total_white_vax <- Dougherty_Vax_Jan$total_mwhite_vax + Dougherty_Vax_Jan$total_fwhite_vax  
+Dougherty_Vax_Feb$total_white_vax <- Dougherty_Vax_Feb$total_mwhite_vax + Dougherty_Vax_Feb$total_fwhite_vax  
 ```
@@
 #summing all residents in the age range of 05-09
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
 
 #summing all residents in the age range of 10-17
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
 
 #summing all residents in the age range of 18-44
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
 
 #summing all residents in the age range of 45-64
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
 
 #summing all residents in the age range of 65+ 
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
 
@@
 #Extract necessary variables only  
-Dougherty_Vax_Jan_sums <- Dougherty_Vax_Jan[,c(1:4,185:204)]  
+Dougherty_Vax_Feb_sums <- Dougherty_Vax_Feb[,c(1:4,185:204)]  
 ```
@@
 #All of the following FIPS or census tracts correspond to the 31701 zip code
-Zip_31701_Jan <-filter(Dougherty_Vax_Jan_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
+Zip_31701_Feb <-filter(Dougherty_Vax_Feb_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
 
@@
 #All of the following FIPS or census tracts correspond to the 31705 zip code
-Zip_31705_Jan <- filter(Dougherty_Vax_Jan_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
+Zip_31705_Feb <- filter(Dougherty_Vax_Feb_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
 
@@
 #All of the following FIPS or census tracts correspond to the 31707 zip code
-Zip_31707_Jan <- filter(Dougherty_Vax_Jan_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
+Zip_31707_Feb <- filter(Dougherty_Vax_Feb_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
 
@@
 
-### Change the column names for 31701 Jan
+### Change the column names for 31701 Feb
 
 ```{r}
-colnames(Zip_31701_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")
+colnames(Zip_31701_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")
 
-#change column names for 31705 Jan  
-colnames(Zip_31705_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")
+#change column names for 31705 Feb  
+colnames(Zip_31705_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")
 
-#change column names for 31707 Jan  
-colnames(Zip_31707_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")
+#change column names for 31707 Feb  
+colnames(Zip_31707_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")
 
@@
 
-### Deleting excess environment objects and saving data as processed_data_Jan
+### Deleting excess environment objects and saving data as processed_data_Feb
+
 ```{r}
 #Only need to save processed data for the three zip codes
-rm(Dougherty_Vax_Jan,Jan_26, codebook, Dougherty_Vax_Jan_sums)
+rm(Dougherty_Vax_Feb,Feb_26, codebook, Dougherty_Vax_Feb_sums)
 
 #saving clean data for 31701 to processed_data folder
-save_data_location = here("data", "processed_data","Jan", "processed_data_31701_Jan.rds")
+save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Feb.rds")
 
-saveRDS(Zip_31701_Jan, file = save_data_location)
+saveRDS(Zip_31701_Feb, file = save_data_location)
 
 #saving clean data for 31705 to processed_data folder
-save_data_location2 = here("data", "processed_data", "Jan", "processed_data_31705_Jan.rds")
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Feb.rds")
 
-saveRDS(Zip_31705_Jan, file = save_data_location2)
+saveRDS(Zip_31705_Feb, file = save_data_location2)
 
 #saving clean data for 31707 to processed_data folder
-save_data_location3 = here("data", "processed_data", "Jan", "processed_data_31707_Jan.rds")
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Feb.rds")
 
-saveRDS(Zip_31707_Jan, file = save_data_location3)
+saveRDS(Zip_31707_Feb, file = save_data_location3)
 

### 2023-03-19 products/supplement.qmd
Part 3

@@
 
-        -   `Sept` folder: Contains processed data for September
+        -   `Zip_31701` folder: Contains processed data for the 31701 zip code
 
-            -   `processed_data_31701_Sept.rds`: Contains processed data for the 31701 zip code
+            -   `processed_data_31701_Sept.rds`: Contains processed data for the 31701 zip code during September, 2022
 
-            -   `processed_data_31705_Sept.rds`: Contains processed data for the 31705 zip code
+            -   `processed_data_31701_Oct.rds`: Contains processed data for the 31701 zip code during October, 2022
 
-            -   `processed_data_31707_Sept.rds`: Contains processed data for the 31707 zip code
+            -   `processed_data_31701_Nov.rds`: Contains processed data for the 31701 zip code during November, 2022
 
-        -   `Oct` folder: Contains processed data for October
+            -   `processed_data_31701_Dec.rds`: Contains processed data for the 31701 zip code during December, 2022
 
-            -   `processed_data_31701_Oct.rds`: Contains processed data for the 31701 zip code
+            -   `processed_data_31701_Jan.rds`: Contains processed data for the 31701 zip code during January, 2023
 
-            -   `processed_data_31705_Oct.rds`: Contains processed data for the 31705 zip code
+            -   `processed_data_31701_Feb.rds`: Contains processed data for the 31701 zip code during February, 2023
 
-            -   `processed_data_31707_Oct.rds`: Contains processed data for the 31707 zip code
+            -   `processed_combined_31701.rds`: Contains all vaccinations with all variables by month (September-February) for African Americans in the 31701 zip
 
-        -   `Nov` folder: Contains processed data for November
+        -   `Zip_31705` folder: Contains processed data for the 31705 zip code
 
-            -   `processed_data_31701_Nov.rds`: Contains processed data for the 31701 zip code
+            -   `processed_data_31705_Sept.rds`: Contains processed data for the 31705 zip code during September, 2022
 
-            -   `processed_data_31705_Nov.rds`: Contains processed data for the 31705 zip code
+            -   `processed_data_31705_Oct.rds`: Contains processed data for the 31705 zip code during October, 2022
 
-            -   `processed_data_31707_Nov.rds`: Contains processed data for the 31707 zip code
+            -   `processed_data_31705_Nov.rds`: Contains processed data for the 31705 zip code during November, 2022
 
-        -   `Dec` folder: Contains processed data for December
+            -   `processed_data_31705_Dec.rds`: Contains processed data for the 31705 zip code during December, 2022
 
-            -   `processed_data_31701_Dec.rds`: Contains processed data for the 31701 zip code
+            -   `processed_data_31705_Jan.rds`: Contains processed data for the 31705 zip code during January, 2022
 
-            -   `processed_data_31705_Dec.rds`: Contains processed data for the 31705 zip code
+            -   `processed_data_31705_Feb.rds`: Contains processed data for the 31705 zip code during Feb ruary, 2022
 
-            -   `processed_data_31707_Dec.rds`: Contains processed data for the 31707 zip code
+        -   `Zip_31707` folder: Contains processed data for the 31707 zip code
 
-        -   `Jan` folder: Contains processed data for January
+            -   `processed_data_31707_Sept.rds`: Contains processed data for the 31707 zip code during September, 2022
 
-            -   `processed_data_31701_Jan.rds`: Contains processed data for the 31701 zip code
+            -   `processed_data_31707_Oct.rds`: Contains processed data for the 31707 zip code during October, 2022
 
-            -   `processed_data_31705_Jan.rds`: Contains processed data for the 31705 zip code
+            -   `processed_data_31707_Nov.rds`: Contains processed data for the 31707 zip code during November, 2022
 
-            -   `processed_data_31707_Jan.rds`: Contains processed data for the 31707 zip code
+            -   `processed_data_31707_Dec.rds`: Contains processed data for the 31707 zip code during December, 2022
 
-        -   `Feb` folder: Contains processed data for February
+            -   `processed_data_31707_Jan.rds`: Contains processed data for the 31707 zip code during January, 2023
 
-            -   `processed_data_31701_Feb.rds`: Contains processed data for the 31701 zip code
-
-            -   `processed_data_31705_Feb.rds`: Contains processed data for the 31705 zip code
-
-            -   `processed_data_31707_Feb.rds`: Contains processed data for the 31707 zip code
+            -   `processed_data_31707_Feb.rds`: Contains processed data for the 31707 zip code during February, 2023
 
@@
 
-        -   `analysiscode1.R`: Contains the code an analysis of the total amount of African Americans with at least one COVID-19 vaccine in the 31701 zip code
+        -   `analysiscode1.qmd`: Contains the code an analysis of the total amount of African Americans with at least one COVID-19 vaccine in all three target zip codes
+        -   `analysiscode1.html`: Contains the output document from rendering `analysiscode1.qmd`
         -   `readme.md`: provides a brief description of the analysis files
@@
     -   `Manuscript.qmd` file: Contains the manuscript for this project
+
     -   `My_Library.bib`: Bibliography for this project
@@
 
--   `results` folder:
+-   `results` folder: Contains the graphs produced from rendering the analysis `analysiscode1.qmd`
+
+    -   `Black_Vax_Total.png`: Contains a line graph with all three zip codes plotted for the African American populations
+
+    -   `Black_Vax_Total_31701.png` : A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31701 zip code
+
+    -   `Black_Vax_Total_31705.png` : A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31705 zip code
+
+    -   `Black_Vax_Total_31707.png`: A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31707 zip code

### 2023-03-19 code/analysis_code/analysiscode1.R

@@ -1,70 +1,0 @@
-# Analysis script for 31701 zip code for the total African American Population
-
-#library
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr)
-library(ggplot2)
-
-# Loading the data
-
-# Loading September 31701
-data_location1 = here("data", "processed_data", "Sept", "processed_data_31701_Sept.rds")
-processed_data_31701_Sept = readRDS(data_location1)
-
-# Loading October 31701 and joining september and october 
-data_location2 = here("data", "processed_data", "Oct", "processed_data_31701_Oct.rds")
-processed_data_31701_Oct = readRDS(data_location2)
-processed_combined <- left_join(processed_data_31701_Sept, processed_data_31701_Oct, by = c("FIPS" = "FIPS"))
-
-# Loading November 31701
-data_location3 = here("data", "processed_data", "Nov", "processed_data_31701_Nov.rds")
-processed_data_31701_Nov = readRDS(data_location3)
-processed_combined <- left_join(processed_combined, processed_data_31701_Nov, by = c("FIPS" = "FIPS"))
-
-# Loading December 31701
-data_location4 = here("data", "processed_data", "Dec", "processed_data_31701_Dec.rds")
-processed_data_31701_Dec = readRDS(data_location4)
-processed_combined <- left_join(processed_combined, processed_data_31701_Dec, by = c("FIPS" = "FIPS"))
-
-# Loading January 31701
-data_location5 = here("data", "processed_data", "Jan", "processed_data_31701_Jan.rds")
-processed_data_31701_Jan = readRDS(data_location5)
-processed_combined <- left_join(processed_combined, processed_data_31701_Jan, by = c("FIPS" = "FIPS"))
-
-# Loading February 31701
-data_location6 = here("data", "processed_data", "Feb", "processed_data_31701_Feb.rds")
-processed_data_31701_Feb = readRDS(data_location6)
-processed_combined_31701 <- left_join(processed_combined, processed_data_31701_Feb, by = c("FIPS" = "FIPS"))
-
-# taking a look at the combined data that contains vaccinations by month for African Americans from September through February
-View(processed_combined)
-processed_combined_31701
-
-# saving the combined data for African Americans for future use 
-save_data_location = here("data", "processed_data","Sept", "processed_combined_31701.rds")
[…]

### 2023-03-19 code/processing_code/processingcode1.R

@@ -1,199 +1,0 @@
-###############################
-# processing script
-
-#library
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
-
-# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
-rawdata_location = here("data","raw_data","Sept_17.xlsx")
-Sept_17 <- read_excel(rawdata_location)
-
-
-#loading data code book for reference 
-codebook_location = here("data","raw_data","Codebook.xlsx")
-codebook = read_excel(codebook_location)
-
-#Summary of data
-glimpse(Sept_17)
-summary(Sept_17)
-
-
-
-
-########################################################################################cleaning1
-#filtering data by county because the data contains all counties of Georgia and for this 
-#project I am interested in Dougherty Count, which is where the city of Albany is 
-Dougherty_Vax_Sept <- filter(Sept_17, COUNTY_NAME == "Dougherty County")  
-
-#replacing negative values with 0
-Dougherty_Vax_Sept [Dougherty_Vax_Sept < 0] <- 0  
-
-
-
-
-
-
-########################################################################################cleaning2
-##organizing the data then separate into zips and add a zip code column   
-
-#Adding a column of the total vax for each census tract (for each observation)  
-Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_vax_Sept = sum(across(ends_with("VAX")), na.rm = T))  
-
-#total vax column for male  
-Dougherty_Vax_Sept <- Dougherty_Vax_Sept  %>% rowwise() %>%   
-  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
-
-#total vax column for female 
-Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
[…]

### 2023-03-19 code/processing_code/processingcode2.R

@@ -1,198 +1,0 @@
-# processing code 2 
-
-#library
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
-
-# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
-rawdata_location = here("data","raw_data","Oct_28.xlsx")
-Oct_28 <- read_excel(rawdata_location)
-
-
-#loading data code book for reference 
-codebook_location = here("data","raw_data","Codebook.xlsx")
-codebook = read_excel(codebook_location)
-
-#Summary of data
-glimpse(Oct_28)
-summary(Oct_28)
-
-
-
-
-########################################################################################cleaning1
-#filtering data by county because the data contains all counties of Georgia and for this 
-#project I am interested in Dougherty Count, which is where the city of Albany is 
-Dougherty_Vax_Oct <- filter(Oct_28, COUNTY_NAME == "Dougherty County")  
-
-#replacing negative values with 0
-Dougherty_Vax_Oct [Dougherty_Vax_Oct < 0] <- 0  
-
-
-
-
-
-
-########################################################################################cleaning2
-##organizing the data then separate into zips and add a zip code column   
-
-#Adding a column of the total vax for each census tract (for each observation)  
-Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_vax_Oct = sum(across(ends_with("VAX")), na.rm = T))  
-
-#total vax column for male  
-Dougherty_Vax_Oct <- Dougherty_Vax_Oct  %>% rowwise() %>%   
-  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
-
-#total vax column for female 
-Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
[…]

### 2023-03-19 code/processing_code/processingcode3.R

@@ -1,198 +1,0 @@
-# processing code 3 
-
-#library
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
-
-# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
-rawdata_location = here("data","raw_data","Nov_24.xlsx")
-Nov_24 <- read_excel(rawdata_location)
-
-
-#loading data code book for reference 
-codebook_location = here("data","raw_data","Codebook.xlsx")
-codebook = read_excel(codebook_location)
-
-#Summary of data
-glimpse(Nov_24)
-summary(Nov_24)
-
-
-
-
-########################################################################################cleaning1
-#filtering data by county because the data contains all counties of Georgia and for this 
-#project I am interested in Dougherty Count, which is where the city of Albany is 
-Dougherty_Vax_Nov <- filter(Nov_24, COUNTY_NAME == "Dougherty County")  
-
-#replacing negative values with 0
-Dougherty_Vax_Nov [Dougherty_Vax_Nov < 0] <- 0  
-
-
-
-
-
-
-########################################################################################cleaning2
-##organizing the data then separate into zips and add a zip code column   
-
-#Adding a column of the total vax for each census tract (for each observation)  
-Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_vax_Nov = sum(across(ends_with("VAX")), na.rm = T))  
-
-#total vax column for male  
-Dougherty_Vax_Nov <- Dougherty_Vax_Nov  %>% rowwise() %>%   
-  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
-
-#total vax column for female 
-Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
[…]

### 2023-03-19 code/processing_code/processingcode4.R

@@ -1,198 +1,0 @@
-#processing code 4
-
-#library
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
-
-# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
-rawdata_location = here("data","raw_data","Dec_29.xlsx")
-Dec_29 <- read_excel(rawdata_location)
-
-
-#loading data code book for reference 
-codebook_location = here("data","raw_data","Codebook.xlsx")
-codebook = read_excel(codebook_location)
-
-#Summary of data
-glimpse(Dec_29)
-summary(Dec_29)
-
-
-
-
-########################################################################################cleaning1
-#filtering data by county because the data contains all counties of Georgia and for this 
-#project I am interested in Dougherty Count, which is where the city of Albany is 
-Dougherty_Vax_Dec <- filter(Dec_29, COUNTY_NAME == "Dougherty County")  
-
-#replacing negative values with 0
-Dougherty_Vax_Dec [Dougherty_Vax_Dec < 0] <- 0  
-
-
-
-
-
-
-########################################################################################cleaning2
-##organizing the data then separate into zips and add a zip code column   
-
-#Adding a column of the total vax for each census tract (for each observation)  
-Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_vax_Dec = sum(across(ends_with("VAX")), na.rm = T))  
-
-#total vax column for male  
-Dougherty_Vax_Dec <- Dougherty_Vax_Dec  %>% rowwise() %>%   
-  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
-
-#total vax column for female 
-Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
-  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))  
[…]

### 2023-03-19 code/processing_code/processingcode5.R

@@ -1,198 +1,0 @@
-#processing code 5
-
-#library
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
-
-# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
-rawdata_location = here("data","raw_data","Jan_26.xlsx")
-Jan_26 <- read_excel(rawdata_location)
-
-
-#loading data code book for reference 
-codebook_location = here("data","raw_data","Codebook.xlsx")
-codebook = read_excel(codebook_location)
-
-#Summary of data
-glimpse(Jan_26)
-summary(Jan_26)
-
-
-
-
-########################################################################################cleaning1
-#filtering data by county because the data contains all counties of Georgia and for this 
-#project I am interested in Dougherty Count, which is where the city of Albany is 
-Dougherty_Vax_Jan <- filter(Jan_26, COUNTY_NAME == "Dougherty County")  
-
-#replacing negative values with 0
-Dougherty_Vax_Jan [Dougherty_Vax_Jan < 0] <- 0  
-
-
-
-
-
-
-########################################################################################cleaning2
-##organizing the data then separate into zips and add a zip code column   
-
-#Adding a column of the total vax for each census tract (for each observation)  
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_vax_Jan = sum(across(ends_with("VAX")), na.rm = T))  
-
-#total vax column for male  
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan  %>% rowwise() %>%   
-  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
-
-#total vax column for female 
-Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
-  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))  
[…]

### 2023-03-19 code/processing_code/processingcode6.R

@@ -1,198 +1,0 @@
-#processing code 6
-
-#library
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning and looking at the data using glimpse()
-
-# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
-rawdata_location = here("data","raw_data","Feb_26.xlsx")
-Feb_26 <- read_excel(rawdata_location)
-
-
-#loading data code book for reference 
-codebook_location = here("data","raw_data","Codebook.xlsx")
-codebook = read_excel(codebook_location)
-
-#Summary of data
-glimpse(Feb_26)
-summary(Feb_26)
-
-
-
-
-########################################################################################cleaning1
-#filtering data by county because the data contains all counties of Georgia and for this 
-#project I am interested in Dougherty Count, which is where the city of Albany is 
-Dougherty_Vax_Feb <- filter(Feb_26, COUNTY_NAME == "Dougherty County")  
-
-#replacing negative values with 0
-Dougherty_Vax_Feb [Dougherty_Vax_Feb < 0] <- 0  
-
-
-
-
-
-
-########################################################################################cleaning2
-##organizing the data then separate into zips and add a zip code column   
-
-#Adding a column of the total vax for each census tract (for each observation)  
-Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_vax_Feb = sum(across(ends_with("VAX")), na.rm = T))  
-
-#total vax column for male  
-Dougherty_Vax_Feb <- Dougherty_Vax_Feb  %>% rowwise() %>%   
-  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))  
-
-#total vax column for female 
-Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
-  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))  
[…]

### 2023-04-05 code/analysis_code/analysiscode.qmd
Part 4

--- 2023-03-19 code/analysis_code/analysiscode1.qmd
@@
 ---
-title: "Total Amount of Vaccinations Among African Americans in the Target Zip Codes"
-format: html
-editor: visual
+title: "Total Amount of Vaccinations Among Black population in the Target Zip Codes"
+output: 
+  html_document
 echo: FALSE 
@@
 
-# Analysis of the Amount of Vaccinations Among African Americans in Each Target Zip Code
+# Analysis of the Amount of Vaccinations Among Black population in Each Target Zip Code
 
@@
 
-### Saving the combined data for African Americans for future use. Will be useful when stratifying the African American population by age and gender
+### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender
 
@@
 
-### Dropping all variables except for total black variables in order to obtain vaccination rates of all African Americans in the 31701 zip code
+### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31701 zip code
 
-#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of African Americans with at least one dose of the COVID-19 vaccine by month
+#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
 
@@
 ### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 31701 zip code
+
 ```{r}
@@
 plot_31701_total_black_vax = plot_31701_total_black_vax + 
-  ggtitle("Vaccination Totals for African Americans in the 31701 Zipcode") + 
+  ggtitle("Vaccination Totals for Black population in the 31701 Zipcode") + 
   theme(plot.title = 
@@
 ### Loading the processed data in order to begin analysis for the 31705 zip code
+
 ```{r}
@@
 
-### Saving the combined data for African Americans for future use 
+### Saving the combined data for Black population for future use
+
 ```{r}
@@
 
-### Dropping all variables except for total black variables in order to obtain vaccination rates of all African Americans in the 31705 zip code 
+### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31705 zip code
+
 ```{r}
@@
 ### Summing each column for each month
+
 ```{r}
@@
 ### Total contains the vaccination totals for each month
+
 ```{r}
@@
 ### ggplot creating a line plot for the vaccination totals by month for the 31705 zip code African American population
+
 ```{r}
@@
 plot_31705_total_black_vax = plot_31705_total_black_vax + 
-  ggtitle("Vaccination Totals for African Americans in the 31705 Zipcode") + 
+  ggtitle("Vaccination Totals for Black population in the 31705 Zipcode") + 
   theme(plot.title = 
@@
 ### Loading the processed data in order to begin analysis for the 31707 zip code
+
 ```{r}
@@
 
-### Dropping all variables except for total black variables in order to obtain vaccination rates of all African Americans in the 31707 zip code 
+### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31707 zip code
+
 ```{r}
@@
 ### Separating sums by month into a data fram consisting of two variables: Date and Total
+
 ### Date contains months September-February
+
 ### Total contains the vaccination totals for each month
+
 ```{r}
@@
 ###ggplot: creating a line plot for the vaccination totals by month for the 31707 zip code African American population
+
 ```{r}
@@
 plot_31707_total_black_vax = plot_31707_total_black_vax + 
-  ggtitle("Vaccination Totals for African Americans in the 31707 Zipcode") + 
+  ggtitle("Vaccination Totals for Black population in the 31707 Zipcode") + 
   theme(plot.title = 
@@
 ### Taking a look at each of the line plots for the different zip codes and comparing the total vaccinations by month in the African American population
+
 ```{r}
@@
 ### Combining all plots into one
+
 ```{r}
@@
 ### Combining the new dataframe "Total_black_vax" with the 31707 zip vaccination totals so all three target zip codes are in the same data frame
+
 ```{r}
@@
 ### Renaming column totals to match the zip
+
 ### Haing to do this because the the total variable names are Total.x, Total.y and Total
+
 ### Need to specificy each total variable by zip using the following code
+
 ```{r}
@@
 ### Converting data to long format in order to create a line plot using ggplot
+
 ```{r}
@@
 ### Changing the variable name "variable" to "Zip_code" and "value" to "Vaccination_total" for clarity
+
 ### Doing this because when the data was converted to a long format the column names we changed to "variable" which refers to the zip code and "value" which refers to the amount of vaccinations
+
 ### Changing "variable" to "Zip_code" and "value" to "Vaccination_total"
+
 ```{r}
@@
 ### ggplot: Making a line graph that compares the vaccination totals for each African American population in each zip code
+
 ```{r}
@@
 plot_Total_black_vax = plot_Total_black_vax + 
-  ggtitle("Vaccination Totals for African Americans in Target Zip Codes") + 
+  ggtitle("Vaccination Totals for Black population in Target Zip Codes") + 
   theme(plot.title = 
@@
 ### Turning ggplot into an interactive plot
+
 ```{r}

### 2023-04-05 code/analysis_code/analysiscode2.qmd
Part 4

---
title: "Total Amount of Vaccinations by Gender Among Black population in the Target Zip Codes"
output: 
  html_document
echo: FALSE 
---

# Library

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr)
library(ggplot2)
library(reshape2) #reshaping data from wide to long format
library(plotly) #creating an interactive line plot 
```

# Total Amount of vaccinations by gender among Black population in the target zip codes

------------------------------------------------------------------------

## Zip code: 31701

### Loading the `processed_combined_31701.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
processed_combined_31701 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31701

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31701_Gender_black = processed_combined_31701 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31701_total_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31701_total_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31701_total_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31701_total_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31701_total_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31701_total_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31701_total_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31701_total_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31701_total_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31701_total_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31701_total_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31701_total_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420))
```

### Converting the Gender_Sum_by_month_31701 to long data in order to plot
```{r}
Gender_Sum_by_month_31701_long = reshape2::melt(Gender_Sum_by_month_31701, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31701_long <- Gender_Sum_by_month_31701_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31701 zip code

```{r}
plot_Gender_black_vax = Gender_Sum_by_month_31701_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31701") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
plotly_plot_Gender_black_vax
```

------------------------------------------------------------------------

## Zip code: 31705

### Loading the `processed_combined_31705.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
processed_combined_31705 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31705

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31705_total_black = processed_combined_31705 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31705_total_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31705_total_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31705_total_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31705_total_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31705_total_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31705_total_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31705_total_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31705_total_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31705_total_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31705_total_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31705_total_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31705_total_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949))
```

### Converting the Gender_Sum_by_month_31705 to long data in order to plot
```{r}
Gender_Sum_by_month_31705_long = reshape2::melt(Gender_Sum_by_month_31705, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31705_long <- Gender_Sum_by_month_31705_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31705 zip code

```{r}
plot_Gender_black_vax_31705 = Gender_Sum_by_month_31705_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31705") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax_31705
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax_31705 = ggplotly(plot_Gender_black_vax_31705)
plotly_plot_Gender_black_vax_31705
```

------------------------------------------------------------------------

## Zip code: 31707

### Loading the `processed_combined_31707.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
processed_combined_31707 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31707

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31707_total_black = processed_combined_31707 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31707_total_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31707_total_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31707_total_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31707_total_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31707_total_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31707_total_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31707_total_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31707_total_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31707_total_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31707_total_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31707_total_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31707_total_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361))
```

### Converting the Gender_Sum_by_month_31707 to long data in order to plot
```{r}
Gender_Sum_by_month_31707_long = reshape2::melt(Gender_Sum_by_month_31707, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31707_long <- Gender_Sum_by_month_31707_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31707 zip code

```{r}
plot_Gender_black_vax_31707 = Gender_Sum_by_month_31707_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31707") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax_31707
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax_31707 = ggplotly(plot_Gender_black_vax_31707)
plotly_plot_Gender_black_vax_31707
```

# Thoughts regarding exploratory analysis 
In each zip code there is a large gap between females that are Black and males that are Black. I am pretty taken back by how large the gap is.

### 2023-04-05 code/analysis_code/analysiscode3.qmd
Part 4

--- 2023-04-05 code/analysis_code/analysiscode2.qmd
@@
 ---
-title: "Total Amount of Vaccinations by Gender Among Black population in the Target Zip Codes"
+title: "Total Amount of Vaccinations by Age Among Black population in the Target Zip Codes"
 output: 
@@
 ```{r}
-processed_combined_31701_Gender_black = processed_combined_31701 %>% select(matches("^total_(mblack|fblack)_vax"))
+processed_combined_31701_Age_black = processed_combined_31701 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
 ```
@@
 ```
-
-# Thoughts regarding exploratory analysis 
-In each zip code there is a large gap between females that are Black and males that are Black. I am pretty taken back by how large the gap is.

### 2023-04-05 code/processing_code/processingcode.qmd
Part 4

---
title: "Processing Code for September Data"
output: 
  html_document
---

# Processing 1

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept_17 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Sept_17)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Sept <- filter(Sept_17, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Sept [Dougherty_Vax_Sept < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_vax_Sept = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Sept <- Dougherty_Vax_Sept  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Sept<- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Sept$total_asian_vax <- Dougherty_Vax_Sept$total_masian_vax + Dougherty_Vax_Sept$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Sept$total_black_vax <- Dougherty_Vax_Sept$total_mblack_vax + Dougherty_Vax_Sept$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Sept$total_orace_vax <- Dougherty_Vax_Sept$total_morace_vax + Dougherty_Vax_Sept$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Sept$total_white_vax <- Dougherty_Vax_Sept$total_mwhite_vax + Dougherty_Vax_Sept$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Sept <- Dougherty_Vax_Sept %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Sept_sums <- Dougherty_Vax_Sept[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Sept <-filter(Dougherty_Vax_Sept_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Sept <- filter(Dougherty_Vax_Sept_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Sept <- filter(Dougherty_Vax_Sept_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Sept

```{r}
colnames(Zip_31701_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")

#change column names for 31705 Sept  
colnames(Zip_31705_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")

#change column names for 31707 Sept  
colnames(Zip_31707_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")

```

### Deleting excess environment objects and saving data as processed_data_sept
```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Sept,Sept_17, codebook, Dougherty_Vax_Sept_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data", "Zip_31701", "processed_data_31701_Sept.rds")

saveRDS(Zip_31701_Sept, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Sept.rds")

saveRDS(Zip_31705_Sept, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Sept.rds")

saveRDS(Zip_31707_Sept, file = save_data_location3)

```

# Processing 2

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Oct_28.xlsx")
Oct_28 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Oct_28)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Oct <- filter(Oct_28, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Oct [Dougherty_Vax_Oct < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_vax_Oct = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Oct <- Dougherty_Vax_Oct  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Oct<- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Oct$total_asian_vax <- Dougherty_Vax_Oct$total_masian_vax + Dougherty_Vax_Oct$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Oct$total_black_vax <- Dougherty_Vax_Oct$total_mblack_vax + Dougherty_Vax_Oct$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Oct$total_orace_vax <- Dougherty_Vax_Oct$total_morace_vax + Dougherty_Vax_Oct$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Oct$total_white_vax <- Dougherty_Vax_Oct$total_mwhite_vax + Dougherty_Vax_Oct$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Oct <- Dougherty_Vax_Oct %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Oct_sums <- Dougherty_Vax_Oct[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Oct <-filter(Dougherty_Vax_Oct_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Oct <- filter(Dougherty_Vax_Oct_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Oct <- filter(Dougherty_Vax_Oct_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Oct

```{r}
colnames(Zip_31701_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")

#change column names for 31705 Oct  
colnames(Zip_31705_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")

#change column names for 31707 Oct  
colnames(Zip_31707_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")

```

# Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Oct

```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Oct,Oct_28, codebook, Dougherty_Vax_Oct_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Oct.rds")

saveRDS(Zip_31701_Oct, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Oct.rds")

saveRDS(Zip_31705_Oct, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Oct.rds")

saveRDS(Zip_31707_Oct, file = save_data_location3)

```

# Processing 3

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Nov_24.xlsx")
Nov_24 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Nov_24)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Nov <- filter(Nov_24, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Nov [Dougherty_Vax_Nov < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_vax_Nov = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Nov <- Dougherty_Vax_Nov  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Nov<- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Nov$total_asian_vax <- Dougherty_Vax_Nov$total_masian_vax + Dougherty_Vax_Nov$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Nov$total_black_vax <- Dougherty_Vax_Nov$total_mblack_vax + Dougherty_Vax_Nov$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Nov$total_orace_vax <- Dougherty_Vax_Nov$total_morace_vax + Dougherty_Vax_Nov$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Nov$total_white_vax <- Dougherty_Vax_Nov$total_mwhite_vax + Dougherty_Vax_Nov$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Nov <- Dougherty_Vax_Nov %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Nov_sums <- Dougherty_Vax_Nov[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Nov <-filter(Dougherty_Vax_Nov_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Nov <- filter(Dougherty_Vax_Nov_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Nov <- filter(Dougherty_Vax_Nov_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Nov

```{r}
colnames(Zip_31701_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

#change column names for 31705 Nov  
colnames(Zip_31705_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

#change column names for 31707 Nov  
colnames(Zip_31707_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

```

# Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Nov

```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Nov,Nov_24, codebook, Dougherty_Vax_Nov_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Nov.rds")

saveRDS(Zip_31701_Nov, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Nov.rds")

saveRDS(Zip_31705_Nov, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Nov.rds")

saveRDS(Zip_31707_Nov, file = save_data_location3)

```

# Processing 4

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Dec_29.xlsx")
Dec_29 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Dec_29)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Dec <- filter(Dec_29, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Dec [Dougherty_Vax_Dec < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_vax_Dec = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Dec <- Dougherty_Vax_Dec  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Dec<- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Dec$total_asian_vax <- Dougherty_Vax_Dec$total_masian_vax + Dougherty_Vax_Dec$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Dec$total_black_vax <- Dougherty_Vax_Dec$total_mblack_vax + Dougherty_Vax_Dec$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Dec$total_orace_vax <- Dougherty_Vax_Dec$total_morace_vax + Dougherty_Vax_Dec$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Dec$total_white_vax <- Dougherty_Vax_Dec$total_mwhite_vax + Dougherty_Vax_Dec$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Dec <- Dougherty_Vax_Dec %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Dec_sums <- Dougherty_Vax_Dec[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Dec <-filter(Dougherty_Vax_Dec_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Dec <- filter(Dougherty_Vax_Dec_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Dec <- filter(Dougherty_Vax_Dec_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Dec

```{r}
colnames(Zip_31701_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

#change column names for 31705 Dec  
colnames(Zip_31705_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

#change column names for 31707 Dec  
colnames(Zip_31707_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

```

# Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Dec

```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Dec,Dec_29, codebook, Dougherty_Vax_Dec_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Dec.rds")

saveRDS(Zip_31701_Dec, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Dec.rds")

saveRDS(Zip_31705_Dec, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Dec.rds")

saveRDS(Zip_31707_Dec, file = save_data_location3)

```

# Processing script 5

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Jan_26.xlsx")
Jan_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Jan_26)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Jan <- filter(Jan_26, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Jan [Dougherty_Vax_Jan < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_vax_Jan = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Jan <- Dougherty_Vax_Jan  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Jan<- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Jan$total_asian_vax <- Dougherty_Vax_Jan$total_masian_vax + Dougherty_Vax_Jan$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Jan$total_black_vax <- Dougherty_Vax_Jan$total_mblack_vax + Dougherty_Vax_Jan$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Jan$total_orace_vax <- Dougherty_Vax_Jan$total_morace_vax + Dougherty_Vax_Jan$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Jan$total_white_vax <- Dougherty_Vax_Jan$total_mwhite_vax + Dougherty_Vax_Jan$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Jan <- Dougherty_Vax_Jan %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Jan_sums <- Dougherty_Vax_Jan[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Jan <-filter(Dougherty_Vax_Jan_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Jan <- filter(Dougherty_Vax_Jan_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Jan <- filter(Dougherty_Vax_Jan_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Jan

```{r}
colnames(Zip_31701_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

#change column names for 31705 Jan  
colnames(Zip_31705_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

#change column names for 31707 Jan  
colnames(Zip_31707_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

```

# Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Jan

```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Jan,Jan_26, codebook, Dougherty_Vax_Jan_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Jan.rds")

saveRDS(Zip_31701_Jan, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Jan.rds")

saveRDS(Zip_31705_Jan, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Jan.rds")

saveRDS(Zip_31707_Jan, file = save_data_location3)

```

# Processing 6

Load needed packages. make sure they are installed.

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
```

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Feb_26.xlsx")
Feb_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Feb_26)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
Dougherty_Vax_Feb <- filter(Feb_26, COUNTY_NAME == "Dougherty County")

#replacing negative values with 0
Dougherty_Vax_Feb [Dougherty_Vax_Feb < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_vax_Feb = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Dougherty_Vax_Feb <- Dougherty_Vax_Feb  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Dougherty_Vax_Feb<- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Dougherty_Vax_Feb$total_asian_vax <- Dougherty_Vax_Feb$total_masian_vax + Dougherty_Vax_Feb$total_fasian_vax

#Vax total for African American Residents 
Dougherty_Vax_Feb$total_black_vax <- Dougherty_Vax_Feb$total_mblack_vax + Dougherty_Vax_Feb$total_fblack_vax

#Vax total for Other Race Residents
Dougherty_Vax_Feb$total_orace_vax <- Dougherty_Vax_Feb$total_morace_vax + Dougherty_Vax_Feb$total_forace_vax

#Vax total for White Residents
Dougherty_Vax_Feb$total_white_vax <- Dougherty_Vax_Feb$total_mwhite_vax + Dougherty_Vax_Feb$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Dougherty_Vax_Feb <- Dougherty_Vax_Feb %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Dougherty_Vax_Feb_sums <- Dougherty_Vax_Feb[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Feb <-filter(Dougherty_Vax_Feb_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Feb <- filter(Dougherty_Vax_Feb_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Feb <- filter(Dougherty_Vax_Feb_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Feb

```{r}
colnames(Zip_31701_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")

#change column names for 31705 Feb  
colnames(Zip_31705_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")

#change column names for 31707 Feb  
colnames(Zip_31707_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")

```

# Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Feb

```{r}
#Only need to save processed data for the three zip codes
rm(Dougherty_Vax_Feb,Feb_26, codebook, Dougherty_Vax_Feb_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Feb.rds")

saveRDS(Zip_31701_Feb, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Feb.rds")

saveRDS(Zip_31705_Feb, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Feb.rds")

saveRDS(Zip_31707_Feb, file = save_data_location3)

```

### 2023-04-06 code/analysis_code/analysiscode2.qmd
Part 4

@@
 
-#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
 
@@
 #Summing for black males by month 
-sum(processed_combined_31701_total_black$total_mblack_vax_Sept_17) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Oct_28) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Nov_24) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Dec_29) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Jan_26) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Feb_26)
+sum(processed_combined_31701_Gender_black$total_mblack_vax_Sept_17) 
+sum(processed_combined_31701_Gender_black$total_mblack_vax_Oct_28) 
+sum(processed_combined_31701_Gender_black$total_mblack_vax_Nov_24) 
+sum(processed_combined_31701_Gender_black$total_mblack_vax_Dec_29) 
+sum(processed_combined_31701_Gender_black$total_mblack_vax_Jan_26) 
+sum(processed_combined_31701_Gender_black$total_mblack_vax_Feb_26)
 
 #summing for black females by month 
-sum(processed_combined_31701_total_black$total_fblack_vax_Sept_17) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Oct_28) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Nov_24) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Dec_29) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Jan_26) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Feb_26) 
+sum(processed_combined_31701_Gender_black$total_fblack_vax_Sept_17) 
+sum(processed_combined_31701_Gender_black$total_fblack_vax_Oct_28) 
+sum(processed_combined_31701_Gender_black$total_fblack_vax_Nov_24) 
+sum(processed_combined_31701_Gender_black$total_fblack_vax_Dec_29) 
+sum(processed_combined_31701_Gender_black$total_fblack_vax_Jan_26) 
+sum(processed_combined_31701_Gender_black$total_fblack_vax_Feb_26) 
 ```
@@
 #adding a title to the plot 
-plot_Total_black_vax = plot_Total_black_vax + 
+plot_Gender_black_vax = plot_Gender_black_vax + 
   ggtitle("Vaccination Totals for Black population by Gender in 31701") + 
@@
 
-#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
 
 ```{r}
-processed_combined_31705_total_black = processed_combined_31705 %>% select(matches("^total_(mblack|fblack)_vax"))
+processed_combined_31705_Gender_black = processed_combined_31705 %>% select(matches("^total_(mblack|fblack)_vax"))
 ```
@@
 #Summing for black males by month 
-sum(processed_combined_31705_total_black$total_mblack_vax_Sept_17) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Oct_28) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Nov_24) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Dec_29) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Jan_26) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Feb_26)
+sum(processed_combined_31705_Gender_black$total_mblack_vax_Sept_17) 
+sum(processed_combined_31705_Gender_black$total_mblack_vax_Oct_28) 
+sum(processed_combined_31705_Gender_black$total_mblack_vax_Nov_24) 
+sum(processed_combined_31705_Gender_black$total_mblack_vax_Dec_29) 
+sum(processed_combined_31705_Gender_black$total_mblack_vax_Jan_26) 
+sum(processed_combined_31705_Gender_black$total_mblack_vax_Feb_26)
 
 #summing for black females by month 
-sum(processed_combined_31705_total_black$total_fblack_vax_Sept_17) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Oct_28) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Nov_24) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Dec_29) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Jan_26) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Feb_26) 
+sum(processed_combined_31705_Gender_black$total_fblack_vax_Sept_17) 
+sum(processed_combined_31705_Gender_black$total_fblack_vax_Oct_28) 
+sum(processed_combined_31705_Gender_black$total_fblack_vax_Nov_24) 
+sum(processed_combined_31705_Gender_black$total_fblack_vax_Dec_29) 
+sum(processed_combined_31705_Gender_black$total_fblack_vax_Jan_26) 
+sum(processed_combined_31705_Gender_black$total_fblack_vax_Feb_26) 
 ```
@@
 #adding a title to the plot 
-plot_Total_black_vax = plot_Total_black_vax + 
+plot_Gender_black_vax = plot_Gender_black_vax + 
   ggtitle("Vaccination Totals for Black population by Gender in 31705") + 
@@
 
-#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
 
 ```{r}
-processed_combined_31707_total_black = processed_combined_31707 %>% select(matches("^total_(mblack|fblack)_vax"))
+processed_combined_31707_Gender_black = processed_combined_31707 %>% select(matches("^total_(mblack|fblack)_vax"))
 ```
@@
 #Summing for black males by month 
-sum(processed_combined_31707_total_black$total_mblack_vax_Sept_17) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Oct_28) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Nov_24) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Dec_29) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Jan_26) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Feb_26)
+sum(processed_combined_31707_Gender_black$total_mblack_vax_Sept_17) 
+sum(processed_combined_31707_Gender_black$total_mblack_vax_Oct_28) 
+sum(processed_combined_31707_Gender_black$total_mblack_vax_Nov_24) 
+sum(processed_combined_31707_Gender_black$total_mblack_vax_Dec_29) 
+sum(processed_combined_31707_Gender_black$total_mblack_vax_Jan_26) 
+sum(processed_combined_31707_Gender_black$total_mblack_vax_Feb_26)
 
 #summing for black females by month 
-sum(processed_combined_31707_total_black$total_fblack_vax_Sept_17) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Oct_28) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Nov_24) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Dec_29) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Jan_26) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Feb_26) 
+sum(processed_combined_31707_Gender_black$total_fblack_vax_Sept_17) 
+sum(processed_combined_31707_Gender_black$total_fblack_vax_Oct_28) 
+sum(processed_combined_31707_Gender_black$total_fblack_vax_Nov_24) 
+sum(processed_combined_31707_Gender_black$total_fblack_vax_Dec_29) 
+sum(processed_combined_31707_Gender_black$total_fblack_vax_Jan_26) 
+sum(processed_combined_31707_Gender_black$total_fblack_vax_Feb_26) 
 ```
@@
 #adding a title to the plot 
-plot_Total_black_vax = plot_Total_black_vax + 
+plot_Gender_black_vax = plot_Gender_black_vax + 
   ggtitle("Vaccination Totals for Black population by Gender in 31707") + 

### 2023-04-06 code/analysis_code/analysiscode3.qmd
Part 4

@@
 
-# Total Amount of vaccinations by gender among Black population in the target zip codes
+# Total Amount of vaccinations by Age among Black population in the target zip codes
 
@@
 
-### Dropping all variables except for black males and black females within 31701
+### Dropping all variables except for black age groups within 31701
 
-#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Using the select() function to subset the data frame by columns that start with an age group
 
@@
 
-### Summing each Month variable by census tract
+### Summing each Month variable by age group by month
 ```{r}
-#Summing for black males by month 
-sum(processed_combined_31701_total_black$total_mblack_vax_Sept_17) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Oct_28) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Nov_24) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Dec_29) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Jan_26) 
-sum(processed_combined_31701_total_black$total_mblack_vax_Feb_26)
+#Summing for ages 5-9 by month 
+sum(processed_combined_31701_Age_black$total_05_09_vax_Sept_17) 
+sum(processed_combined_31701_Age_black$total_05_09_vax_Oct_28) 
+sum(processed_combined_31701_Age_black$total_05_09_vax_Nov_24) 
+sum(processed_combined_31701_Age_black$total_05_09_vax_Dec_29) 
+sum(processed_combined_31701_Age_black$total_05_09_vax_Jan_26) 
+sum(processed_combined_31701_Age_black$total_05_09_vax_Feb_26)
 
-#summing for black females by month 
-sum(processed_combined_31701_total_black$total_fblack_vax_Sept_17) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Oct_28) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Nov_24) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Dec_29) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Jan_26) 
-sum(processed_combined_31701_total_black$total_fblack_vax_Feb_26) 
+#Summing for ages 10-17 by month 
+sum(processed_combined_31701_Age_black$total_10_17_vax_Sept_17) 
+sum(processed_combined_31701_Age_black$total_10_17_vax_Oct_28) 
+sum(processed_combined_31701_Age_black$total_10_17_vax_Nov_24) 
+sum(processed_combined_31701_Age_black$total_10_17_vax_Dec_29) 
+sum(processed_combined_31701_Age_black$total_10_17_vax_Jan_26) 
+sum(processed_combined_31701_Age_black$total_10_17_vax_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_combined_31701_Age_black$total_18_44_vax_Sept_17) 
+sum(processed_combined_31701_Age_black$total_18_44_vax_Oct_28) 
+sum(processed_combined_31701_Age_black$total_18_44_vax_Nov_24) 
+sum(processed_combined_31701_Age_black$total_18_44_vax_Dec_29) 
+sum(processed_combined_31701_Age_black$total_18_44_vax_Jan_26) 
+sum(processed_combined_31701_Age_black$total_18_44_vax_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_combined_31701_Age_black$total_45_64_vax_Sept_17) 
+sum(processed_combined_31701_Age_black$total_45_64_vax_Oct_28) 
+sum(processed_combined_31701_Age_black$total_45_64_vax_Nov_24) 
+sum(processed_combined_31701_Age_black$total_45_64_vax_Dec_29) 
+sum(processed_combined_31701_Age_black$total_45_64_vax_Jan_26) 
+sum(processed_combined_31701_Age_black$total_45_64_vax_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_combined_31701_Age_black$total_65Plus_vax_Sept_17) 
+sum(processed_combined_31701_Age_black$total_65Plus_vax_Oct_28) 
+sum(processed_combined_31701_Age_black$total_65Plus_vax_Nov_24) 
+sum(processed_combined_31701_Age_black$total_65Plus_vax_Dec_29) 
+sum(processed_combined_31701_Age_black$total_65Plus_vax_Jan_26) 
+sum(processed_combined_31701_Age_black$total_65Plus_vax_Feb_26)
+
 ```
 
-#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
+### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
 
 ```{r}
-Gender_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
-                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420))
+Age_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          Age_05_09 = c(123, 126, 126, 127, 127, 128),
+                          Age_10_17 = c(642, 635, 633, 637, 630, 629),
+                          Age_18_44 = c(3310, 3316, 3313, 3323, 3325, 3331),
+                          Age_45_64 = c(3229, 3244, 3239, 3259, 3264, 3258),
+                          Age_65Plus = c(2576, 2570, 2569, 2574, 2568, 2572))
+
 ```
 
-### Converting the Gender_Sum_by_month_31701 to long data in order to plot
+### Converting the Age_Sum_by_month_31701 to long data in order to plot
 ```{r}
-Gender_Sum_by_month_31701_long = reshape2::melt(Gender_Sum_by_month_31701, id = "Date")
+Age_Sum_by_month_31701_long = reshape2::melt(Age_Sum_by_month_31701, id = "Date")
 ```
 
-### Changing "variable" to "Gender" and "value" to "Vaccination_total"
+### Changing "variable" to "Age" and "value" to "Vaccination_total"
 
 ```{r}
-Gender_Sum_by_month_31701_long <- Gender_Sum_by_month_31701_long %>% 
-  rename("Gender" = "variable") %>% 
+Age_Sum_by_month_31701_long <- Age_Sum_by_month_31701_long %>% 
+  rename("Age" = "variable") %>% 
   rename("Vaccination_total" = "value")
@@
 
-### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31701 zip code
+### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 31701 zip code
 
 ```{r}
-plot_Gender_black_vax = Gender_Sum_by_month_31701_long %>% 
+plot_Age_black_vax = Age_Sum_by_month_31701_long %>% 
                      ggplot(aes(Date, Vaccination_total, 
-                                group = Gender, 
-                                color = Gender)) + 
+                                group = Age, 
+                                color = Age)) + 
                      geom_point() + geom_line() + 
@@
 #adding a title to the plot 
-plot_Total_black_vax = plot_Total_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Gender in 31701") + 
+plot_Age_black_vax = plot_Age_black_vax + 
+  ggtitle("Vaccination Totals for Black population by Age in 31701") + 
   theme(plot.title = 
@@
   labs(y = "Vaccination Total")
-plot_Gender_black_vax
+plot_Age_black_vax
 ```
@@
 ```{r}
-plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
-plotly_plot_Gender_black_vax
+plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
+plotly_plot_Age_black_vax
 ```
@@
 ------------------------------------------------------------------------
-
 ## Zip code: 31705
@@
 
-### Dropping all variables except for black males and black females within 31705
+### Dropping all variables except for black age groups within 31705
 
-#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Using the select() function to subset the data frame by columns that start with an age group
 
 ```{r}
-processed_combined_31705_total_black = processed_combined_31705 %>% select(matches("^total_(mblack|fblack)_vax"))
+processed_combined_31705_Age_black = processed_combined_31705 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
 ```
 
-### Summing each Month variable by census tract
+### Summing each Month variable by age group by month
 ```{r}
-#Summing for black males by month 
-sum(processed_combined_31705_total_black$total_mblack_vax_Sept_17) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Oct_28) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Nov_24) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Dec_29) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Jan_26) 
-sum(processed_combined_31705_total_black$total_mblack_vax_Feb_26)
+#Summing for ages 5-9 by month 
+sum(processed_combined_31705_Age_black$total_05_09_vax_Sept_17) 
+sum(processed_combined_31705_Age_black$total_05_09_vax_Oct_28) 
+sum(processed_combined_31705_Age_black$total_05_09_vax_Nov_24) 
+sum(processed_combined_31705_Age_black$total_05_09_vax_Dec_29) 
+sum(processed_combined_31705_Age_black$total_05_09_vax_Jan_26) 
+sum(processed_combined_31705_Age_black$total_05_09_vax_Feb_26)
 
-#summing for black females by month 
-sum(processed_combined_31705_total_black$total_fblack_vax_Sept_17) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Oct_28) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Nov_24) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Dec_29) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Jan_26) 
-sum(processed_combined_31705_total_black$total_fblack_vax_Feb_26) 
+#Summing for ages 10-17 by month 
+sum(processed_combined_31705_Age_black$total_10_17_vax_Sept_17) 
+sum(processed_combined_31705_Age_black$total_10_17_vax_Oct_28) 
+sum(processed_combined_31705_Age_black$total_10_17_vax_Nov_24) 
+sum(processed_combined_31705_Age_black$total_10_17_vax_Dec_29) 
+sum(processed_combined_31705_Age_black$total_10_17_vax_Jan_26) 
+sum(processed_combined_31705_Age_black$total_10_17_vax_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_combined_31705_Age_black$total_18_44_vax_Sept_17) 
+sum(processed_combined_31705_Age_black$total_18_44_vax_Oct_28) 
+sum(processed_combined_31705_Age_black$total_18_44_vax_Nov_24) 
+sum(processed_combined_31705_Age_black$total_18_44_vax_Dec_29) 
+sum(processed_combined_31705_Age_black$total_18_44_vax_Jan_26) 
+sum(processed_combined_31705_Age_black$total_18_44_vax_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_combined_31705_Age_black$total_45_64_vax_Sept_17) 
+sum(processed_combined_31705_Age_black$total_45_64_vax_Oct_28) 
+sum(processed_combined_31705_Age_black$total_45_64_vax_Nov_24) 
+sum(processed_combined_31705_Age_black$total_45_64_vax_Dec_29) 
+sum(processed_combined_31705_Age_black$total_45_64_vax_Jan_26) 
+sum(processed_combined_31705_Age_black$total_45_64_vax_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_combined_31705_Age_black$total_65Plus_vax_Sept_17) 
+sum(processed_combined_31705_Age_black$total_65Plus_vax_Oct_28) 
+sum(processed_combined_31705_Age_black$total_65Plus_vax_Nov_24) 
+sum(processed_combined_31705_Age_black$total_65Plus_vax_Dec_29) 
+sum(processed_combined_31705_Age_black$total_65Plus_vax_Jan_26) 
+sum(processed_combined_31705_Age_black$total_65Plus_vax_Feb_26)
+
 ```
 
-#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
+### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February 
+### Total contains the vaccination totals for each month
 
 ```{r}
-Gender_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
-                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949))
+Age_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          Age_05_09 = c(210, 211, 209, 201, 202, 202),
+                          Age_10_17 = c(1159, 1171, 1169, 1163, 1168, 1167),
+                          Age_18_44 = c(5201, 5215, 5239, 5263, 5278, 5281), #pretty notable increase for ages 18-44 
+                          Age_45_64 = c(4380, 4381, 4366, 4364, 4367, 4367),
+                          Age_65Plus = c(2983, 3000, 2997, 2997, 2994, 2989))
+
 ```
 
-### Converting the Gender_Sum_by_month_31705 to long data in order to plot
+### Converting the Age_Sum_by_month_31705 to long data in order to plot
 ```{r}
-Gender_Sum_by_month_31705_long = reshape2::melt(Gender_Sum_by_month_31705, id = "Date")
+Age_Sum_by_month_31705_long = reshape2::melt(Age_Sum_by_month_31705, id = "Date")
 ```
 
-### Changing "variable" to "Gender" and "value" to "Vaccination_total"
+### Changing "variable" to "Age" and "value" to "Vaccination_total"
 
 ```{r}
-Gender_Sum_by_month_31705_long <- Gender_Sum_by_month_31705_long %>% 
-  rename("Gender" = "variable") %>% 
+Age_Sum_by_month_31705_long <- Age_Sum_by_month_31705_long %>% 
+  rename("Age" = "variable") %>% 
   rename("Vaccination_total" = "value")
@@
 
-### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31705 zip code
+### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 31705 zip code
 
 ```{r}
-plot_Gender_black_vax_31705 = Gender_Sum_by_month_31705_long %>% 
+plot_Age_black_vax = Age_Sum_by_month_31705_long %>% 
                      ggplot(aes(Date, Vaccination_total, 
-                                group = Gender, 
-                                color = Gender)) + 
+                                group = Age, 
+                                color = Age)) + 
                      geom_point() + geom_line() + 
@@
 #adding a title to the plot 
-plot_Total_black_vax = plot_Total_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Gender in 31705") + 
+plot_Age_black_vax = plot_Age_black_vax + 
+  ggtitle("Vaccination Totals for Black population by Age in 31705") + 
   theme(plot.title = 
@@
   labs(y = "Vaccination Total")
-plot_Gender_black_vax_31705
+plot_Age_black_vax
 ```
@@
 ```{r}
-plotly_plot_Gender_black_vax_31705 = ggplotly(plot_Gender_black_vax_31705)
-plotly_plot_Gender_black_vax_31705
+plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
+plotly_plot_Age_black_vax
 ```
@@
 
-### Dropping all variables except for black males and black females within 31707
+### Dropping all variables except for black age groups within 31707
 
-#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Using the select() function to subset the data frame by columns that start with an age group
 
 ```{r}
-processed_combined_31707_total_black = processed_combined_31707 %>% select(matches("^total_(mblack|fblack)_vax"))
+processed_combined_31707_Age_black = processed_combined_31707 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
 ```
 
-### Summing each Month variable by census tract
+### Summing each Month variable by age group by month
 ```{r}
-#Summing for black males by month 
-sum(processed_combined_31707_total_black$total_mblack_vax_Sept_17) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Oct_28) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Nov_24) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Dec_29) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Jan_26) 
-sum(processed_combined_31707_total_black$total_mblack_vax_Feb_26)
+#Summing for ages 5-9 by month 
+sum(processed_combined_31707_Age_black$total_05_09_vax_Sept_17) 
+sum(processed_combined_31707_Age_black$total_05_09_vax_Oct_28) 
+sum(processed_combined_31707_Age_black$total_05_09_vax_Nov_24) 
+sum(processed_combined_31707_Age_black$total_05_09_vax_Dec_29) 
+sum(processed_combined_31707_Age_black$total_05_09_vax_Jan_26) 
+sum(processed_combined_31707_Age_black$total_05_09_vax_Feb_26)
 
-#summing for black females by month 
-sum(processed_combined_31707_total_black$total_fblack_vax_Sept_17) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Oct_28) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Nov_24) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Dec_29) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Jan_26) 
-sum(processed_combined_31707_total_black$total_fblack_vax_Feb_26) 
+#Summing for ages 10-17 by month 
+sum(processed_combined_31707_Age_black$total_10_17_vax_Sept_17) 
+sum(processed_combined_31707_Age_black$total_10_17_vax_Oct_28) 
+sum(processed_combined_31707_Age_black$total_10_17_vax_Nov_24) 
+sum(processed_combined_31707_Age_black$total_10_17_vax_Dec_29) 
+sum(processed_combined_31707_Age_black$total_10_17_vax_Jan_26) 
+sum(processed_combined_31707_Age_black$total_10_17_vax_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_combined_31707_Age_black$total_18_44_vax_Sept_17) 
+sum(processed_combined_31707_Age_black$total_18_44_vax_Oct_28) 
+sum(processed_combined_31707_Age_black$total_18_44_vax_Nov_24) 
+sum(processed_combined_31707_Age_black$total_18_44_vax_Dec_29) 
+sum(processed_combined_31707_Age_black$total_18_44_vax_Jan_26) 
+sum(processed_combined_31707_Age_black$total_18_44_vax_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_combined_31707_Age_black$total_45_64_vax_Sept_17) 
+sum(processed_combined_31707_Age_black$total_45_64_vax_Oct_28) 
+sum(processed_combined_31707_Age_black$total_45_64_vax_Nov_24) 
+sum(processed_combined_31707_Age_black$total_45_64_vax_Dec_29) 
+sum(processed_combined_31707_Age_black$total_45_64_vax_Jan_26) 
+sum(processed_combined_31707_Age_black$total_45_64_vax_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_combined_31707_Age_black$total_65Plus_vax_Sept_17) 
+sum(processed_combined_31707_Age_black$total_65Plus_vax_Oct_28) 
+sum(processed_combined_31707_Age_black$total_65Plus_vax_Nov_24) 
+sum(processed_combined_31707_Age_black$total_65Plus_vax_Dec_29) 
+sum(processed_combined_31707_Age_black$total_65Plus_vax_Jan_26) 
+sum(processed_combined_31707_Age_black$total_65Plus_vax_Feb_26)
+
 ```
 
-#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
+### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February 
+### Total contains the vaccination totals for each month
 
 ```{r}
-Gender_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
-                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361))
+Age_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          Age_05_09 = c(275, 277, 280, 282, 284, 283),
+                          Age_10_17 = c(1109, 1120, 1125, 1134, 1136, 1142),
+                          Age_18_44 = c(5903, 5931, 5943, 5947, 5960, 5958),
+                          Age_45_64 = c(5130, 5120, 5125, 5123, 5133, 5140),
+                          Age_65Plus = c(4200, 4204, 4195, 4206, 4202, 4189))
+
 ```
 
-### Converting the Gender_Sum_by_month_31707 to long data in order to plot
+### Converting the Age_Sum_by_month_31707 to long data in order to plot
 ```{r}
-Gender_Sum_by_month_31707_long = reshape2::melt(Gender_Sum_by_month_31707, id = "Date")
+Age_Sum_by_month_31707_long = reshape2::melt(Age_Sum_by_month_31707, id = "Date")
 ```
 
-### Changing "variable" to "Gender" and "value" to "Vaccination_total"
+### Changing "variable" to "Age" and "value" to "Vaccination_total"
 
 ```{r}
-Gender_Sum_by_month_31707_long <- Gender_Sum_by_month_31707_long %>% 
-  rename("Gender" = "variable") %>% 
+Age_Sum_by_month_31707_long <- Age_Sum_by_month_31707_long %>% 
+  rename("Age" = "variable") %>% 
   rename("Vaccination_total" = "value")
@@
 
-### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31707 zip code
+### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 31707 zip code
 
 ```{r}
-plot_Gender_black_vax_31707 = Gender_Sum_by_month_31707_long %>% 
+plot_Age_black_vax = Age_Sum_by_month_31707_long %>% 
                      ggplot(aes(Date, Vaccination_total, 
-                                group = Gender, 
-                                color = Gender)) + 
+                                group = Age, 
+                                color = Age)) + 
                      geom_point() + geom_line() + 
@@
 #adding a title to the plot 
-plot_Total_black_vax = plot_Total_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Gender in 31707") + 
+plot_Age_black_vax = plot_Age_black_vax + 
+  ggtitle("Vaccination Totals for Black population by Age in 31707") + 
   theme(plot.title = 
@@
   labs(y = "Vaccination Total")
-plot_Gender_black_vax_31707
+plot_Age_black_vax
 ```
@@
 ```{r}
-plotly_plot_Gender_black_vax_31707 = ggplotly(plot_Gender_black_vax_31707)
-plotly_plot_Gender_black_vax_31707
+plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
+plotly_plot_Age_black_vax
 ```

### 2023-04-06 code/analysis_code/UntitledQMD.qmd
Part 4

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept_17 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Richmond_vax_Sept <- filter(Sept_17, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_vax_Sept [Richmond_vax_Sept < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_vax_Sept = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Richmond_vax_Sept <- Richmond_vax_Sept  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Richmond_vax_Sept<- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asian residents.

```{r}
#Vax total Asian Residents
Richmond_vax_Sept$total_asian_vax <- Richmond_vax_Sept$total_masian_vax + Richmond_vax_Sept$total_fasian_vax

#Vax total for African American Residents 
Richmond_vax_Sept$total_black_vax <- Richmond_vax_Sept$total_mblack_vax + Richmond_vax_Sept$total_fblack_vax

#Vax total for Other Race Residents
Richmond_vax_Sept$total_orace_vax <- Richmond_vax_Sept$total_morace_vax + Richmond_vax_Sept$total_forace_vax

#Vax total for White Residents
Richmond_vax_Sept$total_white_vax <- Richmond_vax_Sept$total_mwhite_vax + Richmond_vax_Sept$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_vax_Sept_sums <- Richmond_vax_Sept[,c(1:4,185:204)]  
```

### Filtering the data by zip code

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Sept <-filter(Richmond_vax_Sept_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Sept <- filter(Richmond_vax_Sept_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Sept <- filter(Richmond_vax_Sept_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Sept

```{r}
colnames(Zip_31701_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")

#change column names for 31705 Sept  
colnames(Zip_31705_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")

#change column names for 31707 Sept  
colnames(Zip_31707_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")

```

### Deleting excess environment objects and saving data as processed_data_sept
```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_vax_Sept,Sept_17, codebook, Richmond_vax_Sept_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data", "Zip_31701", "processed_data_31701_Sept.rds")

saveRDS(Zip_31701_Sept, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Sept.rds")

saveRDS(Zip_31705_Sept, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Sept.rds")

saveRDS(Zip_31707_Sept, file = save_data_location3)

```

## Processing 2: Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of October

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Oct_28.xlsx")
Oct_28 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Oct_28)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Richmond_vax_Oct <- filter(Oct_28, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_vax_Oct [Richmond_vax_Oct < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_vax_Oct = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Richmond_vax_Oct <- Richmond_vax_Oct  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Richmond_vax_Oct<- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Richmond_vax_Oct$total_asian_vax <- Richmond_vax_Oct$total_masian_vax + Richmond_vax_Oct$total_fasian_vax

#Vax total for African American Residents 
Richmond_vax_Oct$total_black_vax <- Richmond_vax_Oct$total_mblack_vax + Richmond_vax_Oct$total_fblack_vax

#Vax total for Other Race Residents
Richmond_vax_Oct$total_orace_vax <- Richmond_vax_Oct$total_morace_vax + Richmond_vax_Oct$total_forace_vax

#Vax total for White Residents
Richmond_vax_Oct$total_white_vax <- Richmond_vax_Oct$total_mwhite_vax + Richmond_vax_Oct$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_vax_Oct_sums <- Richmond_vax_Oct[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Oct <-filter(Richmond_vax_Oct_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Oct <- filter(Richmond_vax_Oct_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Oct <- filter(Richmond_vax_Oct_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Oct

```{r}
colnames(Zip_31701_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")

#change column names for 31705 Oct  
colnames(Zip_31705_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")

#change column names for 31707 Oct  
colnames(Zip_31707_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")

```

## Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Oct

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_vax_Oct,Oct_28, codebook, Richmond_vax_Oct_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Oct.rds")

saveRDS(Zip_31701_Oct, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Oct.rds")

saveRDS(Zip_31705_Oct, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Oct.rds")

saveRDS(Zip_31707_Oct, file = save_data_location3)

```

## Processing 3: : Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of November

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Nov_24.xlsx")
Nov_24 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Nov_24)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Richmond_vax_Nov <- filter(Nov_24, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_vax_Nov [Richmond_vax_Nov < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_vax_Nov = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Richmond_vax_Nov <- Richmond_vax_Nov  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Richmond_vax_Nov<- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Richmond_vax_Nov$total_asian_vax <- Richmond_vax_Nov$total_masian_vax + Richmond_vax_Nov$total_fasian_vax

#Vax total for African American Residents 
Richmond_vax_Nov$total_black_vax <- Richmond_vax_Nov$total_mblack_vax + Richmond_vax_Nov$total_fblack_vax

#Vax total for Other Race Residents
Richmond_vax_Nov$total_orace_vax <- Richmond_vax_Nov$total_morace_vax + Richmond_vax_Nov$total_forace_vax

#Vax total for White Residents
Richmond_vax_Nov$total_white_vax <- Richmond_vax_Nov$total_mwhite_vax + Richmond_vax_Nov$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_vax_Nov_sums <- Richmond_vax_Nov[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Nov <-filter(Richmond_vax_Nov_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Nov <- filter(Richmond_vax_Nov_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Nov <- filter(Richmond_vax_Nov_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Nov

```{r}
colnames(Zip_31701_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

#change column names for 31705 Nov  
colnames(Zip_31705_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

#change column names for 31707 Nov  
colnames(Zip_31707_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")

```

## Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Nov

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_vax_Nov,Nov_24, codebook, Richmond_vax_Nov_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Nov.rds")

saveRDS(Zip_31701_Nov, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Nov.rds")

saveRDS(Zip_31705_Nov, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Nov.rds")

saveRDS(Zip_31707_Nov, file = save_data_location3)

```

## Processing 4: Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of December

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Dec_29.xlsx")
Dec_29 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Dec_29)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Richmond_vax_Dec <- filter(Dec_29, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_vax_Dec [Richmond_vax_Dec < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_vax_Dec = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Richmond_vax_Dec <- Richmond_vax_Dec  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Richmond_vax_Dec<- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Richmond_vax_Dec$total_asian_vax <- Richmond_vax_Dec$total_masian_vax + Richmond_vax_Dec$total_fasian_vax

#Vax total for African American Residents 
Richmond_vax_Dec$total_black_vax <- Richmond_vax_Dec$total_mblack_vax + Richmond_vax_Dec$total_fblack_vax

#Vax total for Other Race Residents
Richmond_vax_Dec$total_orace_vax <- Richmond_vax_Dec$total_morace_vax + Richmond_vax_Dec$total_forace_vax

#Vax total for White Residents
Richmond_vax_Dec$total_white_vax <- Richmond_vax_Dec$total_mwhite_vax + Richmond_vax_Dec$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_vax_Dec_sums <- Richmond_vax_Dec[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Dec <-filter(Richmond_vax_Dec_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Dec <- filter(Richmond_vax_Dec_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Dec <- filter(Richmond_vax_Dec_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Dec

```{r}
colnames(Zip_31701_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

#change column names for 31705 Dec  
colnames(Zip_31705_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

#change column names for 31707 Dec  
colnames(Zip_31707_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")

```

## Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Dec

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_vax_Dec,Dec_29, codebook, Richmond_vax_Dec_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Dec.rds")

saveRDS(Zip_31701_Dec, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Dec.rds")

saveRDS(Zip_31705_Dec, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Dec.rds")

saveRDS(Zip_31707_Dec, file = save_data_location3)

```

## Processing 5: Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of January

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Jan_26.xlsx")
Jan_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Jan_26)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Richmond_vax_Jan <- filter(Jan_26, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_vax_Jan [Richmond_vax_Jan < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_vax_Jan = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Richmond_vax_Jan <- Richmond_vax_Jan  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Richmond_vax_Jan<- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Richmond_vax_Jan$total_asian_vax <- Richmond_vax_Jan$total_masian_vax + Richmond_vax_Jan$total_fasian_vax

#Vax total for African American Residents 
Richmond_vax_Jan$total_black_vax <- Richmond_vax_Jan$total_mblack_vax + Richmond_vax_Jan$total_fblack_vax

#Vax total for Other Race Residents
Richmond_vax_Jan$total_orace_vax <- Richmond_vax_Jan$total_morace_vax + Richmond_vax_Jan$total_forace_vax

#Vax total for White Residents
Richmond_vax_Jan$total_white_vax <- Richmond_vax_Jan$total_mwhite_vax + Richmond_vax_Jan$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_vax_Jan_sums <- Richmond_vax_Jan[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Jan <-filter(Richmond_vax_Jan_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Jan <- filter(Richmond_vax_Jan_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Jan <- filter(Richmond_vax_Jan_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Jan

```{r}
colnames(Zip_31701_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

#change column names for 31705 Jan  
colnames(Zip_31705_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

#change column names for 31707 Jan  
colnames(Zip_31707_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")

```

## Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Jan

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_vax_Jan,Jan_26, codebook, Richmond_vax_Jan_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Jan.rds")

saveRDS(Zip_31701_Jan, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Jan.rds")

saveRDS(Zip_31705_Jan, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Jan.rds")

saveRDS(Zip_31707_Jan, file = save_data_location3)

```

## Processing 6: : Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of February

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Feb_26.xlsx")
Feb_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Feb_26)
```

## Cleaning

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 31701, 31705 and 31707
#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
Richmond_vax_Feb <- filter(Feb_26, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_vax_Feb [Richmond_vax_Feb < 0] <- 0  
```

### Adding new columns: total vax sum, total vax of all males, total vax of all females

```{r}
#Adding a column of the total vax for each census tract (for each observation)  
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_vax_Feb = sum(across(ends_with("VAX")), na.rm = T))

#total vax column for all males
Richmond_vax_Feb <- Richmond_vax_Feb  %>% rowwise() %>%   
  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))

#total vax column for all females
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total vax column for Asian male residents 
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American male residents
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race male residents
Richmond_vax_Feb<- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White male residents 
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total vax column for Asian female residents    
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))

#total vax column for African American female residents
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))

#total vax column for other race female residents
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))

#total vax column for White female residents 
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
```

### Adding new columns for Vax totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.

For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.

```{r}
#Vax total Asian Residents
Richmond_vax_Feb$total_asian_vax <- Richmond_vax_Feb$total_masian_vax + Richmond_vax_Feb$total_fasian_vax

#Vax total for African American Residents 
Richmond_vax_Feb$total_black_vax <- Richmond_vax_Feb$total_mblack_vax + Richmond_vax_Feb$total_fblack_vax

#Vax total for Other Race Residents
Richmond_vax_Feb$total_orace_vax <- Richmond_vax_Feb$total_morace_vax + Richmond_vax_Feb$total_forace_vax

#Vax total for White Residents
Richmond_vax_Feb$total_white_vax <- Richmond_vax_Feb$total_mwhite_vax + Richmond_vax_Feb$total_fwhite_vax  
```

### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_vax_Feb_sums <- Richmond_vax_Feb[,c(1:4,185:204)]  
```

### Filtering the data by zip code and creating a

```{r}
#filtering data by zip: 31701  
#All of the following FIPS or census tracts correspond to the 31701 zip code
Zip_31701_Feb <-filter(Richmond_vax_Feb_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")

#filtering data by zip: 31705  
#All of the following FIPS or census tracts correspond to the 31705 zip code
Zip_31705_Feb <- filter(Richmond_vax_Feb_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")

#filtering data by zip: 31707  
#All of the following FIPS or census tracts correspond to the 31707 zip code
Zip_31707_Feb <- filter(Richmond_vax_Feb_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")

```

### Change the column names for 31701 Feb

```{r}
colnames(Zip_31701_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")

#change column names for 31705 Feb  
colnames(Zip_31705_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")

#change column names for 31707 Feb  
colnames(Zip_31707_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")

```

## Cleaning is done, now lets save the processed data

### Deleting excess environment objects and saving data as processed_data_Feb

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_vax_Feb,Feb_26, codebook, Richmond_vax_Feb_sums)

#saving clean data for 31701 to processed_data folder
save_data_location = here("data", "processed_data","Zip_31701", "processed_data_31701_Feb.rds")

saveRDS(Zip_31701_Feb, file = save_data_location)

#saving clean data for 31705 to processed_data folder
save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Feb.rds")

saveRDS(Zip_31705_Feb, file = save_data_location2)

#saving clean data for 31707 to processed_data folder
save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Feb.rds")

saveRDS(Zip_31707_Feb, file = save_data_location3)

```

### 2023-04-06 code/processing_code/processingcode.qmd
Part 4

@@
 ---
-title: "Processing Code for September Data"
+title: "Processing Code"
 output: 
@@
 
-# Processing 1
+# Processing Section for Vaccination Data in Albany, GA
+
+## Processing 1: Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of September
 
@@
 
-## Cleaning
-
 ### Separating the data by zip code and removing negative values
@@
 
-### Filtering the data by zip code and creating a
+### Filtering the data by zip code
 
@@
 
-# Processing 2
-
-Load needed packages. make sure they are installed.
-
-```{r}
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-```
+## Processing 2: Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of October
 
@@
 
-# Cleaning is done, now lets save the processed data
+ is done, now lets save the processed data
 
@@
 
-# Processing 3
-
-Load needed packages. make sure they are installed.
-
-```{r}
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-```
+## Processing 3: : Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of November
 
@@
 
-# Processing 4
-
-Load needed packages. make sure they are installed.
-
-```{r}
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-```
+## Processing 4: Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of December
 
@@
 
-# Processing script 5
-
-Load needed packages. make sure they are installed.
-
-```{r}
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-```
+## Processing 5: Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of January
 
@@
 
-# Processing 6
-
-Load needed packages. make sure they are installed.
-
-```{r}
-library(here)
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-```
+## Processing 6: : Getting Vaccination Totals by Target Zips 31701, 31705 & 31707 for the Month of February
 
@@
 ```
+
+# Processing for Vaccination Data for Our Control Richmond County
+
+## Processing 1: Getting Vaccination Totals by Control Zip 31701

### 2023-04-07 code/analysis_code/exploratoryanalysis.qmd
Part 4

---
title: "Total Amount of Vaccinations Among Black population in the Target Zip Codes"
output: 
  html_document
echo: FALSE 
---

# Library

```{r}
library(here)
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr)
library(ggplot2)
library(reshape2) #reshaping data from wide to long format
library(plotly) #creating an interactive line plot 
```

# Analysis of the Amount of Vaccinations Among Black population in Each Target Zip Code

## Zip code: 31701

### Loading the processed data in order to begin analysis for the 31701 zip code

```{r}
# Loading September 31701
data_location1 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Sept.rds")
processed_data_31701_Sept = readRDS(data_location1)

# Loading October 31701 and joining september and october 
data_location2 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Oct.rds")
processed_data_31701_Oct = readRDS(data_location2)
processed_combined <- left_join(processed_data_31701_Sept, processed_data_31701_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31701
data_location3 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Nov.rds")
processed_data_31701_Nov = readRDS(data_location3)
processed_combined <- left_join(processed_combined, processed_data_31701_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31701
data_location4 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Dec.rds")
processed_data_31701_Dec = readRDS(data_location4)
processed_combined <- left_join(processed_combined, processed_data_31701_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31701
data_location5 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Jan.rds")
processed_data_31701_Jan = readRDS(data_location5)
processed_combined <- left_join(processed_combined, processed_data_31701_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31701
data_location6 = here("data", "processed_data", "Zip_31701", "processed_data_31701_Feb.rds")
processed_data_31701_Feb = readRDS(data_location6)
processed_combined_31701 <- left_join(processed_combined, processed_data_31701_Feb, by = c("FIPS" = "FIPS"))
```

### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
saveRDS(processed_combined_31701, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31701 zip code

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31701_total_black = processed_combined_31701 %>% select(starts_with("total_black_vax"))
```

### Summing each Month variable by census tract

```{r}
#Summing each column for each month 
sum(processed_combined_31701_total_black$total_black_vax_Sept_17) 
sum(processed_combined_31701_total_black$total_black_vax_Oct_28) 
sum(processed_combined_31701_total_black$total_black_vax_Nov_24) 
sum(processed_combined_31701_total_black$total_black_vax_Dec_29) 
sum(processed_combined_31701_total_black$total_black_vax_Jan_26) 
sum(processed_combined_31701_total_black$total_black_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Total = c(7797, 7780, 7755, 7773, 7792, 7817)) 
```

### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 31701 zip code

```{r}
plot_31701_total_black_vax = Sum_by_month_31701 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot 
plot_31701_total_black_vax = plot_31701_total_black_vax + 
  ggtitle("Vaccination Totals for Black population in the 31701 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
plot_31701_total_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_31701_total_black_vax = ggplotly(plot_31701_total_black_vax)
plotly_31701_total_black_vax
```

## Zip code: 31705

### Loading the processed data in order to begin analysis for the 31705 zip code

```{r}
#Loading September 31705
data_location7 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Sept.rds")
processed_data_31705_Sept = readRDS(data_location7)

# Loading October 31705 and joining september and october 
data_location8 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Oct.rds")
processed_data_31705_Oct = readRDS(data_location8)
processed_combined <- left_join(processed_data_31705_Sept, processed_data_31705_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31705
data_location9 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Nov.rds")
processed_data_31705_Nov = readRDS(data_location9)
processed_combined <- left_join(processed_combined, processed_data_31705_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31705
data_location10= here("data", "processed_data", "Zip_31705", "processed_data_31705_Dec.rds")
processed_data_31705_Dec = readRDS(data_location10)
processed_combined <- left_join(processed_combined, processed_data_31705_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31705
data_location11 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Jan.rds")
processed_data_31705_Jan = readRDS(data_location11)
processed_combined <- left_join(processed_combined, processed_data_31705_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31705
data_location12 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Feb.rds")
processed_data_31705_Feb = readRDS(data_location12)
processed_combined_31705 <- left_join(processed_combined, processed_data_31705_Feb, by = c("FIPS" = "FIPS"))

```

### Saving the combined data for Black population for future use

```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
saveRDS(processed_combined_31705, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31705 zip code

```{r}
processed_combined_31705_total_black = processed_combined_31705 %>% select(starts_with("total_black_vax"))
```

### Summing each column for each month

```{r}

sum(processed_combined_31705_total_black$total_black_vax_Sept_17)

sum(processed_combined_31705_total_black$total_black_vax_Oct_28)

sum(processed_combined_31705_total_black$total_black_vax_Nov_24)

sum(processed_combined_31705_total_black$total_black_vax_Dec_29)

sum(processed_combined_31705_total_black$total_black_vax_Jan_26)

sum(processed_combined_31705_total_black$total_black_vax_Feb_26)
```

### Separating sums by month into a data fram consisting of two variables: Date and Total

### Date contains months September-February

### Total contains the vaccination totals for each month

```{r}
Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                                Total = c(10383, 10327, 10303, 10301, 10337, 10348)) 
```

### ggplot creating a line plot for the vaccination totals by month for the 31705 zip code African American population

```{r}
plot_31705_total_black_vax = Sum_by_month_31705 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot 
plot_31705_total_black_vax = plot_31705_total_black_vax + 
  ggtitle("Vaccination Totals for Black population in the 31705 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
```

## Zip code: 31707

### Loading the processed data in order to begin analysis for the 31707 zip code

```{r}
# Loading September 31707
data_location13 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Sept.rds")
processed_data_31707_Sept = readRDS(data_location13)

# Loading October 31707 and joining september and october 
data_location14 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Oct.rds")
processed_data_31707_Oct = readRDS(data_location14)
processed_combined <- left_join(processed_data_31707_Sept, processed_data_31707_Oct, by = c("FIPS" = "FIPS"))

# Loading November 31707
data_location15 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Nov.rds")
processed_data_31707_Nov = readRDS(data_location15)
processed_combined <- left_join(processed_combined, processed_data_31707_Nov, by = c("FIPS" = "FIPS"))

# Loading December 31707
data_location16= here("data", "processed_data", "Zip_31707", "processed_data_31707_Dec.rds")
processed_data_31707_Dec = readRDS(data_location16)
processed_combined <- left_join(processed_combined, processed_data_31707_Dec, by = c("FIPS" = "FIPS"))

# Loading January 31707
data_location17 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Jan.rds")
processed_data_31707_Jan = readRDS(data_location17)
processed_combined <- left_join(processed_combined, processed_data_31707_Jan, by = c("FIPS" = "FIPS"))

# Loading February 31707
data_location18 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Feb.rds")
processed_data_31707_Feb = readRDS(data_location18)
processed_combined_31707 <- left_join(processed_combined, processed_data_31707_Feb, by = c("FIPS" = "FIPS"))
```

### Saving the combined data for Black population for future use

```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
saveRDS(processed_combined_31707, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31707 zip code

```{r}
processed_combined_31707_total_black = processed_combined_31707 %>% select(starts_with("total_black_vax"))

```

### Summing each column for each month

```{r}
sum(processed_combined_31707_total_black$total_black_vax_Sept_17) 
sum(processed_combined_31707_total_black$total_black_vax_Oct_28) 
sum(processed_combined_31707_total_black$total_black_vax_Nov_24) 
sum(processed_combined_31707_total_black$total_black_vax_Dec_29) 
sum(processed_combined_31707_total_black$total_black_vax_Jan_26) 
sum(processed_combined_31707_total_black$total_black_vax_Feb_26)
```

### Separating sums by month into a data fram consisting of two variables: Date and Total

### Date contains months September-February

### Total contains the vaccination totals for each month

```{r}
Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                                Total = c(9035, 8985, 8997, 9005, 9035, 9042))

```

###ggplot: creating a line plot for the vaccination totals by month for the 31707 zip code African American population

```{r}
plot_31707_total_black_vax = Sum_by_month_31707 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot
plot_31707_total_black_vax = plot_31707_total_black_vax + 
  ggtitle("Vaccination Totals for Black population in the 31707 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
```

## Comparing plots the 3 target zip codes

### Taking a look at each of the line plots for the different zip codes and comparing the total vaccinations by month in the African American population

```{r}
plot_31701_total_black_vax
plot_31705_total_black_vax
plot_31707_total_black_vax
```

### Combining all plots into one

```{r}
## Creating a new dataframe called "Total_black_vax" by combining the vaccination totals from 31701 and 3105 by using leftjoin() by the Date variable 
Total_black_vax = left_join(Sum_by_month_31701, Sum_by_month_31705, by = c("Date" = "Date"))

```

### Combining the new dataframe "Total_black_vax" with the 31707 zip vaccination totals so all three target zip codes are in the same data frame

```{r}
Total_black_vax = left_join(Total_black_vax, Sum_by_month_31707, by = c("Date" = "Date"))
```

### Renaming column totals to match the zip

### Haing to do this because the the total variable names are Total.x, Total.y and Total

### Need to specificy each total variable by zip using the following code

```{r}
Total_black_vax <- Total_black_vax %>% 
  rename("31701" = "Total.x") %>% 
  rename("31705" = "Total.y") %>%
  rename("31707" = "Total")
```

### Converting data to long format in order to create a line plot using ggplot

```{r}
Total_black_vax_long = reshape2::melt(Total_black_vax, id = "Date")
```

### Changing the variable name "variable" to "Zip_code" and "value" to "Vaccination_total" for clarity

### Doing this because when the data was converted to a long format the column names we changed to "variable" which refers to the zip code and "value" which refers to the amount of vaccinations

### Changing "variable" to "Zip_code" and "value" to "Vaccination_total"

```{r}
Total_black_vax_long <- Total_black_vax_long %>% 
  rename("Zip_code" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### ggplot: Making a line graph that compares the vaccination totals for each African American population in each zip code

```{r}
plot_Total_black_vax = Total_black_vax_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Zip_code, 
                                color = Zip_code)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Total_black_vax = plot_Total_black_vax + 
  ggtitle("Vaccination Totals for Black population in Target Zip Codes") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Total_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_Total_black_vax = ggplotly(plot_Total_black_vax)
plotly_Total_black_vax
```

# Total Amount of vaccinations by Gender Among Black Population in the Target Zip Codes

## Zip code: 31701

### Loading the `processed_combined_31701.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
processed_combined_31701 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31701

#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31701_Gender_black = processed_combined_31701 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31701_Gender_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31701_Gender_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420))
```

### Converting the Gender_Sum_by_month_31701 to long data in order to plot
```{r}
Gender_Sum_by_month_31701_long = reshape2::melt(Gender_Sum_by_month_31701, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31701_long <- Gender_Sum_by_month_31701_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31701 zip code

```{r}
plot_Gender_black_vax = Gender_Sum_by_month_31701_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Gender_black_vax = plot_Gender_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31701") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
plotly_plot_Gender_black_vax
```

## Zip code: 31705

### Loading the `processed_combined_31705.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
processed_combined_31705 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31705

#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31705_Gender_black = processed_combined_31705 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31705_Gender_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31705_Gender_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949))
```

### Converting the Gender_Sum_by_month_31705 to long data in order to plot
```{r}
Gender_Sum_by_month_31705_long = reshape2::melt(Gender_Sum_by_month_31705, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31705_long <- Gender_Sum_by_month_31705_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31705 zip code

```{r}
plot_Gender_black_vax_31705 = Gender_Sum_by_month_31705_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Gender_black_vax = plot_Gender_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31705") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax_31705
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax_31705 = ggplotly(plot_Gender_black_vax_31705)
plotly_plot_Gender_black_vax_31705
```

## Zip code: 31707

### Loading the `processed_combined_31707.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
processed_combined_31707 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 31707

#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_31707_Gender_black = processed_combined_31707 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Sept_17) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Oct_28) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Nov_24) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Dec_29) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Jan_26) 
sum(processed_combined_31707_Gender_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Sept_17) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Oct_28) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Nov_24) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Dec_29) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Jan_26) 
sum(processed_combined_31707_Gender_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361))
```

### Converting the Gender_Sum_by_month_31707 to long data in order to plot
```{r}
Gender_Sum_by_month_31707_long = reshape2::melt(Gender_Sum_by_month_31707, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_31707_long <- Gender_Sum_by_month_31707_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 31707 zip code

```{r}
plot_Gender_black_vax_31707 = Gender_Sum_by_month_31707_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Gender_black_vax = plot_Gender_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 31707") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax_31707
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax_31707 = ggplotly(plot_Gender_black_vax_31707)
plotly_plot_Gender_black_vax_31707
```

# Thoughts regarding exploratory analysis 
In each zip code there is a large gap between females that are Black and males that are Black. I am pretty taken back by how large the gap is.

# Total Amount of vaccinations by Age among Black population in the target zip codes

## Zip code: 31701

### Loading the `processed_combined_31701.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31701", "processed_combined_31701.rds")
processed_combined_31701 = readRDS(save_data_location)
```

### Dropping all variables except for black age groups within 31701

#### Using the select() function to subset the data frame by columns that start with an age group

```{r}
processed_combined_31701_Age_black = processed_combined_31701 %>% 
  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
```

### Summing each Month variable by age group by month
```{r}
#Summing for ages 5-9 by month 
sum(processed_combined_31701_Age_black$total_05_09_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_05_09_vax_Feb_26)

#Summing for ages 10-17 by month 
sum(processed_combined_31701_Age_black$total_10_17_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_10_17_vax_Feb_26)

#Summing for ages 18-44 by month 
sum(processed_combined_31701_Age_black$total_18_44_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_18_44_vax_Feb_26)

#Summing for ages 45-64 by month 
sum(processed_combined_31701_Age_black$total_45_64_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_45_64_vax_Feb_26)

#Summing for ages 65+ by month 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Sept_17) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Oct_28) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Nov_24) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Dec_29) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Jan_26) 
sum(processed_combined_31701_Age_black$total_65Plus_vax_Feb_26)

```

### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Age_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Age_05_09 = c(123, 126, 126, 127, 127, 128),
                          Age_10_17 = c(642, 635, 633, 637, 630, 629),
                          Age_18_44 = c(3310, 3316, 3313, 3323, 3325, 3331),
                          Age_45_64 = c(3229, 3244, 3239, 3259, 3264, 3258),
                          Age_65Plus = c(2576, 2570, 2569, 2574, 2568, 2572))

```

### Converting the Age_Sum_by_month_31701 to long data in order to plot
```{r}
Age_Sum_by_month_31701_long = reshape2::melt(Age_Sum_by_month_31701, id = "Date")
```

### Changing "variable" to "Age" and "value" to "Vaccination_total"

```{r}
Age_Sum_by_month_31701_long <- Age_Sum_by_month_31701_long %>% 
  rename("Age" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 31701 zip code

```{r}
plot_Age_black_vax = Age_Sum_by_month_31701_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Age, 
                                color = Age)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Age_black_vax = plot_Age_black_vax + 
  ggtitle("Vaccination Totals for Black population by Age in 31701") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Age_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
plotly_plot_Age_black_vax
```

------------------------------------------------------------------------
## Zip code: 31705

### Loading the `processed_combined_31705.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31705", "processed_combined_31705.rds")
processed_combined_31705 = readRDS(save_data_location)
```

### Dropping all variables except for black age groups within 31705

#### Using the select() function to subset the data frame by columns that start with an age group

```{r}
processed_combined_31705_Age_black = processed_combined_31705 %>% 
  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
```

### Summing each Month variable by age group by month
```{r}
#Summing for ages 5-9 by month 
sum(processed_combined_31705_Age_black$total_05_09_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_05_09_vax_Feb_26)

#Summing for ages 10-17 by month 
sum(processed_combined_31705_Age_black$total_10_17_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_10_17_vax_Feb_26)

#Summing for ages 18-44 by month 
sum(processed_combined_31705_Age_black$total_18_44_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_18_44_vax_Feb_26)

#Summing for ages 45-64 by month 
sum(processed_combined_31705_Age_black$total_45_64_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_45_64_vax_Feb_26)

#Summing for ages 65+ by month 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Sept_17) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Oct_28) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Nov_24) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Dec_29) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Jan_26) 
sum(processed_combined_31705_Age_black$total_65Plus_vax_Feb_26)

```

### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February 
### Total contains the vaccination totals for each month

```{r}
Age_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Age_05_09 = c(210, 211, 209, 201, 202, 202),
                          Age_10_17 = c(1159, 1171, 1169, 1163, 1168, 1167),
                          Age_18_44 = c(5201, 5215, 5239, 5263, 5278, 5281), #pretty notable increase for ages 18-44 
                          Age_45_64 = c(4380, 4381, 4366, 4364, 4367, 4367),
                          Age_65Plus = c(2983, 3000, 2997, 2997, 2994, 2989))

```

### Converting the Age_Sum_by_month_31705 to long data in order to plot
```{r}
Age_Sum_by_month_31705_long = reshape2::melt(Age_Sum_by_month_31705, id = "Date")
```

### Changing "variable" to "Age" and "value" to "Vaccination_total"

```{r}
Age_Sum_by_month_31705_long <- Age_Sum_by_month_31705_long %>% 
  rename("Age" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 31705 zip code

```{r}
plot_Age_black_vax = Age_Sum_by_month_31705_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Age, 
                                color = Age)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Age_black_vax = plot_Age_black_vax + 
  ggtitle("Vaccination Totals for Black population by Age in 31705") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Age_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
plotly_plot_Age_black_vax
```

------------------------------------------------------------------------

## Zip code: 31707

### Loading the `processed_combined_31707.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_31707", "processed_combined_31707.rds")
processed_combined_31707 = readRDS(save_data_location)
```

### Dropping all variables except for black age groups within 31707

#### Using the select() function to subset the data frame by columns that start with an age group

```{r}
processed_combined_31707_Age_black = processed_combined_31707 %>% 
  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
```

### Summing each Month variable by age group by month
```{r}
#Summing for ages 5-9 by month 
sum(processed_combined_31707_Age_black$total_05_09_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_05_09_vax_Feb_26)

#Summing for ages 10-17 by month 
sum(processed_combined_31707_Age_black$total_10_17_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_10_17_vax_Feb_26)

#Summing for ages 18-44 by month 
sum(processed_combined_31707_Age_black$total_18_44_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_18_44_vax_Feb_26)

#Summing for ages 45-64 by month 
sum(processed_combined_31707_Age_black$total_45_64_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_45_64_vax_Feb_26)

#Summing for ages 65+ by month 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Sept_17) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Oct_28) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Nov_24) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Dec_29) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Jan_26) 
sum(processed_combined_31707_Age_black$total_65Plus_vax_Feb_26)

```

# The Amount of Vaccinations Among Black population in the Control Zip Code

## Zip code: 30815

### Loading the processed data in order to begin analysis for the 30815 zip code

```{r}
# Loading September 30815
data_location1 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Sept.rds")
processed_data_30815_Sept = readRDS(data_location1)

# Loading October 30815 and joining september and october 
data_location2 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Oct.rds")
processed_data_30815_Oct = readRDS(data_location2)
processed_combined <- left_join(processed_data_30815_Sept, processed_data_30815_Oct, by = c("FIPS" = "FIPS"))

# Loading November 30815
data_location3 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Nov.rds")
processed_data_30815_Nov = readRDS(data_location3)
processed_combined <- left_join(processed_combined, processed_data_30815_Nov, by = c("FIPS" = "FIPS"))

# Loading December 30815
data_location4 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Dec.rds")
processed_data_30815_Dec = readRDS(data_location4)
processed_combined <- left_join(processed_combined, processed_data_30815_Dec, by = c("FIPS" = "FIPS"))

# Loading January 30815
data_location5 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Jan.rds")
processed_data_30815_Jan = readRDS(data_location5)
processed_combined <- left_join(processed_combined, processed_data_30815_Jan, by = c("FIPS" = "FIPS"))

# Loading February 30815
data_location6 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Feb.rds")
processed_data_30815_Feb = readRDS(data_location6)
processed_combined_30815 <- left_join(processed_combined, processed_data_30815_Feb, by = c("FIPS" = "FIPS"))
```

### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender

```{r}
save_data_location = here("data", "processed_data","Zip_30815", "processed_combined_30815.rds")
saveRDS(processed_combined_30815, file = save_data_location)
```

### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 30815 zip code

#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_30815_total_black = processed_combined_30815 %>% select(starts_with("total_black_vax"))
```

### Summing each Month variable by census tract

```{r}
#Summing each column for each month 
sum(processed_combined_30815_total_black$total_black_vax_Sept_17) 
sum(processed_combined_30815_total_black$total_black_vax_Oct_28) 
sum(processed_combined_30815_total_black$total_black_vax_Nov_24) 
sum(processed_combined_30815_total_black$total_black_vax_Dec_29) 
sum(processed_combined_30815_total_black$total_black_vax_Jan_26) 
sum(processed_combined_30815_total_black$total_black_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Total = c(8758, 8688, 8639, 8593, 8623, 8666)) 
```

### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 30815 zip code

```{r}
plot_30815_total_black_vax = Sum_by_month_30815 %>% 
  ggplot(aes(Date, Total, group=1)) + 
  geom_point() + geom_line() + 
  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))

#Adding a customization title for the plot 
plot_30815_total_black_vax = plot_30815_total_black_vax + 
  ggtitle("Vaccination Totals for Black population in the 30815 Zipcode") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic"))
plot_30815_total_black_vax
```
### Turning ggplot into an interactive plot

```{r}
plotly_30815_total_black_vax = ggplotly(plot_30815_total_black_vax)
plotly_30815_total_black_vax
```

# The Amount of Vaccinations by Gender Among Black Population in the Control Zip Code

## Zip code: 30815

### Loading the `processed_combined_30815.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_30815", "processed_combined_30815.rds")
processed_combined_30815 = readRDS(save_data_location)
```

### Dropping all variables except for black males and black females within 30815

#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month

```{r}
processed_combined_30815_Gender_black = processed_combined_30815 %>% select(matches("^total_(mblack|fblack)_vax"))
```

### Summing each Month variable by census tract
```{r}
#Summing for black males by month 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Sept_17) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Oct_28) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Nov_24) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Dec_29) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Jan_26) 
sum(processed_combined_30815_Gender_black$total_mblack_vax_Feb_26)

#summing for black females by month 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Sept_17) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Oct_28) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Nov_24) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Dec_29) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Jan_26) 
sum(processed_combined_30815_Gender_black$total_fblack_vax_Feb_26) 
```

#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Gender_Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Male_Total = c(3526, 3514, 3503, 3484, 3489, 3498),
                          Female_Total = c(5232, 5174, 5136, 5109, 5134, 5168))
```

### Converting the Gender_Sum_by_month_30815 to long data in order to plot
```{r}
Gender_Sum_by_month_30815_long = reshape2::melt(Gender_Sum_by_month_30815, id = "Date")
```

### Changing "variable" to "Gender" and "value" to "Vaccination_total"

```{r}
Gender_Sum_by_month_30815_long <- Gender_Sum_by_month_30815_long %>% 
  rename("Gender" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 30815 zip code

```{r}
plot_Gender_black_vax = Gender_Sum_by_month_30815_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Gender, 
                                color = Gender)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Gender_black_vax = plot_Gender_black_vax + 
  ggtitle("Vaccination Totals for Black population by Gender in 30815") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Gender_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
plotly_plot_Gender_black_vax
```

# Total Amount of vaccinations by Age among Black population in the Control Zip Code

## Zip code: 30815

### Loading the `processed_combined_30815.rds` file from `analysiscode.qmd`

```{r}
save_data_location = here("data", "processed_data","Zip_30815", "processed_combined_30815.rds")
processed_combined_30815 = readRDS(save_data_location)
```

### Dropping all variables except for black age groups within 30815

#### Using the select() function to subset the data frame by columns that start with an age group

```{r}
processed_combined_30815_Age_black = processed_combined_30815 %>% 
  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
```

### Summing each Month variable by age group by month
```{r}
#Summing for ages 5-9 by month 
sum(processed_combined_30815_Age_black$total_05_09_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_05_09_vax_Feb_26)

#Summing for ages 10-17 by month 
sum(processed_combined_30815_Age_black$total_10_17_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_10_17_vax_Feb_26)

#Summing for ages 18-44 by month 
sum(processed_combined_30815_Age_black$total_18_44_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_18_44_vax_Feb_26)

#Summing for ages 45-64 by month 
sum(processed_combined_30815_Age_black$total_45_64_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_45_64_vax_Feb_26)

#Summing for ages 65+ by month 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Sept_17) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Oct_28) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Nov_24) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Dec_29) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Jan_26) 
sum(processed_combined_30815_Age_black$total_65Plus_vax_Feb_26)

```

### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month

```{r}
Age_Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Age_05_09 = c(201, 213, 211, 215, 211, 211),
                          Age_10_17 = c(1116, 1113, 1115, 1118, 1115, 1114),
                          Age_18_44 = c(5001, 5009, 5020, 5026, 5025, 5032),
                          Age_45_64 = c(5026, 5017, 5020, 5009, 5013, 5020),
                          Age_65Plus = c(2816, 2825, 2827, 2814, 2828, 2832))

```

### Converting the Age_Sum_by_month_30815 to long data in order to plot
```{r}
Age_Sum_by_month_30815_long = reshape2::melt(Age_Sum_by_month_30815, id = "Date")
```

### Changing "variable" to "Age" and "value" to "Vaccination_total"

```{r}
Age_Sum_by_month_30815_long <- Age_Sum_by_month_30815_long %>% 
  rename("Age" = "variable") %>% 
  rename("Vaccination_total" = "value")
```

### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 30815 zip code

```{r}
plot_Age_black_vax = Age_Sum_by_month_30815_long %>% 
                     ggplot(aes(Date, Vaccination_total, 
                                group = Age, 
                                color = Age)) + 
                     geom_point() + geom_line() + 
                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
#adding a title to the plot 
plot_Age_black_vax = plot_Age_black_vax + 
  ggtitle("Vaccination Totals for Black population by Age in 30815") + 
  theme(plot.title = 
          element_text(color="red", size=10, face="bold.italic")) + 
  labs(y = "Vaccination Total")
plot_Age_black_vax
```

### Turning ggplot into an interactive plot

```{r}
plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
plotly_plot_Age_black_vax
```

### 2023-04-07 code/processing_code/processingcode.qmd
Part 4

@@
 
-# Processing Section for Vaccination Data in Albany, GA
+# Processing Section for Target Zip Codes
 
@@
 
- is done, now lets save the processed data
-
 ### Deleting excess environment objects and saving data as processed_data_Oct
@@
 
- is done, now lets save the processed data
-
 ### Deleting excess environment objects and saving data as processed_data_Nov
@@
 
- is done, now lets save the processed data
-
 ### Deleting excess environment objects and saving data as processed_data_Dec
@@
 
- is done, now lets save the processed data
-
 ### Deleting excess environment objects and saving data as processed_data_Jan
@@
 
- is done, now lets save the processed data
-
 ### Deleting excess environment objects and saving data as processed_data_Feb
@@
 
-# Processing for Vaccination Data for Our Control Richmond County
-
-## Processing 1: Getting Vaccination Totals by Control Zip 31701
+# Processing Section for Control Zip Code
+
+## Processing 1: Getting Vaccination Totals by Control Zip Code 30815 in the Month of September
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Sept_17.xlsx")
+Sept_17 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Richmond_vax_Sept <- filter(Sept_17, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_vax_Sept [Richmond_vax_Sept < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_vax_Sept = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Richmond_vax_Sept <- Richmond_vax_Sept  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Richmond_vax_Sept<- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asian residents.
+
+```{r}
+#Vax total Asian Residents
+Richmond_vax_Sept$total_asian_vax <- Richmond_vax_Sept$total_masian_vax + Richmond_vax_Sept$total_fasian_vax
+
+#Vax total for African American Residents 
+Richmond_vax_Sept$total_black_vax <- Richmond_vax_Sept$total_mblack_vax + Richmond_vax_Sept$total_fblack_vax
+
+#Vax total for Other Race Residents
+Richmond_vax_Sept$total_orace_vax <- Richmond_vax_Sept$total_morace_vax + Richmond_vax_Sept$total_forace_vax
+
+#Vax total for White Residents
+Richmond_vax_Sept$total_white_vax <- Richmond_vax_Sept$total_mwhite_vax + Richmond_vax_Sept$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_vax_Sept <- Richmond_vax_Sept %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_vax_Sept_sums <- Richmond_vax_Sept[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+#All of the following FIPS or census tracts correspond to the 30815 zip code
+Zip_30815_Sept <-filter(Richmond_vax_Sept_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+```
+Not that the new data set `Zip_30815_Sept` only has 7 observations because the orinigal data set only had 7 of the 13 census tracts found within the zip code 30815
+
+### Change the column names for 30815 Sept
+```{r}
+colnames(Zip_30815_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")
+```
+
+### Deleting excess environment objects and saving data as processed_data_sept
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_vax_Sept,Sept_17, codebook, Richmond_vax_Sept_sums)
+
+#saving clean data for 30815 to processed_data folder
+save_data_location4 = here("data", "processed_data", "Zip_30815", "processed_data_30815_Sept.rds")
+
+saveRDS(Zip_30815_Sept, save_data_location4)
+```
+
+## Processing 2: Getting Vaccination Totals by Target Zips 30815 for the Month of October
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Oct_28.xlsx")
+Oct_28 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Richmond_vax_Oct <- filter(Oct_28, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_vax_Oct [Richmond_vax_Oct < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_vax_Oct = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Richmond_vax_Oct <- Richmond_vax_Oct  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Richmond_vax_Oct<- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.
+
+```{r}
+#Vax total Asian Residents
+Richmond_vax_Oct$total_asian_vax <- Richmond_vax_Oct$total_masian_vax + Richmond_vax_Oct$total_fasian_vax
+
+#Vax total for African American Residents 
+Richmond_vax_Oct$total_black_vax <- Richmond_vax_Oct$total_mblack_vax + Richmond_vax_Oct$total_fblack_vax
+
+#Vax total for Other Race Residents
+Richmond_vax_Oct$total_orace_vax <- Richmond_vax_Oct$total_morace_vax + Richmond_vax_Oct$total_forace_vax
+
+#Vax total for White Residents
+Richmond_vax_Oct$total_white_vax <- Richmond_vax_Oct$total_mwhite_vax + Richmond_vax_Oct$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_vax_Oct <- Richmond_vax_Oct %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_vax_Oct_sums <- Richmond_vax_Oct[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+#All of the following FIPS or census tracts correspond to the 30815 zip code
+
+Zip_30815_Oct <-filter(Richmond_vax_Oct_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+```
+
+### Change the column names for 30815 Oct
+
+```{r}
+colnames(Zip_30815_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")
+```
+
+### Deleting excess environment objects and saving data as processed_data_Oct
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_vax_Oct,Oct_28, codebook, Richmond_vax_Oct_sums)
+
+#saving clean data for 30815 to processed_data folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_30815_Oct.rds")
+
+saveRDS(Zip_30815_Oct, file = save_data_location4)
+```
+
+## Processing 3: : Getting Vaccination Totals by Target Zips 30815 for the Month of November
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Nov_24.xlsx")
+Nov_24 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+## Summary of data
+
+```{r}
+glimpse(Nov_24)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Richmond_vax_Nov <- filter(Nov_24, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_vax_Nov [Richmond_vax_Nov < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_vax_Nov = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Richmond_vax_Nov <- Richmond_vax_Nov  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Richmond_vax_Nov<- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.
+
+```{r}
+#Vax total Asian Residents
+Richmond_vax_Nov$total_asian_vax <- Richmond_vax_Nov$total_masian_vax + Richmond_vax_Nov$total_fasian_vax
+
+#Vax total for African American Residents 
+Richmond_vax_Nov$total_black_vax <- Richmond_vax_Nov$total_mblack_vax + Richmond_vax_Nov$total_fblack_vax
+
+#Vax total for Other Race Residents
+Richmond_vax_Nov$total_orace_vax <- Richmond_vax_Nov$total_morace_vax + Richmond_vax_Nov$total_forace_vax
+
+#Vax total for White Residents
+Richmond_vax_Nov$total_white_vax <- Richmond_vax_Nov$total_mwhite_vax + Richmond_vax_Nov$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_vax_Nov <- Richmond_vax_Nov %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_vax_Nov_sums <- Richmond_vax_Nov[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+#All of the following FIPS or census tracts correspond to the 30815 zip code
+
+Zip_30815_Nov <-filter(Richmond_vax_Nov_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+
+```
+
+### Change the column names for 30815 Nov
+
+```{r}
+colnames(Zip_30815_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")
+```
+
+### Deleting excess environment objects and saving data as processed_data_Nov
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_vax_Nov,Nov_24, codebook, Richmond_vax_Nov_sums)
+
+#saving clean data for 30815 to processed_data folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_30815_Nov.rds")
+
+saveRDS(Zip_30815_Nov, file = save_data_location4)
+
+```
+
+## Processing 4: Getting Vaccination Totals by Target Zips 30815 for the Month of December
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Dec_29.xlsx")
+Dec_29 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+## Summary of data
+
+```{r}
+glimpse(Dec_29)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Richmond_vax_Dec <- filter(Dec_29, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_vax_Dec [Richmond_vax_Dec < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_vax_Dec = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Richmond_vax_Dec <- Richmond_vax_Dec  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Richmond_vax_Dec<- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.
+
+```{r}
+#Vax total Asian Residents
+Richmond_vax_Dec$total_asian_vax <- Richmond_vax_Dec$total_masian_vax + Richmond_vax_Dec$total_fasian_vax
+
+#Vax total for African American Residents 
+Richmond_vax_Dec$total_black_vax <- Richmond_vax_Dec$total_mblack_vax + Richmond_vax_Dec$total_fblack_vax
+
+#Vax total for Other Race Residents
+Richmond_vax_Dec$total_orace_vax <- Richmond_vax_Dec$total_morace_vax + Richmond_vax_Dec$total_forace_vax
+
+#Vax total for White Residents
+Richmond_vax_Dec$total_white_vax <- Richmond_vax_Dec$total_mwhite_vax + Richmond_vax_Dec$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_vax_Dec <- Richmond_vax_Dec %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_vax_Dec_sums <- Richmond_vax_Dec[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+Zip_30815_Dec <-filter(Richmond_vax_Dec_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+
+```
+
+### Change the column names for 30815 zip code data
+
+```{r}
+colnames(Zip_30815_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")
+
+```
+
+### Deleting excess environment objects and saving data as processed_data_Dec
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_vax_Dec,Dec_29, codebook, Richmond_vax_Dec_sums)
+
+#saving clean data for 30815 to processed_data folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_30815_Dec.rds")
+
+saveRDS(Zip_30815_Dec, file = save_data_location4)
+```
+
+## Processing 5: Getting Vaccination Totals by Target Zips 30815 for the Month of January
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Jan_26.xlsx")
+Jan_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+## Summary of data
+
+```{r}
+glimpse(Jan_26)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Richmond_vax_Jan <- filter(Jan_26, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_vax_Jan [Richmond_vax_Jan < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_vax_Jan = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Richmond_vax_Jan <- Richmond_vax_Jan  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Richmond_vax_Jan<- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.
+
+```{r}
+#Vax total Asian Residents
+Richmond_vax_Jan$total_asian_vax <- Richmond_vax_Jan$total_masian_vax + Richmond_vax_Jan$total_fasian_vax
+
+#Vax total for African American Residents 
+Richmond_vax_Jan$total_black_vax <- Richmond_vax_Jan$total_mblack_vax + Richmond_vax_Jan$total_fblack_vax
+
+#Vax total for Other Race Residents
+Richmond_vax_Jan$total_orace_vax <- Richmond_vax_Jan$total_morace_vax + Richmond_vax_Jan$total_forace_vax
+
+#Vax total for White Residents
+Richmond_vax_Jan$total_white_vax <- Richmond_vax_Jan$total_mwhite_vax + Richmond_vax_Jan$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_vax_Jan <- Richmond_vax_Jan %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_vax_Jan_sums <- Richmond_vax_Jan[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+Zip_30815_Jan <-filter(Richmond_vax_Jan_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+
+```
+
+### Change the column names for 30815 Jan
+
+```{r}
+colnames(Zip_30815_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")
+```
+
+### Deleting excess environment objects and saving data as processed_data_Jan
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_vax_Jan,Jan_26, codebook, Richmond_vax_Jan_sums)
+
+#saving clean data for 30815 to processed_data folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_30815_Jan.rds")
+
+saveRDS(Zip_30815_Jan, file = save_data_location4)
+```
+
+## Processing 6: : Getting Vaccination Totals by Target Zips 30815 for the Month of February
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+## Summary of data
+
+```{r}
+glimpse(Feb_26)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Richmond_vax_Feb <- filter(Feb_26, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_vax_Feb [Richmond_vax_Feb < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_vax_Feb = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Richmond_vax_Feb <- Richmond_vax_Feb  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Richmond_vax_Feb<- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asain residents.
+
+```{r}
+#Vax total Asian Residents
+Richmond_vax_Feb$total_asian_vax <- Richmond_vax_Feb$total_masian_vax + Richmond_vax_Feb$total_fasian_vax
+
+#Vax total for African American Residents 
+Richmond_vax_Feb$total_black_vax <- Richmond_vax_Feb$total_mblack_vax + Richmond_vax_Feb$total_fblack_vax
+
+#Vax total for Other Race Residents
+Richmond_vax_Feb$total_orace_vax <- Richmond_vax_Feb$total_morace_vax + Richmond_vax_Feb$total_forace_vax
+
+#Vax total for White Residents
+Richmond_vax_Feb$total_white_vax <- Richmond_vax_Feb$total_mwhite_vax + Richmond_vax_Feb$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_vax_Feb <- Richmond_vax_Feb %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_vax_Feb_sums <- Richmond_vax_Feb[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+Zip_30815_Feb <-filter(Richmond_vax_Feb_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+
+```
+
+### Change the column names for 30815 Feb
+
+```{r}
+colnames(Zip_30815_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")
+```
+
+### Deleting excess environment objects and saving data as processed_data_Feb
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_vax_Feb,Feb_26, codebook, Richmond_vax_Feb_sums)
+
+#saving clean data for 30815 to processed_data folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_30815_Feb.rds")
+
+saveRDS(Zip_30815_Feb, file = save_data_location4)
+```

### 2023-04-10 code/analysis_code/exploratoryanalysis.qmd
Part 4

@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31701_total_black_vax.png"), plot=plot_31701_total_black_vax)
+```
+
 ### Turning ggplot into an interactive plot
@@
           element_text(color="red", size=10, face="bold.italic"))
+```
+
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31705_total_black_vax.png"), plot=plot_31705_total_black_vax)
+```
+
+### Turning ggplot into an interactive plot
+
+```{r}
+plotly_31705_total_black_vax = ggplotly(plot_31705_total_black_vax)
+plotly_31705_total_black_vax
 ```
@@
           element_text(color="red", size=10, face="bold.italic"))
+```
+
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31707_total_black_vax.png"), plot=plot_31707_total_black_vax)
 ```
@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","total_black_vax.png"), plot=plot_Total_black_vax)
+```
+
 ### Turning ggplot into an interactive plot
@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31701_Gender_black_vax.png"), plot=plot_Gender_black_vax)
+```
+
 ### Turning ggplot into an interactive plot
@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31705_Gender_black_vax.png"), plot=plot_Gender_black_vax_31705)
+```
+
 ### Turning ggplot into an interactive plot
@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31707_Gender_black_vax.png"), plot=plot_Gender_black_vax_31707)
+```
+
 ### Turning ggplot into an interactive plot
@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31701_Age_black_vax.png"), plot=plot_Age_black_vax)
+```
+
 ### Turning ggplot into an interactive plot
@@
 
-------------------------------------------------------------------------
 ## Zip code: 31705
@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31705_Age_black_vax.png"), plot=plot_Age_black_vax)
+```
+
 ### Turning ggplot into an interactive plot
@@
 ```
-
-------------------------------------------------------------------------
 
@@
 
+### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February 
+### Total contains the vaccination totals for each month
+
+```{r}
+Age_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          Age_05_09 = c(275, 277, 280, 282, 284, 283),
+                          Age_10_17 = c(1109, 1120, 1125, 1134, 1136, 1142),
+                          Age_18_44 = c(5903, 5931, 5943, 5947, 5960, 5958), #pretty notable increase for ages 18-44 
+                          Age_45_64 = c(5130, 5120, 5125, 5123, 5133, 5140),
+                          Age_65Plus = c(4200, 4204, 4195, 4206, 4202, 4189))
+
+```
+
+### Converting the Age_Sum_by_month_31707 to long data in order to plot
+```{r}
+Age_Sum_by_month_31707_long = reshape2::melt(Age_Sum_by_month_31707, id = "Date")
+```
+
+### Changing "variable" to "Age" and "value" to "Vaccination_total"
+
+```{r}
+Age_Sum_by_month_31707_long <- Age_Sum_by_month_31707_long %>% 
+  rename("Age" = "variable") %>% 
+  rename("Vaccination_total" = "value")
+```
+
+### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 31707 zip code
+
+```{r}
+plot_Age_black_vax = Age_Sum_by_month_31707_long %>% 
+                     ggplot(aes(Date, Vaccination_total, 
+                                group = Age, 
+                                color = Age)) + 
+                     geom_point() + geom_line() + 
+                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
+#adding a title to the plot 
+plot_Age_black_vax = plot_Age_black_vax + 
+  ggtitle("Vaccination Totals for Black population by Age in 31707") + 
+  theme(plot.title = 
+          element_text(color="red", size=10, face="bold.italic")) + 
+  labs(y = "Vaccination Total")
+plot_Age_black_vax
+```
+
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","31707_Age_black_vax.png"), plot=plot_Age_black_vax)
+```
+
+### Turning ggplot into an interactive plot
+
+```{r}
+plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
+plotly_plot_Age_black_vax
+```
+
 # The Amount of Vaccinations Among Black population in the Control Zip Code
@@
 
-#Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February #Total contains the vaccination totals for each month
+### Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February ### Total contains the vaccination totals for each month
 
@@
 ```
+
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","30815_total_black_vax.png"), plot=plot_30815_total_black_vax)
+```
+
 ### Turning ggplot into an interactive plot
@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","30815_Gender_black_vax.png"), plot=plot_Gender_black_vax)
+```
+
 ### Turning ggplot into an interactive plot
@@
 
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","30815_Age_black_vax.png"), plot=plot_Gender_black_vax)
+```
+
 ### Turning ggplot into an interactive plot
@@
 ```
+
+# The Amount of Vaccinations Among Black population in the Control Zip Code
+
+## Zip code: 30907
+
+### Loading the processed data in order to begin analysis for the 30907 zip code
+
+```{r}
+# Loading September 30907
+data_location1 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Sept.rds")
+processed_data_30907_Sept = readRDS(data_location1)
+
+# Loading October 30907 and joining september and october 
+data_location2 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Oct.rds")
+processed_data_30907_Oct = readRDS(data_location2)
+processed_combined <- left_join(processed_data_30907_Sept, processed_data_30907_Oct, by = c("FIPS" = "FIPS"))
+
+# Loading November 30907
+data_location3 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Nov.rds")
+processed_data_30907_Nov = readRDS(data_location3)
+processed_combined <- left_join(processed_combined, processed_data_30907_Nov, by = c("FIPS" = "FIPS"))
+
+# Loading December 30907
+data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Dec.rds")
+processed_data_30907_Dec = readRDS(data_location4)
+processed_combined <- left_join(processed_combined, processed_data_30907_Dec, by = c("FIPS" = "FIPS"))
+
+# Loading January 30907
+data_location5 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Jan.rds")
+processed_data_30907_Jan = readRDS(data_location5)
+processed_combined <- left_join(processed_combined, processed_data_30907_Jan, by = c("FIPS" = "FIPS"))
+
+# Loading February 30907
+data_location6 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Feb.rds")
+processed_data_30907_Feb = readRDS(data_location6)
+processed_combined_30907 <- left_join(processed_combined, processed_data_30907_Feb, by = c("FIPS" = "FIPS"))
+```
+
+### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender
+
+```{r}
+save_data_location = here("data", "processed_data","Zip_30907", "processed_combined_30907.rds")
+saveRDS(processed_combined_30907, file = save_data_location)
+```
+
+### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 30907 zip code
+
+#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+
+```{r}
+processed_combined_30907_total_black = processed_combined_30907 %>% select(starts_with("total_black_vax"))
+```
+
+### Summing each Month variable by census tract
+
+```{r}
+#Summing each column for each month 
+sum(processed_combined_30907_total_black$total_black_vax_Sept_17) 
+sum(processed_combined_30907_total_black$total_black_vax_Oct_28) 
+sum(processed_combined_30907_total_black$total_black_vax_Nov_24) 
+sum(processed_combined_30907_total_black$total_black_vax_Dec_29) 
+sum(processed_combined_30907_total_black$total_black_vax_Jan_26) 
+sum(processed_combined_30907_total_black$total_black_vax_Feb_26) 
+```
+
+### Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February ### Total contains the vaccination totals for each month
+
+```{r}
+Sum_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          Total = c(4956, 4917, 4888, 4851, 4866, 4894)) 
+```
+
+### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 30907 zip code
+
+```{r}
+plot_30907_total_black_vax = Sum_by_month_30907 %>% 
+  ggplot(aes(Date, Total, group=1)) + 
+  geom_point() + geom_line() + 
+  scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
+
+#Adding a customization title for the plot 
+plot_30907_total_black_vax = plot_30907_total_black_vax + 
+  ggtitle("Vaccination Totals for Black population in the 30907 Zipcode") + 
+  theme(plot.title = 
+          element_text(color="red", size=10, face="bold.italic"))
+plot_30907_total_black_vax
+```
+
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","30907_total_black_vax.png"), plot=plot_30907_total_black_vax)
+```
+
+### Turning ggplot into an interactive plot
+
+```{r}
+plotly_30907_total_black_vax = ggplotly(plot_30907_total_black_vax)
+plotly_30907_total_black_vax
+```
+
+# The Amount of Vaccinations by Gender Among Black Population in the Control Zip Code
+
+## Zip code: 30907
+
+### Loading the `processed_combined_30907.rds` file from `analysiscode.qmd`
+
+```{r}
+save_data_location = here("data", "processed_data","Zip_30907", "processed_combined_30907.rds")
+processed_combined_30907 = readRDS(save_data_location)
+```
+
+### Dropping all variables except for black males and black females within 30907
+
+#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+
+```{r}
+processed_combined_30907_Gender_black = processed_combined_30907 %>% select(matches("^total_(mblack|fblack)_vax"))
+```
+
+### Summing each Month variable by census tract
+```{r}
+#Summing for black males by month 
+sum(processed_combined_30907_Gender_black$total_mblack_vax_Sept_17) 
+sum(processed_combined_30907_Gender_black$total_mblack_vax_Oct_28) 
+sum(processed_combined_30907_Gender_black$total_mblack_vax_Nov_24) 
+sum(processed_combined_30907_Gender_black$total_mblack_vax_Dec_29) 
+sum(processed_combined_30907_Gender_black$total_mblack_vax_Jan_26) 
+sum(processed_combined_30907_Gender_black$total_mblack_vax_Feb_26)
+
+#summing for black females by month 
+sum(processed_combined_30907_Gender_black$total_fblack_vax_Sept_17) 
+sum(processed_combined_30907_Gender_black$total_fblack_vax_Oct_28) 
+sum(processed_combined_30907_Gender_black$total_fblack_vax_Nov_24) 
+sum(processed_combined_30907_Gender_black$total_fblack_vax_Dec_29) 
+sum(processed_combined_30907_Gender_black$total_fblack_vax_Jan_26) 
+sum(processed_combined_30907_Gender_black$total_fblack_vax_Feb_26) 
+```
+
+#Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
+
+```{r}
+Gender_Sum_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          Male_Total = c(1947, 1943, 1926, 1900, 1913, 1919),
+                          Female_Total = c(3009, 2974, 2962, 2951, 2953, 2975))
+```
+
+### Converting the Gender_Sum_by_month_30907 to long data in order to plot
+```{r}
+Gender_Sum_by_month_30907_long = reshape2::melt(Gender_Sum_by_month_30907, id = "Date")
+```
+
+### Changing "variable" to "Gender" and "value" to "Vaccination_total"
+
+```{r}
+Gender_Sum_by_month_30907_long <- Gender_Sum_by_month_30907_long %>% 
+  rename("Gender" = "variable") %>% 
+  rename("Vaccination_total" = "value")
+```
+
+### gglot2: Creating a line plot for the vaccination totals by month for the Black population by gender in the 30907 zip code
+
+```{r}
+plot_Gender_black_vax = Gender_Sum_by_month_30907_long %>% 
+                     ggplot(aes(Date, Vaccination_total, 
+                                group = Gender, 
+                                color = Gender)) + 
+                     geom_point() + geom_line() + 
+                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
+#adding a title to the plot 
+plot_Gender_black_vax = plot_Gender_black_vax + 
+  ggtitle("Vaccination Totals for Black population by Gender in 30907") + 
+  theme(plot.title = 
+          element_text(color="red", size=10, face="bold.italic")) + 
+  labs(y = "Vaccination Total")
+plot_Gender_black_vax
+```
+
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","30907_Gender_black_vax.png"), plot=plot_Gender_black_vax)
+```
+
+### Turning ggplot into an interactive plot
+
+```{r}
+plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
+plotly_plot_Gender_black_vax
+```
+
+# Total Amount of vaccinations by Age among Black population in the Control Zip Code
+
+## Zip code: 30907
+
+### Loading the `processed_combined_30907.rds` file from `analysiscode.qmd`
+
+```{r}
+save_data_location = here("data", "processed_data","Zip_30907", "processed_combined_30907.rds")
+processed_combined_30907 = readRDS(save_data_location)
+```
+
+### Dropping all variables except for black age groups within 30907
+
+#### Using the select() function to subset the data frame by columns that start with an age group
+
+```{r}
+processed_combined_30907_Age_black = processed_combined_30907 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
+```
+
+### Summing each Month variable by age group by month
+```{r}
+#Summing for ages 5-9 by month 
+sum(processed_combined_30907_Age_black$total_05_09_vax_Sept_17) 
+sum(processed_combined_30907_Age_black$total_05_09_vax_Oct_28) 
+sum(processed_combined_30907_Age_black$total_05_09_vax_Nov_24) 
+sum(processed_combined_30907_Age_black$total_05_09_vax_Dec_29) 
+sum(processed_combined_30907_Age_black$total_05_09_vax_Jan_26) 
+sum(processed_combined_30907_Age_black$total_05_09_vax_Feb_26)
+
+#Summing for ages 10-17 by month 
+sum(processed_combined_30907_Age_black$total_10_17_vax_Sept_17) 
+sum(processed_combined_30907_Age_black$total_10_17_vax_Oct_28) 
+sum(processed_combined_30907_Age_black$total_10_17_vax_Nov_24) 
+sum(processed_combined_30907_Age_black$total_10_17_vax_Dec_29) 
+sum(processed_combined_30907_Age_black$total_10_17_vax_Jan_26) 
+sum(processed_combined_30907_Age_black$total_10_17_vax_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_combined_30907_Age_black$total_18_44_vax_Sept_17) 
+sum(processed_combined_30907_Age_black$total_18_44_vax_Oct_28) 
+sum(processed_combined_30907_Age_black$total_18_44_vax_Nov_24) 
+sum(processed_combined_30907_Age_black$total_18_44_vax_Dec_29) 
+sum(processed_combined_30907_Age_black$total_18_44_vax_Jan_26) 
+sum(processed_combined_30907_Age_black$total_18_44_vax_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_combined_30907_Age_black$total_45_64_vax_Sept_17) 
+sum(processed_combined_30907_Age_black$total_45_64_vax_Oct_28) 
+sum(processed_combined_30907_Age_black$total_45_64_vax_Nov_24) 
+sum(processed_combined_30907_Age_black$total_45_64_vax_Dec_29) 
+sum(processed_combined_30907_Age_black$total_45_64_vax_Jan_26) 
+sum(processed_combined_30907_Age_black$total_45_64_vax_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_combined_30907_Age_black$total_65Plus_vax_Sept_17) 
+sum(processed_combined_30907_Age_black$total_65Plus_vax_Oct_28) 
+sum(processed_combined_30907_Age_black$total_65Plus_vax_Nov_24) 
+sum(processed_combined_30907_Age_black$total_65Plus_vax_Dec_29) 
+sum(processed_combined_30907_Age_black$total_65Plus_vax_Jan_26) 
+sum(processed_combined_30907_Age_black$total_65Plus_vax_Feb_26)
+
+```
+
+### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
+
+```{r}
+Age_Sum_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          Age_05_09 = c(244, 239, 241, 236, 240, 242),
+                          Age_10_17 = c(1344, 1358, 1357, 1341, 1343, 1335),
+                          Age_18_44 = c(9203, 9217, 9238, 9245, 9253, 9246),
+                          Age_45_64 = c(6740, 6700, 6667, 6626, 6617, 6624),
+                          Age_65Plus = c(6632, 6633, 6639, 6635, 6633, 6623))
+
+```
+
+### Converting the Age_Sum_by_month_30907 to long data in order to plot
+```{r}
+Age_Sum_by_month_30907_long = reshape2::melt(Age_Sum_by_month_30907, id = "Date")
+```
+
+### Changing "variable" to "Age" and "value" to "Vaccination_total"
+
+```{r}
+Age_Sum_by_month_30907_long <- Age_Sum_by_month_30907_long %>% 
+  rename("Age" = "variable") %>% 
+  rename("Vaccination_total" = "value")
+```
+
+### gglot2: Creating a line plot for the vaccination totals by month for the Black population by Age in the 30907 zip code
+
+```{r}
+plot_Age_black_vax = Age_Sum_by_month_30907_long %>% 
+                     ggplot(aes(Date, Vaccination_total, 
+                                group = Age, 
+                                color = Age)) + 
+                     geom_point() + geom_line() + 
+                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
+#adding a title to the plot 
+plot_Age_black_vax = plot_Age_black_vax + 
+  ggtitle("Vaccination Totals for Black population by Age in 30907") + 
+  theme(plot.title = 
+          element_text(color="red", size=10, face="bold.italic")) + 
+  labs(y = "Vaccination Total")
+plot_Age_black_vax
+```
+
+### Save plot
+```{r}
+#save figure
+ggsave(filename = here("results","30907_Age_black_vax.png"), plot=plot_Age_black_vax)
+```
+
+### Turning ggplot into an interactive plot
+
+```{r}
+plotly_plot_Age_black_vax = ggplotly(plot_Age_black_vax)
+plotly_plot_Age_black_vax
+```

### 2023-04-10 code/analysis_code/statisticalanalysis.qmd
Part 4

### Library
```{r}
library(tidymodels)
library(vip) #most important variables in LASSO Model
library(forcats)
library(tidyverse)
```

### creating a data frame that contains the variables that will be included in new linear regression mode. Just copying and pasting all of the totals generated from the `exploratoryanalysis.qmd` file
```{r}
#combined all relevant variables into 5 different data sets for each zip code. Got the black population totals for each zip code from  https://www.unitedstateszipcodes.org/31701/

Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Total = c(7797, 7780, 7755, 7773, 7792, 7817),
                          Intervention = c(1, 1, 1, 1, 1, 1),
                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420),
                          Age_05_09 = c(123, 126, 126, 127, 127, 128),
                          Age_10_17 = c(642, 635, 633, 637, 630, 629),
                          Age_18_44 = c(3310, 3316, 3313, 3323, 3325, 3331),
                          Age_45_64 = c(3229, 3244, 3239, 3259, 3264, 3258),
                          Age_65Plus = c(2576, 2570, 2569, 2574, 2568, 2572),
                          Pop = c(16475, 16475, 16475, 16475, 16475, 16475))

Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                                Total = c(10383, 10327, 10303, 10301, 10337, 10348),                                 Intervention = c(1, 1, 1, 1, 1, 1),
                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949),
                          Age_05_09 = c(210, 211, 209, 201, 202, 202),
                          Age_10_17 = c(1159, 1171, 1169, 1163, 1168, 1167),
                          Age_18_44 = c(5201, 5215, 5239, 5263, 5278, 5281),
                          Age_45_64 = c(4380, 4381, 4366, 4364, 4367, 4367),
                          Age_65Plus = c(2983, 3000, 2997, 2997, 2994, 2989),
                          Pop = c(24779, 24779, 24779, 24779, 24779, 24779))

Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                                Total = c(9035, 8985, 8997, 9005, 9035, 9042),                                 Intervention = c(1, 1, 1, 1, 1, 1),
                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361),
                          Age_05_09 = c(275, 277, 280, 282, 284, 283),
                          Age_10_17 = c(1109, 1120, 1125, 1134, 1136, 1142),
                          Age_18_44 = c(5903, 5931, 5943, 5947, 5960, 5958),  
                          Age_45_64 = c(5130, 5120, 5125, 5123, 5133, 5140),
                          Age_65Plus = c(4200, 4204, 4195, 4206, 4202, 4189),
                          Pop = c(13884, 13884, 13884, 13884, 13884, 13884))

Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          Total = c(8758, 8688, 8639, 8593, 8623, 8666),                                      Intervention = c(0, 0, 0, 0, 0, 0),
                          Male_Total = c(3526, 3514, 3503, 3484, 3489, 3498),
                          Female_Total = c(5232, 5174, 5136, 5109, 5134, 5168),
                          Age_05_09 = c(201, 213, 211, 215, 211, 211),
                          Age_10_17 = c(1116, 1113, 1115, 1118, 1115, 1114),
                          Age_18_44 = c(5001, 5009, 5020, 5026, 5025, 5032),
                          Age_45_64 = c(5026, 5017, 5020, 5009, 5013, 5020),
                          Age_65Plus = c(2816, 2825, 2827, 2814, 2828, 2832),
                          Pop = c(23481, 23481, 23481, 23481, 23481, 23481))

#um_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
                          #Total = c(4956, 4917, 4888, 4851, 4866, 4894),                                      Zip = c(30907, 30907, 30907, 30907, 30907, 30907),
                          #Male_Total = c(1947, 1943, 1926, 1900, 1913, 1919),
                          #Female_Total = c(3009, 2974, 2962, 2951, 2953, 2975),
                          #Age_05_09 = c(244, 239, 241, 236, 240, 242),
                          #Age_10_17 = c(1344, 1358, 1357, 1341, 1343, 1335),
                          #Age_18_44 = c(9203, 9217, 9238, 9245, 9253, 9246),
                          #Age_45_64 = c(6740, 6700, 6667, 6626, 6617, 6624),
                          #Age_65Plus = c(6632, 6633, 6639, 6635, 6633, 6623),
                          #Pop = c(7744, 7744, 7744, 7744, 7744, 7744)) 
#the 30907 zip code population of residents that are black does not match the vaccination data. also this zip code does not have a majority of residents that are black like other other 4 zip codes so I am going to leave it out of the statistical analysis

#combining all of the zip codes into one data set except for the 30907 data 
combined = rbind(Sum_by_month_31701, Sum_by_month_31705, Sum_by_month_31707, Sum_by_month_30815)
```

### creating a vaccination rate variable as the outcome 
```{r}
combined$Vax_rate = (combined$Total / combined$Pop) * 100
```

### Creating a linear regression model 
```{r}
lm_model2 <- lm(Vax_rate ~ Intervention + Male_Total + Female_Total + Age_05_09 + Age_10_17 + Age_18_44 + Age_45_64 + Age_65Plus, data = combined)

summary(lm_model2)
```
The results indicate that Intervention is a significant predictor of Vax_rate, as indicated by its estimate of 16.23 and its marginally significant p-value of 0.067. This means that there is some evidence that the intervention has an effect on the vaccination rate

Male_Total is a significant negative predictor of Vax_rate, with a coefficient of -0.0133 and a significant p-value of 0.0054. This suggests that a higher male population is associated with a lower vaccination rate

Among the age groups, Age_65Plus is a significant positive predictor of Vax_rate, with a coefficient of 0.0114 and a marginally significant p-value of 0.055. This means that a higher proportion of the population in the age group 65 and above is associated with a higher vaccination rate

Overall, the model has a high adjusted R-squared value of 0.9995, indicating that the predictor variables account for a large proportion of the variation in Vax_rate. However, it is important to note that with such a high R-squared value, overfitting may be a concern, and it may be necessary to perform additional checks, such as cross-validation or regularization

### evaluating this model using RMSE 
```{r}
# Split the data into training and testing sets
train <- combined[1:15,]
test <- combined[16:24,]

# Train the linear regression model
model <- lm(Vax_rate ~ Intervention + Male_Total + Female_Total + 
             Age_05_09 + Age_10_17 + Age_18_44 + Age_45_64 + Age_65Plus, 
           data = train)

# Generate predictions for the test data
predictions <- predict(model, newdata = test)

# Calculate RMSE
rmse <- sqrt(mean((test$Vax_rate - predictions)^2))

rmse
```

### Train a LASSO model 
```{r}
Vax_split <- initial_split(combined, strata = Intervention)
Vax_train <- training(Vax_split)
Vax_test <- testing(Vax_split)
```

```{r}
Vax_rec <- recipe(Vax_rate ~ Intervention + Male_Total + Female_Total + 
             Age_05_09 + Age_10_17 + Age_18_44 + Age_45_64 + Age_65Plus, data = Vax_train) %>%
  step_zv(all_numeric(), -all_outcomes()) %>%
  step_normalize(all_numeric(), -all_outcomes()) #centering and scaling data 
```

```{r}
lasso_Vax <- linear_reg(penalty = 0.1, mixture = 1) %>%
  set_engine("glmnet")

wf <- workflow() %>%
  add_recipe(Vax_rec)

lasso_fit <- wf %>%
  add_model(lasso_Vax) %>%
  fit(data = Vax_train)

lasso_fit %>%
  pull_workflow_fit() %>%
  tidy()
```

### Tuning LASSO Parameters
```{r}
set.seed(1234)
Vax_boot <- bootstraps(Vax_train, strata = Intervention)

tune_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lambda_grid <- grid_regular(penalty(), levels = 50)
```

### Tune the grid
```{r}
set.seed(2020)
lasso_grid <- tune_grid(
  wf %>% add_model(tune_spec),
  resamples = Vax_boot,
  grid = lambda_grid
)
```

### Results 
```{r}
lasso_grid %>%
  collect_metrics()
```

### Visualizing the LASSO results 
```{r}
lasso_grid %>%
  collect_metrics()

lasso_grid_plot = lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err
  ),
  alpha = 0.5
  ) +
  geom_line(size = 1.5) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none")

lasso_grid_plot
```

### Save plot
```{r}
#save figure
ggsave(filename = here("results","lasso_grid_plot.png"), plot=lasso_grid_plot)
```

### Lets pick the LASSO model with the lowest RMSE 
```{r}
lowest_rmse <- lasso_grid %>%
  select_best("rmse", maximize = FALSE)

final_lasso <- finalize_workflow(
  wf %>% add_model(tune_spec),
  lowest_rmse
)
```

### Final fit and viewing most important variables 
```{r}
final_lasso_plot = final_lasso %>%
  fit(Vax_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = lowest_rmse$penalty) %>%
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
final_lasso_plot
```

### Looking at final fit 
```{r}
last_fit(
  final_lasso,
  Vax_split
) %>%
  collect_metrics()
```

### Save plot
```{r}
#save figure
ggsave(filename = here("results","final_lasso_plot.png"), plot=final_lasso_plot)
```

### 2023-04-10 code/processing_code/processingcode.qmd
Part 4

@@
 
-Load needed packages. make sure they are installed.
-
+### Load needed packages. make sure they are installed.
 ```{r}
@@
 ```
+
+# Processing Section for Control Zip Code: 30907
+
+## Processing 1: Getting Vaccination Totals by Control Zip Code 30907 in the Month of September
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Sept_17.xlsx")
+Sept_17 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Zip_30907_vax_Sept <-filter(Sept_17, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_vax_Sept)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_vax_Sept [Zip_30907_vax_Sept < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_vax_Sept = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Zip_30907_vax_Sept<- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asian residents.
+
+```{r}
+#Vax total Asian Residents
+Zip_30907_vax_Sept$total_asian_vax <- Zip_30907_vax_Sept$total_masian_vax + Zip_30907_vax_Sept$total_fasian_vax
+
+#Vax total for African American Residents 
+Zip_30907_vax_Sept$total_black_vax <- Zip_30907_vax_Sept$total_mblack_vax + Zip_30907_vax_Sept$total_fblack_vax
+
+#Vax total for Other Race Residents
+Zip_30907_vax_Sept$total_orace_vax <- Zip_30907_vax_Sept$total_morace_vax + Zip_30907_vax_Sept$total_forace_vax
+
+#Vax total for White Residents
+Zip_30907_vax_Sept$total_white_vax <- Zip_30907_vax_Sept$total_mwhite_vax + Zip_30907_vax_Sept$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_vax_Sept <- Zip_30907_vax_Sept %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_vax_Sept_sums <- Zip_30907_vax_Sept[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Sept
+```{r}
+colnames(Zip_30907_vax_Sept_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Sept_17", "total_male_vax_Sept_17", "total_female_vax_Sept_17", "total_masian_vax_Sept_17", "total_mblack_vax_Sept_17", "total_morace_vax_Sept_17", "total_mwhite_vax_Sept_17", "total_fasian_vax_Sept_17", "total_fblack_vax_Sept_17", "total_forace_vax_Sept_17", "total_fwhite_vax_Sept_17", "total_asian_vax_Sept_17","total_black_vax_Sept_17", "total_orace_vax_Sept_17", "total_white_vax_Sept_17", "total_05_09_vax_Sept_17", "total_10_17_vax_Sept_17", "total_18_44_vax_Sept_17", "total_45_64_vax_Sept_17", "total_65Plus_vax_Sept_17")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Sept.rds")
+
+saveRDS(Zip_30907_vax_Sept_sums, save_data_location4)
+```
+
+## Processing 2: Getting Vaccination Totals by Control Zip Code 30907 in the Month of October
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Oct_28.xlsx")
+Oct_28 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Zip_30907_vax_Oct <-filter(Oct_28, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_vax_Oct)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_vax_Oct [Zip_30907_vax_Oct < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_vax_Oct = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Zip_30907_vax_Oct<- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asian residents.
+
+```{r}
+#Vax total Asian Residents
+Zip_30907_vax_Oct$total_asian_vax <- Zip_30907_vax_Oct$total_masian_vax + Zip_30907_vax_Oct$total_fasian_vax
+
+#Vax total for African American Residents 
+Zip_30907_vax_Oct$total_black_vax <- Zip_30907_vax_Oct$total_mblack_vax + Zip_30907_vax_Oct$total_fblack_vax
+
+#Vax total for Other Race Residents
+Zip_30907_vax_Oct$total_orace_vax <- Zip_30907_vax_Oct$total_morace_vax + Zip_30907_vax_Oct$total_forace_vax
+
+#Vax total for White Residents
+Zip_30907_vax_Oct$total_white_vax <- Zip_30907_vax_Oct$total_mwhite_vax + Zip_30907_vax_Oct$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_vax_Oct <- Zip_30907_vax_Oct %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_vax_Oct_sums <- Zip_30907_vax_Oct[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Oct
+```{r}
+colnames(Zip_30907_vax_Oct_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Oct_28", "total_male_vax_Oct_28", "total_female_vax_Oct_28", "total_masian_vax_Oct_28", "total_mblack_vax_Oct_28", "total_morace_vax_Oct_28", "total_mwhite_vax_Oct_28", "total_fasian_vax_Oct_28", "total_fblack_vax_Oct_28", "total_forace_vax_Oct_28", "total_fwhite_vax_Oct_28", "total_asian_vax_Oct_28","total_black_vax_Oct_28", "total_orace_vax_Oct_28", "total_white_vax_Oct_28", "total_05_09_vax_Oct_28", "total_10_17_vax_Oct_28", "total_18_44_vax_Oct_28", "total_45_64_vax_Oct_28", "total_65Plus_vax_Oct_28")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Oct.rds")
+
+saveRDS(Zip_30907_vax_Oct_sums, save_data_location4)
+```
+
+## Processing 3: : Getting Vaccination Totals by Control Zip Code 30907 in the Month of November
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Nov_24.xlsx")
+Nov_24 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Zip_30907_vax_Nov <-filter(Nov_24, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_vax_Nov)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_vax_Nov [Zip_30907_vax_Nov < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_vax_Nov = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Zip_30907_vax_Nov<- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asian residents.
+
+```{r}
+#Vax total Asian Residents
+Zip_30907_vax_Nov$total_asian_vax <- Zip_30907_vax_Nov$total_masian_vax + Zip_30907_vax_Nov$total_fasian_vax
+
+#Vax total for African American Residents 
+Zip_30907_vax_Nov$total_black_vax <- Zip_30907_vax_Nov$total_mblack_vax + Zip_30907_vax_Nov$total_fblack_vax
+
+#Vax total for Other Race Residents
+Zip_30907_vax_Nov$total_orace_vax <- Zip_30907_vax_Nov$total_morace_vax + Zip_30907_vax_Nov$total_forace_vax
+
+#Vax total for White Residents
+Zip_30907_vax_Nov$total_white_vax <- Zip_30907_vax_Nov$total_mwhite_vax + Zip_30907_vax_Nov$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_vax_Nov <- Zip_30907_vax_Nov %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_vax_Nov_sums <- Zip_30907_vax_Nov[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Nov
+```{r}
+colnames(Zip_30907_vax_Nov_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Nov_24", "total_male_vax_Nov_24", "total_female_vax_Nov_24", "total_masian_vax_Nov_24", "total_mblack_vax_Nov_24", "total_morace_vax_Nov_24", "total_mwhite_vax_Nov_24", "total_fasian_vax_Nov_24", "total_fblack_vax_Nov_24", "total_forace_vax_Nov_24", "total_fwhite_vax_Nov_24", "total_asian_vax_Nov_24","total_black_vax_Nov_24", "total_orace_vax_Nov_24", "total_white_vax_Nov_24", "total_05_09_vax_Nov_24", "total_10_17_vax_Nov_24", "total_18_44_vax_Nov_24", "total_45_64_vax_Nov_24", "total_65Plus_vax_Nov_24")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Nov.rds")
+
+saveRDS(Zip_30907_vax_Nov_sums, save_data_location4)
+```
+
+## Processing 4: : Getting Vaccination Totals by Control Zip Code 30907 in the Month of December
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Dec_29.xlsx")
+Dec_29 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Zip_30907_vax_Dec <-filter(Dec_29, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_vax_Dec)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_vax_Dec [Zip_30907_vax_Dec < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_vax_Dec = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Zip_30907_vax_Dec<- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asian residents.
+
+```{r}
+#Vax total Asian Residents
+Zip_30907_vax_Dec$total_asian_vax <- Zip_30907_vax_Dec$total_masian_vax + Zip_30907_vax_Dec$total_fasian_vax
+
+#Vax total for African American Residents 
+Zip_30907_vax_Dec$total_black_vax <- Zip_30907_vax_Dec$total_mblack_vax + Zip_30907_vax_Dec$total_fblack_vax
+
+#Vax total for Other Race Residents
+Zip_30907_vax_Dec$total_orace_vax <- Zip_30907_vax_Dec$total_morace_vax + Zip_30907_vax_Dec$total_forace_vax
+
+#Vax total for White Residents
+Zip_30907_vax_Dec$total_white_vax <- Zip_30907_vax_Dec$total_mwhite_vax + Zip_30907_vax_Dec$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_vax_Dec <- Zip_30907_vax_Dec %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_vax_Dec_sums <- Zip_30907_vax_Dec[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Dec
+```{r}
+colnames(Zip_30907_vax_Dec_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Dec_29", "total_male_vax_Dec_29", "total_female_vax_Dec_29", "total_masian_vax_Dec_29", "total_mblack_vax_Dec_29", "total_morace_vax_Dec_29", "total_mwhite_vax_Dec_29", "total_fasian_vax_Dec_29", "total_fblack_vax_Dec_29", "total_forace_vax_Dec_29", "total_fwhite_vax_Dec_29", "total_asian_vax_Dec_29","total_black_vax_Dec_29", "total_orace_vax_Dec_29", "total_white_vax_Dec_29", "total_05_09_vax_Dec_29", "total_10_17_vax_Dec_29", "total_18_44_vax_Dec_29", "total_45_64_vax_Dec_29", "total_65Plus_vax_Dec_29")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Dec.rds")
+
+saveRDS(Zip_30907_vax_Dec_sums, save_data_location4)
+```
+
+## Processing 5: : Getting Vaccination Totals by Control Zip Code 30907 in the Month of January
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Jan_26.xlsx")
+Jan_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Zip_30907_vax_Jan <-filter(Jan_26, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_vax_Jan)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_vax_Jan [Zip_30907_vax_Jan < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>% mutate(total_vax_Jan = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Zip_30907_vax_Jan<- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asian residents.
+
+```{r}
+#Vax total Asian Residents
+Zip_30907_vax_Jan$total_asian_vax <- Zip_30907_vax_Jan$total_masian_vax + Zip_30907_vax_Jan$total_fasian_vax
+
+#Vax total for African American Residents 
+Zip_30907_vax_Jan$total_black_vax <- Zip_30907_vax_Jan$total_mblack_vax + Zip_30907_vax_Jan$total_fblack_vax
+
+#Vax total for Other Race Residents
+Zip_30907_vax_Jan$total_orace_vax <- Zip_30907_vax_Jan$total_morace_vax + Zip_30907_vax_Jan$total_forace_vax
+
+#Vax total for White Residents
+Zip_30907_vax_Jan$total_white_vax <- Zip_30907_vax_Jan$total_mwhite_vax + Zip_30907_vax_Jan$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_vax_Jan <- Zip_30907_vax_Jan %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_vax_Jan_sums <- Zip_30907_vax_Jan[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Jan
+```{r}
+colnames(Zip_30907_vax_Jan_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Jan_26", "total_male_vax_Jan_26", "total_female_vax_Jan_26", "total_masian_vax_Jan_26", "total_mblack_vax_Jan_26", "total_morace_vax_Jan_26", "total_mwhite_vax_Jan_26", "total_fasian_vax_Jan_26", "total_fblack_vax_Jan_26", "total_forace_vax_Jan_26", "total_fwhite_vax_Jan_26", "total_asian_vax_Jan_26","total_black_vax_Jan_26", "total_orace_vax_Jan_26", "total_white_vax_Jan_26", "total_05_09_vax_Jan_26", "total_10_17_vax_Jan_26", "total_18_44_vax_Jan_26", "total_45_64_vax_Jan_26", "total_65Plus_vax_Jan_26")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Jan.rds")
+
+saveRDS(Zip_30907_vax_Jan_sums, save_data_location4)
+```
+
+## Processing 6: : Getting Vaccination Totals by Control Zip Code 30907 in the Month of February
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with vaccinations from Richmond County 
+Zip_30907_vax_Feb <-filter(Feb_26, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_vax_Feb)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_vax_Feb [Zip_30907_vax_Feb < 0] <- 0  
+```
+
+### Adding new columns: total vax sum, total vax of all males, total vax of all females
+
+```{r}
+#Adding a column of the total vax for each census tract (for each observation)  
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>% mutate(total_vax_Feb = sum(across(ends_with("VAX")), na.rm = T))
+
+#total vax column for all males
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb  %>% rowwise() %>%   
+  mutate(total_male_vax = sum(across(starts_with("M") & ends_with("VAX")), na.rm = T))
+
+#total vax column for all females
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_female_vax = sum(across(starts_with("F") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total vax column for Asian male residents 
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_masian_vax = sum(across(starts_with("MAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American male residents
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_mblack_vax = sum(across(starts_with("MBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race male residents
+Zip_30907_vax_Feb<- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_morace_vax = sum(across(starts_with("MOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White male residents 
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_mwhite_vax = sum(across(starts_with("MWhite") & ends_with("VAX")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total vax column for Asian female residents    
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_fasian_vax = sum(across(starts_with("FAsianNHPI") & ends_with("VAX")), na.rm = T))
+
+#total vax column for African American female residents
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_fblack_vax = sum(across(starts_with("FBlack") & ends_with("VAX")), na.rm = T))
+
+#total vax column for other race female residents
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_forace_vax = sum(across(starts_with("FOtherRace") & ends_with("VAX")), na.rm = T))
+
+#total vax column for White female residents 
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>%   
+  mutate(total_fwhite_vax = sum(across(starts_with("FWhite") & ends_with("VAX")), na.rm = T))  
+```
+
+### Adding new columns for Vax totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident vaccinations by adding all males and females of a specific race to get the the overall amount of resident vaccinations by race.
+
+For example(see below): I get the total Asian resident vaccinations by summing male Asian residents + female Asian residents.
+
+```{r}
+#Vax total Asian Residents
+Zip_30907_vax_Feb$total_asian_vax <- Zip_30907_vax_Feb$total_masian_vax + Zip_30907_vax_Feb$total_fasian_vax
+
+#Vax total for African American Residents 
+Zip_30907_vax_Feb$total_black_vax <- Zip_30907_vax_Feb$total_mblack_vax + Zip_30907_vax_Feb$total_fblack_vax
+
+#Vax total for Other Race Residents
+Zip_30907_vax_Feb$total_orace_vax <- Zip_30907_vax_Feb$total_morace_vax + Zip_30907_vax_Feb$total_forace_vax
+
+#Vax total for White Residents
+Zip_30907_vax_Feb$total_white_vax <- Zip_30907_vax_Feb$total_mwhite_vax + Zip_30907_vax_Feb$total_fwhite_vax  
+```
+
+### Adding new columns for Vax totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>% mutate(total_05_09_vax = sum(FAsianNHPI_05_09_VAX, MAsianNHPI_05_09_VAX, UAsianNHPI_05_09_VAX,FBlack_05_09_VAX, MBlack_05_09_VAX, UBlack_05_09_VAX,FNone_05_09_VAX, MNone_05_09_VAX, UNone_05_09_VAX,FOtherRace_05_09_VAX, MOtherRace_05_09_VAX, UOtherRace_05_09_VAX,FWhite_05_09_VAX, MWhite_05_09_VAX, UWhite_05_09_VAX, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>% mutate(total_10_17_vax = sum(FAsianNHPI_10_17_VAX, MAsianNHPI_10_17_VAX, UAsianNHPI_10_17_VAX,FBlack_10_17_VAX, MBlack_10_17_VAX, UBlack_10_17_VAX,FNone_10_17_VAX, MNone_10_17_VAX, UNone_10_17_VAX,FOtherRace_10_17_VAX, MOtherRace_10_17_VAX, UOtherRace_10_17_VAX,FWhite_10_17_VAX, MWhite_10_17_VAX, UWhite_10_17_VAX, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>% mutate(total_18_44_vax = sum(FAsianNHPI_18_44_VAX, MAsianNHPI_18_44_VAX, UAsianNHPI_18_44_VAX,FBlack_18_44_VAX, MBlack_18_44_VAX, UBlack_18_44_VAX,FNone_18_44_VAX, MNone_18_44_VAX, UNone_18_44_VAX,FOtherRace_18_44_VAX, MOtherRace_18_44_VAX, UOtherRace_18_44_VAX,FWhite_18_44_VAX, MWhite_18_44_VAX, UWhite_18_44_VAX, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>% mutate(total_45_64_vax = sum(FAsianNHPI_45_64_VAX, MAsianNHPI_45_64_VAX, UAsianNHPI_45_64_VAX,FBlack_45_64_VAX, MBlack_45_64_VAX, UBlack_45_64_VAX,FNone_45_64_VAX, MNone_45_64_VAX, UNone_45_64_VAX,FOtherRace_45_64_VAX, MOtherRace_45_64_VAX, UOtherRace_45_64_VAX,FWhite_45_64_VAX, MWhite_45_64_VAX, UWhite_45_64_VAX, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_vax_Feb <- Zip_30907_vax_Feb %>% rowwise() %>% mutate(total_65Plus_vax = sum(FAsianNHPI_65Plus_VAX, MAsianNHPI_65Plus_VAX, UAsianNHPI_65Plus_VAX,FBlack_65Plus_VAX, MBlack_65Plus_VAX, UBlack_65Plus_VAX,FNone_65Plus_VAX, MNone_65Plus_VAX, UNone_65Plus_VAX,FOtherRace_65Plus_VAX, MOtherRace_65Plus_VAX, UOtherRace_65Plus_VAX,FWhite_65Plus_VAX, MWhite_65Plus_VAX, UWhite_65Plus_VAX, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_vax_Feb_sums <- Zip_30907_vax_Feb[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Feb
+```{r}
+colnames(Zip_30907_vax_Feb_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_vax_Feb_26", "total_male_vax_Feb_26", "total_female_vax_Feb_26", "total_masian_vax_Feb_26", "total_mblack_vax_Feb_26", "total_morace_vax_Feb_26", "total_mwhite_vax_Feb_26", "total_fasian_vax_Feb_26", "total_fblack_vax_Feb_26", "total_forace_vax_Feb_26", "total_fwhite_vax_Feb_26", "total_asian_vax_Feb_26","total_black_vax_Feb_26", "total_orace_vax_Feb_26", "total_white_vax_Feb_26", "total_05_09_vax_Feb_26", "total_10_17_vax_Feb_26", "total_18_44_vax_Feb_26", "total_45_64_vax_Feb_26", "total_65Plus_vax_Feb_26")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_30907_Feb.rds")
+
+saveRDS(Zip_30907_vax_Feb_sums, save_data_location4)
+```

### 2023-04-10 products/Manuscript.qmd
Part 4

@@
     highlight-style: github
-bibliography: ../My_Library.bib
-csl: ../apa.csl
+bibliography: ./My_Library.bib
+csl: ./apa.csl
 ---
@@
 There are 187 variables each are a different combination of gender, race and age.
-The data is uploaded weekly.
+The data is uploaded monthly.
 
@@
 This allowed me to work only with census tracts from Dougherty county.
-I then summed the data by demographic to get demographic totals
+I then summed the data by demographic to get demographic totals.
 
 # Results
+
+## Vaccination Trends of the Black Community for Each Zip Code 
+```{r}
+knitr::include_graphics("../results/31701_total_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/31705_total_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/31707_total_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/30815_total_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/30907_total_black_vax.png")
+```
+
+## Vaccination Trends of the Black Community for Each Zip Code by Gender
+```{r}
+knitr::include_graphics("../results/31701_Gender_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/31705_Gender_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/31707_Gender_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/30815_Gender_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/30907_Gender_black_vax.png")
+```
+
+## Vaccination Trends of the Black Community for Each Zip Code by Gender
+```{r}
+knitr::include_graphics("../results/31701_Age_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/31705_Age_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/31707_Age_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/30815_Age_black_vax.png")
+```
+
+```{r}
+knitr::include_graphics("../results/30907_Age_black_vax.png")
+```
+
+## LASSO Results Visual 
+```{r}
+knitr::include_graphics("../results/lasso_grid_plot.png")
+```
+
+## Final LASSO Results Variable Importance 
+```{r}
+knitr::include_graphics("../results/final_lasso_plot.png")
+```
 

### 2023-04-10 products/supplement.qmd
Part 4

@@
 
-        -   `processingcode1.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in September 2022
-
-        -   `processingcode1.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in September 2022
-
-        -   `processingcode2.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022
-
-        -   `processingcode2.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022
-
-        -   `processingcode3.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in November 2022
-
-        -   `processingcode3.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in October 2022
-
-        -   `processingcode4.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in December 2022
-
-        -   `processingcode4.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in December 2022
-
-        -   `processingcode5.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in January 2023
-
-        -   `processingcode5.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in January 2022
-
-        -   `processingcode6.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in February 2023
-
-        -   `processingcode6.R`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in February 2023
+        -   `processingcode.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in September 2022
 
@@
 
-        -   `analysiscode1.qmd`: Contains the code an analysis of the total amount of African Americans with at least one COVID-19 vaccine in all three target zip codes
-        -   `analysiscode1.html`: Contains the output document from rendering `analysiscode1.qmd`
+        -   `exploratoryanalysis.qmd`: Contains the code to an analysis of the total amount of African Americans with at least one COVID-19 vaccine in all three target zip codes and two control zip codes
+
+        -   `statisticalanalysis.qmd`: Contains the code to a statistical analysis of the total amount of African Americans with at least one COVID-19 vaccine in all three target zip codes and two control zip codes
+
         -   `readme.md`: provides a brief description of the analysis files
@@
 
--   `results` folder: Contains the graphs produced from rendering the analysis `analysiscode1.qmd`
+-   `results` folder: Contains the graphs produced from rendering the analysis
 

### 2023-04-18 code/processing_code/UntitledQMD2.qmd
Population sums by zip

# Processing Section for Control Zip Code

## Processing 1: Getting Population Totals by Control Zip Code 30815 in the Month of September

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept_17 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Sept <- filter(Sept_17, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Sept [Richmond_POP_Sept < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_POP_Sept = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Sept <- Richmond_POP_Sept  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Sept<- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Richmond_POP_Sept$total_asian_POP <- Richmond_POP_Sept$total_masian_POP + Richmond_POP_Sept$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Sept$total_black_POP <- Richmond_POP_Sept$total_mblack_POP + Richmond_POP_Sept$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Sept$total_orace_POP <- Richmond_POP_Sept$total_morace_POP + Richmond_POP_Sept$total_forace_POP

#POP total for White Residents
Richmond_POP_Sept$total_white_POP <- Richmond_POP_Sept$total_mwhite_POP + Richmond_POP_Sept$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Sept_sums <- Richmond_POP_Sept[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
#All of the following FIPS or census tracts correspond to the 30815 zip code
Zip_30815_POP_Sept <-filter(Richmond_POP_Sept_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
```
Not that the new data set `Zip_30815_POP_Sept` only has 7 observations because the orinigal data set only had 7 of the 13 census tracts found within the zip code 30815

### Change the column names for 30815 Sept
```{r}
colnames(Zip_30815_POP_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17") 
```

### Deleting excess environment objects and saving data as processed_data_POP_sept
```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Sept,Sept_17, codebook, Richmond_POP_Sept_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Sept.rds")

saveRDS(Zip_30815_POP_Sept, save_data_location4)
```

## Processing 2: Getting Population Totals by Target Zips 30815 for the Month of October

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Oct_28.xlsx")
Oct_28 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Oct <- filter(Oct_28, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Oct [Richmond_POP_Oct < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_POP_Oct = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Oct <- Richmond_POP_Oct  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Oct<- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Oct$total_asian_POP <- Richmond_POP_Oct$total_masian_POP + Richmond_POP_Oct$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Oct$total_black_POP <- Richmond_POP_Oct$total_mblack_POP + Richmond_POP_Oct$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Oct$total_orace_POP <- Richmond_POP_Oct$total_morace_POP + Richmond_POP_Oct$total_forace_POP

#POP total for White Residents
Richmond_POP_Oct$total_white_POP <- Richmond_POP_Oct$total_mwhite_POP + Richmond_POP_Oct$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Oct_sums <- Richmond_POP_Oct[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
#All of the following FIPS or census tracts correspond to the 30815 zip code

Zip_30815_POP_Oct <-filter(Richmond_POP_Oct_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
```

### Change the column names for 30815 Oct

```{r}
colnames(Zip_30815_POP_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28") 
```

### Deleting excess environment objects and saving data as processed_data_POP_Oct

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Oct,Oct_28, codebook, Richmond_POP_Oct_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Oct.rds")

saveRDS(Zip_30815_POP_Oct, file = save_data_location4)
```

## Processing 3: : Getting Population Totals by Target Zips 30815 for the Month of November

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Nov_24.xlsx")
Nov_24 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Nov_24)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Nov <- filter(Nov_24, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Nov [Richmond_POP_Nov < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_POP_Nov = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Nov <- Richmond_POP_Nov  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Nov<- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Nov$total_asian_POP <- Richmond_POP_Nov$total_masian_POP + Richmond_POP_Nov$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Nov$total_black_POP <- Richmond_POP_Nov$total_mblack_POP + Richmond_POP_Nov$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Nov$total_orace_POP <- Richmond_POP_Nov$total_morace_POP + Richmond_POP_Nov$total_forace_POP

#POP total for White Residents
Richmond_POP_Nov$total_white_POP <- Richmond_POP_Nov$total_mwhite_POP + Richmond_POP_Nov$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Nov_sums <- Richmond_POP_Nov[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
#All of the following FIPS or census tracts correspond to the 30815 zip code

Zip_30815_POP_Nov <-filter(Richmond_POP_Nov_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")

```

### Change the column names for 30815 Nov

```{r}
colnames(Zip_30815_POP_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24") 
```

### Deleting excess environment objects and saving data as processed_data_POP_Nov

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Nov,Nov_24, codebook, Richmond_POP_Nov_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Nov.rds")

saveRDS(Zip_30815_POP_Nov, file = save_data_location4)

```

## Processing 4: Getting Population Totals by Target Zips 30815 for the Month of December

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Dec_29.xlsx")
Dec_29 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Dec_29)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Dec <- filter(Dec_29, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Dec [Richmond_POP_Dec < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_POP_Dec = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Dec <- Richmond_POP_Dec  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Dec<- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Dec$total_asian_POP <- Richmond_POP_Dec$total_masian_POP + Richmond_POP_Dec$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Dec$total_black_POP <- Richmond_POP_Dec$total_mblack_POP + Richmond_POP_Dec$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Dec$total_orace_POP <- Richmond_POP_Dec$total_morace_POP + Richmond_POP_Dec$total_forace_POP

#POP total for White Residents
Richmond_POP_Dec$total_white_POP <- Richmond_POP_Dec$total_mwhite_POP + Richmond_POP_Dec$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Dec_sums <- Richmond_POP_Dec[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
Zip_30815_POP_Dec <-filter(Richmond_POP_Dec_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")

```

### Change the column names for 30815 zip code data

```{r}
colnames(Zip_30815_POP_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29")

```

### Deleting excess environment objects and saving data as processed_data_POP_Dec

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Dec,Dec_29, codebook, Richmond_POP_Dec_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Dec.rds")

saveRDS(Zip_30815_POP_Dec, file = save_data_location4)
```

## Processing 5: Getting Population Totals by Target Zips 30815 for the Month of January

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Jan_26.xlsx")
Jan_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Jan_26)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Jan <- filter(Jan_26, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Jan [Richmond_POP_Jan < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_POP_Jan = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Jan <- Richmond_POP_Jan  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Jan<- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Jan$total_asian_POP <- Richmond_POP_Jan$total_masian_POP + Richmond_POP_Jan$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Jan$total_black_POP <- Richmond_POP_Jan$total_mblack_POP + Richmond_POP_Jan$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Jan$total_orace_POP <- Richmond_POP_Jan$total_morace_POP + Richmond_POP_Jan$total_forace_POP

#POP total for White Residents
Richmond_POP_Jan$total_white_POP <- Richmond_POP_Jan$total_mwhite_POP + Richmond_POP_Jan$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Jan_sums <- Richmond_POP_Jan[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
Zip_30815_POP_Jan <-filter(Richmond_POP_Jan_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")

```

### Change the column names for 30815 Jan

```{r}
colnames(Zip_30815_POP_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Jan_26", "total_male_POP_Jan_26", "total_female_POP_Jan_26", "total_masian_POP_Jan_26", "total_mblack_POP_Jan_26", "total_morace_POP_Jan_26", "total_mwhite_POP_Jan_26", "total_fasian_POP_Jan_26", "total_fblack_POP_Jan_26", "total_forace_POP_Jan_26", "total_fwhite_POP_Jan_26", "total_asian_POP_Jan_26","total_black_POP_Jan_26", "total_orace_POP_Jan_26", "total_white_POP_Jan_26", "total_05_09_POP_Jan_26", "total_10_17_POP_Jan_26", "total_18_44_POP_Jan_26", "total_45_64_POP_Jan_26", "total_65Plus_POP_Jan_26") 
```

### Deleting excess environment objects and saving data as processed_data_POP_Jan

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Jan,Jan_26, codebook, Richmond_POP_Jan_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Jan.rds")

saveRDS(Zip_30815_POP_Jan, file = save_data_location4)
```

## Processing 6: : Getting Population Totals by Target Zips 30815 for the Month of February

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Feb_26.xlsx")
Feb_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

## Summary of data

```{r}
glimpse(Feb_26)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Richmond_POP_Feb <- filter(Feb_26, COUNTY_NAME == "Richmond County")

#replacing negative values with 0
Richmond_POP_Feb [Richmond_POP_Feb < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_POP_Feb = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Richmond_POP_Feb <- Richmond_POP_Feb  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Richmond_POP_Feb<- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.

```{r}
#POP total Asian Residents
Richmond_POP_Feb$total_asian_POP <- Richmond_POP_Feb$total_masian_POP + Richmond_POP_Feb$total_fasian_POP

#POP total for African American Residents 
Richmond_POP_Feb$total_black_POP <- Richmond_POP_Feb$total_mblack_POP + Richmond_POP_Feb$total_fblack_POP

#POP total for Other Race Residents
Richmond_POP_Feb$total_orace_POP <- Richmond_POP_Feb$total_morace_POP + Richmond_POP_Feb$total_forace_POP

#POP total for White Residents
Richmond_POP_Feb$total_white_POP <- Richmond_POP_Feb$total_mwhite_POP + Richmond_POP_Feb$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Richmond_POP_Feb_sums <- Richmond_POP_Feb[,c(1:4,185:204)]  
```

### Filtering the data by census tract within the 30815 zip code

```{r}
#filtering data by zip: 30815  
Zip_30815_POP_Feb <-filter(Richmond_POP_Feb_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")

```

### Change the column names for 30815 Feb

```{r}
colnames(Zip_30815_POP_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26") 
```

### Deleting excess environment objects and saving data as processed_data_POP_Feb

```{r}
#Only need to save processed data for the three zip codes
rm(Richmond_POP_Feb,Feb_26, codebook, Richmond_POP_Feb_sums)

#saving clean data for 30815 to processed_data_POP folder
save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Feb.rds")

saveRDS(Zip_30815_POP_Feb, file = save_data_location4)
```

# Processing Section for Control Zip Code: 30907

## Processing 1: Getting Population Totals by Control Zip Code 30907 in the Month of September

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Sept_17.xlsx")
Sept_17 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Sept <-filter(Sept_17, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Sept)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Sept [Zip_30907_POP_Sept < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_POP_Sept = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Sept <- Zip_30907_POP_Sept  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Sept<- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Sept$total_asian_POP <- Zip_30907_POP_Sept$total_masian_POP + Zip_30907_POP_Sept$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Sept$total_black_POP <- Zip_30907_POP_Sept$total_mblack_POP + Zip_30907_POP_Sept$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Sept$total_orace_POP <- Zip_30907_POP_Sept$total_morace_POP + Zip_30907_POP_Sept$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Sept$total_white_POP <- Zip_30907_POP_Sept$total_mwhite_POP + Zip_30907_POP_Sept$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Sept_sums <- Zip_30907_POP_Sept[,c(1:4,185:204)]  
```

### Change the column names for 30907 Sept
```{r}
colnames(Zip_30907_POP_Sept_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Sept.rds")

saveRDS(Zip_30907_POP_Sept_sums, save_data_location4)
```

## Processing 2: Getting Population Totals by Control Zip Code 30907 in the Month of October

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Oct_28.xlsx")
Oct_28 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Oct <-filter(Oct_28, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Oct)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Oct [Zip_30907_POP_Oct < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_POP_Oct = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Oct <- Zip_30907_POP_Oct  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Oct<- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Oct$total_asian_POP <- Zip_30907_POP_Oct$total_masian_POP + Zip_30907_POP_Oct$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Oct$total_black_POP <- Zip_30907_POP_Oct$total_mblack_POP + Zip_30907_POP_Oct$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Oct$total_orace_POP <- Zip_30907_POP_Oct$total_morace_POP + Zip_30907_POP_Oct$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Oct$total_white_POP <- Zip_30907_POP_Oct$total_mwhite_POP + Zip_30907_POP_Oct$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Oct_sums <- Zip_30907_POP_Oct[,c(1:4,185:204)]  
```

### Change the column names for 30907 Oct
```{r}
colnames(Zip_30907_POP_Oct_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Oct.rds")

saveRDS(Zip_30907_POP_Oct_sums, save_data_location4)
```

## Processing 3: : Getting Population Totals by Control Zip Code 30907 in the Month of November

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Nov_24.xlsx")
Nov_24 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Nov <-filter(Nov_24, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Nov)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Nov [Zip_30907_POP_Nov < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_POP_Nov = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Nov <- Zip_30907_POP_Nov  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Nov<- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Nov$total_asian_POP <- Zip_30907_POP_Nov$total_masian_POP + Zip_30907_POP_Nov$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Nov$total_black_POP <- Zip_30907_POP_Nov$total_mblack_POP + Zip_30907_POP_Nov$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Nov$total_orace_POP <- Zip_30907_POP_Nov$total_morace_POP + Zip_30907_POP_Nov$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Nov$total_white_POP <- Zip_30907_POP_Nov$total_mwhite_POP + Zip_30907_POP_Nov$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Nov_sums <- Zip_30907_POP_Nov[,c(1:4,185:204)]  
```

### Change the column names for 30907 Nov
```{r}
colnames(Zip_30907_POP_Nov_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Nov.rds")

saveRDS(Zip_30907_POP_Nov_sums, save_data_location4)
```

## Processing 4: : Getting Population Totals by Control Zip Code 30907 in the Month of December

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Dec_29.xlsx")
Dec_29 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Dec <-filter(Dec_29, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Dec)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Dec [Zip_30907_POP_Dec < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_POP_Dec = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Dec <- Zip_30907_POP_Dec  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Dec<- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Dec$total_asian_POP <- Zip_30907_POP_Dec$total_masian_POP + Zip_30907_POP_Dec$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Dec$total_black_POP <- Zip_30907_POP_Dec$total_mblack_POP + Zip_30907_POP_Dec$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Dec$total_orace_POP <- Zip_30907_POP_Dec$total_morace_POP + Zip_30907_POP_Dec$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Dec$total_white_POP <- Zip_30907_POP_Dec$total_mwhite_POP + Zip_30907_POP_Dec$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Dec_sums <- Zip_30907_POP_Dec[,c(1:4,185:204)]  
```

### Change the column names for 30907 Dec
```{r}
colnames(Zip_30907_POP_Dec_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Dec.rds")

saveRDS(Zip_30907_POP_Dec_sums, save_data_location4)
```

## Processing 5: : Getting Population Totals by Control Zip Code 30907 in the Month of January

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Jan_26.xlsx")
Jan_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Jan <-filter(Jan_26, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Jan)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Jan [Zip_30907_POP_Jan < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_POP_Jan = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Jan <- Zip_30907_POP_Jan  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Jan<- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Jan$total_asian_POP <- Zip_30907_POP_Jan$total_masian_POP + Zip_30907_POP_Jan$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Jan$total_black_POP <- Zip_30907_POP_Jan$total_mblack_POP + Zip_30907_POP_Jan$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Jan$total_orace_POP <- Zip_30907_POP_Jan$total_morace_POP + Zip_30907_POP_Jan$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Jan$total_white_POP <- Zip_30907_POP_Jan$total_mwhite_POP + Zip_30907_POP_Jan$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Jan_sums <- Zip_30907_POP_Jan[,c(1:4,185:204)]  
```

### Change the column names for 30907 Jan
```{r}
colnames(Zip_30907_POP_Jan_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Jan_26", "total_male_POP_Jan_26", "total_female_POP_Jan_26", "total_masian_POP_Jan_26", "total_mblack_POP_Jan_26", "total_morace_POP_Jan_26", "total_mwhite_POP_Jan_26", "total_fasian_POP_Jan_26", "total_fblack_POP_Jan_26", "total_forace_POP_Jan_26", "total_fwhite_POP_Jan_26", "total_asian_POP_Jan_26","total_black_POP_Jan_26", "total_orace_POP_Jan_26", "total_white_POP_Jan_26", "total_05_09_POP_Jan_26", "total_10_17_POP_Jan_26", "total_18_44_POP_Jan_26", "total_45_64_POP_Jan_26", "total_65Plus_POP_Jan_26") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Jan.rds")

saveRDS(Zip_30907_POP_Jan_sums, save_data_location4)
```

## Processing 6: : Getting Population Totals by Control Zip Code 30907 in the Month of February

## Raw Data & Codebook loading

```{r}
# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
rawdata_location = here("data","raw_data","Feb_26.xlsx")
Feb_26 <- read_excel(rawdata_location)

#loading data code book for reference 
codebook_location = here("data","raw_data","Codebook.xlsx")
codebook = read_excel(codebook_location)
```

### Separating the data by zip code and removing negative values

```{r}
#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
#by using filter() I am able to create a new dataset with Populations from Richmond County 
Zip_30907_POP_Feb <-filter(Feb_26, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")

#taking a look at the data
head(Zip_30907_POP_Feb)

#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county. This is okay because census tracts are not completely contained by one county or zip code.

#replacing negative values with 0
Zip_30907_POP_Feb [Zip_30907_POP_Feb < 0] <- 0  
```

### Adding new columns: total POP sum, total POP of all males, total POP of all females

```{r}
#Adding a column of the total POP for each census tract (for each observation)  
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_POP_Feb = sum(across(ends_with("POP")), na.rm = T))

#total POP column for all males
Zip_30907_POP_Feb <- Zip_30907_POP_Feb  %>% rowwise() %>%   
  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))

#total POP column for all females
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Male Residents by race

```{r}
#total POP column for Asian male residents 
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American male residents
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race male residents
Zip_30907_POP_Feb<- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White male residents 
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))

```

### Adding new columns for Female Residents by race

```{r}
#total POP column for Asian female residents    
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))

#total POP column for African American female residents
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))

#total POP column for other race female residents
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))

#total POP column for White female residents 
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
```

### Adding new columns for POP totals by race

Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.

For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.

```{r}
#POP total Asian Residents
Zip_30907_POP_Feb$total_asian_POP <- Zip_30907_POP_Feb$total_masian_POP + Zip_30907_POP_Feb$total_fasian_POP

#POP total for African American Residents 
Zip_30907_POP_Feb$total_black_POP <- Zip_30907_POP_Feb$total_mblack_POP + Zip_30907_POP_Feb$total_fblack_POP

#POP total for Other Race Residents
Zip_30907_POP_Feb$total_orace_POP <- Zip_30907_POP_Feb$total_morace_POP + Zip_30907_POP_Feb$total_forace_POP

#POP total for White Residents
Zip_30907_POP_Feb$total_white_POP <- Zip_30907_POP_Feb$total_mwhite_POP + Zip_30907_POP_Feb$total_fwhite_POP  
```

### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+

```{r}
#summing all residents in the age range of 05-09
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))

#summing all residents in the age range of 10-17
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))

#summing all residents in the age range of 18-44
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))

#summing all residents in the age range of 45-64
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))

#summing all residents in the age range of 65+ 
Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))

```

### Creating a data set with only the sums columns I just created via the previous code

```{r}
#Extract necessary variables only  
Zip_30907_POP_Feb_sums <- Zip_30907_POP_Feb[,c(1:4,185:204)]  
```

### Change the column names for 30907 Feb
```{r}
colnames(Zip_30907_POP_Feb_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26") 
```

### Saving data
```{r}
#saving clean data for 30907 to processed_data_POP folder
save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Feb.rds")

saveRDS(Zip_30907_POP_Feb_sums, save_data_location4)
```

### 2023-04-18 code/processing_code/processingcode2.qmd
Population sums by zip

--- 2023-03-07 code/processing_code/processingcode2.qmd
@@
 ---
-title: "Processing script 2"
-author: "Jacob Matta"
-date: "2023-01-03"
-output: html_document
+title: "Processing Code for Populations of Zips"
+output: 
+  html_document
 ---
-# Processing script 2: Understanding the rawdata and data source
+# Processing Section for Populations of Target Zip Codes
 
-## Determining the count of residents in GA with at least one dose of the covid-19 vaccines
-
-The count of residents in GA with at least one dose of the Covid-19 vaccines as of Feb 26th, 2023 is 6,935,491.
-This was found from the Georgia Department of [Public Health Vaccine Distribution Dashboard](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c) on Feb 26th/
-
-## This processing script will look to see how the raw data used for this analysis compares to the overall number of 6,935,491 as of Feb 26th, 2023.
-
-The overall sum of residents in GA with at least one dose of the covid-19 vaccines as of Feb 26th, 2023 should equal 6,935,491
+## Processing 1: Getting Population Totals by Target Zips 31701, 31705 & 31707 for the Month of September
 
 ### Load needed packages. make sure they are installed.
-
 ```{r}
-library(readxl) #for loading Excel files
-library(here) #used for setting a path 
+library(here)
 library(readxl) #for loading Excel files
@@
 
-### Raw Data loading
+## Raw Data & Codebook loading
 
 ```{r}
-rawdata_location = here("data","raw_data","Feb_26.xlsx")
-Feb_26 <- read_excel(rawdata_location)
-```
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Sept_17.xlsx")
+Sept_17 <- read_excel(rawdata_location)
 
-### Codebook loading
-
-```{r}
 #loading data code book for reference 
@@
 
-### Cleaning
+### Separating the data by zip code and removing negative values
 
 ```{r}
-#dropping columns that i don't need
-#keeping county name to group data together 
-Feb_26 <- Feb_26[-c(1:3)]
+#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
+Dougherty_POP_Sept <- filter(Sept_17, COUNTY_NAME == "Dougherty County")
 
-#removing all of the population data 
-Feb_26 = Feb_26[-c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21, 23, 25, 27, 29, 31, 33, 35, 37, 39, 41, 43, 45, 47, 49, 51, 53, 55, 57, 59, 61, 63, 65, 67, 69, 71, 73, 75, 77, 79, 81, 83, 85, 87, 89, 91, 93, 95, 97, 99, 101, 103, 105, 107, 109, 111, 113, 115, 117, 119, 121, 123, 125, 127, 129, 131, 133, 135, 137, 139, 141, 143, 145, 147, 149, 151, 153, 155, 157, 159, 161, 163, 165, 167, 169, 171, 173, 175, 177, 179, 181.)]
-Feb_26[-c(1)]
+#replacing negative values with 0
+Dougherty_POP_Sept [Dougherty_POP_Sept < 0] <- 0  
+```
 
-#removing the negative values 
-Feb_26 [Feb_26 < 0] <- 0
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_POP_Sept = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Dougherty_POP_Sept <- Dougherty_POP_Sept  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
 
@@
 
-### Summing the vaccinations in each column
+### Adding new columns for Male Residents by race
 
 ```{r}
-Feb_26_sum_by_col = colSums(Feb_26[ , c(2:91)])     
-Feb_26_sum_by_col
+#total POP column for Asian male residents 
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Dougherty_POP_Sept<- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
 ```
 
-### Turning the sums by column into a tibble that puts all of the sums of each column into one column
+### Adding new columns for Female Residents by race
 
 ```{r}
-Feb_26_sum_tib = tibble(Feb_26_sum_by_col)
-Feb_26_sum_tib
+#total POP column for Asian female residents    
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
 ```
 
-### Summing all of the vaccinations together
+### Adding new columns for POP totals by race
 
 ```{r}
-Overall_sum = sum(Feb_26_sum_tib) 
-Overall_sum
+#POP total Asian Residents
+Dougherty_POP_Sept$total_asian_POP <- Dougherty_POP_Sept$total_masian_POP + Dougherty_POP_Sept$total_fasian_POP
+
+#POP total for African American Residents 
+Dougherty_POP_Sept$total_black_POP <- Dougherty_POP_Sept$total_mblack_POP + Dougherty_POP_Sept$total_fblack_POP
+
+#POP total for Other Race Residents
+Dougherty_POP_Sept$total_orace_POP <- Dougherty_POP_Sept$total_morace_POP + Dougherty_POP_Sept$total_forace_POP
+
+#POP total for White Residents
+Dougherty_POP_Sept$total_white_POP <- Dougherty_POP_Sept$total_mwhite_POP + Dougherty_POP_Sept$total_fwhite_POP  
 ```
 
-### Thoughts
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
 
-The overall total of residents in GA with at least one dose of the covid-19 vaccines is 5,654,635 which is more than 1 million should of the total about of GA residents with at least one dose of a Covid-19 vaccine that the Georgia Department of Public Health Vaccine Distribution Dashboard reported on Feb, 26th, 2023.
+```{r}
+#summing all residents in the age range of 05-09
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
 
-The data that I am using for this project possibly excludes certain vaccines brands (eg.
-Pfizer, Moderna, Johnson & Johnson) whereas the overall vaccination count represented on the website includes all vaccination brands.
-This is a question I will attempt to figure out.
-For now, it is my understanding that there are 5,654,635 Ga residents by census tract who currently have one dose of the Covid-19 vaccine.
-The brands of Covid-19 vaccines included in the census tract data has yet to be specified.
+#summing all residents in the age range of 10-17
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
 
-### Explanation for difference Covid-19 Vaccination totals
+#summing all residents in the age range of 18-44
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
 
-1.  The data I downloaded is by census tract
+#summing all residents in the age range of 45-64
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
 
-2.
-The data displayed on the Georgia Department of [Public Health Vaccine Distribution Dashboard](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c) on Feb 26th is by county level which does not align with census tracts
+#summing all residents in the age range of 65+ 
+Dougherty_POP_Sept <- Dougherty_POP_Sept %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
 
-3.
-Going forward for details regarding the raw data you should go to arcGIS insights dashboard for Covid-19 vaccinations among GA Residents with at least one dose found [here](https://insights.arcgis.com/#/view/08d2c1f995524053b1f9669dd0b0b995)
+```
 
-4.
-The data dashboard for Covid-19 vaccinations among GA Residents with at least one dose (link above) states that the data downloaded by census tract will not add up to the county level and state level data
+### Creating a data set with only the sums columns I just created via the previous code
 
-### Moving forward
-If the vaccinations are not rates (eg.
-per 100,000) why are there negative values and why does it state that the where "-1 is indicated, either the numerator is less than five or the denominator is less than 25".
--   What is the numerator and what is the denominator?
+```{r}
+#Extract necessary variables only  
+Dougherty_POP_Sept_sums <- Dougherty_POP_Sept[,c(1:4,185:204)]  
+```
+
+### Filtering the data by zip code
+
+```{r}
+#filtering data by zip: 31701  
+#All of the following FIPS or census tracts correspond to the 31701 zip code
+Zip_31701_Sept_POP <-filter(Dougherty_POP_Sept_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
+
+#filtering data by zip: 31705  
+#All of the following FIPS or census tracts correspond to the 31705 zip code
+Zip_31705_Sept_POP <- filter(Dougherty_POP_Sept_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
+
+#filtering data by zip: 31707  
+#All of the following FIPS or census tracts correspond to the 31707 zip code
+Zip_31707_Sep_POP <- filter(Dougherty_POP_Sept_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
+
+```
+
+### Change the column names for 31701 Sept
+
+```{r}
+colnames(Zip_31701_Sept_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17")
+
+#change column names for 31705 Sept  
+colnames(Zip_31705_Sept_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17")
+
+#change column names for 31707 Sept  
+colnames(Zip_31707_Sep_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17")
+
+```
+
+### Deleting excess environment objects and saving data as processed_data_sept
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Dougherty_POP_Sept,Sept_17, codebook, Dougherty_POP_Sept_sums)
+
+#saving clean data for 31701 to processed_data folder
+save_data_location = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Sept.rds")
+
+saveRDS(Zip_31701_Sept_POP, file = save_data_location)
+
+#saving clean data for 31705 to processed_data folder
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Sept.rds")
+
+saveRDS(Zip_31705_Sept_POP, file = save_data_location2)
+
+#saving clean data for 31707 to processed_data folder
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Sept.rds")
+
+saveRDS(Zip_31707_Sep_POP, file = save_data_location3)
+
+```
+
+## Processing 2: Getting Population Totals by Target Zips 31701, 31705 & 31707 for the Month of October 
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Oct_28.xlsx")
+Oct_28 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Octarating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
+Dougherty_POP_Oct <- filter(Oct_28, COUNTY_NAME == "Dougherty County")
+
+#replacing negative values with 0
+Dougherty_POP_Oct [Dougherty_POP_Oct < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_POP_Oct = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Dougherty_POP_Oct <- Dougherty_POP_Oct  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Dougherty_POP_Oct<- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+```{r}
+#POP total Asian Residents
+Dougherty_POP_Oct$total_asian_POP <- Dougherty_POP_Oct$total_masian_POP + Dougherty_POP_Oct$total_fasian_POP
+
+#POP total for African American Residents 
+Dougherty_POP_Oct$total_black_POP <- Dougherty_POP_Oct$total_mblack_POP + Dougherty_POP_Oct$total_fblack_POP
+
+#POP total for Other Race Residents
+Dougherty_POP_Oct$total_orace_POP <- Dougherty_POP_Oct$total_morace_POP + Dougherty_POP_Oct$total_forace_POP
+
+#POP total for White Residents
+Dougherty_POP_Oct$total_white_POP <- Dougherty_POP_Oct$total_mwhite_POP + Dougherty_POP_Oct$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Dougherty_POP_Oct <- Dougherty_POP_Oct %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Dougherty_POP_Oct_sums <- Dougherty_POP_Oct[,c(1:4,185:204)]  
+```
+
+### Filtering the data by zip code
+
+```{r}
+#filtering data by zip: 31701  
+#All of the following FIPS or census tracts correspond to the 31701 zip code
+Zip_31701_Oct_POP <-filter(Dougherty_POP_Oct_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
+
+#filtering data by zip: 31705  
+#All of the following FIPS or census tracts correspond to the 31705 zip code
+Zip_31705_Oct_POP <- filter(Dougherty_POP_Oct_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
+
+#filtering data by zip: 31707  
+#All of the following FIPS or census tracts correspond to the 31707 zip code
+Zip_31707_Oct_POP <- filter(Dougherty_POP_Oct_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
+
+```
+
+### Change the column names for 31701 Oct
+
+```{r}
+colnames(Zip_31701_Oct_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28")
+
+#change column names for 31705 Oct  
+colnames(Zip_31705_Oct_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28")
+
+#change column names for 31707 Oct  
+colnames(Zip_31707_Oct_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28")
+
+```
+
+### Deleting excess environment objects and saving data as processed_data_Oct
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Dougherty_POP_Oct,Oct_28, codebook, Dougherty_POP_Oct_sums)
+
+#saving clean data for 31701 to processed_data folder
+save_data_location = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Oct.rds")
+
+saveRDS(Zip_31701_Oct_POP, file = save_data_location)
+
+#saving clean data for 31705 to processed_data folder
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Oct.rds")
+
+saveRDS(Zip_31705_Oct_POP, file = save_data_location2)
+
+#saving clean data for 31707 to processed_data folder
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Oct.rds")
+
+saveRDS(Zip_31707_Oct_POP, file = save_data_location3)
+
+```
+
+## Processing 3: Getting Population Totals by Target Zips 31701, 31705 & 31707 for the Month of November
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Nov_24.xlsx")
+Nov_24 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Novarating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
+Dougherty_POP_Nov <- filter(Nov_24, COUNTY_NAME == "Dougherty County")
+
+#replacing negative values with 0
+Dougherty_POP_Nov [Dougherty_POP_Nov < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>% mutate(total_POP_Nov = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Dougherty_POP_Nov <- Dougherty_POP_Nov  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Dougherty_POP_Nov<- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+```{r}
+#POP total Asian Residents
+Dougherty_POP_Nov$total_asian_POP <- Dougherty_POP_Nov$total_masian_POP + Dougherty_POP_Nov$total_fasian_POP
+
+#POP total for African American Residents 
+Dougherty_POP_Nov$total_black_POP <- Dougherty_POP_Nov$total_mblack_POP + Dougherty_POP_Nov$total_fblack_POP
+
+#POP total for Other Race Residents
+Dougherty_POP_Nov$total_orace_POP <- Dougherty_POP_Nov$total_morace_POP + Dougherty_POP_Nov$total_forace_POP
+
+#POP total for White Residents
+Dougherty_POP_Nov$total_white_POP <- Dougherty_POP_Nov$total_mwhite_POP + Dougherty_POP_Nov$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Dougherty_POP_Nov <- Dougherty_POP_Nov %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Dougherty_POP_Nov_sums <- Dougherty_POP_Nov[,c(1:4,185:204)]  
+```
+
+### Filtering the data by zip code
+
+```{r}
+#filtering data by zip: 31701  
+#All of the following FIPS or census tracts correspond to the 31701 zip code
+Zip_31701_Nov_POP <-filter(Dougherty_POP_Nov_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
+
+#filtering data by zip: 31705  
+#All of the following FIPS or census tracts correspond to the 31705 zip code
+Zip_31705_Nov_POP <- filter(Dougherty_POP_Nov_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
+
+#filtering data by zip: 31707  
+#All of the following FIPS or census tracts correspond to the 31707 zip code
+Zip_31707_Nov_POP <- filter(Dougherty_POP_Nov_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
+
+```
+
+### Change the column names for 31701 Nov
+
+```{r}
+colnames(Zip_31701_Nov_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24")
+
+#change column names for 31705 Nov  
+colnames(Zip_31705_Nov_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24")
+
+#change column names for 31707 Nov  
+colnames(Zip_31707_Nov_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24")
+
+```
+
+### Deleting excess environment objects and saving data as processed_data_Nov
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Dougherty_POP_Nov,Nov_24, codebook, Dougherty_POP_Nov_sums)
+
+#saving clean data for 31701 to processed_data folder
+save_data_location = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Nov.rds")
+
+saveRDS(Zip_31701_Nov_POP, file = save_data_location)
+
+#saving clean data for 31705 to processed_data folder
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Nov.rds")
+
+saveRDS(Zip_31705_Nov_POP, file = save_data_location2)
+
+#saving clean data for 31707 to processed_data folder
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Nov.rds")
+
+saveRDS(Zip_31707_Nov_POP, file = save_data_location3)
+
+```
+
+## Processing 4: Getting Population Totals by Target Zips 31701, 31705 & 31707 for the Month of December
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Dec_29.xlsx")
+Dec_29 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Decarating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
+Dougherty_POP_Dec <- filter(Dec_29, COUNTY_NAME == "Dougherty County")
+
+#replacing negative values with 0
+Dougherty_POP_Dec [Dougherty_POP_Dec < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>% mutate(total_POP_Dec = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Dougherty_POP_Dec <- Dougherty_POP_Dec  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Dougherty_POP_Dec<- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+```{r}
+#POP total Asian Residents
+Dougherty_POP_Dec$total_asian_POP <- Dougherty_POP_Dec$total_masian_POP + Dougherty_POP_Dec$total_fasian_POP
+
+#POP total for African American Residents 
+Dougherty_POP_Dec$total_black_POP <- Dougherty_POP_Dec$total_mblack_POP + Dougherty_POP_Dec$total_fblack_POP
+
+#POP total for Other Race Residents
+Dougherty_POP_Dec$total_orace_POP <- Dougherty_POP_Dec$total_morace_POP + Dougherty_POP_Dec$total_forace_POP
+
+#POP total for White Residents
+Dougherty_POP_Dec$total_white_POP <- Dougherty_POP_Dec$total_mwhite_POP + Dougherty_POP_Dec$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Dougherty_POP_Dec <- Dougherty_POP_Dec %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Dougherty_POP_Dec_sums <- Dougherty_POP_Dec[,c(1:4,185:204)]  
+```
+
+### Filtering the data by zip code
+
+```{r}
+#filtering data by zip: 31701  
+#All of the following FIPS or census tracts correspond to the 31701 zip code
+Zip_31701_Dec_POP <-filter(Dougherty_POP_Dec_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
+
+#filtering data by zip: 31705  
+#All of the following FIPS or census tracts correspond to the 31705 zip code
+Zip_31705_Dec_POP <- filter(Dougherty_POP_Dec_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
+
+#filtering data by zip: 31707  
+#All of the following FIPS or census tracts correspond to the 31707 zip code
+Zip_31707_Dec_POP <- filter(Dougherty_POP_Dec_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
+
+```
+
+### Change the column names for 31701 Dec
+
+```{r}
+colnames(Zip_31701_Dec_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29")
+
+#change column names for 31705 Dec  
+colnames(Zip_31705_Dec_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29")
+
+#change column names for 31707 Dec  
+colnames(Zip_31707_Dec_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29")
+
+```
+
+### Deleting excess environment objects and saving data as processed_data_Dec
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Dougherty_POP_Dec,Dec_29, codebook, Dougherty_POP_Dec_sums)
+
+#saving clean data for 31701 to processed_data folder
+save_data_location = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Dec.rds")
+
+saveRDS(Zip_31701_Dec_POP, file = save_data_location)
+
+#saving clean data for 31705 to processed_data folder
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Dec.rds")
+
+saveRDS(Zip_31705_Dec_POP, file = save_data_location2)
+
+#saving clean data for 31707 to processed_data folder
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Dec.rds")
+
+saveRDS(Zip_31707_Dec_POP, file = save_data_location3)
+
+```
+
+## Processing 6: Getting Population Totals by Target Zips 31701, 31705 & 31707 for the Month of February
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Febarating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Vaccination from Dougherty County which contains the target zip codes of 31701, 31705 and 31707
+#by using filter() I am able to create a new dataset with vaccinations from Dougherty County 
+Dougherty_POP_Feb <- filter(Feb_26, COUNTY_NAME == "Dougherty County")
+
+#replacing negative values with 0
+Dougherty_POP_Feb [Dougherty_POP_Feb < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>% mutate(total_POP_Feb = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Dougherty_POP_Feb <- Dougherty_POP_Feb  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Dougherty_POP_Feb<- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+```{r}
+#POP total Asian Residents
+Dougherty_POP_Feb$total_asian_POP <- Dougherty_POP_Feb$total_masian_POP + Dougherty_POP_Feb$total_fasian_POP
+
+#POP total for African American Residents 
+Dougherty_POP_Feb$total_black_POP <- Dougherty_POP_Feb$total_mblack_POP + Dougherty_POP_Feb$total_fblack_POP
+
+#POP total for Other Race Residents
+Dougherty_POP_Feb$total_orace_POP <- Dougherty_POP_Feb$total_morace_POP + Dougherty_POP_Feb$total_forace_POP
+
+#POP total for White Residents
+Dougherty_POP_Feb$total_white_POP <- Dougherty_POP_Feb$total_mwhite_POP + Dougherty_POP_Feb$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Dougherty_POP_Feb <- Dougherty_POP_Feb %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Dougherty_POP_Feb_sums <- Dougherty_POP_Feb[,c(1:4,185:204)]  
+```
+
+### Filtering the data by zip code
+
+```{r}
+#filtering data by zip: 31701  
+#All of the following FIPS or census tracts correspond to the 31701 zip code
+Zip_31701_Feb_POP <-filter(Dougherty_POP_Feb_sums,FIPS=="13095000700" | FIPS=="13095000800" | FIPS=="13095001403" | FIPS=="13095001500" | FIPS=="13095010601" | FIPS=="13095011300" | FIPS=="13095011400" | FIPS=="13095010602")
+
+#filtering data by zip: 31705  
+#All of the following FIPS or census tracts correspond to the 31705 zip code
+Zip_31705_Feb_POP <- filter(Dougherty_POP_Feb_sums, FIPS=="13095000100" | FIPS=="13095000200" | FIPS=="13095010302" | FIPS=="13095010700" | FIPS=="13095010900" | FIPS=="13095011000" | FIPS=="13095011200" | FIPS=="13095011600")
+
+#filtering data by zip: 31707  
+#All of the following FIPS or census tracts correspond to the 31707 zip code
+Zip_31707_Feb_POP <- filter(Dougherty_POP_Feb_sums, FIPS=="13095000400" | FIPS=="13095000501" | FIPS=="13095000502" | FIPS=="13095000600" | FIPS=="13095000900" | FIPS=="13095001000" | FIPS=="13095001100")
+
+```
+
+### Change the column names for 31701 Feb
+
+```{r}
+colnames(Zip_31701_Feb_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26")
+
+#change column names for 31705 Feb  
+colnames(Zip_31705_Feb_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26")
+
+#change column names for 31707 Feb  
+colnames(Zip_31707_Feb_POP) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26")
+
+```
+
+### Deleting excess environment objects and saving data as processed_data_Feb
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Dougherty_POP_Feb,Feb_26, codebook, Dougherty_POP_Feb_sums)
+
+#saving clean data for 31701 to processed_data folder
+save_data_location = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Feb.rds")
+
+saveRDS(Zip_31701_Feb_POP, file = save_data_location)
+
+#saving clean data for 31705 to processed_data folder
+save_data_location2 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Feb.rds")
+
+saveRDS(Zip_31705_Feb_POP, file = save_data_location2)
+
+#saving clean data for 31707 to processed_data folder
+save_data_location3 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Feb.rds")
+
+saveRDS(Zip_31707_Feb_POP, file = save_data_location3)
+
+```
+
+# Processing Section for Control Zip Code
+
+## Processing 1: Getting Population Totals by Control Zip Code 30815 in the Month of September
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Sept_17.xlsx")
+Sept_17 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Richmond_POP_Sept <- filter(Sept_17, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_POP_Sept [Richmond_POP_Sept < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_POP_Sept = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Richmond_POP_Sept <- Richmond_POP_Sept  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Richmond_POP_Sept<- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.
+
+```{r}
+#POP total Asian Residents
+Richmond_POP_Sept$total_asian_POP <- Richmond_POP_Sept$total_masian_POP + Richmond_POP_Sept$total_fasian_POP
+
+#POP total for African American Residents 
+Richmond_POP_Sept$total_black_POP <- Richmond_POP_Sept$total_mblack_POP + Richmond_POP_Sept$total_fblack_POP
+
+#POP total for Other Race Residents
+Richmond_POP_Sept$total_orace_POP <- Richmond_POP_Sept$total_morace_POP + Richmond_POP_Sept$total_forace_POP
+
+#POP total for White Residents
+Richmond_POP_Sept$total_white_POP <- Richmond_POP_Sept$total_mwhite_POP + Richmond_POP_Sept$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_POP_Sept <- Richmond_POP_Sept %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_POP_Sept_sums <- Richmond_POP_Sept[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+#All of the following FIPS or census tracts correspond to the 30815 zip code
+Zip_30815_POP_Sept <-filter(Richmond_POP_Sept_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+```
+Not that the new data set `Zip_30815_POP_Sept` only has 7 observations because the orinigal data set only had 7 of the 13 census tracts found within the zip code 30815
+
+### Change the column names for 30815 Sept
+```{r}
+colnames(Zip_30815_POP_Sept) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17")
+```
+
+### Deleting excess environment objects and saving data as processed_data_POP_sept
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_POP_Sept,Sept_17, codebook, Richmond_POP_Sept_sums)
+
+#saving clean data for 30815 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Sept.rds")
+
+saveRDS(Zip_30815_POP_Sept, save_data_location4)
+```
+
+## Processing 2: Getting Population Totals by Target Zips 30815 for the Month of October
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Oct_28.xlsx")
+Oct_28 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Richmond_POP_Oct <- filter(Oct_28, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_POP_Oct [Richmond_POP_Oct < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_POP_Oct = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Richmond_POP_Oct <- Richmond_POP_Oct  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Richmond_POP_Oct<- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.
+
+```{r}
+#POP total Asian Residents
+Richmond_POP_Oct$total_asian_POP <- Richmond_POP_Oct$total_masian_POP + Richmond_POP_Oct$total_fasian_POP
+
+#POP total for African American Residents 
+Richmond_POP_Oct$total_black_POP <- Richmond_POP_Oct$total_mblack_POP + Richmond_POP_Oct$total_fblack_POP
+
+#POP total for Other Race Residents
+Richmond_POP_Oct$total_orace_POP <- Richmond_POP_Oct$total_morace_POP + Richmond_POP_Oct$total_forace_POP
+
+#POP total for White Residents
+Richmond_POP_Oct$total_white_POP <- Richmond_POP_Oct$total_mwhite_POP + Richmond_POP_Oct$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_POP_Oct <- Richmond_POP_Oct %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_POP_Oct_sums <- Richmond_POP_Oct[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+#All of the following FIPS or census tracts correspond to the 30815 zip code
+
+Zip_30815_POP_Oct <-filter(Richmond_POP_Oct_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+```
+
+### Change the column names for 30815 Oct
+
+```{r}
+colnames(Zip_30815_POP_Oct) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28")
+```
+
+### Deleting excess environment objects and saving data as processed_data_POP_Oct
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_POP_Oct,Oct_28, codebook, Richmond_POP_Oct_sums)
+
+#saving clean data for 30815 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Oct.rds")
+
+saveRDS(Zip_30815_POP_Oct, file = save_data_location4)
+```
+
+## Processing 3: : Getting Population Totals by Target Zips 30815 for the Month of November
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Nov_24.xlsx")
+Nov_24 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+## Summary of data
+
+```{r}
+glimpse(Nov_24)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Richmond_POP_Nov <- filter(Nov_24, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_POP_Nov [Richmond_POP_Nov < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_POP_Nov = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Richmond_POP_Nov <- Richmond_POP_Nov  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Richmond_POP_Nov<- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.
+
+```{r}
+#POP total Asian Residents
+Richmond_POP_Nov$total_asian_POP <- Richmond_POP_Nov$total_masian_POP + Richmond_POP_Nov$total_fasian_POP
+
+#POP total for African American Residents 
+Richmond_POP_Nov$total_black_POP <- Richmond_POP_Nov$total_mblack_POP + Richmond_POP_Nov$total_fblack_POP
+
+#POP total for Other Race Residents
+Richmond_POP_Nov$total_orace_POP <- Richmond_POP_Nov$total_morace_POP + Richmond_POP_Nov$total_forace_POP
+
+#POP total for White Residents
+Richmond_POP_Nov$total_white_POP <- Richmond_POP_Nov$total_mwhite_POP + Richmond_POP_Nov$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_POP_Nov <- Richmond_POP_Nov %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_POP_Nov_sums <- Richmond_POP_Nov[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+#All of the following FIPS or census tracts correspond to the 30815 zip code
+
+Zip_30815_POP_Nov <-filter(Richmond_POP_Nov_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+
+```
+
+### Change the column names for 30815 Nov
+
+```{r}
+colnames(Zip_30815_POP_Nov) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24")
+```
+
+### Deleting excess environment objects and saving data as processed_data_POP_Nov
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_POP_Nov,Nov_24, codebook, Richmond_POP_Nov_sums)
+
+#saving clean data for 30815 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Nov.rds")
+
+saveRDS(Zip_30815_POP_Nov, file = save_data_location4)
+
+```
+
+## Processing 4: Getting Population Totals by Target Zips 30815 for the Month of December
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Dec_29.xlsx")
+Dec_29 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+## Summary of data
+
+```{r}
+glimpse(Dec_29)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Richmond_POP_Dec <- filter(Dec_29, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_POP_Dec [Richmond_POP_Dec < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_POP_Dec = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Richmond_POP_Dec <- Richmond_POP_Dec  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Richmond_POP_Dec<- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.
+
+```{r}
+#POP total Asian Residents
+Richmond_POP_Dec$total_asian_POP <- Richmond_POP_Dec$total_masian_POP + Richmond_POP_Dec$total_fasian_POP
+
+#POP total for African American Residents 
+Richmond_POP_Dec$total_black_POP <- Richmond_POP_Dec$total_mblack_POP + Richmond_POP_Dec$total_fblack_POP
+
+#POP total for Other Race Residents
+Richmond_POP_Dec$total_orace_POP <- Richmond_POP_Dec$total_morace_POP + Richmond_POP_Dec$total_forace_POP
+
+#POP total for White Residents
+Richmond_POP_Dec$total_white_POP <- Richmond_POP_Dec$total_mwhite_POP + Richmond_POP_Dec$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_POP_Dec <- Richmond_POP_Dec %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_POP_Dec_sums <- Richmond_POP_Dec[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+Zip_30815_POP_Dec <-filter(Richmond_POP_Dec_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+
+```
+
+### Change the column names for 30815 zip code data
+
+```{r}
+colnames(Zip_30815_POP_Dec) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29")
+
+```
+
+### Deleting excess environment objects and saving data as processed_data_POP_Dec
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_POP_Dec,Dec_29, codebook, Richmond_POP_Dec_sums)
+
+#saving clean data for 30815 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Dec.rds")
+
+saveRDS(Zip_30815_POP_Dec, file = save_data_location4)
+```
+
+## Processing 5: Getting Population Totals by Target Zips 30815 for the Month of January
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Jan_26.xlsx")
+Jan_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+## Summary of data
+
+```{r}
+glimpse(Jan_26)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Richmond_POP_Jan <- filter(Jan_26, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_POP_Jan [Richmond_POP_Jan < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_POP_Jan = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Richmond_POP_Jan <- Richmond_POP_Jan  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Richmond_POP_Jan<- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.
+
+```{r}
+#POP total Asian Residents
+Richmond_POP_Jan$total_asian_POP <- Richmond_POP_Jan$total_masian_POP + Richmond_POP_Jan$total_fasian_POP
+
+#POP total for African American Residents 
+Richmond_POP_Jan$total_black_POP <- Richmond_POP_Jan$total_mblack_POP + Richmond_POP_Jan$total_fblack_POP
+
+#POP total for Other Race Residents
+Richmond_POP_Jan$total_orace_POP <- Richmond_POP_Jan$total_morace_POP + Richmond_POP_Jan$total_forace_POP
+
+#POP total for White Residents
+Richmond_POP_Jan$total_white_POP <- Richmond_POP_Jan$total_mwhite_POP + Richmond_POP_Jan$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_POP_Jan <- Richmond_POP_Jan %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_POP_Jan_sums <- Richmond_POP_Jan[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+Zip_30815_POP_Jan <-filter(Richmond_POP_Jan_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+
+```
+
+### Change the column names for 30815 Jan
+
+```{r}
+colnames(Zip_30815_POP_Jan) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Jan_26", "total_male_POP_Jan_26", "total_female_POP_Jan_26", "total_masian_POP_Jan_26", "total_mblack_POP_Jan_26", "total_morace_POP_Jan_26", "total_mwhite_POP_Jan_26", "total_fasian_POP_Jan_26", "total_fblack_POP_Jan_26", "total_forace_POP_Jan_26", "total_fwhite_POP_Jan_26", "total_asian_POP_Jan_26","total_black_POP_Jan_26", "total_orace_POP_Jan_26", "total_white_POP_Jan_26", "total_05_09_POP_Jan_26", "total_10_17_POP_Jan_26", "total_18_44_POP_Jan_26", "total_45_64_POP_Jan_26", "total_65Plus_POP_Jan_26")
+```
+
+### Deleting excess environment objects and saving data as processed_data_POP_Jan
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_POP_Jan,Jan_26, codebook, Richmond_POP_Jan_sums)
+
+#saving clean data for 30815 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Jan.rds")
+
+saveRDS(Zip_30815_POP_Jan, file = save_data_location4)
+```
+
+## Processing 6: : Getting Population Totals by Target Zips 30815 for the Month of February
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+## Summary of data
+
+```{r}
+glimpse(Feb_26)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from Richmond County which contains the target zip codes of 30815, 31705 and 31707
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Richmond_POP_Feb <- filter(Feb_26, COUNTY_NAME == "Richmond County")
+
+#replacing negative values with 0
+Richmond_POP_Feb [Richmond_POP_Feb < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_POP_Feb = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Richmond_POP_Feb <- Richmond_POP_Feb  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Richmond_POP_Feb<- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asain residents.
+
+```{r}
+#POP total Asian Residents
+Richmond_POP_Feb$total_asian_POP <- Richmond_POP_Feb$total_masian_POP + Richmond_POP_Feb$total_fasian_POP
+
+#POP total for African American Residents 
+Richmond_POP_Feb$total_black_POP <- Richmond_POP_Feb$total_mblack_POP + Richmond_POP_Feb$total_fblack_POP
+
+#POP total for Other Race Residents
+Richmond_POP_Feb$total_orace_POP <- Richmond_POP_Feb$total_morace_POP + Richmond_POP_Feb$total_forace_POP
+
+#POP total for White Residents
+Richmond_POP_Feb$total_white_POP <- Richmond_POP_Feb$total_mwhite_POP + Richmond_POP_Feb$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Richmond_POP_Feb <- Richmond_POP_Feb %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Richmond_POP_Feb_sums <- Richmond_POP_Feb[,c(1:4,185:204)]  
+```
+
+### Filtering the data by census tract within the 30815 zip code
+
+```{r}
+#filtering data by zip: 30815  
+Zip_30815_POP_Feb <-filter(Richmond_POP_Feb_sums, FIPS=="13245010904" | FIPS=="13245010907" | FIPS=="13245010908" | FIPS=="13245010713" | FIPS=="13245010714" | FIPS=="13245010711" | FIPS=="13245010708" | FIPS=="13245010707" | FIPS=="13245010709" | FIPS=="13245010710" | FIPS=="13033950101" | FIPS=="13245010905" | FIPS=="13033950200")
+
+```
+
+### Change the column names for 30815 Feb
+
+```{r}
+colnames(Zip_30815_POP_Feb) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26")
+```
+
+### Deleting excess environment objects and saving data as processed_data_POP_Feb
+
+```{r}
+#Only need to save processed data for the three zip codes
+rm(Richmond_POP_Feb,Feb_26, codebook, Richmond_POP_Feb_sums)
+
+#saving clean data for 30815 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data","Zip_30815", "processed_data_POP_30815_Feb.rds")
+
+saveRDS(Zip_30815_POP_Feb, file = save_data_location4)
+```
+
+# Processing Section for Control Zip Code: 30907
+
+## Processing 1: Getting Population Totals by Control Zip Code 30907 in the Month of September
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Sept_17.xlsx")
+Sept_17 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Zip_30907_POP_Sept <-filter(Sept_17, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_POP_Sept)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_POP_Sept [Zip_30907_POP_Sept < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_POP_Sept = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Zip_30907_POP_Sept<- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.
+
+```{r}
+#POP total Asian Residents
+Zip_30907_POP_Sept$total_asian_POP <- Zip_30907_POP_Sept$total_masian_POP + Zip_30907_POP_Sept$total_fasian_POP
+
+#POP total for African American Residents 
+Zip_30907_POP_Sept$total_black_POP <- Zip_30907_POP_Sept$total_mblack_POP + Zip_30907_POP_Sept$total_fblack_POP
+
+#POP total for Other Race Residents
+Zip_30907_POP_Sept$total_orace_POP <- Zip_30907_POP_Sept$total_morace_POP + Zip_30907_POP_Sept$total_forace_POP
+
+#POP total for White Residents
+Zip_30907_POP_Sept$total_white_POP <- Zip_30907_POP_Sept$total_mwhite_POP + Zip_30907_POP_Sept$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_POP_Sept <- Zip_30907_POP_Sept %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_POP_Sept_sums <- Zip_30907_POP_Sept[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Sept
+```{r}
+colnames(Zip_30907_POP_Sept_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Sept_17", "total_male_POP_Sept_17", "total_female_POP_Sept_17", "total_masian_POP_Sept_17", "total_mblack_POP_Sept_17", "total_morace_POP_Sept_17", "total_mwhite_POP_Sept_17", "total_fasian_POP_Sept_17", "total_fblack_POP_Sept_17", "total_forace_POP_Sept_17", "total_fwhite_POP_Sept_17", "total_asian_POP_Sept_17","total_black_POP_Sept_17", "total_orace_POP_Sept_17", "total_white_POP_Sept_17", "total_05_09_POP_Sept_17", "total_10_17_POP_Sept_17", "total_18_44_POP_Sept_17", "total_45_64_POP_Sept_17", "total_65Plus_POP_Sept_17")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Sept.rds")
+
+saveRDS(Zip_30907_POP_Sept_sums, save_data_location4)
+```
+
+## Processing 2: Getting Population Totals by Control Zip Code 30907 in the Month of October
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Oct_28.xlsx")
+Oct_28 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Zip_30907_POP_Oct <-filter(Oct_28, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_POP_Oct)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_POP_Oct [Zip_30907_POP_Oct < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_POP_Oct = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Zip_30907_POP_Oct<- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.
+
+```{r}
+#POP total Asian Residents
+Zip_30907_POP_Oct$total_asian_POP <- Zip_30907_POP_Oct$total_masian_POP + Zip_30907_POP_Oct$total_fasian_POP
+
+#POP total for African American Residents 
+Zip_30907_POP_Oct$total_black_POP <- Zip_30907_POP_Oct$total_mblack_POP + Zip_30907_POP_Oct$total_fblack_POP
+
+#POP total for Other Race Residents
+Zip_30907_POP_Oct$total_orace_POP <- Zip_30907_POP_Oct$total_morace_POP + Zip_30907_POP_Oct$total_forace_POP
+
+#POP total for White Residents
+Zip_30907_POP_Oct$total_white_POP <- Zip_30907_POP_Oct$total_mwhite_POP + Zip_30907_POP_Oct$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_POP_Oct <- Zip_30907_POP_Oct %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_POP_Oct_sums <- Zip_30907_POP_Oct[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Oct
+```{r}
+colnames(Zip_30907_POP_Oct_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Oct_28", "total_male_POP_Oct_28", "total_female_POP_Oct_28", "total_masian_POP_Oct_28", "total_mblack_POP_Oct_28", "total_morace_POP_Oct_28", "total_mwhite_POP_Oct_28", "total_fasian_POP_Oct_28", "total_fblack_POP_Oct_28", "total_forace_POP_Oct_28", "total_fwhite_POP_Oct_28", "total_asian_POP_Oct_28","total_black_POP_Oct_28", "total_orace_POP_Oct_28", "total_white_POP_Oct_28", "total_05_09_POP_Oct_28", "total_10_17_POP_Oct_28", "total_18_44_POP_Oct_28", "total_45_64_POP_Oct_28", "total_65Plus_POP_Oct_28")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Oct.rds")
+
+saveRDS(Zip_30907_POP_Oct_sums, save_data_location4)
+```
+
+## Processing 3: : Getting Population Totals by Control Zip Code 30907 in the Month of November
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Nov_24.xlsx")
+Nov_24 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Zip_30907_POP_Nov <-filter(Nov_24, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_POP_Nov)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_POP_Nov [Zip_30907_POP_Nov < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_POP_Nov = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Zip_30907_POP_Nov<- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.
+
+```{r}
+#POP total Asian Residents
+Zip_30907_POP_Nov$total_asian_POP <- Zip_30907_POP_Nov$total_masian_POP + Zip_30907_POP_Nov$total_fasian_POP
+
+#POP total for African American Residents 
+Zip_30907_POP_Nov$total_black_POP <- Zip_30907_POP_Nov$total_mblack_POP + Zip_30907_POP_Nov$total_fblack_POP
+
+#POP total for Other Race Residents
+Zip_30907_POP_Nov$total_orace_POP <- Zip_30907_POP_Nov$total_morace_POP + Zip_30907_POP_Nov$total_forace_POP
+
+#POP total for White Residents
+Zip_30907_POP_Nov$total_white_POP <- Zip_30907_POP_Nov$total_mwhite_POP + Zip_30907_POP_Nov$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_POP_Nov <- Zip_30907_POP_Nov %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_POP_Nov_sums <- Zip_30907_POP_Nov[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Nov
+```{r}
+colnames(Zip_30907_POP_Nov_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Nov_24", "total_male_POP_Nov_24", "total_female_POP_Nov_24", "total_masian_POP_Nov_24", "total_mblack_POP_Nov_24", "total_morace_POP_Nov_24", "total_mwhite_POP_Nov_24", "total_fasian_POP_Nov_24", "total_fblack_POP_Nov_24", "total_forace_POP_Nov_24", "total_fwhite_POP_Nov_24", "total_asian_POP_Nov_24","total_black_POP_Nov_24", "total_orace_POP_Nov_24", "total_white_POP_Nov_24", "total_05_09_POP_Nov_24", "total_10_17_POP_Nov_24", "total_18_44_POP_Nov_24", "total_45_64_POP_Nov_24", "total_65Plus_POP_Nov_24")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Nov.rds")
+
+saveRDS(Zip_30907_POP_Nov_sums, save_data_location4)
+```
+
+## Processing 4: : Getting Population Totals by Control Zip Code 30907 in the Month of December
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Dec_29.xlsx")
+Dec_29 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Zip_30907_POP_Dec <-filter(Dec_29, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_POP_Dec)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_POP_Dec [Zip_30907_POP_Dec < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_POP_Dec = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Zip_30907_POP_Dec<- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.
+
+```{r}
+#POP total Asian Residents
+Zip_30907_POP_Dec$total_asian_POP <- Zip_30907_POP_Dec$total_masian_POP + Zip_30907_POP_Dec$total_fasian_POP
+
+#POP total for African American Residents 
+Zip_30907_POP_Dec$total_black_POP <- Zip_30907_POP_Dec$total_mblack_POP + Zip_30907_POP_Dec$total_fblack_POP
+
+#POP total for Other Race Residents
+Zip_30907_POP_Dec$total_orace_POP <- Zip_30907_POP_Dec$total_morace_POP + Zip_30907_POP_Dec$total_forace_POP
+
+#POP total for White Residents
+Zip_30907_POP_Dec$total_white_POP <- Zip_30907_POP_Dec$total_mwhite_POP + Zip_30907_POP_Dec$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_POP_Dec <- Zip_30907_POP_Dec %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_POP_Dec_sums <- Zip_30907_POP_Dec[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Dec
+```{r}
+colnames(Zip_30907_POP_Dec_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Dec_29", "total_male_POP_Dec_29", "total_female_POP_Dec_29", "total_masian_POP_Dec_29", "total_mblack_POP_Dec_29", "total_morace_POP_Dec_29", "total_mwhite_POP_Dec_29", "total_fasian_POP_Dec_29", "total_fblack_POP_Dec_29", "total_forace_POP_Dec_29", "total_fwhite_POP_Dec_29", "total_asian_POP_Dec_29","total_black_POP_Dec_29", "total_orace_POP_Dec_29", "total_white_POP_Dec_29", "total_05_09_POP_Dec_29", "total_10_17_POP_Dec_29", "total_18_44_POP_Dec_29", "total_45_64_POP_Dec_29", "total_65Plus_POP_Dec_29")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Dec.rds")
+
+saveRDS(Zip_30907_POP_Dec_sums, save_data_location4)
+```
+
+## Processing 5: : Getting Population Totals by Control Zip Code 30907 in the Month of January
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Jan_26.xlsx")
+Jan_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Zip_30907_POP_Jan <-filter(Jan_26, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_POP_Jan)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_POP_Jan [Zip_30907_POP_Jan < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_POP_Jan = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Zip_30907_POP_Jan<- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.
+
+```{r}
+#POP total Asian Residents
+Zip_30907_POP_Jan$total_asian_POP <- Zip_30907_POP_Jan$total_masian_POP + Zip_30907_POP_Jan$total_fasian_POP
+
+#POP total for African American Residents 
+Zip_30907_POP_Jan$total_black_POP <- Zip_30907_POP_Jan$total_mblack_POP + Zip_30907_POP_Jan$total_fblack_POP
+
+#POP total for Other Race Residents
+Zip_30907_POP_Jan$total_orace_POP <- Zip_30907_POP_Jan$total_morace_POP + Zip_30907_POP_Jan$total_forace_POP
+
+#POP total for White Residents
+Zip_30907_POP_Jan$total_white_POP <- Zip_30907_POP_Jan$total_mwhite_POP + Zip_30907_POP_Jan$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_POP_Jan <- Zip_30907_POP_Jan %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_POP_Jan_sums <- Zip_30907_POP_Jan[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Jan
+```{r}
+colnames(Zip_30907_POP_Jan_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Jan_26", "total_male_POP_Jan_26", "total_female_POP_Jan_26", "total_masian_POP_Jan_26", "total_mblack_POP_Jan_26", "total_morace_POP_Jan_26", "total_mwhite_POP_Jan_26", "total_fasian_POP_Jan_26", "total_fblack_POP_Jan_26", "total_forace_POP_Jan_26", "total_fwhite_POP_Jan_26", "total_asian_POP_Jan_26","total_black_POP_Jan_26", "total_orace_POP_Jan_26", "total_white_POP_Jan_26", "total_05_09_POP_Jan_26", "total_10_17_POP_Jan_26", "total_18_44_POP_Jan_26", "total_45_64_POP_Jan_26", "total_65Plus_POP_Jan_26")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Jan.rds")
+
+saveRDS(Zip_30907_POP_Jan_sums, save_data_location4)
+```
+
+## Processing 6: : Getting Population Totals by Control Zip Code 30907 in the Month of February
+
+## Raw Data & Codebook loading
+
+```{r}
+# Loading the rawdata containing GA residents with at least one dose of the Covid-19 vaccine
+rawdata_location = here("data","raw_data","Feb_26.xlsx")
+Feb_26 <- read_excel(rawdata_location)
+
+#loading data code book for reference 
+codebook_location = here("data","raw_data","Codebook.xlsx")
+codebook = read_excel(codebook_location)
+```
+
+### Separating the data by zip code and removing negative values
+
+```{r}
+#I am only interested in Covid-19 Population from the 30907 zip code within which contains the following cenus tracts below 
+#by using filter() I am able to create a new dataset with Populations from Richmond County 
+Zip_30907_POP_Feb <-filter(Feb_26, FIPS=="13073030107" | FIPS=="13073030108" | FIPS=="13245010104" | FIPS=="13245010105" | FIPS=="13245010205" | FIPS=="13073030202" | FIPS=="13073030313" | FIPS=="13073030312" | FIPS=="13073030306" | FIPS=="13073030201" | FIPS=="13073030603" | FIPS=="13073030106" | FIPS=="13245010203"| FIPS=="13245010201")
+
+#taking a look at the data
+head(Zip_30907_POP_Feb)
+
+#notice that although the control zip 30907 is in Richmond county, there are some observartions from Columbia county.
+This is okay because census tracts are not completely contained by one county or zip code.
+
+#replacing negative values with 0
+Zip_30907_POP_Feb [Zip_30907_POP_Feb < 0] <- 0  
+```
+
+### Adding new columns: total POP sum, total POP of all males, total POP of all females
+
+```{r}
+#Adding a column of the total POP for each census tract (for each observation)  
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_POP_Feb = sum(across(ends_with("POP")), na.rm = T))
+
+#total POP column for all males
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb  %>% rowwise() %>%   
+  mutate(total_male_POP = sum(across(starts_with("M") & ends_with("POP")), na.rm = T))
+
+#total POP column for all females
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_female_POP = sum(across(starts_with("F") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Male Residents by race
+
+```{r}
+#total POP column for Asian male residents 
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_masian_POP = sum(across(starts_with("MAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American male residents
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_mblack_POP = sum(across(starts_with("MBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race male residents
+Zip_30907_POP_Feb<- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_morace_POP = sum(across(starts_with("MOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White male residents 
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_mwhite_POP = sum(across(starts_with("MWhite") & ends_with("POP")), na.rm = T))
+
+```
+
+### Adding new columns for Female Residents by race
+
+```{r}
+#total POP column for Asian female residents    
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_fasian_POP = sum(across(starts_with("FAsianNHPI") & ends_with("POP")), na.rm = T))
+
+#total POP column for African American female residents
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_fblack_POP = sum(across(starts_with("FBlack") & ends_with("POP")), na.rm = T))
+
+#total POP column for other race female residents
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_forace_POP = sum(across(starts_with("FOtherRace") & ends_with("POP")), na.rm = T))
+
+#total POP column for White female residents 
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>%   
+  mutate(total_fwhite_POP = sum(across(starts_with("FWhite") & ends_with("POP")), na.rm = T))  
+```
+
+### Adding new columns for POP totals by race
+
+Note that these columns are different than the previous columns because here I am summing the total amount of resident Populations by adding all males and females of a specific race to get the the overall amount of resident Populations by race.
+
+For example(see below): I get the total Asian resident Populations by summing male Asian residents + female Asian residents.
+
+```{r}
+#POP total Asian Residents
+Zip_30907_POP_Feb$total_asian_POP <- Zip_30907_POP_Feb$total_masian_POP + Zip_30907_POP_Feb$total_fasian_POP
+
+#POP total for African American Residents 
+Zip_30907_POP_Feb$total_black_POP <- Zip_30907_POP_Feb$total_mblack_POP + Zip_30907_POP_Feb$total_fblack_POP
+
+#POP total for Other Race Residents
+Zip_30907_POP_Feb$total_orace_POP <- Zip_30907_POP_Feb$total_morace_POP + Zip_30907_POP_Feb$total_forace_POP
+
+#POP total for White Residents
+Zip_30907_POP_Feb$total_white_POP <- Zip_30907_POP_Feb$total_mwhite_POP + Zip_30907_POP_Feb$total_fwhite_POP  
+```
+
+### Adding new columns for POP totals by age ranges: 05-09, 10-17, 18-44, 45-64, 65+
+
+```{r}
+#summing all residents in the age range of 05-09
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_05_09_POP = sum(FAsianNHPI_05_09_POP, MAsianNHPI_05_09_POP, UAsianNHPI_05_09_POP,FBlack_05_09_POP, MBlack_05_09_POP, UBlack_05_09_POP,FNone_05_09_POP, MNone_05_09_POP, UNone_05_09_POP,FOtherRace_05_09_POP, MOtherRace_05_09_POP, UOtherRace_05_09_POP,FWhite_05_09_POP, MWhite_05_09_POP, UWhite_05_09_POP, na.rm = T))
+
+#summing all residents in the age range of 10-17
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_10_17_POP = sum(FAsianNHPI_10_17_POP, MAsianNHPI_10_17_POP, UAsianNHPI_10_17_POP,FBlack_10_17_POP, MBlack_10_17_POP, UBlack_10_17_POP,FNone_10_17_POP, MNone_10_17_POP, UNone_10_17_POP,FOtherRace_10_17_POP, MOtherRace_10_17_POP, UOtherRace_10_17_POP,FWhite_10_17_POP, MWhite_10_17_POP, UWhite_10_17_POP, na.rm = T))
+
+#summing all residents in the age range of 18-44
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_18_44_POP = sum(FAsianNHPI_18_44_POP, MAsianNHPI_18_44_POP, UAsianNHPI_18_44_POP,FBlack_18_44_POP, MBlack_18_44_POP, UBlack_18_44_POP,FNone_18_44_POP, MNone_18_44_POP, UNone_18_44_POP,FOtherRace_18_44_POP, MOtherRace_18_44_POP, UOtherRace_18_44_POP,FWhite_18_44_POP, MWhite_18_44_POP, UWhite_18_44_POP, na.rm = T))
+
+#summing all residents in the age range of 45-64
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_45_64_POP = sum(FAsianNHPI_45_64_POP, MAsianNHPI_45_64_POP, UAsianNHPI_45_64_POP,FBlack_45_64_POP, MBlack_45_64_POP, UBlack_45_64_POP,FNone_45_64_POP, MNone_45_64_POP, UNone_45_64_POP,FOtherRace_45_64_POP, MOtherRace_45_64_POP, UOtherRace_45_64_POP,FWhite_45_64_POP, MWhite_45_64_POP, UWhite_45_64_POP, na.rm = T))
+
+#summing all residents in the age range of 65+ 
+Zip_30907_POP_Feb <- Zip_30907_POP_Feb %>% rowwise() %>% mutate(total_65Plus_POP = sum(FAsianNHPI_65Plus_POP, MAsianNHPI_65Plus_POP, UAsianNHPI_65Plus_POP,FBlack_65Plus_POP, MBlack_65Plus_POP, UBlack_65Plus_POP,FNone_65Plus_POP, MNone_65Plus_POP, UNone_65Plus_POP,FOtherRace_65Plus_POP, MOtherRace_65Plus_POP, UOtherRace_65Plus_POP,FWhite_65Plus_POP, MWhite_65Plus_POP, UWhite_65Plus_POP, na.rm = T))
+
+```
+
+### Creating a data set with only the sums columns I just created via the previous code
+
+```{r}
+#Extract necessary variables only  
+Zip_30907_POP_Feb_sums <- Zip_30907_POP_Feb[,c(1:4,185:204)]  
+```
+
+### Change the column names for 30907 Feb
+```{r}
+colnames(Zip_30907_POP_Feb_sums) <- c("FIPS", "GEONAME", "COUNTY_ID", "COUNTY_NAME", "total_POP_Feb_26", "total_male_POP_Feb_26", "total_female_POP_Feb_26", "total_masian_POP_Feb_26", "total_mblack_POP_Feb_26", "total_morace_POP_Feb_26", "total_mwhite_POP_Feb_26", "total_fasian_POP_Feb_26", "total_fblack_POP_Feb_26", "total_forace_POP_Feb_26", "total_fwhite_POP_Feb_26", "total_asian_POP_Feb_26","total_black_POP_Feb_26", "total_orace_POP_Feb_26", "total_white_POP_Feb_26", "total_05_09_POP_Feb_26", "total_10_17_POP_Feb_26", "total_18_44_POP_Feb_26", "total_45_64_POP_Feb_26", "total_65Plus_POP_Feb_26")
+```
+
+### Saving data
+```{r}
+#saving clean data for 30907 to processed_data_POP folder
+save_data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Feb.rds")
+
+saveRDS(Zip_30907_POP_Feb_sums, save_data_location4)
+```

### 2023-04-19 code/analysis_code/exploratoryanalysis.qmd
Population sums by zip
Vax Rates

@@
 
-### Loading the processed data in order to begin analysis for the 31701 zip code
-
+### Combining Vaccination data from the 31701 zip from Sept-Feb
 ```{r}
@@
 
+### Combining Population data from the 31701 zip from Sept-Feb
+```{r}
+# Loading September 31701
+data_location1 = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Sept.rds")
+processed_POP_data_31701_Sept = readRDS(data_location1)
+
+# Loading October 31701 and joining september and october 
+data_location2 = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Oct.rds")
+processed_POP_data_31701_Oct = readRDS(data_location2)
+processed_POP_combined <- left_join(processed_POP_data_31701_Sept, processed_POP_data_31701_Oct, by = c("FIPS" = "FIPS"))
+
+# Loading November 31701
+data_location3 = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Nov.rds")
+processed_POP_data_31701_Nov = readRDS(data_location3)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31701_Nov, by = c("FIPS" = "FIPS"))
+
+# Loading December 31701
+data_location4 = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Dec.rds")
+processed_POP_data_31701_Dec = readRDS(data_location4)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31701_Dec, by = c("FIPS" = "FIPS"))
+
+# Loading January 31701
+data_location5 = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Jan.rds")
+processed_POP_data_31701_Jan = readRDS(data_location5)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31701_Jan, by = c("FIPS" = "FIPS"))
+
+# Loading February 31701
+data_location6 = here("data", "processed_data", "Zip_31701", "processed_POP_data_31701_Feb.rds")
+processed_POP_data_31701_Feb = readRDS(data_location6)
+processed_POP_combined_31701 <- left_join(processed_POP_combined, processed_POP_data_31701_Feb, by = c("FIPS" = "FIPS"))
+```
+
 ### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender
@@
 
+```{r}
+save_data_location = here("data", "processed_data","Zip_31701", "processed_POP_combined_31701.rds")
+saveRDS(processed_POP_combined_31701, file = save_data_location)
+```
+
 ### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31701 zip code
@@
 
-### Summing each Month variable by census tract
-
+```{r}
+processed_combined_31701_total_black_POP = processed_POP_combined_31701 %>% select(starts_with("total_black_POP"))
+```
+
+### Summing vaccination numbers
 ```{r}
@@
 
-#Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February #Total contains the vaccination totals for each month
+### Summing population numbers
+```{r}
+#Summing each column for each month 
+sum(processed_combined_31701_total_black_POP$total_black_POP_Sept_17) 
+sum(processed_combined_31701_total_black_POP$total_black_POP_Oct_28) 
+sum(processed_combined_31701_total_black_POP$total_black_POP_Nov_24) 
+sum(processed_combined_31701_total_black_POP$total_black_POP_Dec_29) 
+sum(processed_combined_31701_total_black_POP$total_black_POP_Jan_26) 
+sum(processed_combined_31701_total_black_POP$total_black_POP_Feb_26) 
+```
+
+#Separating sums by month into a data fram consisting of two variables: Date and VAX_RATE #Date contains months September-February #Total contains the vaccination totals for each month
 
@@
 Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Total = c(7797, 7780, 7755, 7773, 7792, 7817)) 
+                               VAX_RATE = c(7797/12768, 7780/12768, 7755/12897, 7773/12897, 7792/12897, 7817/12897)) 
 ```
@@
 ```{r}
-plot_31701_total_black_vax = Sum_by_month_31701 %>% 
-  ggplot(aes(Date, Total, group=1)) + 
+plot_31701_total_black_vax_rate = Sum_by_month_31701 %>% 
+  ggplot(aes(Date, VAX_RATE, group=1)) + 
   geom_point() + geom_line() + 
@@
 #Adding a customization title for the plot 
-plot_31701_total_black_vax = plot_31701_total_black_vax + 
-  ggtitle("Vaccination Totals for Black population in the 31701 Zipcode") + 
+plot_31701_total_black_vax_rate = plot_31701_total_black_vax_rate + 
+  ggtitle("Vaccination Rates for Black Population in the 31701 Zipcode") + 
   theme(plot.title = 
           element_text(color="red", size=10, face="bold.italic"))
-plot_31701_total_black_vax
+plot_31701_total_black_vax_rate
 ```
@@
 #save figure
-ggsave(filename = here("results","31701_total_black_vax.png"), plot=plot_31701_total_black_vax)
+ggsave(filename = here("results","plot_31701_total_black_vax_rate.png"), plot=plot_31701_total_black_vax_rate)
 ```
@@
 ```{r}
-plotly_31701_total_black_vax = ggplotly(plot_31701_total_black_vax)
-plotly_31701_total_black_vax
+plotly_31701_total_black_vax_rate = ggplotly(plot_31701_total_black_vax_rate)
+plotly_31701_total_black_vax_rate
 ```
@@
 
-### Loading the processed data in order to begin analysis for the 31705 zip code
-
-```{r}
-#Loading September 31705
-data_location7 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Sept.rds")
-processed_data_31705_Sept = readRDS(data_location7)
+### Combining Vaccination data from the 31705 zip from Sept-Feb
+```{r}
+# Loading September 31705
+data_location1 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Sept.rds")
+processed_data_31705_Sept = readRDS(data_location1)
 
 # Loading October 31705 and joining september and october 
-data_location8 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Oct.rds")
-processed_data_31705_Oct = readRDS(data_location8)
+data_location2 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Oct.rds")
+processed_data_31705_Oct = readRDS(data_location2)
 processed_combined <- left_join(processed_data_31705_Sept, processed_data_31705_Oct, by = c("FIPS" = "FIPS"))
@@
 # Loading November 31705
-data_location9 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Nov.rds")
-processed_data_31705_Nov = readRDS(data_location9)
+data_location3 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Nov.rds")
+processed_data_31705_Nov = readRDS(data_location3)
 processed_combined <- left_join(processed_combined, processed_data_31705_Nov, by = c("FIPS" = "FIPS"))
@@
 # Loading December 31705
-data_location10= here("data", "processed_data", "Zip_31705", "processed_data_31705_Dec.rds")
-processed_data_31705_Dec = readRDS(data_location10)
+data_location4 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Dec.rds")
+processed_data_31705_Dec = readRDS(data_location4)
 processed_combined <- left_join(processed_combined, processed_data_31705_Dec, by = c("FIPS" = "FIPS"))
@@
 # Loading January 31705
-data_location11 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Jan.rds")
-processed_data_31705_Jan = readRDS(data_location11)
+data_location5 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Jan.rds")
+processed_data_31705_Jan = readRDS(data_location5)
 processed_combined <- left_join(processed_combined, processed_data_31705_Jan, by = c("FIPS" = "FIPS"))
@@
 # Loading February 31705
-data_location12 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Feb.rds")
-processed_data_31705_Feb = readRDS(data_location12)
+data_location6 = here("data", "processed_data", "Zip_31705", "processed_data_31705_Feb.rds")
+processed_data_31705_Feb = readRDS(data_location6)
 processed_combined_31705 <- left_join(processed_combined, processed_data_31705_Feb, by = c("FIPS" = "FIPS"))
-
-```
-
-### Saving the combined data for Black population for future use
+```
+
+### Combining Population data from the 31705 zip from Sept-Feb
+```{r}
+# Loading September 31705
+data_location1 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Sept.rds")
+processed_POP_data_31705_Sept = readRDS(data_location1)
+
+# Loading October 31705 and joining september and october 
+data_location2 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Oct.rds")
+processed_POP_data_31705_Oct = readRDS(data_location2)
+processed_POP_combined <- left_join(processed_POP_data_31705_Sept, processed_POP_data_31705_Oct, by = c("FIPS" = "FIPS"))
+
+# Loading November 31705
+data_location3 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Nov.rds")
+processed_POP_data_31705_Nov = readRDS(data_location3)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31705_Nov, by = c("FIPS" = "FIPS"))
+
+# Loading December 31705
+data_location4 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Dec.rds")
+processed_POP_data_31705_Dec = readRDS(data_location4)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31705_Dec, by = c("FIPS" = "FIPS"))
+
+# Loading January 31705
+data_location5 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Jan.rds")
+processed_POP_data_31705_Jan = readRDS(data_location5)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31705_Jan, by = c("FIPS" = "FIPS"))
+
+# Loading February 31705
+data_location6 = here("data", "processed_data", "Zip_31705", "processed_POP_data_31705_Feb.rds")
+processed_POP_data_31705_Feb = readRDS(data_location6)
+processed_POP_combined_31705 <- left_join(processed_POP_combined, processed_POP_data_31705_Feb, by = c("FIPS" = "FIPS"))
+```
+
+### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender
 
@@
 
+```{r}
+save_data_location = here("data", "processed_data","Zip_31705", "processed_POP_combined_31705.rds")
+saveRDS(processed_POP_combined_31705, file = save_data_location)
+```
+
 ### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31705 zip code
 
+#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+
 ```{r}
@@
 
-### Summing each column for each month
-
-```{r}
-
+```{r}
+processed_combined_31705_total_black_POP = processed_POP_combined_31705 %>% select(starts_with("total_black_POP"))
+```
+
+### Summing vaccination numbers
+```{r}
+#Summing each column for each month 
 sum(processed_combined_31705_total_black$total_black_vax_Sept_17) 
-
 sum(processed_combined_31705_total_black$total_black_vax_Oct_28) 
-
 sum(processed_combined_31705_total_black$total_black_vax_Nov_24) 
-
 sum(processed_combined_31705_total_black$total_black_vax_Dec_29) 
-
 sum(processed_combined_31705_total_black$total_black_vax_Jan_26) 
-
 sum(processed_combined_31705_total_black$total_black_vax_Feb_26) 
@@
 
-### Separating sums by month into a data fram consisting of two variables: Date and Total
-
-### Date contains months September-February
-
-### Total contains the vaccination totals for each month
+### Summing population numbers
+```{r}
+#Summing each column for each month 
+sum(processed_combined_31705_total_black_POP$total_black_POP_Sept_17) 
+sum(processed_combined_31705_total_black_POP$total_black_POP_Oct_28) 
+sum(processed_combined_31705_total_black_POP$total_black_POP_Nov_24) 
+sum(processed_combined_31705_total_black_POP$total_black_POP_Dec_29) 
+sum(processed_combined_31705_total_black_POP$total_black_POP_Jan_26) 
+sum(processed_combined_31705_total_black_POP$total_black_POP_Feb_26) 
+```
+
+#Separating sums by month into a data fram consisting of two variables: Date and VAX_RATE #Date contains months September-February #Total contains the vaccination totals for each month
 
@@
 Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                                Total = c(10383, 10327, 10303, 10301, 10337, 10348)) 
-```
-
-### ggplot creating a line plot for the vaccination totals by month for the 31705 zip code African American population
-
-```{r}
-plot_31705_total_black_vax = Sum_by_month_31705 %>% 
-  ggplot(aes(Date, Total, group=1)) + 
+                               VAX_RATE = c(10383/21603, 10327/21572, 10303/21722, 10301/21887, 10337/21887, 10348/21887)) 
+```
+
+### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 31705 zip code
+
+```{r}
+plot_31705_total_black_vax_rate = Sum_by_month_31705 %>% 
+  ggplot(aes(Date, VAX_RATE, group=1)) + 
   geom_point() + geom_line() + 
@@
 #Adding a customization title for the plot 
-plot_31705_total_black_vax = plot_31705_total_black_vax + 
-  ggtitle("Vaccination Totals for Black population in the 31705 Zipcode") + 
+plot_31705_total_black_vax_rate = plot_31705_total_black_vax_rate + 
+  ggtitle("Vaccination Rates for Black Population in the 31705 Zipcode") + 
   theme(plot.title = 
           element_text(color="red", size=10, face="bold.italic"))
+plot_31705_total_black_vax_rate
 ```
@@
 #save figure
-ggsave(filename = here("results","31705_total_black_vax.png"), plot=plot_31705_total_black_vax)
+ggsave(filename = here("results","plot_31705_total_black_vax_rate.png"), plot=plot_31705_total_black_vax_rate)
 ```
@@
 ```{r}
-plotly_31705_total_black_vax = ggplotly(plot_31705_total_black_vax)
-plotly_31705_total_black_vax
+plotly_31705_total_black_vax_rate = ggplotly(plot_31705_total_black_vax_rate)
+plotly_31705_total_black_vax_rate
 ```
@@
 
-### Loading the processed data in order to begin analysis for the 31707 zip code
-
+### Combining Vaccination data from the 31707 zip from Sept-Feb
 ```{r}
 # Loading September 31707
-data_location13 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Sept.rds")
-processed_data_31707_Sept = readRDS(data_location13)
+data_location1 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Sept.rds")
+processed_data_31707_Sept = readRDS(data_location1)
 
 # Loading October 31707 and joining september and october 
-data_location14 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Oct.rds")
-processed_data_31707_Oct = readRDS(data_location14)
+data_location2 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Oct.rds")
+processed_data_31707_Oct = readRDS(data_location2)
 processed_combined <- left_join(processed_data_31707_Sept, processed_data_31707_Oct, by = c("FIPS" = "FIPS"))
@@
 # Loading November 31707
-data_location15 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Nov.rds")
-processed_data_31707_Nov = readRDS(data_location15)
+data_location3 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Nov.rds")
+processed_data_31707_Nov = readRDS(data_location3)
 processed_combined <- left_join(processed_combined, processed_data_31707_Nov, by = c("FIPS" = "FIPS"))
@@
 # Loading December 31707
-data_location16= here("data", "processed_data", "Zip_31707", "processed_data_31707_Dec.rds")
-processed_data_31707_Dec = readRDS(data_location16)
+data_location4 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Dec.rds")
+processed_data_31707_Dec = readRDS(data_location4)
 processed_combined <- left_join(processed_combined, processed_data_31707_Dec, by = c("FIPS" = "FIPS"))
@@
 # Loading January 31707
-data_location17 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Jan.rds")
-processed_data_31707_Jan = readRDS(data_location17)
+data_location5 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Jan.rds")
+processed_data_31707_Jan = readRDS(data_location5)
 processed_combined <- left_join(processed_combined, processed_data_31707_Jan, by = c("FIPS" = "FIPS"))
@@
 # Loading February 31707
-data_location18 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Feb.rds")
-processed_data_31707_Feb = readRDS(data_location18)
+data_location6 = here("data", "processed_data", "Zip_31707", "processed_data_31707_Feb.rds")
+processed_data_31707_Feb = readRDS(data_location6)
 processed_combined_31707 <- left_join(processed_combined, processed_data_31707_Feb, by = c("FIPS" = "FIPS"))
@@
 
-### Saving the combined data for Black population for future use
+### Combining Population data from the 31707 zip from Sept-Feb
+```{r}
+# Loading September 31707
+data_location1 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Sept.rds")
+processed_POP_data_31707_Sept = readRDS(data_location1)
+
+# Loading October 31707 and joining september and october 
+data_location2 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Oct.rds")
+processed_POP_data_31707_Oct = readRDS(data_location2)
+processed_POP_combined <- left_join(processed_POP_data_31707_Sept, processed_POP_data_31707_Oct, by = c("FIPS" = "FIPS"))
+
+# Loading November 31707
+data_location3 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Nov.rds")
+processed_POP_data_31707_Nov = readRDS(data_location3)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31707_Nov, by = c("FIPS" = "FIPS"))
+
+# Loading December 31707
+data_location4 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Dec.rds")
+processed_POP_data_31707_Dec = readRDS(data_location4)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31707_Dec, by = c("FIPS" = "FIPS"))
+
+# Loading January 31707
+data_location5 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Jan.rds")
+processed_POP_data_31707_Jan = readRDS(data_location5)
+processed_POP_combined <- left_join(processed_POP_combined, processed_POP_data_31707_Jan, by = c("FIPS" = "FIPS"))
+
+# Loading February 31707
+data_location6 = here("data", "processed_data", "Zip_31707", "processed_POP_data_31707_Feb.rds")
+processed_POP_data_31707_Feb = readRDS(data_location6)
+processed_POP_combined_31707 <- left_join(processed_POP_combined, processed_POP_data_31707_Feb, by = c("FIPS" = "FIPS"))
+```
+
+### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender
 
@@
 
+```{r}
+save_data_location = here("data", "processed_data","Zip_31707", "processed_POP_combined_31707.rds")
+saveRDS(processed_POP_combined_31707, file = save_data_location)
+```
+
 ### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 31707 zip code
 
+#### Using the select() function to subset the data frame by columns that start with "total_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+
 ```{r}
 processed_combined_31707_total_black = processed_combined_31707 %>% select(starts_with("total_black_vax"))
-
-```
-
-### Summing each column for each month
-
-```{r}
+```
+
+```{r}
+processed_combined_31707_total_black_POP = processed_POP_combined_31707 %>% select(starts_with("total_black_POP"))
+```
+
+### Summing vaccination numbers
+```{r}
+#Summing each column for each month 
 sum(processed_combined_31707_total_black$total_black_vax_Sept_17) 
@@
 
-### Separating sums by month into a data fram consisting of two variables: Date and Total
-
-### Date contains months September-February
-
-### Total contains the vaccination totals for each month
+### Summing population numbers
+```{r}
+#Summing each column for each month 
+sum(processed_combined_31707_total_black_POP$total_black_POP_Sept_17) 
+sum(processed_combined_31707_total_black_POP$total_black_POP_Oct_28) 
+sum(processed_combined_31707_total_black_POP$total_black_POP_Nov_24) 
+sum(processed_combined_31707_total_black_POP$total_black_POP_Dec_29) 
+sum(processed_combined_31707_total_black_POP$total_black_POP_Jan_26) 
+sum(processed_combined_31707_total_black_POP$total_black_POP_Feb_26) 
+```
+
+#Separating sums by month into a data fram consisting of two variables: Date and VAX_RATE #Date contains months September-February #Total contains the vaccination totals for each month
 
@@
 Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                                Total = c(9035, 8985, 8997, 9005, 9035, 9042))
-
-```
-
-###ggplot: creating a line plot for the vaccination totals by month for the 31707 zip code African American population
-
-```{r}
-plot_31707_total_black_vax = Sum_by_month_31707 %>% 
-  ggplot(aes(Date, Total, group=1)) + 
+                               VAX_RATE = c(9035/18045, 8985/18045, 8997/18045, 9005/18081, 9035/18081, 9042/18165)) 
+```
+
+### gglot2: Creating a line plot for the vaccination totals by month for the African American population in the 31707 zip code
+
+```{r}
+plot_31707_total_black_vax_rate = Sum_by_month_31707 %>% 
+  ggplot(aes(Date, VAX_RATE, group=1)) + 
   geom_point() + geom_line() + 
@@
 #Adding a customization title for the plot 
-plot_31707_total_black_vax = plot_31707_total_black_vax + 
-  ggtitle("Vaccination Totals for Black population in the 31707 Zipcode") + 
+plot_31707_total_black_vax_rate = plot_31707_total_black_vax_rate + 
+  ggtitle("Vaccination Rates for Black Population in the 31707 Zipcode") + 
   theme(plot.title = 
           element_text(color="red", size=10, face="bold.italic"))
+plot_31707_total_black_vax_rate
 ```
@@
 #save figure
-ggsave(filename = here("results","31707_total_black_vax.png"), plot=plot_31707_total_black_vax)
-```
-
-## Comparing plots the 3 target zip codes
-
-### Taking a look at each of the line plots for the different zip codes and comparing the total vaccinations by month in the African American population
-
-```{r}
-plot_31701_total_black_vax
-plot_31705_total_black_vax
-plot_31707_total_black_vax
-```
-
-### Combining all plots into one
-
-```{r}
-## Creating a new dataframe called "Total_black_vax" by combining the vaccination totals from 31701 and 3105 by using leftjoin() by the Date variable 
-Total_black_vax = left_join(Sum_by_month_31701, Sum_by_month_31705, by = c("Date" = "Date"))
-
-```
-
-### Combining the new dataframe "Total_black_vax" with the 31707 zip vaccination totals so all three target zip codes are in the same data frame
-
-```{r}
-Total_black_vax = left_join(Total_black_vax, Sum_by_month_31707, by = c("Date" = "Date"))
-```
-
-### Renaming column totals to match the zip
-
-### Haing to do this because the the total variable names are Total.x, Total.y and Total
-
-### Need to specificy each total variable by zip using the following code
-
-```{r}
-Total_black_vax <- Total_black_vax %>% 
-  rename("31701" = "Total.x") %>% 
-  rename("31705" = "Total.y") %>%
-  rename("31707" = "Total")
-```
-
-### Converting data to long format in order to create a line plot using ggplot
-
-```{r}
-Total_black_vax_long = reshape2::melt(Total_black_vax, id = "Date")
-```
-
-### Changing the variable name "variable" to "Zip_code" and "value" to "Vaccination_total" for clarity
-
-### Doing this because when the data was converted to a long format the column names we changed to "variable" which refers to the zip code and "value" which refers to the amount of vaccinations
-
-### Changing "variable" to "Zip_code" and "value" to "Vaccination_total"
-
-```{r}
-Total_black_vax_long <- Total_black_vax_long %>% 
-  rename("Zip_code" = "variable") %>% 
-  rename("Vaccination_total" = "value")
-```
-
-### ggplot: Making a line graph that compares the vaccination totals for each African American population in each zip code
-
-```{r}
-plot_Total_black_vax = Total_black_vax_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
-                                group = Zip_code, 
-                                color = Zip_code)) + 
-                     geom_point() + geom_line() + 
-                     scale_x_discrete(limits = c("Sep","Oct","Nov","Dec","Jan","Feb"))
-#adding a title to the plot 
-plot_Total_black_vax = plot_Total_black_vax + 
-  ggtitle("Vaccination Totals for Black population in Target Zip Codes") + 
-  theme(plot.title = 
-          element_text(color="red", size=10, face="bold.italic")) + 
-  labs(y = "Vaccination Total")
-plot_Total_black_vax
-```
-
-### Save plot
-```{r}
-#save figure
-ggsave(filename = here("results","total_black_vax.png"), plot=plot_Total_black_vax)
+ggsave(filename = here("results","plot_31707_total_black_vax_rate.png"), plot=plot_31707_total_black_vax_rate)
 ```
@@
 ```{r}
-plotly_Total_black_vax = ggplotly(plot_Total_black_vax)
-plotly_Total_black_vax
-```
-
-# Total Amount of vaccinations by Gender Among Black Population in the Target Zip Codes
+plotly_31707_total_black_vax_rate = ggplotly(plot_31707_total_black_vax_rate)
+plotly_31707_total_black_vax_rate
+```
+
+# Vaccinations by Gender Among Black Population in the Target Zip Codes
 
@@
 
-### Loading the `processed_combined_31701.rds` file from `analysiscode.qmd`
-
+### Loading the `processed_combined_31701.rds` 
 ```{r}
@@
 
+### Loading the `processed_POP_combined_31701.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_31701", "processed_POP_combined_31701.rds")
+processed_POP_combined_31701 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black males and black females within 31701
 
-#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Isolating Gender Variables form Vaccination data and Population data
 
@@
 
-### Summing each Month variable by census tract
+```{r}
+processed_POP_combined_31701_Gender_black = processed_POP_combined_31701 %>% select(matches("^total_(mblack|fblack)_POP"))
+```
+
+### Summing each Month for vaccination data 
 ```{r}
@@
 
+### Summing each Month for population data
+```{r}
+#Summing for black males by month 
+sum(processed_POP_combined_31701_Gender_black$total_mblack_POP_Sept_17) 
+sum(processed_POP_combined_31701_Gender_black$total_mblack_POP_Oct_28) 
+sum(processed_POP_combined_31701_Gender_black$total_mblack_POP_Nov_24) 
+sum(processed_POP_combined_31701_Gender_black$total_mblack_POP_Dec_29) 
+sum(processed_POP_combined_31701_Gender_black$total_mblack_POP_Jan_26) 
+sum(processed_POP_combined_31701_Gender_black$total_mblack_POP_Feb_26)
+
+#summing for black females by month 
+sum(processed_POP_combined_31701_Gender_black$total_fblack_POP_Sept_17) 
+sum(processed_POP_combined_31701_Gender_black$total_fblack_POP_Oct_28) 
+sum(processed_POP_combined_31701_Gender_black$total_fblack_POP_Nov_24) 
+sum(processed_POP_combined_31701_Gender_black$total_fblack_POP_Dec_29) 
+sum(processed_POP_combined_31701_Gender_black$total_fblack_POP_Jan_26) 
+sum(processed_POP_combined_31701_Gender_black$total_fblack_POP_Feb_26) 
+```
+
 #Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
@@
 Gender_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
-                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420))
+                          Male_Total = c(3371/5981, 3358/5981, 3368/5981, 3379/5981, 3387/5981, 3397/5981),
+                          Female_Total = c(4426/6787, 4422/6787, 4387/6916, 4394/6916, 4405/6916, 4420/6916))
 ```
@@
   rename("Gender" = "variable") %>% 
-  rename("Vaccination_total" = "value")
+  rename("Vaccination_rate" = "value")
 ```
@@
 plot_Gender_black_vax = Gender_Sum_by_month_31701_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Gender, 
@@
 plot_Gender_black_vax = plot_Gender_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Gender in 31701") + 
+  ggtitle("Vaccination Rates for Black population by Gender in 31701") + 
   theme(plot.title = 
@@
 
-### Loading the `processed_combined_31705.rds` file from `analysiscode.qmd`
-
+### Loading the `processed_combined_31705.rds` 
 ```{r}
@@
 
+### Loading the `processed_POP_combined_31705.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_31705", "processed_POP_combined_31705.rds")
+processed_POP_combined_31705 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black males and black females within 31705
 
-#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Isolating Gender Variables form Vaccination data and Population data
 
@@
 
-### Summing each Month variable by census tract
+```{r}
+processed_POP_combined_31705_Gender_black = processed_POP_combined_31705 %>% select(matches("^total_(mblack|fblack)_POP"))
+```
+
+### Summing each Month for vaccination data 
 ```{r}
@@
 
+### Summing each Month for population data
+```{r}
+#Summing for black males by month 
+sum(processed_POP_combined_31705_Gender_black$total_mblack_POP_Sept_17) 
+sum(processed_POP_combined_31705_Gender_black$total_mblack_POP_Oct_28) 
+sum(processed_POP_combined_31705_Gender_black$total_mblack_POP_Nov_24) 
+sum(processed_POP_combined_31705_Gender_black$total_mblack_POP_Dec_29) 
+sum(processed_POP_combined_31705_Gender_black$total_mblack_POP_Jan_26) 
+sum(processed_POP_combined_31705_Gender_black$total_mblack_POP_Feb_26)
+
+#summing for black females by month 
+sum(processed_POP_combined_31705_Gender_black$total_fblack_POP_Sept_17) 
+sum(processed_POP_combined_31705_Gender_black$total_fblack_POP_Oct_28) 
+sum(processed_POP_combined_31705_Gender_black$total_fblack_POP_Nov_24) 
+sum(processed_POP_combined_31705_Gender_black$total_fblack_POP_Dec_29) 
+sum(processed_POP_combined_31705_Gender_black$total_fblack_POP_Jan_26) 
+sum(processed_POP_combined_31705_Gender_black$total_fblack_POP_Feb_26) 
+```
+
 #Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
@@
 Gender_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
-                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949))
+                          Male_Total = c(4394/9885, 4397/9854, 4399/10004, 4391/10004, 4403/10004, 4399/10004),
+                          Female_Total = c(5989/11718, 5930/11718, 5904/11718, 5910/11883, 5934/11883, 5949/11883))
 ```
@@
 ```{r}
-plot_Gender_black_vax_31705 = Gender_Sum_by_month_31705_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+plot_Gender_black_vax = Gender_Sum_by_month_31705_long %>% 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Gender, 
@@
 plot_Gender_black_vax = plot_Gender_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Gender in 31705") + 
+  ggtitle("Vaccination Rates for Black population by Gender in 31705") + 
   theme(plot.title = 
@@
   labs(y = "Vaccination Total")
-plot_Gender_black_vax_31705
+plot_Gender_black_vax
 ```
@@
 #save figure
-ggsave(filename = here("results","31705_Gender_black_vax.png"), plot=plot_Gender_black_vax_31705)
+ggsave(filename = here("results","31705_Gender_black_vax.png"), plot=plot_Gender_black_vax)
 ```
@@
 ```{r}
-plotly_plot_Gender_black_vax_31705 = ggplotly(plot_Gender_black_vax_31705)
-plotly_plot_Gender_black_vax_31705
+plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
+plotly_plot_Gender_black_vax
 ```
@@
 
-### Loading the `processed_combined_31707.rds` file from `analysiscode.qmd`
-
+### Loading the `processed_combined_31707.rds` 
 ```{r}
@@
 
+### Loading the `processed_POP_combined_31707.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_31707", "processed_POP_combined_31707.rds")
+processed_POP_combined_31707 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black males and black females within 31707
 
-#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Isolating Gender Variables form Vaccination data and Population data
 
@@
 
-### Summing each Month variable by census tract
+```{r}
+processed_POP_combined_31707_Gender_black = processed_POP_combined_31707 %>% select(matches("^total_(mblack|fblack)_POP"))
+```
+
+### Summing each Month for vaccination data 
 ```{r}
@@
 
+### Summing each Month for population data
+```{r}
+#Summing for black males by month 
+sum(processed_POP_combined_31707_Gender_black$total_mblack_POP_Sept_17) 
+sum(processed_POP_combined_31707_Gender_black$total_mblack_POP_Oct_28) 
+sum(processed_POP_combined_31707_Gender_black$total_mblack_POP_Nov_24) 
+sum(processed_POP_combined_31707_Gender_black$total_mblack_POP_Dec_29) 
+sum(processed_POP_combined_31707_Gender_black$total_mblack_POP_Jan_26) 
+sum(processed_POP_combined_31707_Gender_black$total_mblack_POP_Feb_26)
+
+#summing for black females by month 
+sum(processed_POP_combined_31707_Gender_black$total_fblack_POP_Sept_17) 
+sum(processed_POP_combined_31707_Gender_black$total_fblack_POP_Oct_28) 
+sum(processed_POP_combined_31707_Gender_black$total_fblack_POP_Nov_24) 
+sum(processed_POP_combined_31707_Gender_black$total_fblack_POP_Dec_29) 
+sum(processed_POP_combined_31707_Gender_black$total_fblack_POP_Jan_26) 
+sum(processed_POP_combined_31707_Gender_black$total_fblack_POP_Feb_26) 
+```
+
 #Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
@@
 Gender_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
-                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361))
+                          Male_Total = c(3698/8232, 3674/8232, 3678/8232, 3680/8232, 3684/8232, 3681/8232),
+                          Female_Total = c(5337/9813, 5311/9813, 5319/9813, 5325/9849, 5351/9849, 5361/9933))
 ```
@@
 ```{r}
-plot_Gender_black_vax_31707 = Gender_Sum_by_month_31707_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+plot_Gender_black_vax = Gender_Sum_by_month_31707_long %>% 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Gender, 
@@
 plot_Gender_black_vax = plot_Gender_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Gender in 31707") + 
+  ggtitle("Vaccination Rates for Black population by Gender in 31707") + 
   theme(plot.title = 
@@
   labs(y = "Vaccination Total")
-plot_Gender_black_vax_31707
+plot_Gender_black_vax
 ```
@@
 #save figure
-ggsave(filename = here("results","31707_Gender_black_vax.png"), plot=plot_Gender_black_vax_31707)
+ggsave(filename = here("results","31707_Gender_black_vax.png"), plot=plot_Gender_black_vax)
 ```
@@
 ```{r}
-plotly_plot_Gender_black_vax_31707 = ggplotly(plot_Gender_black_vax_31707)
-plotly_plot_Gender_black_vax_31707
-```
-
-# Thoughts regarding exploratory analysis 
-In each zip code there is a large gap between females that are Black and males that are Black. I am pretty taken back by how large the gap is.
+plotly_plot_Gender_black_vax = ggplotly(plot_Gender_black_vax)
+plotly_plot_Gender_black_vax
+```
 
@@
 
+### Loading the `processed_POP_combined_31701.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_31701", "processed_POP_combined_31701.rds")
+processed_POP_combined_31701 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black age groups within 31701
@@
 
-### Summing each Month variable by age group by month
+```{r}
+processed_POP_combined_31701_Age_black = processed_POP_combined_31701 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
+```
+
+### Summing each Month variable by age group by month for vaccinations
 ```{r}
@@
 
+### Summing each Month variable by age group by month for Population
+```{r}
+#Summing for ages 5-9 by month 
+sum(processed_POP_combined_31701_Age_black$total_05_09_POP_Sept_17) 
+sum(processed_POP_combined_31701_Age_black$total_05_09_POP_Oct_28) 
+sum(processed_POP_combined_31701_Age_black$total_05_09_POP_Nov_24) 
+sum(processed_POP_combined_31701_Age_black$total_05_09_POP_Dec_29) 
+sum(processed_POP_combined_31701_Age_black$total_05_09_POP_Jan_26) 
+sum(processed_POP_combined_31701_Age_black$total_05_09_POP_Feb_26)
+
+#Summing for ages 10-17 by month 
+sum(processed_POP_combined_31701_Age_black$total_10_17_POP_Sept_17) 
+sum(processed_POP_combined_31701_Age_black$total_10_17_POP_Oct_28) 
+sum(processed_POP_combined_31701_Age_black$total_10_17_POP_Nov_24) 
+sum(processed_POP_combined_31701_Age_black$total_10_17_POP_Dec_29) 
+sum(processed_POP_combined_31701_Age_black$total_10_17_POP_Jan_26) 
+sum(processed_POP_combined_31701_Age_black$total_10_17_POP_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_POP_combined_31701_Age_black$total_18_44_POP_Sept_17) 
+sum(processed_POP_combined_31701_Age_black$total_18_44_POP_Oct_28) 
+sum(processed_POP_combined_31701_Age_black$total_18_44_POP_Nov_24) 
+sum(processed_POP_combined_31701_Age_black$total_18_44_POP_Dec_29) 
+sum(processed_POP_combined_31701_Age_black$total_18_44_POP_Jan_26) 
+sum(processed_POP_combined_31701_Age_black$total_18_44_POP_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_POP_combined_31701_Age_black$total_45_64_POP_Sept_17) 
+sum(processed_POP_combined_31701_Age_black$total_45_64_POP_Oct_28) 
+sum(processed_POP_combined_31701_Age_black$total_45_64_POP_Nov_24) 
+sum(processed_POP_combined_31701_Age_black$total_45_64_POP_Dec_29) 
+sum(processed_POP_combined_31701_Age_black$total_45_64_POP_Jan_26) 
+sum(processed_POP_combined_31701_Age_black$total_45_64_POP_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_POP_combined_31701_Age_black$total_65Plus_POP_Sept_17) 
+sum(processed_POP_combined_31701_Age_black$total_65Plus_POP_Oct_28) 
+sum(processed_POP_combined_31701_Age_black$total_65Plus_POP_Nov_24) 
+sum(processed_POP_combined_31701_Age_black$total_65Plus_POP_Dec_29) 
+sum(processed_POP_combined_31701_Age_black$total_65Plus_POP_Jan_26) 
+sum(processed_POP_combined_31701_Age_black$total_65Plus_POP_Feb_26)
+
+```
+
 ### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
@@
 Age_Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Age_05_09 = c(123, 126, 126, 127, 127, 128),
-                          Age_10_17 = c(642, 635, 633, 637, 630, 629),
-                          Age_18_44 = c(3310, 3316, 3313, 3323, 3325, 3331),
-                          Age_45_64 = c(3229, 3244, 3239, 3259, 3264, 3258),
-                          Age_65Plus = c(2576, 2570, 2569, 2574, 2568, 2572))
+                          Age_05_09 = c(123/1058, 126/1058, 126/1058, 127/1058, 127/1058, 128/1058),
+                          Age_10_17 = c(642/1560, 635/1560, 633/1560, 637/1560, 630/1560, 629/1560),
+                          Age_18_44 = c(3310/5054, 3316/5054, 3313/5054, 3323/5054, 3325/5054, 3331/5054),
+                          Age_45_64 = c(3229/4628, 3244/4628, 3239/4628, 3259/4628, 3264/4628, 3258/4628),
+                          Age_65Plus = c(2576/2886, 2570/2886, 2569/2886, 2574/2886, 2568/2886, 2572/2886))
 
@@
   rename("Age" = "variable") %>% 
-  rename("Vaccination_total" = "value")
+  rename("Vaccination_rate" = "value")
 ```
@@
 plot_Age_black_vax = Age_Sum_by_month_31701_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Age, 
@@
 plot_Age_black_vax = plot_Age_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Age in 31701") + 
+  ggtitle("Vaccination Rates for Black population by Age in 31701") + 
   theme(plot.title = 
@@
 
+### Loading the `processed_POP_combined_31705.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_31705", "processed_POP_combined_31705.rds")
+processed_POP_combined_31705 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black age groups within 31705
@@
 
-### Summing each Month variable by age group by month
+```{r}
+processed_POP_combined_31705_Age_black = processed_POP_combined_31705 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
+```
+
+### Summing each Month variable by age group by month for vaccinations
 ```{r}
@@
 
-### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February 
-### Total contains the vaccination totals for each month
+### Summing each Month variable by age group by month for Population
+```{r}
+#Summing for ages 5-9 by month 
+sum(processed_POP_combined_31705_Age_black$total_05_09_POP_Sept_17) 
+sum(processed_POP_combined_31705_Age_black$total_05_09_POP_Oct_28) 
+sum(processed_POP_combined_31705_Age_black$total_05_09_POP_Nov_24) 
+sum(processed_POP_combined_31705_Age_black$total_05_09_POP_Dec_29) 
+sum(processed_POP_combined_31705_Age_black$total_05_09_POP_Jan_26) 
+sum(processed_POP_combined_31705_Age_black$total_05_09_POP_Feb_26)
+
+#Summing for ages 10-17 by month 
+sum(processed_POP_combined_31705_Age_black$total_10_17_POP_Sept_17) 
+sum(processed_POP_combined_31705_Age_black$total_10_17_POP_Oct_28) 
+sum(processed_POP_combined_31705_Age_black$total_10_17_POP_Nov_24) 
+sum(processed_POP_combined_31705_Age_black$total_10_17_POP_Dec_29) 
+sum(processed_POP_combined_31705_Age_black$total_10_17_POP_Jan_26) 
+sum(processed_POP_combined_31705_Age_black$total_10_17_POP_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_POP_combined_31705_Age_black$total_18_44_POP_Sept_17) 
+sum(processed_POP_combined_31705_Age_black$total_18_44_POP_Oct_28) 
+sum(processed_POP_combined_31705_Age_black$total_18_44_POP_Nov_24) 
+sum(processed_POP_combined_31705_Age_black$total_18_44_POP_Dec_29) 
+sum(processed_POP_combined_31705_Age_black$total_18_44_POP_Jan_26) 
+sum(processed_POP_combined_31705_Age_black$total_18_44_POP_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_POP_combined_31705_Age_black$total_45_64_POP_Sept_17) 
+sum(processed_POP_combined_31705_Age_black$total_45_64_POP_Oct_28) 
+sum(processed_POP_combined_31705_Age_black$total_45_64_POP_Nov_24) 
+sum(processed_POP_combined_31705_Age_black$total_45_64_POP_Dec_29) 
+sum(processed_POP_combined_31705_Age_black$total_45_64_POP_Jan_26) 
+sum(processed_POP_combined_31705_Age_black$total_45_64_POP_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_POP_combined_31705_Age_black$total_65Plus_POP_Sept_17) 
+sum(processed_POP_combined_31705_Age_black$total_65Plus_POP_Oct_28) 
+sum(processed_POP_combined_31705_Age_black$total_65Plus_POP_Nov_24) 
+sum(processed_POP_combined_31705_Age_black$total_65Plus_POP_Dec_29) 
+sum(processed_POP_combined_31705_Age_black$total_65Plus_POP_Jan_26) 
+sum(processed_POP_combined_31705_Age_black$total_65Plus_POP_Feb_26)
+
+```
+
+### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
 
@@
 Age_Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Age_05_09 = c(210, 211, 209, 201, 202, 202),
-                          Age_10_17 = c(1159, 1171, 1169, 1163, 1168, 1167),
-                          Age_18_44 = c(5201, 5215, 5239, 5263, 5278, 5281), #pretty notable increase for ages 18-44 
-                          Age_45_64 = c(4380, 4381, 4366, 4364, 4367, 4367),
-                          Age_65Plus = c(2983, 3000, 2997, 2997, 2994, 2989))
+                          Age_05_09 = c(210/1639, 211/1639, 209/1639, 201/1639, 202/1639, 202/1639),
+                          Age_10_17 = c(1159/3311, 1171/3311, 1169/3311, 1163/3311, 1168/3311, 1167/3311),
+                          Age_18_44 = c(5201/11978, 5215/11978, 5239/11978, 5263/11978, 5278/11978, 5281/11978),
+                          Age_45_64 = c(4380/6468, 4381/6468, 4366/6468, 4364/6468, 4367/6468, 4367/6468),
+                          Age_65Plus = c(2983/3497, 3000/3497, 2997/3497, 2997/3497, 2994/3497, 2989/3497))
 
@@
 plot_Age_black_vax = Age_Sum_by_month_31705_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Age, 
@@
 plot_Age_black_vax = plot_Age_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Age in 31705") + 
+  ggtitle("Vaccination Rates for Black population by Age in 31705") + 
   theme(plot.title = 
@@
 
+### Loading the `processed_POP_combined_31707.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_31707", "processed_POP_combined_31707.rds")
+processed_POP_combined_31707 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black age groups within 31707
@@
 
-### Summing each Month variable by age group by month
+```{r}
+processed_POP_combined_31707_Age_black = processed_POP_combined_31707 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
+```
+
+### Summing each Month variable by age group by month for vaccinations
 ```{r}
@@
 
-### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February 
-### Total contains the vaccination totals for each month
+### Summing each Month variable by age group by month for Population
+```{r}
+#Summing for ages 5-9 by month 
+sum(processed_POP_combined_31707_Age_black$total_05_09_POP_Sept_17) 
+sum(processed_POP_combined_31707_Age_black$total_05_09_POP_Oct_28) 
+sum(processed_POP_combined_31707_Age_black$total_05_09_POP_Nov_24) 
+sum(processed_POP_combined_31707_Age_black$total_05_09_POP_Dec_29) 
+sum(processed_POP_combined_31707_Age_black$total_05_09_POP_Jan_26) 
+sum(processed_POP_combined_31707_Age_black$total_05_09_POP_Feb_26)
+
+#Summing for ages 10-17 by month 
+sum(processed_POP_combined_31707_Age_black$total_10_17_POP_Sept_17) 
+sum(processed_POP_combined_31707_Age_black$total_10_17_POP_Oct_28) 
+sum(processed_POP_combined_31707_Age_black$total_10_17_POP_Nov_24) 
+sum(processed_POP_combined_31707_Age_black$total_10_17_POP_Dec_29) 
+sum(processed_POP_combined_31707_Age_black$total_10_17_POP_Jan_26) 
+sum(processed_POP_combined_31707_Age_black$total_10_17_POP_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_POP_combined_31707_Age_black$total_18_44_POP_Sept_17) 
+sum(processed_POP_combined_31707_Age_black$total_18_44_POP_Oct_28) 
+sum(processed_POP_combined_31707_Age_black$total_18_44_POP_Nov_24) 
+sum(processed_POP_combined_31707_Age_black$total_18_44_POP_Dec_29) 
+sum(processed_POP_combined_31707_Age_black$total_18_44_POP_Jan_26) 
+sum(processed_POP_combined_31707_Age_black$total_18_44_POP_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_POP_combined_31707_Age_black$total_45_64_POP_Sept_17) 
+sum(processed_POP_combined_31707_Age_black$total_45_64_POP_Oct_28) 
+sum(processed_POP_combined_31707_Age_black$total_45_64_POP_Nov_24) 
+sum(processed_POP_combined_31707_Age_black$total_45_64_POP_Dec_29) 
+sum(processed_POP_combined_31707_Age_black$total_45_64_POP_Jan_26) 
+sum(processed_POP_combined_31707_Age_black$total_45_64_POP_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_POP_combined_31707_Age_black$total_65Plus_POP_Sept_17) 
+sum(processed_POP_combined_31707_Age_black$total_65Plus_POP_Oct_28) 
+sum(processed_POP_combined_31707_Age_black$total_65Plus_POP_Nov_24) 
+sum(processed_POP_combined_31707_Age_black$total_65Plus_POP_Dec_29) 
+sum(processed_POP_combined_31707_Age_black$total_65Plus_POP_Jan_26) 
+sum(processed_POP_combined_31707_Age_black$total_65Plus_POP_Feb_26)
+
+```
+
+### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
 
@@
 Age_Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Age_05_09 = c(275, 277, 280, 282, 284, 283),
-                          Age_10_17 = c(1109, 1120, 1125, 1134, 1136, 1142),
-                          Age_18_44 = c(5903, 5931, 5943, 5947, 5960, 5958), #pretty notable increase for ages 18-44 
-                          Age_45_64 = c(5130, 5120, 5125, 5123, 5133, 5140),
-                          Age_65Plus = c(4200, 4204, 4195, 4206, 4202, 4189))
+                          Age_05_09 = c(275/1583, 277/1702, 280/1702, 282/1702, 284/1702, 283/1702),
+                          Age_10_17 = c(1109/3001, 1120/3001, 1125/3001, 1134/3001, 1136/3001, 1142/3001),
+                          Age_18_44 = c(5903/11434, 5931/11434, 5943/11434, 5947/11434, 5960/11434, 5958/11434),
+                          Age_45_64 = c(5130/6051, 5120/6051, 5125/6051, 5123/6051, 5133/6051, 5140/6051),
+                          Age_65Plus = c(4200/4259, 4204/4259, 4195/4259, 4206/4259, 4202/4259, 4189/4259))
 
@@
 plot_Age_black_vax = Age_Sum_by_month_31707_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Age, 
@@
 plot_Age_black_vax = plot_Age_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Age in 31707") + 
+  ggtitle("Vaccination Rates for Black population by Age in 31707") + 
   theme(plot.title = 
@@
 
-# The Amount of Vaccinations Among Black population in the Control Zip Code
+# Analyzing Control Zip Codes by vaccination rate
 
@@
 
-### Loading the processed data in order to begin analysis for the 30815 zip code
-
+### Combining Vaccination data from the 30815 zip from Sept-Feb
 ```{r}
@@
 
+### Combining Population data from the 30815 zip from Sept-Feb
+```{r}
+# Loading September 30815
+data_location1 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Sept.rds")
+processed_data_POP_30815_Sept = readRDS(data_location1)
+
+# Loading October 30815 and joining september and october 
+data_location2 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Oct.rds")
+processed_data_POP_30815_Oct = readRDS(data_location2)
+processed_POP_combined <- left_join(processed_data_POP_30815_Sept, processed_data_POP_30815_Oct, by = c("FIPS" = "FIPS"))
+
+# Loading November 30815
+data_location3 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Nov.rds")
+processed_data_POP_30815_Nov = readRDS(data_location3)
+processed_POP_combined <- left_join(processed_POP_combined, processed_data_POP_30815_Nov, by = c("FIPS" = "FIPS"))
+
+# Loading December 30815
+data_location4 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Dec.rds")
+processed_data_POP_30815_Dec = readRDS(data_location4)
+processed_POP_combined <- left_join(processed_POP_combined, processed_data_POP_30815_Dec, by = c("FIPS" = "FIPS"))
+
+# Loading January 30815
+data_location5 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Jan.rds")
+processed_data_POP_30815_Jan = readRDS(data_location5)
+processed_POP_combined <- left_join(processed_POP_combined, processed_data_POP_30815_Jan, by = c("FIPS" = "FIPS"))
+
+# Loading February 30815
+data_location6 = here("data", "processed_data", "Zip_30815", "processed_data_POP_30815_Feb.rds")
+processed_data_POP_30815_Feb = readRDS(data_location6)
+processed_POP_combined_30815 <- left_join(processed_POP_combined, processed_data_POP_30815_Feb, by = c("FIPS" = "FIPS"))
+```
+
 ### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender
@@
 
+```{r}
+save_data_location = here("data", "processed_data","Zip_30815", "processed_POP_combined_30815.rds")
+saveRDS(processed_POP_combined_30815, file = save_data_location)
+```
+
 ### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 30815 zip code
@@
 
-### Summing each Month variable by census tract
-
+```{r}
+processed_combined_30815_total_black_POP = processed_POP_combined_30815 %>% select(starts_with("total_black_POP"))
+```
+
+### Summing vaccination numbers
 ```{r}
@@
 
-### Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February ### Total contains the vaccination totals for each month
+### Summing population numbers
+```{r}
+#Summing each column for each month 
+sum(processed_combined_30815_total_black_POP$total_black_POP_Sept_17) 
+sum(processed_combined_30815_total_black_POP$total_black_POP_Oct_28) 
+sum(processed_combined_30815_total_black_POP$total_black_POP_Nov_24) 
+sum(processed_combined_30815_total_black_POP$total_black_POP_Dec_29) 
+sum(processed_combined_30815_total_black_POP$total_black_POP_Jan_26) 
+sum(processed_combined_30815_total_black_POP$total_black_POP_Feb_26) 
+```
+
+#Separating sums by month into a data fram consisting of two variables: Date and VAX_RATE #Date contains months September-February #Total contains the vaccination totals for each month
 
@@
 Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Total = c(8758, 8688, 8639, 8593, 8623, 8666)) 
+                               VAX_RATE = c(8758/19606, 8688/19736, 8639/19736, 8593/19736, 8623/19736, 8666/19736)) 
 ```
@@
 ```{r}
-plot_30815_total_black_vax = Sum_by_month_30815 %>% 
-  ggplot(aes(Date, Total, group=1)) + 
+plot_30815_total_black_vax_rate = Sum_by_month_30815 %>% 
+  ggplot(aes(Date, VAX_RATE, group=1)) + 
   geom_point() + geom_line() + 
@@
 #Adding a customization title for the plot 
-plot_30815_total_black_vax = plot_30815_total_black_vax + 
-  ggtitle("Vaccination Totals for Black population in the 30815 Zipcode") + 
+plot_30815_total_black_vax_rate = plot_30815_total_black_vax_rate + 
+  ggtitle("Vaccination Rates for Black Population in the 30815 Zipcode") + 
   theme(plot.title = 
           element_text(color="red", size=10, face="bold.italic"))
-plot_30815_total_black_vax
+plot_30815_total_black_vax_rate
 ```
@@
 #save figure
-ggsave(filename = here("results","30815_total_black_vax.png"), plot=plot_30815_total_black_vax)
+ggsave(filename = here("results","plot_30815_total_black_vax_rate.png"), plot=plot_30815_total_black_vax_rate)
 ```
@@
 ```{r}
-plotly_30815_total_black_vax = ggplotly(plot_30815_total_black_vax)
-plotly_30815_total_black_vax
-```
-
-# The Amount of Vaccinations by Gender Among Black Population in the Control Zip Code
+plotly_30815_total_black_vax_rate = ggplotly(plot_30815_total_black_vax_rate)
+plotly_30815_total_black_vax_rate
+```
+
+## 30815 by gender
 
@@
 
-### Loading the `processed_combined_30815.rds` file from `analysiscode.qmd`
-
+### Loading the `processed_combined_30815.rds` 
 ```{r}
@@
 
+### Loading the `processed_POP_combined_30815.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_30815", "processed_POP_combined_30815.rds")
+processed_POP_combined_30815 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black males and black females within 30815
 
-#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Isolating Gender Variables form Vaccination data and Population data
 
@@
 
-### Summing each Month variable by census tract
+```{r}
+processed_POP_combined_30815_Gender_black = processed_POP_combined_30815 %>% select(matches("^total_(mblack|fblack)_POP"))
+```
+
+### Summing each Month for vaccination data 
 ```{r}
@@
 
+### Summing each Month for population data
+```{r}
+#Summing for black males by month 
+sum(processed_POP_combined_30815_Gender_black$total_mblack_POP_Sept_17) 
+sum(processed_POP_combined_30815_Gender_black$total_mblack_POP_Oct_28) 
+sum(processed_POP_combined_30815_Gender_black$total_mblack_POP_Nov_24) 
+sum(processed_POP_combined_30815_Gender_black$total_mblack_POP_Dec_29) 
+sum(processed_POP_combined_30815_Gender_black$total_mblack_POP_Jan_26) 
+sum(processed_POP_combined_30815_Gender_black$total_mblack_POP_Feb_26)
+
+#summing for black females by month 
+sum(processed_POP_combined_30815_Gender_black$total_fblack_POP_Sept_17) 
+sum(processed_POP_combined_30815_Gender_black$total_fblack_POP_Oct_28) 
+sum(processed_POP_combined_30815_Gender_black$total_fblack_POP_Nov_24) 
+sum(processed_POP_combined_30815_Gender_black$total_fblack_POP_Dec_29) 
+sum(processed_POP_combined_30815_Gender_black$total_fblack_POP_Jan_26) 
+sum(processed_POP_combined_30815_Gender_black$total_fblack_POP_Feb_26) 
+```
+
 #Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
@@
 Gender_Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Male_Total = c(3526, 3514, 3503, 3484, 3489, 3498),
-                          Female_Total = c(5232, 5174, 5136, 5109, 5134, 5168))
+                          Male_Total = c(3526/9066, 3514/9196, 3503/9196, 3484/9196, 3489/9196, 3498/9196),
+                          Female_Total = c(5232/10540, 5174/10540, 5136/10540, 5109/10540, 5134/10540, 5168/10540))
 ```
@@
 plot_Gender_black_vax = Gender_Sum_by_month_30815_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Gender, 
@@
 plot_Gender_black_vax = plot_Gender_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Gender in 30815") + 
+  ggtitle("Vaccination Rates for Black population by Gender in 30815") + 
   theme(plot.title = 
@@
 
-# Total Amount of vaccinations by Age among Black population in the Control Zip Code
-
-## Zip code: 30815
+## 30815 by age
 
@@
 
+### Loading the `processed_POP_combined_30815.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_30815", "processed_POP_combined_30815.rds")
+processed_POP_combined_30815 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black age groups within 30815
@@
 
-### Summing each Month variable by age group by month
+```{r}
+processed_POP_combined_30815_Age_black = processed_POP_combined_30815 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
+```
+
+### Summing each Month variable by age group by month for vaccinations
 ```{r}
@@
 
+### Summing each Month variable by age group by month for Population
+```{r}
+#Summing for ages 5-9 by month 
+sum(processed_POP_combined_30815_Age_black$total_05_09_POP_Sept_17) 
+sum(processed_POP_combined_30815_Age_black$total_05_09_POP_Oct_28) 
+sum(processed_POP_combined_30815_Age_black$total_05_09_POP_Nov_24) 
+sum(processed_POP_combined_30815_Age_black$total_05_09_POP_Dec_29) 
+sum(processed_POP_combined_30815_Age_black$total_05_09_POP_Jan_26) 
+sum(processed_POP_combined_30815_Age_black$total_05_09_POP_Feb_26)
+
+#Summing for ages 10-17 by month 
+sum(processed_POP_combined_30815_Age_black$total_10_17_POP_Sept_17) 
+sum(processed_POP_combined_30815_Age_black$total_10_17_POP_Oct_28) 
+sum(processed_POP_combined_30815_Age_black$total_10_17_POP_Nov_24) 
+sum(processed_POP_combined_30815_Age_black$total_10_17_POP_Dec_29) 
+sum(processed_POP_combined_30815_Age_black$total_10_17_POP_Jan_26) 
+sum(processed_POP_combined_30815_Age_black$total_10_17_POP_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_POP_combined_30815_Age_black$total_18_44_POP_Sept_17) 
+sum(processed_POP_combined_30815_Age_black$total_18_44_POP_Oct_28) 
+sum(processed_POP_combined_30815_Age_black$total_18_44_POP_Nov_24) 
+sum(processed_POP_combined_30815_Age_black$total_18_44_POP_Dec_29) 
+sum(processed_POP_combined_30815_Age_black$total_18_44_POP_Jan_26) 
+sum(processed_POP_combined_30815_Age_black$total_18_44_POP_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_POP_combined_30815_Age_black$total_45_64_POP_Sept_17) 
+sum(processed_POP_combined_30815_Age_black$total_45_64_POP_Oct_28) 
+sum(processed_POP_combined_30815_Age_black$total_45_64_POP_Nov_24) 
+sum(processed_POP_combined_30815_Age_black$total_45_64_POP_Dec_29) 
+sum(processed_POP_combined_30815_Age_black$total_45_64_POP_Jan_26) 
+sum(processed_POP_combined_30815_Age_black$total_45_64_POP_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_POP_combined_30815_Age_black$total_65Plus_POP_Sept_17) 
+sum(processed_POP_combined_30815_Age_black$total_65Plus_POP_Oct_28) 
+sum(processed_POP_combined_30815_Age_black$total_65Plus_POP_Nov_24) 
+sum(processed_POP_combined_30815_Age_black$total_65Plus_POP_Dec_29) 
+sum(processed_POP_combined_30815_Age_black$total_65Plus_POP_Jan_26) 
+sum(processed_POP_combined_30815_Age_black$total_65Plus_POP_Feb_26)
+
+```
+
 ### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
@@
 Age_Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Age_05_09 = c(201, 213, 211, 215, 211, 211),
-                          Age_10_17 = c(1116, 1113, 1115, 1118, 1115, 1114),
-                          Age_18_44 = c(5001, 5009, 5020, 5026, 5025, 5032),
-                          Age_45_64 = c(5026, 5017, 5020, 5009, 5013, 5020),
-                          Age_65Plus = c(2816, 2825, 2827, 2814, 2828, 2832))
+                          Age_05_09 = c(201/2388, 213/2388, 211/2480, 215/2480, 211/2480, 211/2480),
+                          Age_10_17 = c(1116/3604, 1113/3604, 1115/3604, 1118/3604, 1115/3604, 1114/3604),
+                          Age_18_44 = c(5001/13358, 5009/13358, 5020/13358, 5026/13358, 5025/13358, 5032/13358),
+                          Age_45_64 = c(5026/8431, 5017/8431, 5020/8431, 5009/8431, 5013/8431, 5020/8431),
+                          Age_65Plus = c(2816/3590, 2825/3590, 2827/3590, 2814/3590, 2828/3590, 2832/3590))
 
@@
 plot_Age_black_vax = Age_Sum_by_month_30815_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Age, 
@@
 plot_Age_black_vax = plot_Age_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Age in 30815") + 
+  ggtitle("Vaccination Rates for Black population by Age in 30815") + 
   theme(plot.title = 
@@
 #save figure
-ggsave(filename = here("results","30815_Age_black_vax.png"), plot=plot_Gender_black_vax)
+ggsave(filename = here("results","30815_Age_black_vax.png"), plot=plot_Age_black_vax)
 ```
@@
 
-# The Amount of Vaccinations Among Black population in the Control Zip Code
+## 30907 by black residents
 
@@
 
-### Loading the processed data in order to begin analysis for the 30907 zip code
-
+### Combining Vaccination data from the 30907 zip from Sept-Feb
 ```{r}
@@
 
+### Combining Population data from the 30907 zip from Sept-Feb
+```{r}
+# Loading September 30907
+data_location1 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Sept.rds")
+processed_data_POP_30907_Sept = readRDS(data_location1)
+
+# Loading October 30907 and joining september and october 
+data_location2 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Oct.rds")
+processed_data_POP_30907_Oct = readRDS(data_location2)
+processed_POP_combined <- left_join(processed_data_POP_30907_Sept, processed_data_POP_30907_Oct, by = c("FIPS" = "FIPS"))
+
+# Loading November 30907
+data_location3 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Nov.rds")
+processed_data_POP_30907_Nov = readRDS(data_location3)
+processed_POP_combined <- left_join(processed_POP_combined, processed_data_POP_30907_Nov, by = c("FIPS" = "FIPS"))
+
+# Loading December 30907
+data_location4 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Dec.rds")
+processed_data_POP_30907_Dec = readRDS(data_location4)
+processed_POP_combined <- left_join(processed_POP_combined, processed_data_POP_30907_Dec, by = c("FIPS" = "FIPS"))
+
+# Loading January 30907
+data_location5 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Jan.rds")
+processed_data_POP_30907_Jan = readRDS(data_location5)
+processed_POP_combined <- left_join(processed_POP_combined, processed_data_POP_30907_Jan, by = c("FIPS" = "FIPS"))
+
+# Loading February 30907
+data_location6 = here("data", "processed_data", "Zip_30907", "processed_data_POP_30907_Feb.rds")
+processed_data_POP_30907_Feb = readRDS(data_location6)
+processed_POP_combined_30907 <- left_join(processed_POP_combined, processed_data_POP_30907_Feb, by = c("FIPS" = "FIPS"))
+```
+
 ### Saving the combined data for Black population for future use. Will be useful when stratifying the African American population by age and gender
@@
 
+```{r}
+save_data_location = here("data", "processed_data","Zip_30907", "processed_POP_combined_30907.rds")
+saveRDS(processed_POP_combined_30907, file = save_data_location)
+```
+
 ### Dropping all variables except for total black variables in order to obtain vaccination rates of all Black population in the 30907 zip code
@@
 
-### Summing each Month variable by census tract
-
+```{r}
+processed_combined_30907_total_black_POP = processed_POP_combined_30907 %>% select(starts_with("total_black_POP"))
+```
+
+### Summing vaccination numbers
 ```{r}
@@
 
-### Separating sums by month into a data fram consisting of two variables: Date and Total #Date contains months September-February ### Total contains the vaccination totals for each month
+### Summing population numbers
+```{r}
+#Summing each column for each month 
+sum(processed_combined_30907_total_black_POP$total_black_POP_Sept_17) 
+sum(processed_combined_30907_total_black_POP$total_black_POP_Oct_28) 
+sum(processed_combined_30907_total_black_POP$total_black_POP_Nov_24) 
+sum(processed_combined_30907_total_black_POP$total_black_POP_Dec_29) 
+sum(processed_combined_30907_total_black_POP$total_black_POP_Jan_26) 
+sum(processed_combined_30907_total_black_POP$total_black_POP_Feb_26) 
+```
+
+#Separating sums by month into a data fram consisting of two variables: Date and VAX_RATE #Date contains months September-February #Total contains the vaccination totals for each month
 
@@
 Sum_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Total = c(4956, 4917, 4888, 4851, 4866, 4894)) 
+                               VAX_RATE = c(4956/9122, 4917/9189, 4888/9189, 4851/9189, 4866/9189, 4894/9189)) 
 ```
@@
 ```{r}
-plot_30907_total_black_vax = Sum_by_month_30907 %>% 
-  ggplot(aes(Date, Total, group=1)) + 
+plot_30907_total_black_vax_rate = Sum_by_month_30907 %>% 
+  ggplot(aes(Date, VAX_RATE, group=1)) + 
   geom_point() + geom_line() + 
@@
 #Adding a customization title for the plot 
-plot_30907_total_black_vax = plot_30907_total_black_vax + 
-  ggtitle("Vaccination Totals for Black population in the 30907 Zipcode") + 
+plot_30907_total_black_vax_rate = plot_30907_total_black_vax_rate + 
+  ggtitle("Vaccination Rates for Black Population in the 30907 Zipcode") + 
   theme(plot.title = 
           element_text(color="red", size=10, face="bold.italic"))
-plot_30907_total_black_vax
+plot_30907_total_black_vax_rate
 ```
@@
 #save figure
-ggsave(filename = here("results","30907_total_black_vax.png"), plot=plot_30907_total_black_vax)
+ggsave(filename = here("results","plot_30907_total_black_vax_rate.png"), plot=plot_30907_total_black_vax_rate)
 ```
@@
 ```{r}
-plotly_30907_total_black_vax = ggplotly(plot_30907_total_black_vax)
-plotly_30907_total_black_vax
-```
-
-# The Amount of Vaccinations by Gender Among Black Population in the Control Zip Code
+plotly_30907_total_black_vax_rate = ggplotly(plot_30907_total_black_vax_rate)
+plotly_30907_total_black_vax_rate
+```
+
+## Zip code 30907 by gender among residents that black
 
@@
 
-### Loading the `processed_combined_30907.rds` file from `analysiscode.qmd`
-
+### Loading the `processed_combined_30907.rds` 
 ```{r}
@@
 
+### Loading the `processed_POP_combined_30907.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_30907", "processed_POP_combined_30907.rds")
+processed_POP_combined_30907 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black males and black females within 30907
 
-#### Using the select() function to subset the data frame by columns that start with "Gender_black_vax" which are variables with a total count of Black population with at least one dose of the COVID-19 vaccine by month
+#### Isolating Gender Variables form Vaccination data and Population data
 
@@
 
-### Summing each Month variable by census tract
+```{r}
+processed_POP_combined_30907_Gender_black = processed_POP_combined_30907 %>% select(matches("^total_(mblack|fblack)_POP"))
+```
+
+### Summing each Month for vaccination data 
 ```{r}
@@
 
+### Summing each Month for population data
+```{r}
+#Summing for black males by month 
+sum(processed_POP_combined_30907_Gender_black$total_mblack_POP_Sept_17) 
+sum(processed_POP_combined_30907_Gender_black$total_mblack_POP_Oct_28) 
+sum(processed_POP_combined_30907_Gender_black$total_mblack_POP_Nov_24) 
+sum(processed_POP_combined_30907_Gender_black$total_mblack_POP_Dec_29) 
+sum(processed_POP_combined_30907_Gender_black$total_mblack_POP_Jan_26) 
+sum(processed_POP_combined_30907_Gender_black$total_mblack_POP_Feb_26)
+
+#summing for black females by month 
+sum(processed_POP_combined_30907_Gender_black$total_fblack_POP_Sept_17) 
+sum(processed_POP_combined_30907_Gender_black$total_fblack_POP_Oct_28) 
+sum(processed_POP_combined_30907_Gender_black$total_fblack_POP_Nov_24) 
+sum(processed_POP_combined_30907_Gender_black$total_fblack_POP_Dec_29) 
+sum(processed_POP_combined_30907_Gender_black$total_fblack_POP_Jan_26) 
+sum(processed_POP_combined_30907_Gender_black$total_fblack_POP_Feb_26) 
+```
+
 #Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
@@
 Gender_Sum_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Male_Total = c(1947, 1943, 1926, 1900, 1913, 1919),
-                          Female_Total = c(3009, 2974, 2962, 2951, 2953, 2975))
+                          Male_Total = c(1947/4289, 1943/4289, 1926/4289, 1900/4289, 1913/4289, 1919/4289),
+                          Female_Total = c(3009/4833, 2974/4900, 2962/4900, 2951/4900, 2953/4900, 2975/4900))
 ```
@@
 plot_Gender_black_vax = Gender_Sum_by_month_30907_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Gender, 
@@
 plot_Gender_black_vax = plot_Gender_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Gender in 30907") + 
+  ggtitle("Vaccination Rates for Black population by Gender in 30907") + 
   theme(plot.title = 
@@
 
+### Loading the `processed_POP_combined_30907.rds`
+```{r}
+save_data_location = here("data", "processed_data","Zip_30907", "processed_POP_combined_30907.rds")
+processed_POP_combined_30907 = readRDS(save_data_location)
+```
+
 ### Dropping all variables except for black age groups within 30907
@@
 
-### Summing each Month variable by age group by month
+```{r}
+processed_POP_combined_30907_Age_black = processed_POP_combined_30907 %>% 
+  select(starts_with("total_05_09"), starts_with("total_10_17"), starts_with("total_18_44"), starts_with("total_45_64"), starts_with("total_65Plus"))
+```
+
+### Summing each Month variable by age group by month for vaccinations
 ```{r}
@@
 
+### Summing each Month variable by age group by month for Population
+```{r}
+#Summing for ages 5-9 by month 
+sum(processed_POP_combined_30907_Age_black$total_05_09_POP_Sept_17) 
+sum(processed_POP_combined_30907_Age_black$total_05_09_POP_Oct_28) 
+sum(processed_POP_combined_30907_Age_black$total_05_09_POP_Nov_24) 
+sum(processed_POP_combined_30907_Age_black$total_05_09_POP_Dec_29) 
+sum(processed_POP_combined_30907_Age_black$total_05_09_POP_Jan_26) 
+sum(processed_POP_combined_30907_Age_black$total_05_09_POP_Feb_26)
+
+#Summing for ages 10-17 by month 
+sum(processed_POP_combined_30907_Age_black$total_10_17_POP_Sept_17) 
+sum(processed_POP_combined_30907_Age_black$total_10_17_POP_Oct_28) 
+sum(processed_POP_combined_30907_Age_black$total_10_17_POP_Nov_24) 
+sum(processed_POP_combined_30907_Age_black$total_10_17_POP_Dec_29) 
+sum(processed_POP_combined_30907_Age_black$total_10_17_POP_Jan_26) 
+sum(processed_POP_combined_30907_Age_black$total_10_17_POP_Feb_26)
+
+#Summing for ages 18-44 by month 
+sum(processed_POP_combined_30907_Age_black$total_18_44_POP_Sept_17) 
+sum(processed_POP_combined_30907_Age_black$total_18_44_POP_Oct_28) 
+sum(processed_POP_combined_30907_Age_black$total_18_44_POP_Nov_24) 
+sum(processed_POP_combined_30907_Age_black$total_18_44_POP_Dec_29) 
+sum(processed_POP_combined_30907_Age_black$total_18_44_POP_Jan_26) 
+sum(processed_POP_combined_30907_Age_black$total_18_44_POP_Feb_26)
+
+#Summing for ages 45-64 by month 
+sum(processed_POP_combined_30907_Age_black$total_45_64_POP_Sept_17) 
+sum(processed_POP_combined_30907_Age_black$total_45_64_POP_Oct_28) 
+sum(processed_POP_combined_30907_Age_black$total_45_64_POP_Nov_24) 
+sum(processed_POP_combined_30907_Age_black$total_45_64_POP_Dec_29) 
+sum(processed_POP_combined_30907_Age_black$total_45_64_POP_Jan_26) 
+sum(processed_POP_combined_30907_Age_black$total_45_64_POP_Feb_26)
+
+#Summing for ages 65+ by month 
+sum(processed_POP_combined_30907_Age_black$total_65Plus_POP_Sept_17) 
+sum(processed_POP_combined_30907_Age_black$total_65Plus_POP_Oct_28) 
+sum(processed_POP_combined_30907_Age_black$total_65Plus_POP_Nov_24) 
+sum(processed_POP_combined_30907_Age_black$total_65Plus_POP_Dec_29) 
+sum(processed_POP_combined_30907_Age_black$total_65Plus_POP_Jan_26) 
+sum(processed_POP_combined_30907_Age_black$total_65Plus_POP_Feb_26)
+
+```
+
 ### Separating sums by month into a data fram consisting of two variables: Date and Total Date contains months September-February #Total contains the vaccination totals for each month
@@
 Age_Sum_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Age_05_09 = c(244, 239, 241, 236, 240, 242),
-                          Age_10_17 = c(1344, 1358, 1357, 1341, 1343, 1335),
-                          Age_18_44 = c(9203, 9217, 9238, 9245, 9253, 9246),
-                          Age_45_64 = c(6740, 6700, 6667, 6626, 6617, 6624),
-                          Age_65Plus = c(6632, 6633, 6639, 6635, 6633, 6623))
+                          Age_05_09 = c(244/2174, 239/2174, 241/2174, 236/2174, 240/2174, 242/2174),
+                          Age_10_17 = c(1344/4070, 1358/4070, 1357/4070, 1341/4070, 1343/4070, 1335/4070),
+                          Age_18_44 = c(9203/14571, 9217/14571, 9238/14571, 9245/14571, 9253/14571, 9246/14571),
+                          Age_45_64 = c(6740/10362, 6700/10362, 6667/10362, 6626/10362, 6617/10362, 6624/10362),
+                          Age_65Plus = c(6632/6995, 6633/6995, 6639/6995, 6635/6995, 6633/6995, 6623/6995))
 
@@
 plot_Age_black_vax = Age_Sum_by_month_30907_long %>% 
-                     ggplot(aes(Date, Vaccination_total, 
+                     ggplot(aes(Date, Vaccination_rate, 
                                 group = Age, 
@@
 plot_Age_black_vax = plot_Age_black_vax + 
-  ggtitle("Vaccination Totals for Black population by Age in 30907") + 
+  ggtitle("Vaccination Rates for Black population by Age in 30907") + 
   theme(plot.title = 

### 2023-04-19 code/analysis_code/statisticalanalysis.qmd
Vax Rates

@@
 
-### creating a data frame that contains the variables that will be included in new linear regression mode. Just copying and pasting all of the totals generated from the `exploratoryanalysis.qmd` file
 ```{r}
-#combined all relevant variables into 5 different data sets for each zip code. Got the black population totals for each zip code from  https://www.unitedstateszipcodes.org/31701/
 
 Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Total = c(7797, 7780, 7755, 7773, 7792, 7817),
+                          TOTAL_VAX_RATE = c(7797/12768, 7780/12768, 7755/12897, 7773/12897, 7792/12897, 7817/12897),
                           Intervention = c(1, 1, 1, 1, 1, 1),
-                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
-                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420),
-                          Age_05_09 = c(123, 126, 126, 127, 127, 128),
-                          Age_10_17 = c(642, 635, 633, 637, 630, 629),
-                          Age_18_44 = c(3310, 3316, 3313, 3323, 3325, 3331),
-                          Age_45_64 = c(3229, 3244, 3239, 3259, 3264, 3258),
-                          Age_65Plus = c(2576, 2570, 2569, 2574, 2568, 2572),
-                          Pop = c(16475, 16475, 16475, 16475, 16475, 16475))
+                          Male_VAX_RATE = c(3371/5981, 3358/5981, 3368/5981, 3379/5981, 3387/5981, 3397/5981),
+                          Female_VAX_RATE = c(4426/6787, 4422/6787, 4387/6916, 4394/6916, 4405/6916, 4420/6916),
+                          Age_05_09_VAX_RATE = c(123/1058, 126/1058, 126/1058, 127/1058, 127/1058, 128/1058),
+                          Age_10_17_VAX_RATE = c(642/1560, 635/1560, 633/1560, 637/1560, 630/1560, 629/1560),
+                          Age_18_44_VAX_RATE = c(3310/5054, 3316/5054, 3313/5054, 3323/5054, 3325/5054, 3331/5054),
+                          Age_45_64_VAX_RATE = c(3229/4628, 3244/4628, 3239/4628, 3259/4628, 3264/4628, 3258/4628),
+                          Age_65Plus_VAX_RATE = c(2576/2886, 2570/2886, 2569/2886, 2574/2886, 2568/2886, 2572/2886),
+                          TOTAL_RATE_DIFF = c(0, 7780/12768 - 7797/12768, 7755/12897 - 7780/12768, 7773/12897 - 7755/12897, 7792/12897 - 7773/12897, 7817/12897 - 7792/12897))
 
 Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                                Total = c(10383, 10327, 10303, 10301, 10337, 10348),                                 Intervention = c(1, 1, 1, 1, 1, 1),
-                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
-                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949),
-                          Age_05_09 = c(210, 211, 209, 201, 202, 202),
-                          Age_10_17 = c(1159, 1171, 1169, 1163, 1168, 1167),
-                          Age_18_44 = c(5201, 5215, 5239, 5263, 5278, 5281),
-                          Age_45_64 = c(4380, 4381, 4366, 4364, 4367, 4367),
-                          Age_65Plus = c(2983, 3000, 2997, 2997, 2994, 2989),
-                          Pop = c(24779, 24779, 24779, 24779, 24779, 24779))
+                                TOTAL_VAX_RATE = c(10383/21603, 10327/21572, 10303/21722, 10301/21887, 10337/21887, 10348/21887),                                 Intervention = c(1, 1, 1, 1, 1, 1),
+                          Male_VAX_RATE = c(4394/9885, 4397/9854, 4399/10004, 4391/10004, 4403/10004, 4399/10004),
+                          Female_VAX_RATE = c(5989/11718, 5930/11718, 5904/11718, 5910/11883, 5934/11883, 5949/11883),
+                          Age_05_09_VAX_RATE = c(210/1639, 211/1639, 209/1639, 201/1639, 202/1639, 202/1639),
+                          Age_10_17_VAX_RATE = c(1159/3311, 1171/3311, 1169/3311, 1163/3311, 1168/3311, 1167/3311),
+                          Age_18_44_VAX_RATE = c(5201/11978, 5215/11978, 5239/11978, 5263/11978, 5278/11978, 5281/11978),
+                          Age_45_64_VAX_RATE = c(4380/6468, 4381/6468, 4366/6468, 4364/6468, 4367/6468, 4367/6468),
+                          Age_65Plus_VAX_RATE = c(2983/3497, 3000/3497, 2997/3497, 2997/3497, 2994/3497, 2989/3497), 
+                          TOTAL_RATE_DIFF = c(0, 10327/21572 - 10383/21603, 10303/21722 - 10327/21572, 10301/21887 - 10303/21722, 10337/21887 - 10301/21887, 10348/21887 - 10337/21887))
 
 Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                                Total = c(9035, 8985, 8997, 9005, 9035, 9042),                                 Intervention = c(1, 1, 1, 1, 1, 1),
-                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
-                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361),
-                          Age_05_09 = c(275, 277, 280, 282, 284, 283),
-                          Age_10_17 = c(1109, 1120, 1125, 1134, 1136, 1142),
-                          Age_18_44 = c(5903, 5931, 5943, 5947, 5960, 5958),  
-                          Age_45_64 = c(5130, 5120, 5125, 5123, 5133, 5140),
-                          Age_65Plus = c(4200, 4204, 4195, 4206, 4202, 4189),
-                          Pop = c(13884, 13884, 13884, 13884, 13884, 13884))
+                          TOTAL_VAX_RATE = c(9035/18045, 8985/18045, 8997/18045, 9005/18081, 9035/18081, 9042/18165),                                                       Intervention = c(1, 1, 1, 1, 1, 1),
+                          Male_VAX_RATE = c(3698/8232, 3674/8232, 3678/8232, 3680/8232, 3684/8232, 3681/8232),
+                          Female_VAX_RATE = c(5337/9813, 5311/9813, 5319/9813, 5325/9849, 5351/9849, 5361/9933),
+                          Age_05_09_VAX_RATE = c(275/1583, 277/1702, 280/1702, 282/1702, 284/1702, 283/1702),
+                          Age_10_17_VAX_RATE = c(1109/3001, 1120/3001, 1125/3001, 1134/3001, 1136/3001, 1142/3001),
+                          Age_18_44_VAX_RATE = c(5903/11434, 5931/11434, 5943/11434, 5947/11434, 5960/11434, 5958/11434),
+                          Age_45_64_VAX_RATE = c(5130/6051, 5120/6051, 5125/6051, 5123/6051, 5133/6051, 5140/6051),
+                          Age_65Plus_VAX_RATE = c(4200/4259, 4204/4259, 4195/4259, 4206/4259, 4202/4259, 4189/4259), 
+                          TOTAL_RATE_DIFF = c(0, 8985/18045 - 9035/18045, 8997/18045 - 8985/18045, 9005/18081 - 8997/18045, 9035/18081- 9005/18081, 9042/18165 - 9035/18081))
 
 Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Total = c(8758, 8688, 8639, 8593, 8623, 8666),                                      Intervention = c(0, 0, 0, 0, 0, 0),
-                          Male_Total = c(3526, 3514, 3503, 3484, 3489, 3498),
-                          Female_Total = c(5232, 5174, 5136, 5109, 5134, 5168),
-                          Age_05_09 = c(201, 213, 211, 215, 211, 211),
-                          Age_10_17 = c(1116, 1113, 1115, 1118, 1115, 1114),
-                          Age_18_44 = c(5001, 5009, 5020, 5026, 5025, 5032),
-                          Age_45_64 = c(5026, 5017, 5020, 5009, 5013, 5020),
-                          Age_65Plus = c(2816, 2825, 2827, 2814, 2828, 2832),
-                          Pop = c(23481, 23481, 23481, 23481, 23481, 23481))
+                          TOTAL_VAX_RATE = c(8758/19606, 8688/19736, 8639/19736, 8593/19736, 8623/19736, 8666/19736),                                                       Intervention = c(0, 0, 0, 0, 0, 0),
+                          Male_VAX_RATE = c(3526/9066, 3514/9196, 3503/9196, 3484/9196, 3489/9196, 3498/9196),
+                          Female_VAX_RATE = c(5232/10540, 5174/10540, 5136/10540, 5109/10540, 5134/10540, 5168/10540),
+                          Age_05_09_VAX_RATE = c(201/2388, 213/2388, 211/2480, 215/2480, 211/2480, 211/2480),
+                          Age_10_17_VAX_RATE = c(1116/3604, 1113/3604, 1115/3604, 1118/3604, 1115/3604, 1114/3604),
+                          Age_18_44_VAX_RATE = c(5001/13358, 5009/13358, 5020/13358, 5026/13358, 5025/13358, 5032/13358),
+                          Age_45_64_VAX_RATE = c(5026/8431, 5017/8431, 5020/8431, 5009/8431, 5013/8431, 5020/8431),
+                          Age_65Plus_VAX_RATE = c(2816/3590, 2825/3590, 2827/3590, 2814/3590, 2828/3590, 2832/3590),
+                          TOTAL_RATE_DIFF = c(0, 8688/19736 - 8758/19606, 8639/19736 - 8688/19736, 8593/19736 - 8639/19736, 8623/19736- 8593/19736, 8666/19736 - 8623/19736))
 
-#um_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          #Total = c(4956, 4917, 4888, 4851, 4866, 4894),                                      Zip = c(30907, 30907, 30907, 30907, 30907, 30907),
-                          #Male_Total = c(1947, 1943, 1926, 1900, 1913, 1919),
-                          #Female_Total = c(3009, 2974, 2962, 2951, 2953, 2975),
-                          #Age_05_09 = c(244, 239, 241, 236, 240, 242),
-                          #Age_10_17 = c(1344, 1358, 1357, 1341, 1343, 1335),
-                          #Age_18_44 = c(9203, 9217, 9238, 9245, 9253, 9246),
-                          #Age_45_64 = c(6740, 6700, 6667, 6626, 6617, 6624),
-                          #Age_65Plus = c(6632, 6633, 6639, 6635, 6633, 6623),
-                          #Pop = c(7744, 7744, 7744, 7744, 7744, 7744)) 
-#the 30907 zip code population of residents that are black does not match the vaccination data.
-also this zip code does not have a majority of residents that are black like other other 4 zip codes so I am going to leave it out of the statistical analysis
+Sum_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          TOTAL_VAX_RATE = c(4956/9122, 4917/9189, 4888/9189, 4851/9189, 4866/9189, 4894/9189),                                                                      Intervention = c(0, 0, 0, 0, 0, 0),
+                          Male_VAX_RATE = c(1947/4289, 1943/4289, 1926/4289, 1900/4289, 1913/4289, 1919/4289),
+                          Female_VAX_RATE = c(3009/4833, 2974/4900, 2962/4900, 2951/4900, 2953/4900, 2975/4900),
+                          Age_05_09_VAX_RATE = c(244/2174, 239/2174, 241/2174, 236/2174, 240/2174, 242/2174),
+                          Age_10_17_VAX_RATE = c(1344/4070, 1358/4070, 1357/4070, 1341/4070, 1343/4070, 1335/4070),
+                          Age_18_44_VAX_RATE = c(9203/14571, 9217/14571, 9238/14571, 9245/14571, 9253/14571, 9246/14571),
+                          Age_45_64_VAX_RATE = c(6740/10362, 6700/10362, 6667/10362, 6626/10362, 6617/10362, 6624/10362),
+                          Age_65Plus_VAX_RATE = c(6632/6995, 6633/6995, 6639/6995, 6635/6995, 6633/6995, 6623/6995), 
+                          TOTAL_RATE_DIFF = c(0, 4917/9189 - 4956/9122, 4888/9189 - 4917/9189, 4851/9189 - 4888/9189, 4866/9189 - 4851/9189, 4894/9189 - 4866/9189))
 
 #combining all of the zip codes into one data set except for the 30907 data 
-combined = rbind(Sum_by_month_31701, Sum_by_month_31705, Sum_by_month_31707, Sum_by_month_30815)
-```
-
-### creating a vaccination rate variable as the outcome 
-```{r}
-combined$Vax_rate = (combined$Total / combined$Pop) * 100
+combined = rbind(Sum_by_month_31701, Sum_by_month_31705, Sum_by_month_31707, Sum_by_month_30815,Sum_by_month_30907)
 ```
@@
 ```{r}
-lm_model2 <- lm(Vax_rate ~ Intervention + Male_Total + Female_Total + Age_05_09 + Age_10_17 + Age_18_44 + Age_45_64 + Age_65Plus, data = combined)
+lm_model2 <- lm(TOTAL_RATE_DIFF ~ TOTAL_VAX_RATE + Intervention + Male_VAX_RATE + Female_VAX_RATE + Age_05_09_VAX_RATE + Age_10_17_VAX_RATE + Age_18_44_VAX_RATE + Age_45_64_VAX_RATE + Age_65Plus_VAX_RATE, data = combined)
 
@@
 ```
-The results indicate that Intervention is a significant predictor of Vax_rate, as indicated by its estimate of 16.23 and its marginally significant p-value of 0.067.
-This means that there is some evidence that the intervention has an effect on the vaccination rate
-
-Male_Total is a significant negative predictor of Vax_rate, with a coefficient of -0.0133 and a significant p-value of 0.0054.
-This suggests that a higher male population is associated with a lower vaccination rate
-
-Among the age groups, Age_65Plus is a significant positive predictor of Vax_rate, with a coefficient of 0.0114 and a marginally significant p-value of 0.055.
-This means that a higher proportion of the population in the age group 65 and above is associated with a higher vaccination rate
-
-Overall, the model has a high adjusted R-squared value of 0.9995, indicating that the predictor variables account for a large proportion of the variation in Vax_rate.
-However, it is important to note that with such a high R-squared value, overfitting may be a concern, and it may be necessary to perform additional checks, such as cross-validation or regularization
 

### 2023-04-24 code/analysis_code/statisticalanalysis.qmd
Updates

--- 2023-04-10 code/analysis_code/statisticalanalysis.qmd
@@
 library(tidyverse)
-```
-
-### creating a data frame that contains the variables that will be included in new linear regression mode. Just copying and pasting all of the totals generated from the `exploratoryanalysis.qmd` file
-```{r}
-#combined all relevant variables into 5 different data sets for each zip code. Got the black population totals for each zip code from  https://www.unitedstateszipcodes.org/31701/
-
-Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Total = c(7797, 7780, 7755, 7773, 7792, 7817),
-                          Intervention = c(1, 1, 1, 1, 1, 1),
-                          Male_Total = c(3371, 3358, 3368, 3379, 3387, 3397),
-                          Female_Total = c(4426, 4422, 4387, 4394, 4405, 4420),
-                          Age_05_09 = c(123, 126, 126, 127, 127, 128),
-                          Age_10_17 = c(642, 635, 633, 637, 630, 629),
-                          Age_18_44 = c(3310, 3316, 3313, 3323, 3325, 3331),
-                          Age_45_64 = c(3229, 3244, 3239, 3259, 3264, 3258),
-                          Age_65Plus = c(2576, 2570, 2569, 2574, 2568, 2572),
-                          Pop = c(16475, 16475, 16475, 16475, 16475, 16475))
-
-Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                                Total = c(10383, 10327, 10303, 10301, 10337, 10348),                                 Intervention = c(1, 1, 1, 1, 1, 1),
-                          Male_Total = c(4394, 4397, 4399, 4391, 4403, 4399),
-                          Female_Total = c(5989, 5930, 5904, 5910, 5934, 5949),
-                          Age_05_09 = c(210, 211, 209, 201, 202, 202),
-                          Age_10_17 = c(1159, 1171, 1169, 1163, 1168, 1167),
-                          Age_18_44 = c(5201, 5215, 5239, 5263, 5278, 5281),
-                          Age_45_64 = c(4380, 4381, 4366, 4364, 4367, 4367),
-                          Age_65Plus = c(2983, 3000, 2997, 2997, 2994, 2989),
-                          Pop = c(24779, 24779, 24779, 24779, 24779, 24779))
-
-Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                                Total = c(9035, 8985, 8997, 9005, 9035, 9042),                                 Intervention = c(1, 1, 1, 1, 1, 1),
-                          Male_Total = c(3698, 3674, 3678, 3680, 3684, 3681),
-                          Female_Total = c(5337, 5311, 5319, 5325, 5351, 5361),
-                          Age_05_09 = c(275, 277, 280, 282, 284, 283),
-                          Age_10_17 = c(1109, 1120, 1125, 1134, 1136, 1142),
-                          Age_18_44 = c(5903, 5931, 5943, 5947, 5960, 5958),  
-                          Age_45_64 = c(5130, 5120, 5125, 5123, 5133, 5140),
-                          Age_65Plus = c(4200, 4204, 4195, 4206, 4202, 4189),
-                          Pop = c(13884, 13884, 13884, 13884, 13884, 13884))
-
-Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          Total = c(8758, 8688, 8639, 8593, 8623, 8666),                                      Intervention = c(0, 0, 0, 0, 0, 0),
-                          Male_Total = c(3526, 3514, 3503, 3484, 3489, 3498),
-                          Female_Total = c(5232, 5174, 5136, 5109, 5134, 5168),
-                          Age_05_09 = c(201, 213, 211, 215, 211, 211),
-                          Age_10_17 = c(1116, 1113, 1115, 1118, 1115, 1114),
-                          Age_18_44 = c(5001, 5009, 5020, 5026, 5025, 5032),
-                          Age_45_64 = c(5026, 5017, 5020, 5009, 5013, 5020),
-                          Age_65Plus = c(2816, 2825, 2827, 2814, 2828, 2832),
-                          Pop = c(23481, 23481, 23481, 23481, 23481, 23481))
-
-#um_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
-                          #Total = c(4956, 4917, 4888, 4851, 4866, 4894),                                      Zip = c(30907, 30907, 30907, 30907, 30907, 30907),
-                          #Male_Total = c(1947, 1943, 1926, 1900, 1913, 1919),
-                          #Female_Total = c(3009, 2974, 2962, 2951, 2953, 2975),
-                          #Age_05_09 = c(244, 239, 241, 236, 240, 242),
-                          #Age_10_17 = c(1344, 1358, 1357, 1341, 1343, 1335),
-                          #Age_18_44 = c(9203, 9217, 9238, 9245, 9253, 9246),
-                          #Age_45_64 = c(6740, 6700, 6667, 6626, 6617, 6624),
-                          #Age_65Plus = c(6632, 6633, 6639, 6635, 6633, 6623),
-                          #Pop = c(7744, 7744, 7744, 7744, 7744, 7744)) 
-#the 30907 zip code population of residents that are black does not match the vaccination data.
-also this zip code does not have a majority of residents that are black like other other 4 zip codes so I am going to leave it out of the statistical analysis
-
-#combining all of the zip codes into one data set except for the 30907 data 
-combined = rbind(Sum_by_month_31701, Sum_by_month_31705, Sum_by_month_31707, Sum_by_month_30815)
-```
-
-### creating a vaccination rate variable as the outcome 
-```{r}
-combined$Vax_rate = (combined$Total / combined$Pop) * 100
-```
-
-### Creating a linear regression model 
-```{r}
-lm_model2 <- lm(Vax_rate ~ Intervention + Male_Total + Female_Total + Age_05_09 + Age_10_17 + Age_18_44 + Age_45_64 + Age_65Plus, data = combined)
-
-summary(lm_model2)
-```
-The results indicate that Intervention is a significant predictor of Vax_rate, as indicated by its estimate of 16.23 and its marginally significant p-value of 0.067.
-This means that there is some evidence that the intervention has an effect on the vaccination rate
-
-Male_Total is a significant negative predictor of Vax_rate, with a coefficient of -0.0133 and a significant p-value of 0.0054.
-This suggests that a higher male population is associated with a lower vaccination rate
-
-Among the age groups, Age_65Plus is a significant positive predictor of Vax_rate, with a coefficient of 0.0114 and a marginally significant p-value of 0.055.
-This means that a higher proportion of the population in the age group 65 and above is associated with a higher vaccination rate
-
-Overall, the model has a high adjusted R-squared value of 0.9995, indicating that the predictor variables account for a large proportion of the variation in Vax_rate.
-However, it is important to note that with such a high R-squared value, overfitting may be a concern, and it may be necessary to perform additional checks, such as cross-validation or regularization
-
-### evaluating this model using RMSE 
-```{r}
-# Split the data into training and testing sets
-train <- combined[1:15,]
-test <- combined[16:24,]
-
-# Train the linear regression model
-model <- lm(Vax_rate ~ Intervention + Male_Total + Female_Total + 
-             Age_05_09 + Age_10_17 + Age_18_44 + Age_45_64 + Age_65Plus, 
-           data = train)
-
-# Generate predictions for the test data
-predictions <- predict(model, newdata = test)
-
-# Calculate RMSE
-rmse <- sqrt(mean((test$Vax_rate - predictions)^2))
-
-rmse
-```
-
-### Train a LASSO model 
-```{r}
-Vax_split <- initial_split(combined, strata = Intervention)
-Vax_train <- training(Vax_split)
-Vax_test <- testing(Vax_split)
+library(Metrics) #RMSE
+library(rsq) #rsqured 
+library(yardstick)
+library(MASS) #used for AIC forward selection 
+library(glmnet) #for ridge regression 
 ```
@@
 ```{r}
-Vax_rec <- recipe(Vax_rate ~ Intervention + Male_Total + Female_Total + 
-             Age_05_09 + Age_10_17 + Age_18_44 + Age_45_64 + Age_65Plus, data = Vax_train) %>%
-  step_zv(all_numeric(), -all_outcomes()) %>%
-  step_normalize(all_numeric(), -all_outcomes()) #centering and scaling data 
+
+Sum_by_month_31701 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          TOTAL_VAX_RATE = c(7797/12768, 7780/12768, 7755/12897, 7773/12897, 7792/12897, 7817/12897),
+                          Intervention = c(1, 1, 1, 1, 1, 1),
+                          Male_VAX_RATE = c(3371/5981, 3358/5981, 3368/5981, 3379/5981, 3387/5981, 3397/5981),
+                          Female_VAX_RATE = c(4426/6787, 4422/6787, 4387/6916, 4394/6916, 4405/6916, 4420/6916),
+                          Age_05_09_VAX_RATE = c(123/1058, 126/1058, 126/1058, 127/1058, 127/1058, 128/1058),
+                          Age_10_17_VAX_RATE = c(642/1560, 635/1560, 633/1560, 637/1560, 630/1560, 629/1560),
+                          Age_18_44_VAX_RATE = c(3310/5054, 3316/5054, 3313/5054, 3323/5054, 3325/5054, 3331/5054),
+                          Age_45_64_VAX_RATE = c(3229/4628, 3244/4628, 3239/4628, 3259/4628, 3264/4628, 3258/4628),
+                          Age_65Plus_VAX_RATE = c(2576/2886, 2570/2886, 2569/2886, 2574/2886, 2568/2886, 2572/2886),
+                          TOTAL_RATE_DIFF = c(0, 7780/12768 - 7797/12768, 7755/12897 - 7780/12768, 7773/12897 - 7755/12897, 7792/12897 - 7773/12897, 7817/12897 - 7792/12897), 
+                          POP_n = c(12768, 12768, 12897, 12897, 12897, 12897),
+                          Vac_n = c(7797, 7780, 7755, 7773, 7792, 7817))
+
+Sum_by_month_31705 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                                TOTAL_VAX_RATE = c(10383/21603, 10327/21572, 10303/21722, 10301/21887, 10337/21887, 10348/21887),                                 Intervention = c(1, 1, 1, 1, 1, 1),
+                          Male_VAX_RATE = c(4394/9885, 4397/9854, 4399/10004, 4391/10004, 4403/10004, 4399/10004),
+                          Female_VAX_RATE = c(5989/11718, 5930/11718, 5904/11718, 5910/11883, 5934/11883, 5949/11883),
+                          Age_05_09_VAX_RATE = c(210/1639, 211/1639, 209/1639, 201/1639, 202/1639, 202/1639),
+                          Age_10_17_VAX_RATE = c(1159/3311, 1171/3311, 1169/3311, 1163/3311, 1168/3311, 1167/3311),
+                          Age_18_44_VAX_RATE = c(5201/11978, 5215/11978, 5239/11978, 5263/11978, 5278/11978, 5281/11978),
+                          Age_45_64_VAX_RATE = c(4380/6468, 4381/6468, 4366/6468, 4364/6468, 4367/6468, 4367/6468),
+                          Age_65Plus_VAX_RATE = c(2983/3497, 3000/3497, 2997/3497, 2997/3497, 2994/3497, 2989/3497), 
+                          TOTAL_RATE_DIFF = c(0, 10327/21572 - 10383/21603, 10303/21722 - 10327/21572, 10301/21887 - 10303/21722, 10337/21887 - 10301/21887, 10348/21887 - 10337/21887), 
+                          POP_n = c(21603, 21572, 21722, 21887, 21887, 21887),
+                          Vac_n = c(10383, 10327, 10303, 10301, 10337, 10348))
+
+Sum_by_month_31707 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          TOTAL_VAX_RATE = c(9035/18045, 8985/18045, 8997/18045, 9005/18081, 9035/18081, 9042/18165),                                                       Intervention = c(1, 1, 1, 1, 1, 1),
+                          Male_VAX_RATE = c(3698/8232, 3674/8232, 3678/8232, 3680/8232, 3684/8232, 3681/8232),
+                          Female_VAX_RATE = c(5337/9813, 5311/9813, 5319/9813, 5325/9849, 5351/9849, 5361/9933),
+                          Age_05_09_VAX_RATE = c(275/1583, 277/1702, 280/1702, 282/1702, 284/1702, 283/1702),
+                          Age_10_17_VAX_RATE = c(1109/3001, 1120/3001, 1125/3001, 1134/3001, 1136/3001, 1142/3001),
+                          Age_18_44_VAX_RATE = c(5903/11434, 5931/11434, 5943/11434, 5947/11434, 5960/11434, 5958/11434),
+                          Age_45_64_VAX_RATE = c(5130/6051, 5120/6051, 5125/6051, 5123/6051, 5133/6051, 5140/6051),
+                          Age_65Plus_VAX_RATE = c(4200/4259, 4204/4259, 4195/4259, 4206/4259, 4202/4259, 4189/4259), 
+                          TOTAL_RATE_DIFF = c(0, 8985/18045 - 9035/18045, 8997/18045 - 8985/18045, 9005/18081 - 8997/18045, 9035/18081- 9005/18081, 9042/18165 - 9035/18081),
+                          POP_n = c(18045, 18045, 18045, 18081, 18081, 18165),
+                          Vac_n = c(9035, 8985, 8997, 9005, 9035, 9042))
+
+Sum_by_month_30815 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          TOTAL_VAX_RATE = c(8758/19606, 8688/19736, 8639/19736, 8593/19736, 8623/19736, 8666/19736),                                                       Intervention = c(0, 0, 0, 0, 0, 0),
+                          Male_VAX_RATE = c(3526/9066, 3514/9196, 3503/9196, 3484/9196, 3489/9196, 3498/9196),
+                          Female_VAX_RATE = c(5232/10540, 5174/10540, 5136/10540, 5109/10540, 5134/10540, 5168/10540),
+                          Age_05_09_VAX_RATE = c(201/2388, 213/2388, 211/2480, 215/2480, 211/2480, 211/2480),
+                          Age_10_17_VAX_RATE = c(1116/3604, 1113/3604, 1115/3604, 1118/3604, 1115/3604, 1114/3604),
+                          Age_18_44_VAX_RATE = c(5001/13358, 5009/13358, 5020/13358, 5026/13358, 5025/13358, 5032/13358),
+                          Age_45_64_VAX_RATE = c(5026/8431, 5017/8431, 5020/8431, 5009/8431, 5013/8431, 5020/8431),
+                          Age_65Plus_VAX_RATE = c(2816/3590, 2825/3590, 2827/3590, 2814/3590, 2828/3590, 2832/3590),
+                          TOTAL_RATE_DIFF = c(0, 8688/19736 - 8758/19606, 8639/19736 - 8688/19736, 8593/19736 - 8639/19736, 8623/19736- 8593/19736, 8666/19736 - 8623/19736),
+                          POP_n = c(19606, 19736, 19736, 19736, 19736, 19736),
+                          Vac_n = c(8758, 8688, 8639, 8593, 8623, 8666))
+
+Sum_by_month_30907 = data.frame(Date = c("Sep","Oct","Nov","Dec","Jan","Feb"),
+                          TOTAL_VAX_RATE = c(4956/9122, 4917/9189, 4888/9189, 4851/9189, 4866/9189, 4894/9189),                                                                      Intervention = c(0, 0, 0, 0, 0, 0),
+                          Male_VAX_RATE = c(1947/4289, 1943/4289, 1926/4289, 1900/4289, 1913/4289, 1919/4289),
+                          Female_VAX_RATE = c(3009/4833, 2974/4900, 2962/4900, 2951/4900, 2953/4900, 2975/4900),
+                          Age_05_09_VAX_RATE = c(244/2174, 239/2174, 241/2174, 236/2174, 240/2174, 242/2174),
+                          Age_10_17_VAX_RATE = c(1344/4070, 1358/4070, 1357/4070, 1341/4070, 1343/4070, 1335/4070),
+                          Age_18_44_VAX_RATE = c(9203/14571, 9217/14571, 9238/14571, 9245/14571, 9253/14571, 9246/14571),
+                          Age_45_64_VAX_RATE = c(6740/10362, 6700/10362, 6667/10362, 6626/10362, 6617/10362, 6624/10362),
+                          Age_65Plus_VAX_RATE = c(6632/6995, 6633/6995, 6639/6995, 6635/6995, 6633/6995, 6623/6995), 
+                          TOTAL_RATE_DIFF = c(0, 4917/9189 - 4956/9122, 4888/9189 - 4917/9189, 4851/9189 - 4888/9189, 4866/9189 - 4851/9189, 4894/9189 - 4866/9189),
+                          POP_n = c(9122, 9189, 9189, 9189, 9189, 9189),
+                          Vac_n = c(4956, 4917, 4888, 4851, 4866, 4894))
+
+#combining all of the zip codes into one data 
+combined = rbind(Sum_by_month_31701, Sum_by_month_31705, Sum_by_month_31707, Sum_by_month_30815,Sum_by_month_30907)
+
+#multiply all rates by 100 
+combined <- combined %>% 
+  mutate(TOTAL_VAX_RATE = TOTAL_VAX_RATE * 100,
+         Male_VAX_RATE = Male_VAX_RATE * 100,
+         Female_VAX_RATE = Female_VAX_RATE * 100,
+         Age_05_09_VAX_RATE = Age_05_09_VAX_RATE * 100,
+         Age_10_17_VAX_RATE = Age_10_17_VAX_RATE * 100,
+         Age_18_44_VAX_RATE = Age_18_44_VAX_RATE * 100,
+         Age_45_64_VAX_RATE = Age_45_64_VAX_RATE * 100,
+         Age_65Plus_VAX_RATE = Age_65Plus_VAX_RATE * 100,
+         TOTAL_RATE_DIFF = TOTAL_RATE_DIFF * 100)
+
 ```
 
+### Data splitting
+
 ```{r}
-lasso_Vax <- linear_reg(penalty = 0.1, mixture = 1) %>%
-  set_engine("glmnet")
-
-wf <- workflow() %>%
-  add_recipe(Vax_rec)
-
-lasso_fit <- wf %>%
-  add_model(lasso_Vax) %>%
-  fit(data = Vax_train)
-
-lasso_fit %>%
-  pull_workflow_fit() %>%
-  tidy()
+data_split = initial_split(combined, prop = 7/10, strata = TOTAL_RATE_DIFF)
 ```
 
-### Tuning LASSO Parameters
+### Splitting the data between training and testing
+
 ```{r}
-set.seed(1234)
-Vax_boot <- bootstraps(Vax_train, strata = Intervention)
-
-tune_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
-  set_engine("glmnet")
-
-lambda_grid <- grid_regular(penalty(), levels = 50)
+train_data = training(data_split)
+test_data  = testing(data_split)
 ```
 
-### Tune the grid
+### Doing 5-fold cross validation with 5 repeats and body temperature as the strata
+
 ```{r}
-set.seed(2020)
-lasso_grid <- tune_grid(
-  wf %>% add_model(tune_spec),
-  resamples = Vax_boot,
-  grid = lambda_grid
-)
+CV_fold = vfold_cv(train_data, v = 2, repeats = 5, strata = TOTAL_RATE_DIFF) 
 ```
 
-### Results 
+### Creating a recipe and applying to the training data
+
 ```{r}
-lasso_grid %>%
-  collect_metrics()
+recipe <- recipe(TOTAL_RATE_DIFF ~ TOTAL_VAX_RATE + Intervention + Male_VAX_RATE + Female_VAX_RATE + Age_05_09_VAX_RATE + Age_10_17_VAX_RATE + Age_18_44_VAX_RATE + Age_45_64_VAX_RATE + Age_65Plus_VAX_RATE + Date + POP_n + Vac_n, data = train_data) %>%
+  step_dummy(all_nominal(), one_hot = TRUE) #In this code, we create a recipe using step_dummy to preprocess the data and convert the categorical predictors into numeric predictors
 ```
 
-### Visualizing the LASSO results 
+## Creating a mean null model and evaluating the model for both the training data and test data
+
 ```{r}
-lasso_grid %>%
-  collect_metrics()
+mean_null_model <- function(combined) {
+  outcome <- mean(combined$TOTAL_RATE_DIFF)
+  prediction <- rep(outcome, nrow(combined))
+  return(prediction)}
 
-lasso_grid_plot = lasso_grid %>%
-  collect_metrics() %>%
-  ggplot(aes(penalty, mean, color = .metric)) +
-  geom_errorbar(aes(
-    ymin = mean - std_err,
-    ymax = mean + std_err
-  ),
-  alpha = 0.5
-  ) +
-  geom_line(size = 1.5) +
-  facet_wrap(~.metric, scales = "free", nrow = 2) +
-  scale_x_log10() +
-  theme(legend.position = "none")
+#using model to make predictions based on the training data and test data
+train_pred <- mean_null_model(train_data)
+test_pred <- mean_null_model(test_data)
 
-lasso_grid_plot
+#calculating rmse for the model fit to training data 
+train_rmse <- rmse(train_data$TOTAL_RATE_DIFF, train_pred)
+test_rmse <- rmse(test_data$TOTAL_RATE_DIFF, test_pred)
+
+#taking a look at the rmse for both training and test data
+train_rmse
+test_rmse
 ```
 
-### Save plot
+## Creating a tree model and evaluating the model 
+### Tuning Hyperparameters
+
 ```{r}
-#save figure
-ggsave(filename = here("results","lasso_grid_plot.png"), plot=lasso_grid_plot)
+#tuning: the process of estimating the best values for these values by training many models on resampled data sets and exploring how well all these models perform
+tune_spec <- 
+  decision_tree(
+    cost_complexity = tune(),
+    tree_depth = tune()
+  ) %>% 
+  set_engine("rpart") %>% 
+  set_mode("regression")
 ```
 
-### Lets pick the LASSO model with the lowest RMSE 
+### Creating a grid of values
+
 ```{r}
-lowest_rmse <- lasso_grid %>%
-  select_best("rmse", maximize = FALSE)
-
-final_lasso <- finalize_workflow(
-  wf %>% add_model(tune_spec),
-  lowest_rmse
-)
+tree_grid <- grid_regular(cost_complexity(),
+                          tree_depth(),
+                          levels = 5)
 ```
 
-### Final fit and viewing most important variables 
+### Setting up a tree workflow
+
 ```{r}
-final_lasso_plot = final_lasso %>%
-  fit(Vax_train) %>%
-  pull_workflow_fit() %>%
-  vi(lambda = lowest_rmse$penalty) %>%
-  mutate(
-    Importance = abs(Importance),
-    Variable = fct_reorder(Variable, Importance)
-  ) %>%
-  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
-  geom_col() +
-  scale_x_continuous(expand = c(0, 0)) +
-  labs(y = NULL)
-final_lasso_plot
+tree_wf <- workflow() %>%
+  add_model(tune_spec) %>%
+  add_recipe(recipe)
+
+#getting turning results 
+tree_res <- 
+  tree_wf %>% 
+  tune_grid(
+    resamples = CV_fold,
+    grid = tree_grid
+    )
+
+tree_res %>%  collect_metrics()
 ```
 
-### Looking at final fit 
+### Plotting results of the tree model
+
 ```{r}
-last_fit(
-  final_lasso,
-  Vax_split
-) %>%
-  collect_metrics()
+tree_res %>%
+  collect_metrics() %>%
+  mutate(tree_depth = factor(tree_depth)) %>%
+  ggplot(aes(cost_complexity, mean, color = tree_depth)) +
+  geom_line(linewidth = 1.5, alpha = 0.6) +
+  geom_point(size = 2) +
+  facet_wrap(~ .metric, scales = "free", nrow = 2) +
+  scale_x_log10(labels = scales::label_number()) +
+  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)
 ```
 
-### Save plot
+### Taking a look at the best tree models based on the R\^2 values
+
 ```{r}
-#save figure
-ggsave(filename = here("results","final_lasso_plot.png"), plot=final_lasso_plot)
+tree_res %>%
+  show_best("rmse")
+
+best_tree = tree_res %>%
+  select_best("rmse")
+
+best_tree
 ```
+
+### Finalizing model
+
+```{r}
+final_wf <- 
+  tree_wf %>% 
+  finalize_workflow(best_tree)
+
+final_wf
+```
+
+### Evaluate the final fit for the tree model
+
+```{r}
+final_fit <- 
+  final_wf %>%
+  last_fit(data_split)
+
+final_fit_metrics <- final_fit %>%
+  collect_metrics()
+#final fit metrics
+final_fit_metrics
+
+#creating a tibble for predicted values of model
+predictions = final_fit  %>%
+  collect_predictions
+
+#plotting model predictions vs actual outcome 
+plot1 = final_fit %>%
+  collect_predictions  %>%
+    ggplot(aes(x = .pred, y = TOTAL_RATE_DIFF)) +
+  geom_smooth(method = "lm", se = FALSE) +
+  xlab("Predicted Probability") +
+  ylab("Actual Class") +
+  ggtitle("Predicted Values vs Actual Outcome Values")
+plot1
+
+#plotting the residuals 
+residuals = predictions$TOTAL_RATE_DIFF - predictions$.pred
+
+#creating a residual data frame 
+residuals_df = data.frame(TOTAL_RATE_DIFF = predictions$TOTAL_RATE_DIFF, Predictions = predictions$.pred, Residuals = residuals)
+
+# Create a scatter plot with ggplot for residuals 
+ggplot(residuals_df, aes(x = Predictions , y = Residuals)) +
+  geom_point() +
+  xlab("Predicted values") +
+  ylab("Residuals") +
+  ggtitle("Residual Plot")
+
+```
+
+## Fitting for LASSO model
+
+### Building a model
+
+```{r}
+lm_mod =
+  linear_reg(penalty = tune(), mixture =1) %>%
+  set_engine("glmnet")
+```
+
+### Creating a workflow
+
+```{r}
+lm_workflow =
+  workflow() %>%
+  add_model(lm_mod)  %>%
+  add_recipe(recipe)
+```
+
+### Creating a grid for tuning
+
+```{r}
+lm_reg_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))
+
+# lowest penalty values
+lm_reg_grid %>% top_n(-5)
+
+# highest penalty values 
+lm_reg_grid %>% top_n(5)
+
+```
+
+### Train and tune the model
+
+```{r}
+lasso_res <- lm_workflow %>%
+  tune_grid(resamples = CV_fold, grid = lm_reg_grid)
+```
+
+### Collect metrics and picking the best model
+
+```{r}
+lasso_res %>% show_best("rmse")
+
+best_lasso = lasso_res %>%
+  select_best("rmse")
+best_lasso
+```
+
+### Finalizing Lasso model
+
+```{r}
+final_lasso_wf <- 
+  lm_workflow %>% 
+  finalize_workflow(best_lasso)
+
+final_lasso_wf
+```

### 2023-04-24 products/Manuscript.qmd
Updates

@@
 # Summary
-
-Back in March 2020, COVID-19 began to disproportionately impact the citizens of the United States.
-Among African Americans sub-population in rural South Georgia, there was a call to increase access to health services, promote health literacy as it relates to COVID-19 and decrease vaccine hesitancy.
-One such program includes the partnership between the City of Albany, GA and Albany State University to reduce COVID-19 vaccine hesitancy in Albany through a $3.9 million grant from Health and Human Services and the Office of Minority Health.
-
-This study aims to conduct a secondary analysis of residents in Albany, GA that have received at least one dose of the COVID-19 vaccine and live in the one of the three zip codes 31701, 31705 and 31707.
+This study aims to conduct a secondary analysis of residents of the black community in Albany, GA that have received at least one dose of the COVID-19 vaccine from September 2022 through February 2023 and live in the one of the three zip codes 31701, 31705 and 31707.
+The outcome of interest is the difference in vaccination rates by month from September through February.
+The predictor of interest is the presence of a COVID-19 Vaccine Hesitancy Program Intervention.
 
@@
 ## General Background
+Back in March 2020, COVID-19 greatly impacted the citizens of the Albany, GA.
+Early on in the COVID-19 pandemic Albany, GA was ranked the "forth-worst" city hit by COVID-19 in the United States [@grahamrapierHowSmallGeorgia2020].
+According to U.S Census Bureau, 74.5% of the total population of the City of Albany is made up by residents that are black.
+In a study that examined the health outcomes of residents who tested positive for COVID-19 in Albany, GA it was found that 83.3% of 710 patients that were hospitalized for COVID-19 were black[@].
+Although there was no differences in hospital mortality found between residents that were black compared to residents that were not black but there was a much higher percentage of hospitalizations due to COVID-19 for residents who were black.
+
+In July 2022, Albany was awarded a 3.9 million dollar grant from the Health and Human Services and the Office of Minority Health in an effort to fund COVID-19 vaccine hesitancy and health literacy programs[].
+The Economic Evaluation Research Group (EERG) out of the University of Georgia College of Public Health Department of Health Policy and Management led by Dr.
+Janani Rajbhandari-Thapa is currently conducting an evaluation on the federally funded programs.
+
+This study serves an extension of the evaluation being conducted by EERG through examining the vaccination rates of the zip codes where vaccine hesitancy programs are being conducted.
 
@@
 
-My goal is to compare monthly trends of vaccinations and examine whether there is an increase in vaccinations among African American residents during the time of the health literacy program, which is from September through April.
+My goal is to compare monthly trends of vaccinations and examine whether there is a noticable change in vaccination rates among residents who are black during the time of the health literacy programs, which is from September through February.
 
@@
 
-In preparation for analysis, the raw data was first sub-setted by  Dougherty County, the county in which Albany, Georgia lies.
+In preparation for analysis, the raw data was first subsetted by Dougherty County, the county in which Albany, Georgia lies.
 This allowed me to work only with census tracts from Dougherty county.
-I then summed the data by demographic to get demographic totals.
+I then stratified the data by black residents.
+Of black residents, I summed the vaccination totals and populations by gender, age for each intervention zip code (31701, 31705 and 31707) and the two control zip codes.
 

### 2023-04-24 products/Manuscript.qmd
Updates

@@
 ## Vaccination Trends of the Black Community for Each Zip Code
+
+The was no noticiable increase of vaccination rate for
+
 ```{r}
-knitr::include_graphics("../results/31701_total_black_vax.png")
+knitr::include_graphics("../results/plot_31701_total_black_vax_rate.png")
 ```

### 2023-04-24 code/analysis_code/exploratoryanalysis.qmd
Updates

@@
 #save figure
-ggsave(filename = here("results","plot_31701_total_black_vax_rate.png"), plot=plot_31701_total_black_vax_rate)
+ggsave(filename = here("results","31701_total_black_vax_rate.png"), plot=plot_31701_total_black_vax_rate)
 ```
@@
 #save figure
-ggsave(filename = here("results","plot_31705_total_black_vax_rate.png"), plot=plot_31705_total_black_vax_rate)
+ggsave(filename = here("results","31705_total_black_vax_rate.png"), plot=plot_31705_total_black_vax_rate)
 ```
@@
 #save figure
-ggsave(filename = here("results","plot_31707_total_black_vax_rate.png"), plot=plot_31707_total_black_vax_rate)
+ggsave(filename = here("results","31707_total_black_vax_rate.png"), plot=plot_31707_total_black_vax_rate)
 ```
@@
 #save figure
-ggsave(filename = here("results","plot_30815_total_black_vax_rate.png"), plot=plot_30815_total_black_vax_rate)
+ggsave(filename = here("results","30815_total_black_vax_rate.png"), plot=plot_30815_total_black_vax_rate)
 ```
@@
 #save figure
-ggsave(filename = here("results","plot_30907_total_black_vax_rate.png"), plot=plot_30907_total_black_vax_rate)
+ggsave(filename = here("results","30907_total_black_vax_rate.png"), plot=plot_30907_total_black_vax_rate)
 ```

### 2023-04-24 products/Manuscript.qmd
Updates

@@
 
-{{< pagebreak >}}
-
 # Introduction
@@
 Back in March 2020, COVID-19 greatly impacted the citizens of the Albany, GA.
-Early on in the COVID-19 pandemic Albany, GA was ranked the "forth-worst" city hit by COVID-19 in the United States [@grahamrapierHowSmallGeorgia2020].
+Early on in the COVID-19 pandemic Albany, GA was ranked the "forth-worst" city hit by COVID-19 in the United States [@Rapier2020].
 According to U.S Census Bureau, 74.5% of the total population of the City of Albany is made up by residents that are black.
-In a study that examined the health outcomes of residents who tested positive for COVID-19 in Albany, GA it was found that 83.3% of 710 patients that were hospitalized for COVID-19 were black[@].
+In a study that examined the health outcomes of residents who tested positive for COVID-19 in Albany, GA it was found that 83.3% of 710 patients that were hospitalized for COVID-19 were black[@Racine2022].
 Although there was no differences in hospital mortality found between residents that were black compared to residents that were not black but there was a much higher percentage of hospitalizations due to COVID-19 for residents who were black.
 
-In July 2022, Albany was awarded a 3.9 million dollar grant from the Health and Human Services and the Office of Minority Health in an effort to fund COVID-19 vaccine hesitancy and health literacy programs[].
+In July 2022, Albany was awarded a 3.9 million dollar grant from the Health and Human Services and the Office of Minority Health in an effort to fund COVID-19 vaccine hesitancy and health literacy programs[@TheAlbanyHerald].
 The Economic Evaluation Research Group (EERG) out of the University of Georgia College of Public Health Department of Health Policy and Management led by Dr.
@@
 
-The was no noticiable increase of vaccination rate for
+### Zip code 31701
+The vaccination rate of the the Black population slightly decreased from a rate of 61% to 60.6% \@ref(fig:31701_total_black_vax_rate).
+The vaccination rate for female black residents slightly decreased from 65% to 64% while the male black resident population's vaccination rate increased slightly at 53% to 57%.
+It is important to note the difference between the vaccination rates of female and male black residents, which is around .1 from September to February \@ref(fig:31701_Gender_black_vax).
+When the black population was stratified by age, the rates were the highest for the ages of 65 and older at 89%.
+Black residents aged 45-64 had a steady rate of 70%, black residents aged 18-44 had a steady rate of 65%, blacked residents aged 10-17 had a steady rate of 40% and black residents aged 5-9 had a steady rate of 12% \@ref(fig:31701_Age_black_vax).
 
-```{r}
-knitr::include_graphics("../results/plot_31701_total_black_vax_rate.png")
+```{r 31701_total_black_vax_rate,  fig.cap='\\label{fig::31701_total_black_vax_rate}Vaccination Rates for Black Population in the 31701 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/31701_total_black_vax_rate.png")
 ```
 
-```{r}
-knitr::include_graphics("../results/31705_total_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/31707_total_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/30815_total_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/30907_total_black_vax.png")
-```
-
-## Vaccination Trends of the Black Community for Each Zip Code by Gender
-```{r}
+```{r 31701_Gender_black_vax,  fig.cap='\\label{fig::31701_Gender_black_vax}Vaccination Rates for Black Population by Gender in the 31701 Zipcode', echo=FALSE}
 knitr::include_graphics("../results/31701_Gender_black_vax.png")
@@
 
-```{r}
+```{r 31701_Age_black_vax.png,  fig.cap='\\label{fig::31701_Age_black_vax}Vaccination Rates for Black Population by Age in the 31701 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/31701_Age_black_vax.png")
+```
+
+### Zip code 31705
+The vaccination rate of stayed stable around 48% for the black population in the 31705 zip code \@ref(fig:31701_Age_black_vax).
+The vaccination rate of female black residents dropped from 51.1% to 49.7% from September to December, but remained stable close to 50% until February.
+For male black residents the vaccination rate hovered around 44% \@ref(fig:31705_Gender_black_vax).
+After stratifying by age the vaccination rates for the 65+, 44-64, 18-44, 10-17 and 5-9 ages groups were as follows: 85%, 68%, 43%, 35% and 12% \@ref(fig:31705_Age_black_vax).
+
+```{r 31705_total_black_vax_rate,  fig.cap='\\label{fig::31705_total_black_vax_rate}Vaccination Rates for Black Population in the 31705 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/31705_total_black_vax_rate.png")
+```
+
+```{r 31705_Gender_black_vax,  fig.cap='\\label{fig::31705_Gender_black_vax}Vaccination Rates for Black Population by Gender in the 31705 Zipcode', echo=FALSE}
 knitr::include_graphics("../results/31705_Gender_black_vax.png")
@@
 
-```{r}
+```{r 31705_Age_black_vax,  fig.cap='\\label{fig::31705_Age_black_vax}Vaccination Rates for Black Population by Age in the 31705 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/31705_Age_black_vax.png")
+```
+
+### Zip code 31707
+The total vaccination rate of black residents in the 31707 zip code was around 50% from September through February \@ref(fig:31707_total_black_vax_rate).
+The vaccination rates for the female and male black population were found to be steady around 54% and 44% \@ref(fig:31707_Gender_black_vax).
+The vaccination rates by the age groups 65+, 44-64, 18-44, 10-17 and 5-9 where found to be 98%, 85%, 52%, 37%, and 17% \@ref(fig:31707_Age_black_vax).
+
+```{r 31707_total_black_vax_rate,  fig.cap='\\label{fig::31707_total_black_vax_rate}Vaccination Rates for Black Population in the 31707 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/31707_total_black_vax_rate.png")
+```
+
+```{r 31707_Gender_black_vax,  fig.cap='\\label{fig::31707_Gender_black_vax}Vaccination Rates for Black Population by Gender in the 31707 Zipcode', echo=FALSE}
 knitr::include_graphics("../results/31707_Gender_black_vax.png")
@@
 
-```{r}
+```{r 31707_Age_black_vax,  fig.cap='\\label{fig::31707_Age_black_vax}Vaccination Rates for Black Population by Age in the 31707 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/31707_Age_black_vax.png")
+```
+
+### Zip code 30815
+The total vaccination rate for the black population was 44% in the 30815 zip code from September through February \@ref(fig:30815_total_black_vax_rate).
+The vaccination rates for black female and male residents were 49% and 38% \@ref(fig:30815_Gender_black_vax).
+The vaccination rates by the age groups 65+, 44-64, 18-44, 10-17 and 5-9 where found to be 78%, 60%, 38%, 31%, and 8% \@ref(fig:30815_Age_black_vax).
+
+```{r 30815_total_black_vax_rate,  fig.cap='\\label{fig::30815_total_black_vax_rate}Vaccination Rates for Black Population in the 30815 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/30815_total_black_vax_rate.png")
+```
+
+```{r 30815_Gender_black_vax,  fig.cap='\\label{fig::30815_Gender_black_vax}Vaccination Rates for Black Population by Gender in the 30815 Zipcode', echo=FALSE}
 knitr::include_graphics("../results/30815_Gender_black_vax.png")
@@
 
-```{r}
+```{r 30815_Age_black_vax,  fig.cap='\\label{fig::30815_Age_black_vax}Vaccination Rates for Black Population by Age in the 30815 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/30815_Age_black_vax.png")
+```
+
+### Zip code 30907
+The total vaccination rate for the black population was 54% in September and dropped to around 53% in February in the 30907 zip code \@ref(fig:30907_total_black_vax_rate).
+The vaccination rates for black female and male residents were 60% and 45% \@ref(fig:30907_Gender_black_vax).
+The vaccination rates by the age groups 65+, 44-64, 18-44, 10-17 and 5-9 where found to be 95%, 64%, 63%, 33%, and 11% \@ref(fig:30907_Age_black_vax).
+
+```{r 30907_total_black_vax_rate,  fig.cap='\\label{fig::30907_total_black_vax_rate}Vaccination Rates for Black Population in the 30907 Zipcode', echo=FALSE}
+knitr::include_graphics("../results/30907_total_black_vax_rate.png")
+```
+
+```{r 30907_Gender_black_vax,  fig.cap='\\label{fig::30907_Gender_black_vax}Vaccination Rates for Black Population by Gender in the 30907 Zipcode', echo=FALSE}
 knitr::include_graphics("../results/30907_Gender_black_vax.png")
@@
 
-## Vaccination Trends of the Black Community for Each Zip Code by Gender
-```{r}
-knitr::include_graphics("../results/31701_Age_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/31705_Age_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/31707_Age_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/30815_Age_black_vax.png")
-```
-
-```{r}
+```{r 30907_Age_black_vax,  fig.cap='\\label{fig::30907_Age_black_vax}Vaccination Rates for Black Population by Age in the 30907 Zipcode', echo=FALSE}
 knitr::include_graphics("../results/30907_Age_black_vax.png")
@@
 
-## LASSO Results Visual 
-```{r}
-knitr::include_graphics("../results/lasso_grid_plot.png")
-```
-
-## Final LASSO Results Variable Importance 
-```{r}
-knitr::include_graphics("../results/final_lasso_plot.png")
-```
+### Statistical Analysis
 
@@
 # Discussion
+Albany, Georgia is predominately black

### 2023-04-24 products/supplement.qmd
Updates

@@
 
+            -   processed_POP_data_31701_Sept.rds`: Contains processed data for the 31701 zip code Total Black Population during September, 2022
+
+            -   processed_POP_data_31701_Oct.rds`: Contains processed data for the 31701 zip code Total Black Population during October, 2022
+
+            -   processed_POP_data_31701_Nov.rds`: Contains processed data for the 31701 zip code Total Black Population during November, 2022
+
+            -   processed_POP_data_31701_Dec.rds`: Contains processed data for the 31701 zip code Total Black Population during December, 2022
+
+            -   processed_POP_data_31701_Jan.rds`: Contains processed data for the 31701 zip code Total Black Population during January, 2023
+
+            -   processed_POP_data_31701_Feb.rds`: Contains processed data for the 31701 zip code Total Black Population during February, 2023
+
+            -   `processed_POP_combined_31701.rds`: Contains Total Population with all variables by month (September-February) for African Americans in the 31701 zip
+
         -   `Zip_31705` folder: Contains processed data for the 31705 zip code
@@
 
+            -   processed_POP_data_31705_Sept.rds`: Contains processed data for the 31705 zip code Total Black Population during September, 2022
+
+            -   processed_POP_data_31705_Oct.rds`: Contains processed data for the 31705 zip code Total Black Population during October, 2022
+
+            -   processed_POP_data_31705_Nov.rds`: Contains processed data for the 31705 zip code Total Black Population during November, 2022
+
+            -   processed_POP_data_31705_Dec.rds`: Contains processed data for the 31705 zip code Total Black Population during December, 2022
+
+            -   processed_POP_data_31705_Jan.rds`: Contains processed data for the 31705 zip code Total Black Population during January, 2023
+
+            -   processed_POP_data_31705_Feb.rds`: Contains processed data for the 31705 zip code Total Black Population during February, 2023
+
+            -   `processed_POP_combined_31705.rds`: Contains Total Population with all variables by month (September-February) for African Americans in the 31705 zip
+
         -   `Zip_31707` folder: Contains processed data for the 31707 zip code
@@
             -   `processed_data_31707_Feb.rds`: Contains processed data for the 31707 zip code during February, 2023
+
+            -   processed_POP_data_31707_Sept.rds`: Contains processed data for the 31707 zip code Total Black Population during September, 2022
+
+            -   processed_POP_data_31707_Oct.rds`: Contains processed data for the 31707 zip code Total Black Population during October, 2022
+
+            -   processed_POP_data_31707_Nov.rds`: Contains processed data for the 31707 zip code Total Black Population during November, 2022
+
+            -   processed_POP_data_31707_Dec.rds`: Contains processed data for the 31707 zip code Total Black Population during December, 2022
+
+            -   processed_POP_data_31707_Jan.rds`: Contains processed data for the 31707 zip code Total Black Population during January, 2023
+
+            -   processed_POP_data_31707_Feb.rds`: Contains processed data for the 31707 zip code Total Black Population during February, 2023
+
+            -   `processed_POP_combined_31707.rds`: Contains Total Population with all variables by month (September-February) for African Americans in the 31707 zip
+
+        -   `Zip_30815` folder: Contains processed data for the 31707 zip code
+            -   `processed_data_30815_Sept.rds`: Contains processed data for the 30815 zip code during September, 2022
+
+            -   `processed_data_30815_Oct.rds`: Contains processed data for the 30815 zip code during October, 2022
+
+            -   `processed_data_30815_Nov.rds`: Contains processed data for the 30815 zip code during November, 2022
+
+            -   `processed_data_30815_Dec.rds`: Contains processed data for the 30815 zip code during December, 2022
+
+            -   `processed_data_30815_Jan.rds`: Contains processed data for the 30815 zip code during January, 2023
+
+            -   `processed_data_30815_Feb.rds`: Contains processed data for the 30815 zip code during February, 2023
+
+            -   processed_POP_data_30815_Sept.rds`: Contains processed data for the 30815 zip code Total Black Population during September, 2022
+
+            -   processed_POP_data_30815_Oct.rds`: Contains processed data for the 30815 zip code Total Black Population during October, 2022
+
+            -   processed_POP_data_30815_Nov.rds`: Contains processed data for the 30815 zip code Total Black Population during November, 2022
+
+            -   processed_POP_data_30815_Dec.rds`: Contains processed data for the 30815 zip code Total Black Population during December, 2022
+
+            -   processed_POP_data_30815_Jan.rds`: Contains processed data for the 30815 zip code Total Black Population during January, 2023
+
+            -   processed_POP_data_30815_Feb.rds`: Contains processed data for the 30815 zip code Total Black Population during February, 2023
+
+        -   `Zip_30907` folder: Contains processed data for the 31707 zip code
+
+            -   `processed_data_30907_Sept.rds`: Contains processed data for the 30907 zip code during September, 2022
+
+            -   `processed_data_30907_Oct.rds`: Contains processed data for the 30907 zip code during October, 2022
+
+            -   `processed_data_30907_Nov.rds`: Contains processed data for the 30907 zip code during November, 2022
+
+            -   `processed_data_30907_Dec.rds`: Contains processed data for the 30907 zip code during December, 2022
+
+            -   `processed_data_30907_Jan.rds`: Contains processed data for the 30907 zip code during January, 2023
+
+            -   `processed_data_30907_Feb.rds`: Contains processed data for the 30907 zip code during February, 2023
+            -   processed_POP_data_30907_Sept.rds`: Contains processed data for the 30907 zip code Total Black Population during September, 2022
+
+            -   processed_POP_data_30907_Oct.rds`: Contains processed data for the 30907 zip code Total Black Population during October, 2022
+
+            -   processed_POP_data_30907_Nov.rds`: Contains processed data for the 30907 zip code Total Black Population during November, 2022
+
+            -   processed_POP_data_30907_Dec.rds`: Contains processed data for the 30907 zip code Total Black Population during December, 2022
+
+            -   processed_POP_data_30907_Jan.rds`: Contains processed data for the 30907 zip code Total Black Population during January, 2023
+
+            -   processed_POP_data_30907_Feb.rds`: Contains processed data for the 30907 zip code Total Black Population during February, 2023
 
@@
 
-        -   `processingcode.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 that was downloaded in September 2022
+        -   `processingcode.qmd`: processing code for COVID-19 vaccination data in the three target zip codes 31701, 31705 and 31707 and the two control zip codes 30815 and 30907
+
+        -   `processingcode2.qmd`: processing code for COVID-19 Population data in the three target zip codes 31701, 31705 and 31707 and the two control zip codes 30815 and 30907
 
@@
 
-    -   `Black_Vax_Total.png`: Contains a line graph with all three zip codes plotted for the African American populations
+    -   `31701_total_black_vax_rate.png` : A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31701 zip code
 
-    -   `Black_Vax_Total_31701.png` : A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31701 zip code
+    -   `31705_total_black_vax_rate.png` : A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31705 zip code
 
-    -   `Black_Vax_Total_31705.png` : A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31705 zip code
+    -   `31707_total_black_vax_rate.png`: A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31707 zip code
 
-    -   `Black_Vax_Total_31707.png`: A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 31707 zip code
+    -   `30815_total_black_vax_rate.png` : A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 30815 zip code
+
+    -   `30907_total_black_vax_rate.png` : A line graph of African Americans with at least one dose of the COVID-19 vaccine in the 30907 zip code
+
+    -   `31701_Gender_black_vax.png`: A line graph of African Americans by Gender with at least one dose of the COVID-19 vaccine in the 31701 zip code
+
+    -   `31705_Gender_black_vax.png`: A line graph of African Americans by Gender with at least one dose of the COVID-19 vaccine in the 31705 zip code
+
+    -   `31707_Gender_black_vax.png`: A line graph of African Americans by Gender with at least one dose of the COVID-19 vaccine in the 31707 zip code
+
+    -   `30815_Gender_black_vax.png`: A line graph of African Americans by Gender with at least one dose of the COVID-19 vaccine in the 30815 zip code
+
+    -   `30907_Gender_black_vax.png`: A line graph of African Americans by Gender with at least one dose of the COVID-19 vaccine in the 30907 zip code
+
+    -   `31701_Age_black_vax.png`: A line graph of African Americans by Age with at least one dose of the COVID-19 vaccine in the 31701 zip code
+
+    -   `31705_Age_black_vax.png`: A line graph of African Americans by Age with at least one dose of the COVID-19 vaccine in the 31705 zip code
+
+    -   `31707_Age_black_vax.png`: A line graph of African Americans by Age with at least one dose of the COVID-19 vaccine in the 31707 zip code
+
+    -   `30815_Age_black_vax.png`: A line graph of African Americans by Age with at least one dose of the COVID-19 vaccine in the 30815 zip code
+
+    -   `30907_Age_black_vax.png`: A line graph of African Americans by Age with at least one dose of the COVID-19 vaccine in the 30907 zip code

### 2023-04-27 Hatfield Review.qmd
Kelly Review

---
title: Project Review  
author: Kelly Hatfield
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project:  COVID-19 Vaccination in Albany, GA

Name of project author(s): Jacob Matta

Name of project reviewer: Kelly Hatfield

### Feedback and Comments
 The background and introduction are  thorough and proide good motiviation to the importance of the study.

The abstract could use help with clarity, perhaps reframe the study some: 
The outcome of interest is the change in the proportion of residents vacccinated over time from September - February. The predictor of interest is the COVID-19 vaccine hesitancy program intervention. It would be helpful also to help contextualize the three zip codes of interest.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The hypotheis and questions can be more explicitly stated in the manuscript. For example, in the objectives section you mention data cleaning methods. The objective should be stated more succintly and explicitly:

We aim to assess any trend in vaccine uptake among our targeted community during the time of health literacy programs (September - February 2022)

Your hypothesis is a little unclear to me too. For example, you are looking for an increase in the number of Albany citizens who receive the vaccine in the time period. However, this is a value that can't decrease. So are you looking for an increase in the trend (interrupted tiem series), and increase in the daily rate among unvaccinated (where your denominator will change daily), an increase in the zip codes of interest versus other zip codes where targeted interventions werent applied? The objective/hypothesis might need to be fleshed out a bit more.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data structure is well explained and sources described. It is meddled in the with the importation and cleaning. I would rather encourage the author to separate the findings a bit.

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Data cleaning process is well described. Similar to comment above, it is interwoven with description of the data sources. The manuscript would benefit from separate paragraphs discussing these elements and some editorial review (e.g., ensuring run-on sentences are shortened, fragments removed). Some mentions of "data were cleaned" without explaining exaclty how. An example would be:

"For this analysis, we decided to exclude purina body condition scores as it was implemented later in the study, and most dogs do not have that information. The remaining columns were cleaned to make the data more consistent and in the format required for statistical analysis. "

could be streamlined to: 
"We excluded condition scores because they were not consistently collected and had high numbers of missing values. We recoded XXXX variables from XXX to XXX to facilitate easier statistical analysis.""

### Summary assessment

* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

### Summary assessment
* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
Discussion not present at time of review.

### Summary assessment

* strong, complete and clear discussion

## Further comments

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

### Summary assessment

* strong level of thorougness

## Further comments
---

### 2023-04-27 products/Manuscript.qmd
NG Review Added

@@
 # Summary
+
 This study aims to conduct a secondary analysis of residents of the black community in Albany, GA that have received at least one dose of the COVID-19 vaccine from September 2022 through February 2023 and live in the one of the three zip codes 31701, 31705 and 31707.
@@
 ## General Background
+
 Back in March 2020, COVID-19 greatly impacted the citizens of the Albany, GA.
@@
 ### Zip code 31701
+
 The vaccination rate of the the Black population slightly decreased from a rate of 61% to 60.6% \@ref(fig:31701_total_black_vax_rate).
@@
 ### Zip code 31705
+
 The vaccination rate of stayed stable around 48% for the black population in the 31705 zip code \@ref(fig:31701_Age_black_vax).
@@
 ### Zip code 31707
+
 The total vaccination rate of black residents in the 31707 zip code was around 50% from September through February \@ref(fig:31707_total_black_vax_rate).
@@
 ### Zip code 30815
+
 The total vaccination rate for the black population was 44% in the 30815 zip code from September through February \@ref(fig:30815_total_black_vax_rate).
@@
 ### Zip code 30907
+
 The total vaccination rate for the black population was 54% in September and dropped to around 53% in February in the 30907 zip code \@ref(fig:30907_total_black_vax_rate).
@@
 # Discussion
+
 Albany, Georgia is predominately black

### 2023-04-27 reviews/matta_review_ng.qmd
NG Review Added

---
title: Project Review Jacob Matta 
author: Nathan Greenslit
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Manuscript for MADA Project

Name of project author(s): Jacob Matta

Name of project reviewer: Nathan Greenslit

# Specific project content evaluation

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

Background information is provided with relevant sources, however more may be needed in order to properly provide context, and set up the research question and why it is important. It is generally helpful to think of an "hourglass" format. Where you start of broad, and funnel down to your project and question, and then end with another broad statement of how your research applies to real-world scenarios, or adds to gaps research.

For example, a broader introduction that addresses hesitancy programs, vaccination status among black communities, or variance in mortality rates based on race, may aid in setting up this specific study conducted in Albany Georgia. The question of interest is stated, but it would be beneficial to add a sentence or two of why this vaccination status is important to address. Or what the results of this study can provide.

### Summary assessment

-   some contextualization and motivation

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

Question/Objective is clearly stated: *My goal is to compare monthly trends of vaccinations and examine whether there is a noticeable change in vaccination rates among residents who are black during the time of the health literacy programs, which is from September through February.*

For the hypothesis, can you state specifically what "observable change" you are expecting (increase/decrease).

### Summary assessment

-   question/hypotheses fully clear

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

Data structure is described in depth and links to the data sets are provided. Readme files that pertain to the data go in to good detail of how to interpret the different variables, and how they apply to the project.

### Summary assessment

-   source and overall structure of data well explained

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

#### Data Processing:

Data is extensively processed. Both files ran well with no errors. Each step of processing was very well annotated and organized. The final "processed data" is ready for downstream analysis

#### Exploratory Analysis:

Produced plots are informative with well-written annotation. Additionally, the use of an interactive plot is very cool! Alternatives are not discussed, but the overall workflow makes sense for the project/question.

### Summary assessment

-   essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Based on my knowledge, the analysis was applicable to the data/questions. Both a tree and LASSO model were used to assess the primary question. Some initial cleaning was done in the `statistical analysis` code that may be better in its own cleaning code. It may also be beneficial to explore more simple linear models to backup these more complex ML models.

It would be helpful to add a few sentences in this script that summarize what you are doing and why. And add a conclusion sentence where you evaluate the performance of each model.

### Summary assessment

-   strong and reasonable analysis

## Presentation

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Summary paragraphs are provided for each collection of plots and the figures are fairly interpretable. One optional suggestion is to add `+theme_bw()` to the plots. This produces a cleaner and more professional look. It may be beneficial to to rename the variables within a column (especially with the age ranges) by eliminating "\_" and adding "-". This can be done with the code:

`mutate(column_name = recode(column_name, "Old_Name" = "New-Name"))`

Additionally, you can rename your y axis "VAX_RATE" to "Vaccination Rate"

### Summary assessment

-   results are very well presented

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

No discussion or conclusion was supplied.

### Summary assessment

-   major parts of discussion missing or wrong

## Further comments

See comments at end of review.

# Overall project content evaluation

Evaluate overall features of the project by filling in the sections below.

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Great job on structure! The layout makes it very easy to reproduce. Additionally, your code is incredibly organized and well annotated, making it very easy to follow along. Your figures in your results folder are also well labeled. I would add a note about the \``exploratoryanalysis_files` folder. As well as the `processingcode_files` folder.

### Summary assessment

-   well structured

## Documentation

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Reiterating what I said above, your project is very well documented. I would add some more annotations on the statistical scripts to help others track along. I ended up using the tidymodels webpage to guide my annotations here. It can be useful for summarizing each step.

### Summary assessment

-   decently documented with some gaps

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

All code ran with no issues! In the initial README file, it is clearly stated what must be done to reproduce. This file was efficient in summarizing the structure and workflow of the project (I may end up following a similar format!).

### Summary assessment

-   fully reproducible without issues

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Different ML models were utilized for analysis (LASSO and Tree) that assessed the question. Again, it may be good to run a few simple univariate linear models as well. Adding interpretations/plots to your manuscript would be helpful. Analysis did assess the questions, but the results have not been interpreted in the manuscript.

### Summary assessment

-   strong level of thoroughness

# Further comments

When rendering your manuscript, the citations appeared in the paper but bolded with question marks. This happened with mine as well and the problem is that the script is not recognizing your .bibtex citation keys. This also makes it so you have no references at the end of your manuscript. I would double check your key format and see if they match up.

There is also a potential typo in the raw data readme. Both January 16th and January 26th are mentioned.

Overall, a really interesting question and analysis. I think that you have most of the "meat" of the project, and some factors (comments, annotations, explanations) just need to be fleshed out a little more. The biggest gap is the lack of the abstract, discussion, and conclusion. Additional, some more information would be useful in the introduction in order to set up your project. But all previous analysis (project/data description, cleaning, statistical analysis) is really good!

### 2023-04-28 Review_Troha_Aidan.qmd
Project Review

--- 2023-04-27 Hatfield Review.qmd
@@
 ---
-title: Project Review  
-author: Kelly Hatfield
+title: Project Review Template 
+author: AIDAN TROHA
 date: "`r file.mtime(knitr::current_input())`"
@@
 
-Title of project:  COVID-19 Vaccination in Albany, GA
+Title of project:
 
@@
 
-Name of project reviewer: Kelly Hatfield
+Name of project reviewer: Aidan Troha
+
+## Background, Context and Motivation
+How well is the context of the project described?
+Is a comprehensive background, including summary of previous/related work given?
+Is the project well placed into the context of existing work (including proper referencing of existing work).
+Is it clear why the project was undertaken and what new information it hopes to provide?
 
 ### Feedback and Comments
- The background and introduction are  thorough and proide good motiviation to the importance of the study.
 
-The abstract could use help with clarity, perhaps reframe the study some: 
-The outcome of interest is the change in the proportion of residents vacccinated over time from September - February.
-The predictor of interest is the COVID-19 vaccine hesitancy program intervention.
-It would be helpful also to help contextualize the three zip codes of interest.
+I like the scope of your project here. I am interested to know if dealing with a smaller geographic location made finding data to use harder or easier to obtain.
 
@@
 
-The hypotheis and questions can be more explicitly stated in the manuscript.
-For example, in the objectives section you mention data cleaning methods.
-The objective should be stated more succintly and explicitly:
-
-We aim to assess any trend in vaccine uptake among our targeted community during the time of health literacy programs (September - February 2022)
-
-Your hypothesis is a little unclear to me too.
-For example, you are looking for an increase in the number of Albany citizens who receive the vaccine in the time period.
-However, this is a value that can't decrease.
-So are you looking for an increase in the trend (interrupted tiem series), and increase in the daily rate among unvaccinated (where your denominator will change daily), an increase in the zip codes of interest versus other zip codes where targeted interventions werent applied?
-The objective/hypothesis might need to be fleshed out a bit more.
+As with the introduction, the scope of your questioning is clearly defined. Are there any other covariates you looked into that you did not address in this section?
 
@@
 
-The data structure is well explained and sources described. It is meddled in the with the importation and cleaning. I would rather encourage the author to separate the findings a bit.
+Adding a link to info about the data source is a great way of providing necessary info without taking away from your analyses. I like this!
 
@@
 
-Data cleaning process is well described.
-Similar to comment above, it is interwoven with description of the data sources.
-The manuscript would benefit from separate paragraphs discussing these elements and some editorial review (e.g., ensuring run-on sentences are shortened, fragments removed).
-Some mentions of "data were cleaned" without explaining exaclty how.
-An example would be:
-
-"For this analysis, we decided to exclude purina body condition scores as it was implemented later in the study, and most dogs do not have that information.
-The remaining columns were cleaned to make the data more consistent and in the format required for statistical analysis.
-"
-
-could be streamlined to: 
-"We excluded condition scores because they were not consistently collected and had high numbers of missing values.
-We recoded XXXX variables from XXX to XXX to facilitate easier statistical analysis.""
+The analyses employed make rational sense, and the steps seem easy enough to follow. I counldn't wrap my head around the structure of the README.md files, but that is just a style preference from me.
 
 ### Summary assessment
-
 * some weaknesses in wrangling and exploratory component
@@
 
+From what I can tell, the analyses are done appropriately and correctly. Good implementation of machine learning methods.
+
 ### Summary assessment
-* defensible but not optimal analysis
+* strong and reasonable analysis
 
@@
 
+There are plenty of figures to comment on. As a semantic argument, I wonder if you could find a way to collapse some of these figures into a more concise, paneled figure.
+
 ### Summary assessment
-* results are presented ok, with room for improvement
+* results are very well presented
 
@@
 ### Feedback and Comments
-Discussion not present at time of review.
+
+As of review: no discussion or conclusion statements.
 
 ### Summary assessment
-
-* strong, complete and clear discussion
+* major parts of discussion missing or wrong
 
 ## Further comments
+
+Great work so far!
 
@@
 
+As the default structure, this is sufficient. Personally, I like a more intuitive structure wherein the steps of the analyses follow the order of the manuscript.
+
 ### Summary assessment
-
-* well structured
+* mostly clear, but some confusing parts (e.g. useless files, things in the wrong folders)
 
@@
 
+As per the previous response, it makes it difficult to track down where each file is located.
+
 ### Summary assessment
-
-* fully and well documented
+* decently documented with some gaps
 
@@
 
+In the same vein as the previous, it can be difficult to fix an issue if you can't find the right file. In that, I mean it helps to have clearer readme files and intuitive file structure.
+
 ### Summary assessment
-
-* fully reproducible without issues
+* small parts not reproducible or required manual intervention
 
@@
 
+It is clear that you have put a lot of thought into this project as at each step your methods well argued and appropriate.
+
 ### Summary assessment
-
 * strong level of thorougness
@@
 ## Further comments
----
+
+Great job! I can tell that you have put a lot of work into this project!

### 2023-05-05 code/analysis_code/statisticalanalysis.qmd
Auto stash before merge of "master" and "origin/master"

@@
 library(glmnet) #for ridge regression 
+library(here)
 ```
@@
 
+```
+
+### Bivariate Analysis Through Linear Regression
+
+```{r}
+# fit linear model 
+lm_mod <- linear_reg()
+
+#fit model with total rate difference as outcome and total vax rate as predictor
+total_rate_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ TOTAL_VAX_RATE,
+      data = combined) 
+total_rate <- tidy(total_rate_fit) 
+print(total_rate)
+
+#Save Linear Regression Output Table
+table1= here("results", "table1.RDS")
+saveRDS(total_rate, file = table1) 
+```
+
+```{r}
+#fit model with total rate difference as outcome and male vax rate as predictor
+Male_rate_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ Male_VAX_RATE,
+      data = combined) 
+Male_rate <- tidy(Male_rate_fit) 
+print(Male_rate)
+
+#Save Linear Regression Output Table
+table2= here("results", "table2.RDS")
+saveRDS(Male_rate, file = table2) 
+```
+
+```{r}
+#fit model with total rate difference as outcome and female vax rate as predictor
+Female_rate_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ Female_VAX_RATE,
+      data = combined) 
+Female_rate <- tidy(Female_rate_fit) 
+print(Female_rate)
+
+#Save Linear Regression Output Table
+table3= here("results", "table3.RDS")
+saveRDS(Female_rate, file = table3) 
+```
+
+```{r}
+#fit model with total rate difference as outcome and Age_05_09_VAX_RATE as predictor
+Age_05_09_rate_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ Age_05_09_VAX_RATE,
+      data = combined) 
+Age_05_09_rate <- tidy(Age_05_09_rate_fit) 
+print(Age_05_09_rate)
+
+#Save Linear Regression Output Table
+table4= here("results", "table4.RDS")
+saveRDS(Age_05_09_rate, file = table4) 
+```
+
+```{r}
+#fit model with total rate difference as outcome and Age_10_17_VAX_RATE as predictor
+Age_10_17_rate_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ Age_10_17_VAX_RATE,
+      data = combined) 
+Age_10_17_rate <- tidy(Age_10_17_rate_fit) 
+print(Age_10_17_rate)
+
+#Save Linear Regression Output Table
+table5= here("results", "table5.RDS")
+saveRDS(Age_10_17_rate, file = table5) 
+```
+
+```{r}
+#fit model with total rate difference as outcome and Age_18_44_VAX_ as predictor
+Age_18_44_rate_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ Age_18_44_VAX_RATE,
+      data = combined) 
+Age_18_44_rate <- tidy(Age_18_44_rate_fit) 
+print(Age_18_44_rate)
+
+#Save Linear Regression Output Table
+table6= here("results", "table6.RDS")
+saveRDS(Age_18_44_rate, file = table6) 
+```
+
+```{r}
+#fit model with total rate difference as outcome and Age_45_64_VAX_RATE as predictor
+Age_45_64_rate_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ Age_45_64_VAX_RATE,
+      data = combined) 
+Age_45_64_rate <- tidy(Age_45_64_rate_fit) 
+print(Age_45_64_rate)
+
+#Save Linear Regression Output Table
+table7= here("results", "table7.RDS")
+saveRDS(Age_45_64_rate, file = table7) 
+```
+
+```{r}
+#fit model with total rate difference as outcome and Age_65Plus_VAX_RATE as predictor
+Age_65Plus_rate_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ Age_65Plus_VAX_RATE,
+      data = combined) 
+Age_65Plus_rate <- tidy(Age_65Plus_rate_fit) 
+print(Age_65Plus_rate)
+
+#Save Linear Regression Output Table
+table8= here("results", "table8.RDS")
+saveRDS(Age_65Plus_rate, file = table8) 
+```
+
+```{r}
+#fit model with total rate difference as outcome and Interventione as predictor
+Intervention_fit <- lm_mod %>% 
+  fit( TOTAL_RATE_DIFF ~ Intervention,
+      data = combined) 
+Intervention <- tidy(Intervention_fit) 
+print(Intervention)
+
+#Save Linear Regression Output Table
+table9= here("results", "table9.RDS")
+saveRDS(Intervention, file = table9) 
 ```

### 2023-05-05 code/analysis_code/statisticalanalysis.qmd
Discussion

@@
 # Create a scatter plot with ggplot for residuals 
-ggplot(residuals_df, aes(x = Predictions , y = Residuals)) +
+Res1 = ggplot(residuals_df, aes(x = Predictions , y = Residuals)) +
   geom_point() +
@@
   ggtitle("Residual Plot")
+Res1
+
+#Save Linear Regression Output Table
+table10= here("results", "table10.RDS")
+saveRDS(final_fit_metrics, file = table10)
+
+#save figure of predictions vs outcome values 
+ggsave(filename = here("results","Predictions_vs_OutcomeValues.png"), plot=plot1)
+
+#save figure of residuals 
+ggsave(filename = here("results","Residuals_Plot.png"), plot=Res1)
 
@@
 ```
+
+### Evaluate the final fit for the Lasso model
+
+```{r}
+final_fit2 <- 
+  final_lasso_wf %>%
+  last_fit(data_split)
+
+final_fit_metrics2 <- final_fit2 %>%
+  collect_metrics()
+final_fit_metrics2
+
+#creating a tibble for predicted values of model
+predictions2 = final_fit2  %>%
+  collect_predictions
+
+#plotting model predictions vs actual outcome 
+plot2 = final_fit2 %>%
+  collect_predictions  %>%
+    ggplot(aes(x = .pred, y = TOTAL_RATE_DIFF)) +
+  geom_smooth(method = "lm", se = FALSE) +
+  xlab("Predicted Probability") +
+  ylab("Actual Class") +
+  ggtitle("Predicted Values vs Actual Outcome Values")
+plot2
+
+#plotting the residuals 
+residuals2 = predictions$TOTAL_RATE_DIFF - predictions$.pred
+
+#creating a residual data frame 
+residuals_df2 = data.frame(TOTAL_RATE_DIFF = predictions$TOTAL_RATE_DIFF, Predictions = predictions$.pred, Residuals = residuals)
+
+# Create a scatter plot with ggplot for residuals 
+Res2 = ggplot(residuals_df2, aes(x = Predictions , y = Residuals)) +
+  geom_point() +
+  xlab("Predicted Values") +
+  ylab("Residuals") +
+  ggtitle("Residual Plot")
+
+#Save Linear Regression Output Table
+table11= here("results", "table11.RDS")
+saveRDS(final_fit_metrics2, file = table11)
+
+#save figure of predictions vs outcome values 
+ggsave(filename = here("results","Predictions_vs_OutcomeValues.png"), plot=plot2)
+
+#save figure of residuals 
+ggsave(filename = here("results","Residuals_Plot.png"), plot=Res2)
+```
+
+## Fitting for random forest model
+
+### Building a model
+
+```{r}
+rf_mod =
+  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
+  set_engine("ranger") %>% 
+  set_mode("regression")
+```
+
+## Creating a recipe
+
+```{r}
+recipe2 <- recipe(TOTAL_RATE_DIFF ~ ., data = train_data)
+```
+
+### Creating a workflow
+
+```{r}
+rf_workflow =
+  workflow() %>%
+  add_model(rf_mod)  %>%
+  add_recipe(recipe2)
+```
+
+### Train and tune the model
+
+```{r}
+rf_res <- 
+  rf_workflow %>% 
+  tune_grid(CV_fold)
+```
+
+### Collect metrics and picking the best model
+
+```{r}
+rf_res %>% show_best("rmse")
+
+best_rf = rf_res %>%
+  select_best("rmse")
+best_rf
+```
+
+### Finalizing random forest model
+
+```{r}
+final_rf_wf <- 
+  rf_workflow %>% 
+  finalize_workflow(best_rf)
+
+final_rf_wf
+```
+
+### Evaluate the final fit for the random forest model
+
+```{r}
+final_fit3 <- 
+  final_rf_wf %>%
+  last_fit(data_split)
+
+final_fit_metrics3 <- final_fit3 %>%
+  collect_metrics()
+final_fit_metrics3
+
+#creating a tibble for predicted values of model
+predictions3 = final_fit3  %>%
+  collect_predictions
+
+#plotting model predictions vs actual outcome 
+plot3 = final_fit3 %>%
+  collect_predictions  %>%
+    ggplot(aes(x = .pred, y = TOTAL_RATE_DIFF)) +
+  geom_smooth(method = "lm", se = FALSE) +
+  xlab("Predicted Probability") +
+  ylab("Actual Class") +
+  ggtitle("Predicted Values vs Actual Outcome Values")
+plot3
+
+#plotting the residuals 
+residuals3 = predictions$TOTAL_RATE_DIFF - predictions$.pred
+
+#creating a residual data frame 
+residuals_df3 = data.frame(TOTAL_RATE_DIFF = predictions$TOTAL_RATE_DIFF, Predictions = predictions$.pred, Residuals = residuals)
+
+# Create a scatter plot with ggplot for residuals 
+Res3 = ggplot(residuals_df3, aes(x = Predictions , y = Residuals)) +
+  geom_point() +
+  xlab("Predicted Values") +
+  ylab("Residuals") +
+  ggtitle("Residual Plot")
+
+#Save Linear Regression Output Table
+table12 = here("results", "table12.RDS")
+saveRDS(final_fit_metrics3, file = table12)
+
+#save figure of predictions vs outcome values 
+ggsave(filename = here("results","Predictions_vs_OutcomeValues.png"), plot=plot3)
+
+#save figure of residuals 
+ggsave(filename = here("results","Residuals_Plot.png"), plot=Res3)
+```

### 2023-05-05 products/Manuscript.qmd
Discussion

--- 2023-04-24 products/Manuscript.qmd
@@
 According to U.S Census Bureau, 74.5% of the total population of the City of Albany is made up by residents that are black.
-In a study that examined the health outcomes of residents who tested positive for COVID-19 in Albany, GA it was found that 83.3% of 710 patients that were hospitalized for COVID-19 were black[@].
+In a study that examined the health outcomes of residents who tested positive for COVID-19 in Albany, GA it was found that 83.3% of 710 patients that were hospitalized for COVID-19 were black[@grahamrapierHowSmallGeorgia2020].
 Although there was no differences in hospital mortality found between residents that were black compared to residents that were not black but there was a much higher percentage of hospitalizations due to COVID-19 for residents who were black.
 
-In July 2022, Albany was awarded a 3.9 million dollar grant from the Health and Human Services and the Office of Minority Health in an effort to fund COVID-19 vaccine hesitancy and health literacy programs[].
+In July 2022, Albany was awarded a 3.9 million dollar grant from the Health and Human Services and the Office of Minority Health in an effort to fund COVID-19 vaccine hesitancy and health literacy programs[@thealbanyheraldCityAlbanyReceives].
 The Economic Evaluation Research Group (EERG) out of the University of Georgia College of Public Health Department of Health Policy and Management led by Dr.
@@
 
-The data is downloaded from the Georgia Department of Public Health Distribution Dashboard ([citation for dashboard pasted here](https://experience.arcgis.com/experience/3d8eea39f5c1443db1743a4cb8948a9c)) which is a publicly available dashboard that displays the current vaccination administered and the up-to-date count of Georgia residents that have been vaccinated.
+The data is downloaded from the Georgia Department of Public Health Distribution Dashboard (@georgiadepartmentofpublichealthGADPHVaccine) which is a publicly available dashboard that displays the current vaccination administered and the up-to-date count of Georgia residents that have been vaccinated.
 
 The data of interest from this dashboard is the data containing the count of residents with at least one dose of the COVID-19 vaccine by census tract by demographics.
-You can download the data from [the census tract demographic data](https://www.arcgis.com/home/item.html?id=17a8a4af79e5453a9132aaa29fb4ad22) and a description of the data by census tract demographic can be found on the separate dashboard found [here](https://insights.arcgis.com/#/view/08d2c1f995524053b1f9669dd0b0b995).
+You can download the data from the Georgia Department of Public Health Distribution Dashboard and a description of the data by census tract demographic can be found on the
+COVID-19 Vaccinations Among Georgia Residents with At Least One Dose By Census Tract Dashboard (@georgiadepartmentofpublichealthCOVID19VaccinationsGeorgia).
 
@@
 
-## Vaccination Trends of the Black Community for Each Zip Code
+## Descriptive Analysis: Vaccination Trends of the Black Community for Each Zip Code
 
-The was no noticiable increase of vaccination rate for
+### Zip code 31701
+The vaccination rate of the the Black population slightly decreased from a rate of 61% to 60.6% (@fig-1).
+The vaccination rate for female black residents slightly decreased from 65% to 64% while the male black resident population's vaccination rate increased slightly at 53% to 57%.
+It is important to note the difference between the vaccination rates of female and male black residents, which is around .1 from September to February (@fig-2).
+When the black population was stratified by age, the rates were the highest for the ages of 65 and older at 89%.
+Black residents aged 45-64 had a steady rate of 70%, black residents aged 18-44 had a steady rate of 65%, blacked residents aged 10-17 had a steady rate of 40% and black residents aged 5-9 had a steady rate of 12% (@fig-3).
 
 ```{r}
-knitr::include_graphics("../results/plot_31701_total_black_vax_rate.png")
+#| label: fig-1
+#| fig-cap: "Vaccination Rates for Black Population in the 31701 Zipcode"
+knitr::include_graphics("../results/31701_total_black_vax_rate.png")
 ```
@@
 ```{r}
-knitr::include_graphics("../results/31705_total_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/31707_total_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/30815_total_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/30907_total_black_vax.png")
-```
-
-## Vaccination Trends of the Black Community for Each Zip Code by Gender
-```{r}
+#| label: fig-2
+#| fig-cap: "Vaccination Rates for Black Population by Gender in the 31701 Zipcode"
 knitr::include_graphics("../results/31701_Gender_black_vax.png")
@@
 ```{r}
+#| label: fig-3
+#| fig-cap: "Vaccination Rates for Black Population by Age in the 31701 Zipcode"
+knitr::include_graphics("../results/31701_Age_black_vax.png")
+```
+
+### Zip code 31705
+
+The vaccination rate of stayed stable around 48% for the black population in the 31705 zip code (@fig-4).
+The vaccination rate of female black residents dropped from 51.1% to 49.7% from September to December, but remained stable close to 50% until February.
+For male black residents the vaccination rate hovered around 44% (@fig-5).
+After stratifying by age the vaccination rates for the 65+, 44-64, 18-44, 10-17 and 5-9 ages groups were as follows: 85%, 68%, 43%, 35% and 12% (@fig-6).
+
+```{r}
+#| label: fig-4
+#| fig-cap: "Vaccination Rates for Black Population in the 31705 Zipcode"
+knitr::include_graphics("../results/31705_total_black_vax_rate.png")
+```
+
+```{r}
+#| label: fig-5
+#| fig-cap: "Vaccination Rates for Black Population by Gender in the 31705 Zipcode"
 knitr::include_graphics("../results/31705_Gender_black_vax.png")
@@
 ```{r}
+#| label: fig-6
+#| fig-cap: "Vaccination Rates for Black Population by Age in the 31705 Zipcode"
+knitr::include_graphics("../results/31705_Age_black_vax.png")
+```
+
+### Zip code 31707
+
+The total vaccination rate of black residents in the 31707 zip code was around 50% from September through February (@fig-7).
+The vaccination rates for the female and male black population were found to be steady around 54% and 44% (@fig-8).
+The vaccination rates by the age groups 65+, 44-64, 18-44, 10-17 and 5-9 where found to be 98%, 85%, 52%, 37%, and 17% (@fig-9).
+
+```{r}
+#| label: fig-7
+#| fig-cap: "Vaccination Rates for Black Population in the 31707 Zipcode"
+knitr::include_graphics("../results/31707_total_black_vax_rate.png")
+```
+
+```{r}
+#| label: fig-8
+#| fig-cap: "Vaccination Rates for Black Population by Gender in the 31707 Zipcode"
 knitr::include_graphics("../results/31707_Gender_black_vax.png")
@@
 ```{r}
+#| label: fig-9
+#| fig-cap: "Vaccination Rates for Black Population by Age in the 31707 Zipcode"
+knitr::include_graphics("../results/31707_Age_black_vax.png")
+```
+
+### Zip code 30815
+
+The total vaccination rate for the black population was 44% in the 30815 zip code from September through February (@fig-10).
+The vaccination rates for black female and male residents were 49% and 38% (@fig-11).
+The vaccination rates by the age groups 65+, 44-64, 18-44, 10-17 and 5-9 where found to be 78%, 60%, 38%, 31%, and 8% (@fig-12).
+
+```{r}
+#| label: fig-10
+#| fig-cap: "Vaccination Rates for Black Population in the 30815 Zipcode"
+knitr::include_graphics("../results/30815_total_black_vax_rate.png")
+```
+
+```{r}
+#| label: fig-11
+#| fig-cap: "Vaccination Rates for Black Population by Gender in the 30815 Zipcode"
 knitr::include_graphics("../results/30815_Gender_black_vax.png")
@@
 ```{r}
+#| label: fig-12
+#| fig-cap: "Vaccination Rates for Black Population by Age in the 30815 Zipcode"
+knitr::include_graphics("../results/30815_Age_black_vax.png")
+```
+
+### Zip code 30907
+
+The total vaccination rate for the black population was 54% in September and dropped to around 53% in February in the 30907 zip code (@fig-13).
+The vaccination rates for black female and male residents were 60% and 45% (@fig-14).
+The vaccination rates by the age groups 65+, 44-64, 18-44, 10-17 and 5-9 where found to be 95%, 64%, 63%, 33%, and 11% (@fig-15).
+
+```{r}
+#| label: fig-13
+#| fig-cap: "Vaccination Rates for Black Population in the 30907 Zipcode"
+knitr::include_graphics("../results/30907_total_black_vax_rate.png")
+```
+
+```{r}
+#| label: fig-14
+#| fig-cap: "Vaccination Rates for Black Population by Gender in the 30907 Zipcode"
 knitr::include_graphics("../results/30907_Gender_black_vax.png")
@@
 
-## Vaccination Trends of the Black Community for Each Zip Code by Gender
 ```{r}
-knitr::include_graphics("../results/31701_Age_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/31705_Age_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/31707_Age_black_vax.png")
-```
-
-```{r}
-knitr::include_graphics("../results/30815_Age_black_vax.png")
-```
-
-```{r}
+#| label: fig-15
+#| fig-cap: "Vaccination Rates for Black Population by Age in the 30907 Zipcode"
 knitr::include_graphics("../results/30907_Age_black_vax.png")
@@
 
-## LASSO Results Visual 
+## Statistical Analysis
+
+### Bivariate Analysis Through Simple Linear Regression
+
+#### Demagraphic Vaccination Rates as Predictors
+The following bivariate simple linear regression models were performed on each predictor variable of interest and the main outcome of interest which is total vaccination rate difference by month among the black population in each zip code.
+The following predictors produced results that were found not to be statistically significant predictors at the 0.05 significance level: Total vaccination rate by month among the black population, the male vaccination rate by month among the black population, the female vaccination rate by month among the black population, age ranges of 5-9, 10-17, 18-44, 45-64, 65+ vaccination rate by month among the black population.
+Of these predictors, based on p-values the best predictor of total vaccination rate difference by month was the vaccination rate for the age range of 45-65 which had a p-value of 0.52 (@tbl-1).
+The coefficient estimate for the vaccination rate for the age range of 45-65 was 0.004 which indicates for every one-unit increase of vaccination rate for the age range of 45-65 the total vaccination rate difference by month among the black population increased by 0.004 units (@tbl-1).
+
 ```{r}
-knitr::include_graphics("../results/lasso_grid_plot.png")
+#| label: tbl-1
+#| fig-cap: "Simple Linear Regression Output for the Vaccination Rate of the Age Range of 45-64 Among the Black Population as a Predictor"
+readRDS(file = here("results", "table7.RDS")) 
 ```
 
-## Final LASSO Results Variable Importance 
+#### Intervention as a Predictor
+The Intervention variable is a binary predictor where "1" indicates a zip code hosting the COVID-19 Health Literacy and Vaccine Hesitancy Program and "0" indicates a zip code that is not hosting a program.
+When this was fitted in a simple linear regression model with total vaccination rate difference by month among the black population in each zip code as the outcome, the resulting coefficient estimate was 0.068 and a p-value of 0.55 (@tbl-9).
+The p-value of 0.55 was not significant at the 0.05 level, however, it was among the lowest p-values when compared to the p-values for the demographic vaccination rate predictor variables.
+In addition, the coefficient estimate of 0.068 is the largest positive coefficient among all other predictors, indicating the total vaccination rate difference by month among the black population increases by 0.068 or 6.8% in zip codes with an intervention program.
+Although this result is not supported by the p-value when evaluated for significance at the 0.05 level, it is notable to observe an increase in total vaccination rate by month when intervention is the only predictor being controlled for.
+
 ```{r}
-knitr::include_graphics("../results/final_lasso_plot.png")
+#| label: tbl-9
+#| fig-cap: "Simple Linear Regression Output for Intervention as a Predictor "
+readRDS(file = here("results", "table9.RDS")) 
 ```
 
-# References
+## Full Model Fitting and Evalutation
+
+### Tree Model 
+After running a Regression Tree Model using all the predictors and the main outcome of total vaccination rate difference by month among the black population in each zip code the best fit model was found to have an RMSE value of 0.37584439 and an R-squared value of 0.06730276 (@tbl-10).
+
+```{r}
+#| label: tbl-10 
+#| fig-cap: "Best Fit Tree Model Metrics"
+readRDS(file = here("results", "table10.RDS")) 
+```
+
+### Lasso Model 
+The Lasso Regression Model produced a best fit regression model that had an RMSE value of 0.30444322 and a R-squared of 0.09234508 (@tbl-11).
+
+```{r}
+#| label: tbl-11
+#| fig-cap: "Best Fit LASSO Metrics"
+readRDS(file = here("results", "table11.RDS")) 
+```
+
+### Random Forrest
+The Random Forrest Regression model for the full model produced a best fit model with an RMSE metric of 0.3716112 and an R-squared of 0.2279387 (@tbl-12).
+
+```{r}
+#| label: tbl-12
+#| fig-cap: "Best Fit Random Forrest Metrics"
+readRDS(file = here("results", "table12.RDS")) 
+```
 
 # Discussion
+Since July of 2022, The Health and Human Services and the Office of Minority Health have funded COVID-19 vaccine hesitancy and health literacy prgrams in Albany, GA that have been in action since October of 2022 [@thealbanyheraldCityAlbanyReceives].
+This analysis has produced a data set from the Georgia Department of Public Health COVID-19 vaccination data that contains vaccination rates by demographic and total vaccination rate in zip codes that the COVID-19 vaccine hesitancy and health literacy programs were enrolled compared to a two control zip codes that were matched via SRI and and SVI.
+The intention of this analysis was to examine whether the zip codes with the COVID-19 vaccination and health literacy intervention had a notable increase of vaccination rates from the months of September 2022 through February 2023.
+
+Descriptive statistics were used to examine the vaccination rates among the zip codes of 31701, 31705 and 31707 by month from September 2022 through February 2023. No notable increase were observed
+
+# Limitations 
+# References

### 2023-05-05 reviews/Hatfield Review.qmd
Discussion

--- 2023-04-27 Hatfield Review.qmd
@@
 ### Feedback and Comments
- The background and introduction are  thorough and proide good motiviation to the importance of the study.
+ The background and introduction are  thorough and provide good motiviation to the importance of the study.
 
-The abstract could use help with clarity, perhaps reframe the study some: 
+The abstract could use help with clarity, perhaps even a reframe the study some: 
 The outcome of interest is the change in the proportion of residents vacccinated over time from September - February.
@@
 
-The hypotheis and questions can be more explicitly stated in the manuscript.
-For example, in the objectives section you mention data cleaning methods.
-The objective should be stated more succintly and explicitly:
+The hypothesis and questions can be more explicitly stated in the manuscript. For example, in the objectives section you mention data cleaning methods. The objective should be stated more succinctly :
 
@@
 However, this is a value that can't decrease.
-So are you looking for an increase in the trend (interrupted tiem series), and increase in the daily rate among unvaccinated (where your denominator will change daily), an increase in the zip codes of interest versus other zip codes where targeted interventions werent applied?
+So are you looking for an increase in the trend (interrupted tiem series), and increase in the daily rate among unvaccinated (where your denominator will change daily), an increase in the zip codes of interest versus other zip codes where targeted interventions weren't applied?
 The objective/hypothesis might need to be fleshed out a bit more.
@@
 ### Summary assessment
-* question/hypotheses fully clear
+* question/hypotheses somewhat explained
 
@@
 
-The data structure is well explained and sources described. It is meddled in the with the importation and cleaning. I would rather encourage the author to separate the findings a bit.
+The data structure is well explained and sources described.
+However, the author presents the findings as percents but states that the data source provides counts.
+He does not provide details on where the denominators for calculating percents come from.
+I see some population data in the processed dataset so I would encourage the author to describe the sources of that data too.
 
 ### Summary assessment
-* source and overall structure of data well explained
+* source and overall structure of data somewhat explained
 
@@
 
-Data cleaning process is well described.
-Similar to comment above, it is interwoven with description of the data sources.
-The manuscript would benefit from separate paragraphs discussing these elements and some editorial review (e.g., ensuring run-on sentences are shortened, fragments removed).
-Some mentions of "data were cleaned" without explaining exaclty how.
-An example would be:
+Data processing is explained.
+However, I would be curious as to some of the data points.
+For example, the authors describe a decrease in vaccination rates over time.
+First, I would note that the decreases are small (<1%).
+Second, it is unclear how the decreases could occur (i.e., once a resident is vaccinated, they can't unvaccinated themself).
+This could be due to variation in the data or other factors, but should be addressed.
 
-"For this analysis, we decided to exclude purina body condition scores as it was implemented later in the study, and most dogs do not have that information.
-The remaining columns were cleaned to make the data more consistent and in the format required for statistical analysis.
-"
-
-could be streamlined to: 
-"We excluded condition scores because they were not consistently collected and had high numbers of missing values.
-We recoded XXXX variables from XXX to XXX to facilitate easier statistical analysis.""
+The author also does the analyses over time in separate figures for each zip code.
+It would be helpful to have these on the same graph, perhaps grouped by color for if they are an intervention zip code or control zip code.
 
@@
 
+ There is limited analysis description in the manuscript. Per the analysis code file it appears there is detailed analysis with cross-fold validation. However, this is not described in the text.
+
 ### Summary assessment
-* defensible but not optimal analysis
+* wrong/inadequate analysis
 
@@
 ### Feedback and Comments
+The line graphs showing trends over time in the different zip codes are helpful.
+However, they could be improved with some additioanl work on the visualizations.
+In the manuscript the results for the advanced analysis are limited and not well described.
 
 ### Summary assessment
-* results are presented ok, with room for improvement
+* results are poorly presented, hard to understand, poor quality
 
@@
 
-* strong, complete and clear discussion
+* major parts of discussion missing or wrong
 
@@
 ## Structure
-Is the project well structured?
-Are files in well labeled folders?
-Do files have reasonable names?
-Are all "junk" files not needed for analysis/reproduction removed?
-By just looking at files and folders, can you get an idea of how things fit together?
 
 ### Feedback and Comments
+The project is well organized and the quarto files with code are well documented and easy to file.
+It is unclear to me if processingcode and processingcode2 quarto files need to reamin.
+I would update the readme section in that folder to make it clear which file does what part you are describing.
 
@@
 ### Feedback and Comments
+The project is well documented. Some of the rationale for some of the caclulations is missing in some of the analysis files, but it is easy to understand the analytic parts.
 
@@
 ### Feedback and Comments
+The analysis appears reproducible.
+However, some parts do appear to typed in (e.g., calculation of some rates) where it may be better to program the calculations to ensure typos don't affect your results.
 
@@
 
+The analysis seems thorough in the quarto files. However, from the manuscript file it is unclear.
+
 ### Summary assessment
 
-* strong level of thorougness
+* decent level of thoroughness
 
 ## Further comments
+This study could use a bit more thought and effort into the analysis and interpretation.
+I think the quarto files show a lot of really good exploratory analyses.
+The manuscript could use work to better describe the work and the findings.
+
+I don't think the references are formatted correctly, as they are also not populating in the manuscript. 
 ---

### 2023-05-06 code/analysis_code/statisticalanalysis.qmd
Edits

@@
 ```{r}
-#fit model with total rate difference as outcome and Interventione as predictor
+#fit model with total rate difference as outcome and Intervention as a predictor
 Intervention_fit <- lm_mod %>% 
@@
 #save figure of predictions vs outcome values 
-ggsave(filename = here("results","Predictions_vs_OutcomeValues.png"), plot=plot1)
+ggsave(filename = here("results","Predictions_vs_OutcomeValues1.png"), plot=plot1)
 
 #save figure of residuals 
-ggsave(filename = here("results","Residuals_Plot.png"), plot=Res1)
+ggsave(filename = here("results","Residuals_Plot1.png"), plot=Res1)
 
@@
 #save figure of predictions vs outcome values 
-ggsave(filename = here("results","Predictions_vs_OutcomeValues.png"), plot=plot2)
+ggsave(filename = here("results","Predictions_vs_OutcomeValues2.png"), plot=plot2)
 
 #save figure of residuals 
-ggsave(filename = here("results","Residuals_Plot.png"), plot=Res2)
+ggsave(filename = here("results","Residuals_Plot2.png"), plot=Res2)
 ```
@@
 #save figure of predictions vs outcome values 
-ggsave(filename = here("results","Predictions_vs_OutcomeValues.png"), plot=plot3)
+ggsave(filename = here("results","Predictions_vs_OutcomeValues3.png"), plot=plot3)
 
 #save figure of residuals 
-ggsave(filename = here("results","Residuals_Plot.png"), plot=Res3)
+ggsave(filename = here("results","Residuals_Plot3.png"), plot=Res3)
 ```

### 2023-05-06 products/Manuscript.qmd
Edits

@@
 This study serves an extension of the evaluation being conducted by EERG through examining the vaccination rates of the zip codes where vaccine hesitancy programs are being conducted.
+
+## Intervention Zip Codes vs Control Zip Codes
+The intervention zip codes include 31701, 31705 and 31707 which all are within Albany, GA and Dougherty County.
+These three zip codes will be hosting the COVID-19 vaccine hesitancy and health literacy programs and as a result they will be considered intervention zip codes.
+
+The control zip codes will be both 30907 and 30815 which lie in Richmond County, GA.
+The 30815 zip code was matched with the three intervention zip codes based on the Social Vulnerability Index (SVI) which is in part scaled by a communities vulnerability to disease outbreaks like COVID-19(@agencyfortoxicsubstancesanddiseaseregistryCDCATSDRSocial).
+The 30907 zip code was matched based on Social Risk Index (SRI) which is partly scaled by a communities social risk and ability to recover from disease outbreak (@allianzSocialRiskIndex).
 
@@
 #| label: tbl-9
-#| fig-cap: "Simple Linear Regression Output for Intervention as a Predictor "
+#| fig-cap: "Simple Linear Regression Output for Intervention as a Predictor"
 readRDS(file = here("results", "table9.RDS")) 
@@
 ### Tree Model 
-After running a Regression Tree Model using all the predictors and the main outcome of total vaccination rate difference by month among the black population in each zip code the best fit model was found to have an RMSE value of 0.37584439 and an R-squared value of 0.06730276 (@tbl-10).
+After running a Regression Tree Model using all the predictors and the main outcome of total vaccination rate difference by month among the black population in each zip code the best fit model was found to have an Root Mean Squared Error (RMSE) value of 0.37584439 and an R-squared value of 0.06730276 (@tbl-10).
 
@@
 
-Descriptive statistics were used to examine the vaccination rates among the zip codes of 31701, 31705 and 31707 by month from September 2022 through February 2023. No notable increase were observed
+Descriptive statistics were used to examine the vaccination rates among the zip codes of 31701, 31705 and 31707 by month from September 2022 through February 2023.
+A substantial increase in vaccinate rates was not observed in any of the three zip codes.
+There was a slight increase in COVID-19 vaccination rate of about 1% among the black male population in the 31701 zip code.
+
+The bivariate analysis revealed that all of the demographic variables were not significant predictors of the total vaccination rate difference by month.
+Among the demographic variables, the COVID-19 vaccination rate for the age range of 45-64 proved to be the best predictor according to p-value (@tbl-1).
+According to the linear regression model, for every one-unit increase of the COVID-19 vaccination rate for black members of the community ages 45-64 the monthly difference in COVID-19 vaccination rate for black members of the community increased by 0.004 (@tbl-1).
+Interestingly, the bivariate analysis showed the intervention of the COVID-19 vaccine hesitancy was a relatively good predictor for the monthly difference in COVID-19 vaccination rate for black members of the community.
+The simple linear regression model used monthly difference in COVID-19 vaccination rate for black members of the community as an outcome and intervention as predictor; When the intervention value in the data set was "1" indicating the presence of an intervention in the zip code, the COVID-19 vaccination rate for black members of the community increased by 0.068 (@tbl-9).
+Although the p-value for the Intervention variable was not significant at the 0.05 level (p-value = 0.55), this p-value was among the lower p-values when compared to the other demographic predictors.
+Relatively speaking, Intervention was among the best predictors given the positive magnitude of the coefficient estimate and a low p-value.
+
+A full model was fit using a multiple linear regression model including all predictors and the outcome of monthly difference in COVID-19 vaccination rate for black members of the community.
+The null model produced an RMSE metric of 0.3095648 which indicates the average distance between predicted values of the model and the actual values.
+The Tree Model, produced a model that had an RMSE value of 0.37584439 which is actually slight larger than the null model indicating the predicted values were slightly more off (@tbl-10).
+The LASSO Regression model produced an RMSE of 0.30444322 which is a bit less than the null model (@tbl-11).
+The Random Forest Model produced an RMSE of 0.3716112 which is larger than the LASSO RMSE but lower than the Tree Model RMSE (@tbl-12).
+The best model selection based on RMSE is the LASSO model due to having the lowest RMSE metric.
 
 # Limitations 
-# References
+There are a few issues that resulted from selecting to work with census tract level data for this analysis.
+First, census tracts do not completely align with zip codes.
+In other words, zip codes can share census tracts and this was not accounted for in this analysis.
+The amount of census tracts in each zip code used for this analysis was found through the use of ArcGIS.
+Once the census tracts were obtained via ArcGIS, each census tract was included regardless of the population size and the amount of land they took up in a zip code.
+This becomes an issue because some of the census tracts completely resided in a target zip code while others that were included largly resided in other zip codes.
+By not accounting for proportion of a census tract's population actually resided in the zip code, the vaccination rates and population numbers are not 100% accurate and are most likely larger than expected.
+
+Another limitation is that the final data set used to run the analysis was quite small with few variables (30 Obs and 13 variables).
+This was not accounted for in the beginning of the analysis because the data cleaning and processing was done on fairly large monthly vaccination data sets (1962 Obs and 184 variables).
+However, the variables and observations of interest produced after the cleaning, processing and manipulating were quite small.
+Working with a small data set is not ideal for the machine learning models included in this analysis due to the techniques like data splitting and cross validation.
+Each RMSE metric produced by the machine learning models were relatively low indicating a chance of over fit models.
+To prevent this in the future, producing a larger data set across greater range of months would be helpful for the statistical analysis section of this analysis.
+
+The fact that the machine learning model selection based was primarily based on RMSE value is another limiation of this analysis.
+It is important that RMSE can not primarily determine whether a model is the best fit for a data set.
+Discussing the R-Squared values would have been helpful as it is a metric that indicates the percentage of variability that is explained by the dependent variable.
+
+It is important to note that all of the predictors in the bivariate analysis failed to be statistically significant at the 0.05 level.
+As a result, it is clear that the predictors selected for this analysis are not strong predictors of the outcome of interest.
+New research questions should be analyzed in order to explore additional predictors and other outcome variables to find stronger associations in the data.
+
+# Further Comments 
+It is important to highlight the disparities in the vaccination rates within each zip code.
+Based on demagraphics it appears females are have higher vaccination rates than males and vaccination rate tends to decline as age decreased.
+These findings are not connected to the research question investigated in this analysis.

### 2023-05-06 products/supplement.qmd
Edits

@@
     -   `30907_Age_black_vax.png`: A line graph of African Americans by Age with at least one dose of the COVID-19 vaccine in the 30907 zip code
+
+    -   `table1.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and total vax rate as a predictor
+
+    -   `table2.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and male vax rate as a predictor
+
+    -   `table3.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and female vax rate as a predictor
+
+    -   `table4.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and Age_05_09_VAX_RATE as a predictor
+
+    -   `table5.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and Age_10_17_VAX_RATE as a predictor
+
+    -   `table6.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and Age_18_44_VAX_ as a predictor
+
+    -   `table7.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and Age_45_64_VAX_RATE as a predictor
+
+    -   `table8.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and Age_65Plus_VAX_RATE as a predictor
+
+    -   `table9.RDS`: Simple Linear Regression Output Table for total rate difference as outcome and Intervention as a predictor
+
+    -   `table10.RDS`: Simple Linear Regression Output Table The Tree Model
+
+    -   `table11.RDS`: Simple Linear Regression Output Table The LASSO Model
+
+    -   `table12.RDS`: Simple Linear Regression Output Table The Random Forest Model 
+    -   `Predictions_vs_OutcomeValues1.png`: figure of predictions vs outcome values for the tree model
+
+    -   `Predictions_vs_OutcomeValues2.png`: figure of predictions vs outcome values for the LASSO model
+
+    -   `Predictions_vs_OutcomeValues3.png`: figure of predictions vs outcome values for the Random Forest model
+
+    -   `Residuals_Plot1.png`: figure of residualss for the Tree Model
+
+    -   `Residuals_Plot2.png`: figure of residualss for the LASSO Model
+
+    -   `Residuals_Plot3.png`: figure of residualss for the Random Forest Model
